---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: MOBEAST Preliminary Results
output: 
  html_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE , 
                      warning = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)


library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(GGally)
```

## Load function
```{r}
source(here("functions", "function_ANOVA_nunu.R"))
```


## Load Data
```{r}
meta <- read_delim(here("data", "MOBEAST_metadata.csv"))
DOC <- read_delim(here("data", "raw_data", "DOC", "MOBEAST_NPOC_TN_241010.csv"))
fDOM <- read.csv(here("data", "fDOM", "fDOM_clean.csv"))
```

## Data manipulation 
### Codes of the treatment  
```{r}
treat_code <- tibble(
  treatment = c("A", "K", "C", "R", "Inflow", "Endmember", "Rain"),
  full_name = c("Algae", "Coral", "Control", "Rubble", "Inflow", "Endmember", "Rain")
)
```

### Data clean up 
```{r}
meta <- meta %>% clean_names() %>% drop_na() %>%
  mutate(long_name = str_trim(long_name),
         treatment = as.factor(treatment)) %>% 
  filter(long_name != "Long Name") %>% 
  mutate(date = mdy(date), 
         time = format(strptime(time, format = "%H%M"), "%H:%M"),
         id_number = str_trim(id_number))  
```

```{r}
DOC <- DOC %>% clean_names() 
```

```{r}
fDOM <- fDOM %>% clean_names() %>% 
  rename(id_number = unique_id) %>% 
  select(id_number,coble_a, coble_c, coble_m, bix, hix, m_to_c, fi, lignin)
```

### Joining the data 
```{r}
DOC_join <- left_join(meta, DOC, by = "id_number")
```

```{r}
DOC<- left_join(DOC_join, treat_code, by= "treatment")
```

```{r}
DOC <- DOC %>% select(-treatment) %>% rename(treatment = full_name) %>%
  relocate(treatment,.after = tank_number) %>% 
  mutate(treatment = factor(treatment, levels = c("Inflow", "Control", "Coral", "Algae", "Rubble", "Endmember", "Rain")))
```

```{r}
fDOM <- left_join(meta, fDOM, by="id_number")

```

```{r}
fDOM <- left_join(fDOM, treat_code, by = "treatment")
```

```{r}
fDOM <- fDOM %>% select(-treatment) %>% 
  mutate(treatment = full_name) %>% 
  relocate(treatment, .after = tank_number) %>% 
  mutate(treatment = factor(treatment, levels= c("Inflow", "Control", "Coral", "Algae", "Rubble", "Endmember", "Rain"))) %>% select(-full_name)
```

## Data visualization
```{r}

nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        legend.position = "none",
        panel.background = element_rect(fill = "azure1"))
```

### DOC Visualization
#### Run ANOVA on data and graph plots
```{r}
# apply an ANOVA and Tukey's test that I coded for
anova_nunu(DOC, indep_var = treatment, dep_var = npoc_u_m)

#make a boxplot of the DOC by treatments
DOC_uM <- DOC %>% 
ggplot(aes(treatment, npoc_u_m))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+ #apply the tukey's test results
   labs(x = "Treatment",
       y = "DOC (uM)") + # provide labels 
  nunu_theme #i coded a theme above that will be applied to all graphs
```
```{r}
#same steps applied to chunk above
anova_nunu(DOC, indep_var = treatment, dep_var = npoc_mg_l)

DOC_mgL <- DOC %>% 
ggplot(aes(treatment, npoc_mg_l))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
    labs(title = "Distribution of the DOC per treatment",
       x = "Treatment",
       y = "DOC (uM)",
       fill = "Treatment") +
  nunu_theme
```


```{r}
#same steps applied to chunk above
anova_nunu(DOC, indep_var = treatment, dep_var = tn_mg_l)

TN_mgL<- DOC %>% 
ggplot(aes(treatment, tn_mg_l))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(
       x = "Treatment",
       y = "TN (mg/L)") +
  nunu_theme
```

```{r}
#same steps applied to chunk above
anova_nunu(DOC, indep_var = treatment, dep_var = tn_u_m)

TN_uM <- DOC %>% 
ggplot(aes(treatment, tn_u_m))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(
       x = "Treatment",
       y = "TN (uM)") +
  nunu_theme
```


```{r DOC_summary, fig.width= 12, fig.height=8, fig.path="../outputs/"}
#using patchwork package to arrange the plot as one figure

p_doc_summary <- (DOC_mgL+DOC_uM)/(TN_mgL+TN_uM)
p_doc_summary

#ggsave("p_doc_summary.png", path = here("outputs"))
```


#### DOC Linear Model
```{r}
# the model uses treatment as predictor and the DOC(uM) as dependent variable
# this hypothesize that the treatment will change the DOC concentration in tanks
DOC_lm <- lm(npoc_u_m ~ treatment, data = DOC)
summary(DOC_lm)
```

#### Evaluting the F-test given by the ANOVA
```{r}
Anova(DOC_lm)
```
**Results**:

- F-test is pretty weak (F=1.08) with not significant results (p = 0.38)
- Variation within tank groups cannot be explained by the treatments...

#### Check the "quality" of the data (skewness..etc.)
```{r}
resid_panel(DOC_lm, plots= c("resid", "qq", "lev", "hist"))
```
- Residual plot shows that the data is well distributed on either side of the line and respect the independent distribution assumption.

### fDOM Visualization 


```{r}
#apply personal ANOVA + Tukey's test fucntion
anova_nunu(fDOM, indep_var = treatment, dep_var = coble_a)

# plot a boxplot with the tukey's test results
cobA<- ggplot(fDOM, aes(treatment, coble_a ))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(title = "Distribution of the fDOM index per treatment",
        caption= "cobbleA: Humic, terrestrial ubiquitous FDOM")+ #,
  # adding information on the type of fDOM on the graph
      # x = "Treatment",
      # y = "fDOM",
      # fill = "Treatment") + 
  nunu_theme + # personal theme for these graphs
  theme(plot.caption = element_text(size=11)) # change font of the footnote
```

```{r, include = FALSE}
#same as above
anova_nunu(fDOM, indep_var = treatment, dep_var = coble_c)

cobC<- ggplot(fDOM, aes(treatment, coble_c ))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(caption= "cobleC: Humic-like FDOM, ubiquitous,terrestrial") +
  nunu_theme +
  theme(plot.caption = element_text(size=11))
```

```{r , include = FALSE}
#same as above
anova_nunu(fDOM, indep_var = treatment, dep_var = coble_m)

cobM <-ggplot(fDOM, aes(treatment, coble_m ))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(caption= "cobleM: marine and freshwater autoch prod of new humics") +
  nunu_theme +
  theme(plot.caption = element_text(size=11))
```

```{r, include = FALSE}
#same as above
anova_nunu(fDOM, indep_var = treatment, dep_var = bix)

bix <-ggplot(fDOM, aes(treatment, bix ))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(caption= "BIX: B fluorophore presence, autoch bio activity") +
  nunu_theme +
  theme(plot.caption = element_text(size=11))
```

```{r, include = FALSE}
#same as above
anova_nunu(fDOM, indep_var = treatment, dep_var = hix)

hix <-ggplot(fDOM, aes(treatment, hix ))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(caption= "HIX: humification index, aromatics conc") +
  nunu_theme +
  theme(plot.caption = element_text(size=11))
```

```{r, include = FALSE}
#same as above
anova_nunu(fDOM, indep_var = treatment, dep_var = m_to_c)

m2c <-ggplot(fDOM, aes(treatment, m_to_c ))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(caption= "M:C : marine/terrestrial humics") +
  nunu_theme +
  theme(plot.caption = element_text(size=11))
```

```{r, include = FALSE}
#same as above
anova_nunu(fDOM, indep_var = treatment, dep_var = fi)

fi <-ggplot(fDOM, aes(treatment, fi ))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(caption= "FI: microbL fluvic acid") +
  nunu_theme +
  theme(plot.caption = element_text(size=11))
```

```{r, include = FALSE}
#same as above
anova_nunu(fDOM, indep_var = treatment, dep_var = lignin)

lignin <-ggplot(fDOM, aes(treatment, lignin ))+
  geom_boxplot(aes(fill = treatment))+
  geom_text(data =Tk, aes(x= treatment, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(caption= "lignin: wood DOM ") +
  nunu_theme +
  theme(plot.caption = element_text(size=11))
```


```{r fDOM_summary, fig.height=12, fig.width=8, fig.path= "../outputs/"}
# Using patchwork to consolidated all the plots in one figure
p_fdom_summary <- (cobA+cobC)/(cobM+m2c)/(bix+hix)/(fi+lignin)
p_fdom_summary
#ggsave("p_fdom_summary.png", path = here("outputs"))
```

```{r, include = FALSE}
# plotting all of the relevant fDOM variables in one graph 
# pivoting the dataframe to long format so I can change the fDOM variables as factor and fafcet wrap them

fDOM %>% pivot_longer(cols = coble_a:lignin,
                      names_to = "fdom_type",
                      values_to = "values") %>% 
  ggplot(aes(treatment, values))+
  geom_boxplot(aes(fill = treatment))+
   labs(title = "Distribution of the fDOM index per treatment",
       x = "Treatment",
       y = "fDOM",
       fill = "Treatment") +
  facet_wrap(.~fdom_type, scales = "free")+
  theme_bw() + 
  nunu_theme


```


### fDOM Linear Model

```{r}
cobA_lm <- lm(coble_a ~ treatment, data = fDOM)
```

```{r}
Anova(cobA_lm)
```
```{r}
resid_panel(cobA_lm, plots= c("resid", "qq", "lev", "hist"))
```



## 24h experiment 

```{r}
format <- "%Y-%m-%d %H:%M:%S"

DOC_24h <- DOC %>% 
  mutate(time = paste0(time, ":00"),
         date = as.character(date),
    date_time = as.POSIXct(paste(date, time), format=format),
    date_time = factor(date_time)) %>% 
  filter(date == "2024-06-02" | date == "2024-06-03") %>%

  ggplot(aes(date_time, npoc_mg_l)) +
  geom_boxplot(aes(fill = treatment))+
  facet_wrap(.~treatment, nrow= 1) +
  labs(title= "24h variations in DOC",
       x = "Date-Time",
       y= "DOC (mg/L)",
       caption = "Missing time stamps correspond to 00:00")+
  nunu_theme
```

```{r}

TN_24h <- DOC %>% 
  mutate(time = paste0(time, ":00"),
         date = as.character(date),
    date_time = as.POSIXct(paste(date, time), format=format),
    date_time = factor(date_time)) %>% 
  filter(date == "2024-06-02" | date == "2024-06-03") %>%

  ggplot(aes(date_time, tn_mg_l)) +
  geom_boxplot(aes(fill = treatment))+
  facet_wrap(.~treatment, nrow= 1) +
  labs(title= "24h variations in TN",
       x = "Date-Time",
       y= "TN (mg/L)",
       caption = "Missing time stamps correspond to 00:00")+
  nunu_theme
```



```{r 24h_DOC_TN, fig.width= 12, fig.height=8}
DOC_24h/TN_24h

ggsave("24h_DOC_TN.png", path = here("outputs"))
```






