---
author: 'Keanu Rochette'
title: "MOBEAST DOC V4"
created: '2025-11-10'
edited: 2026-01-30
format: 
  html:
    toc: true
    toc-location: left
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)
library(measurements)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
library(emmeans)
library(performance)
library(multcompView)



```

## Load Data

```{r}
mobeast <- read_csv(here("data", "MOBEAST_full_merged_data.csv"))
```

## Data visualization theme

```{r}
nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=24, face = "bold"), 
        plot.subtitle = element_text(size=20),
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=18, face = "bold"),
        legend.text=element_text(size=16),
        strip.background = element_rect(fill = "white"))
```



```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

## Data Clean Up

```{r}
DOC <- mobeast %>% select(id_number:date_time, npoc_u_m, avg_res_time, avg_flow_rate)
```


### Outlier detection on bulk data
```{r}
# this data is log transformed
## flagging outliers from the data set

#raw_flagged_outliers <- 
DOC_flagged_outlier <- DOC %>%
  filter(treatment != "Endmember") %>% 

  # flagging outliers
  ## grouping by date, time and treatment
  group_by(treatment) %>% 
  ##calculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m, probs = 0.25, na.rm = T),
         quant3 = quantile(npoc_u_m, probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse(npoc_u_m > quant3 + (IQR*1.5) | npoc_u_m < quant1 - (IQR*1.5),
                           "outlier", "not outlier")) %>% 
  ungroup() %>% 
  filter(!is.na(id_number))


```


### Calculating Fluxes 
(([aquarium]-[inflow])*Vol) / (RT*SA ) in [umol/SA/time]. 
```{r}
tank_area = conv_unit(506.25, from = "cm2", to = "m2") # convert area cm2 to m2
tank_vol = 10 # liters
# DOC in uM, i.e. umol/L
# avg flow rate in mL/min converted to L/sec
# residence time in sec
# areal_flux in umol/s/m2


DOC_flux <- DOC_flagged_outlier %>%  
  
  filter(treatment != "Rain") %>% 
  filter(treatment != "Endmember") %>% 
  
  
  mutate(time_string = as.character(time),
         date_string = as.character(date),
         date_time_string = paste(date_string, time_string),
         date_time = ymd_hms(date_time_string)) %>%
  
# Isolate inflow for each date_time and table
  group_by(date_time, inflow_table) %>%
  mutate(inflow = npoc_u_m[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time) %>% 
  #mutate(inflow = ifelse(outlier == "outlier", inflow == NA, inflow)) %>% 
  mutate(inflow = mean(inflow, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(conc_norm = npoc_u_m - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  # calculate res time but convert flow rate units to L/sec
  mutate(resid_time = tank_vol/(avg_flow_rate*0.001/60)) %>%
  # Unsure why we still normalize to the inflow so calculated both 
  # fluxes with raw data and normalized data.
  mutate(areal_flux = conc_norm*tank_vol/ (resid_time*tank_area)) %>% 
  filter(treatment != "Inflow") %>% 
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) 

```




```{r}
tank_vol = 10 #liter 
DOC_flagged_outlier %>%  
  mutate(resid_time = tank_vol/(avg_flow_rate)) %>% 
  select(tank_number, treatment, 
         laurel_res_time = avg_res_time, keannu_res_time =resid_time)
  
```


### XXX Normalizing the data



```{}
# Normalization version 2
## averaging the inflows for normalization 
## remove the outliers a right before the normalization 

## Normalization of the data to the inflow 

DOC_norm <- raw_flagged_outliers %>% 
  mutate(time_string = as.character(time),
         date_string = as.character(date),
         date_time_string = paste(date_string, time_string),
         date_time = ymd_hms(date_time_string)) %>% 
  select(-IQR, -quant1, -quant3) %>% 
  
  filter(treatment != "Rain") %>% 
  
  ## testing the log10 data transformation
  #mutate(conc = log10(conc)) %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table, nutrient) %>%
  mutate(inflow = conc[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time, nutrient) %>% 
  mutate(inflow = ifelse(outlier == "outlier", inflow == NA, inflow)) %>% 
  mutate(inflow = mean(inflow, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(conc_norm = conc - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```



## XXX Areal Flux Rate 
(([aquarium]-[inflow])*Vol) / (RT*SA ) in [umol/SA/time]. 


```{}
DOC_norm <- DOC %>% filter(!is.na(id_number)) %>% 
  mutate(time_string = as.character(time),
         date_string = as.character(date),
         date_time_string = paste(date_string, time_string),
         date_time = ymd_hms(date_time_string)) %>% 
  
  filter(treatment != "Rain") %>% 
  filter(treatment != "Endmember") %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table) %>%
  mutate(inflow = npoc_u_m[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time) %>% 
  #mutate(inflow = ifelse(outlier == "outlier", inflow == NA, inflow)) %>% 
  mutate(inflow = mean(inflow, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(conc_norm = npoc_u_m - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```

## Calculating DOC fluxes

```{}
tank_area = conv_unit(506.25, from = "cm2", to = "m2") # convert area cm2 to m2
tank_vol = 10 # liters
# DOC in uM, i.e. umol/L
# avg flow rate in mL/min converted to L/sec
# residence time in sec
# areal_flux in umol/s/m2


DOC_flux <- DOC_norm %>% 
  # calculate res time but convert flow rate units to L/sec
  mutate(resid_time = tank_vol/(avg_flow_rate*0.001/60)) %>%
  # Unsure why we still normalize to the inflow so calculated both 
  # fluxes with raw data and normalized data.
  mutate(areal_flux_norm = conc_norm*tank_vol/ (resid_time*tank_area),
         areal_flux = npoc_u_m*tank_vol/ (resid_time*tank_area)) %>% 
  filter(treatment != "Inflow") 
```

### XXX Removing outlier values 
```{}
# Outlier detection on normalize fluxes 
## Norm = Tank-Inflow Avg

DOC_flux_clean  <- DOC_flux %>% 
  # flagging outliers
  group_by(treatment) %>% 
  ## calculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(areal_flux_norm, probs = 0.25, na.rm = T),
         quant3 = quantile(areal_flux_norm, probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse(areal_flux_norm > quant3 + (IQR*1.5) | areal_flux_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup()
  
DOC_flux_clean %>% #filter(time_string == "12:00:00") %>% 
  group_by( outlier) %>% 
  summarise(count = n()) 

# There are a lot of data points that are flagged as outliers, specifically at 00:00.

```

```{}
# Outlier detection on non-normalized fluxes 

DOC_flux_clean <- DOC_flux %>%  
  # flagging outliers
  group_by(treatment) %>% 
  ## calculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(areal_flux, probs = 0.25, na.rm = T),
         quant3 = quantile(areal_flux, probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse(areal_flux > quant3 + (IQR*1.5) | areal_flux < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() 

DOC_flux_clean %>% #filter(time_string == "12:00:00") %>% 
  group_by( outlier) %>% 
  summarise(count = n()) 
  
DOC_flux_clean %>% 
  filter(outlier != "outlier") %>% 
  ggplot(aes(x = treatment, y = areal_flux, fill = treatment))+
  geom_violin()
  

# There are a lot less data points that are flagged as outliers
```


## 12PM data
```{r}
# removing outliers from the data set
DOC_12PM <- DOC_flux %>%
  filter(time_string == "12:00:00") %>% 
  filter(treatment != "Inflow") %>% 
  select(-quant1, -quant3, -IQR) #%>% 
  
  ## remove the outliers 
  #filter(outlier == "not outlier") 

```

```{}
DOC_12PM %>% filter(areal_flux>0.15) %>% 
  select(id_number, treatment, npoc_u_m,outlier, conc_norm, areal_flux) 
```


```{r}
mod_12PM <- lmer(areal_flux~ treatment + (1|tank_number), data = DOC_12PM)
```

```{r}
summary(mod_12PM)
anova(mod_12PM, type = 2)
emmeans(mod_12PM, specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[[1]], ~ treatment)
```

```{r}
resid_panel(mod_12PM)

```

```{r}
 DOC_12PM %>%  
    ggplot(aes(x= treatment, y= areal_flux, fill = treatment))+
    geom_violin()+
    geom_jitter(aes(color = outlier), size = 2)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("12PM", "DOC", "Production"),
       x= "Treatment",
       y = "Areal Flux (µmol/s/m2)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
  nunu_theme #+
  #theme(legend.position = "none")
```


### Iterative graphing

```{}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## --> In the for-loop, filter the df by nutrient and make a graph that will display nutrient data
## Save the outputs of the model in the list such that each nutrient has its stats

DOC_loop <- DOC_12PM %>% select(nutrient) %>% distinct() %>% pull()

graphs_12PM <- vector(mode = "list", length = length(DOC_loop))
names(graphs_12PM) <- DOC_loop

for (var in DOC_loop) {
  
  df <- DOC_12PM %>% filter(nutrient == var)
  
  graphs_12PM[[var]] <- df %>%  
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("12PM", var, "Production"),
       x= "Treatment",
       y = "Production (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none")
}

summary(models_12PM[["DOC"]])
anova(models_12PM[["DOC"]], type = 2)
emmeans(models_12PM[["DOC"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[[1]], ~ treatment)

resid_panel(models_12PM[["DOC"]])


```


#### Data Visualization 

```{, fig.width= 12, fig.height = 6}
graphs_12PM[["DOC"]]#+ 
  #geom_jitter(aes(color = inflow_table), size = 2)+
  #scale_color_manual(values = c("red", "blue"))+
  #facet_wrap(date~.)+
  #theme(legend.position = "right")

#ggsave(DOC_12PM.png, path = here("outputs", "Nov_com_meeting", "DOC_12PM.png"))
```



```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
DOC_12PM %>%
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(areal_flux, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  ) %>%  select(-t_test)

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```


## Hypothesis 2: There are variation in exudation on a diel cycle 24h

```{r}
## removing outliers from the data set
DOC_24h <- DOC_flux %>%
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  #filter(treatment != "Inflow") %>% 
  ## remove the outliers 
  #filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00"))) #%>% 
  #filter(time_string != "00:00:00") 
```



#### 24h Model setup and results
```{r}
mod <- lmer(areal_flux ~ treatment*time_fac + (1|tank_number), data = DOC_24h)
```

```{r}
summary(mod)
anova(mod, type = 2)
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
#emmeans(models_24h[["DOC"]], ~ treatment)
```

```{r}
resid_panel(mod)
```

#### Data Visualization 
```{r}
DOC_24h %>% 
    ggplot(aes(x= treatment, y= areal_flux, fill = treatment))+
    geom_violin()+
    geom_jitter(aes(color = outlier), size = 2)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_wrap(~time_fac, ncol = 4 )+
    labs(title = paste("24h DOC Production"),
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Production (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18))#,
        #legend.position = "none") #+
  #geom_jitter(aes(color = inflow_table), size = 2)+
  #scale_color_manual(values = c("red", "blue"))+
  #theme(legend.position = "right")
```



### Pairwise comparison


```{}
control_24h <- DOC_24h %>% filter(treatment == "Coral") %>% 
  filter(time_string != "12:00:00") %>% 
  mutate(time_US = ifelse(time_string == "18:00:00", "six_pm", "six_am")) %>% 
  select(tank_number, time_US, conc_norm) %>% 
  pivot_wider(names_from = time_US, values_from = conc_norm) 

t.test(control_24h$six_pm, control_24h$six_am, paired = TRUE)
t.test(control_24h$six_pm, control_24h$six_am, paired = TRUE)$estimate
t.test(control_24h$six_pm, control_24h$six_am, paired = TRUE)$p.value
```



```{r}
nunu_pairwise <- function(data, time_point = 6){

  ## Get a list of the treatments for the for loop   
  treatment <- data %>% mutate(treatment = as.character(treatment)) %>%
  select(treatment) %>% distinct() %>% pull()
  
  ## Create un empty data frame
pairwise_results <- tibble(
  time_point = vector(),
  treatment= vector(),
  mean = vector(),
  p_value = vector())

# this is the pairwise comparison for the 6 pairs (6AM-6PM)
  if(time_point == 6){
    for(var in treatment){
      
      ## only keep the 6AM and 6PM
      df <- data %>% filter(treatment == var) %>% 
  filter(time_string == "18:00:00" | time_string == "06:00:00") %>% 
        ## Rename the times so they can become column names
  mutate(time_US = ifelse(time_string == "18:00:00", "T18", "T06")) %>% 
        ## Only keep the column of interest for the stats
  select(tank_number, time_US, areal_flux) %>% 
  pivot_wider(names_from = time_US, values_from = areal_flux) 
      
      ## Pairwise function 
      paired <- t.test(df$T18, df$T06, paired = TRUE, na.action = na.pass)
      
      ## Save the result in at data frame
      df2 <- tibble(
        time_point = "06AM-06PM",
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      # Code will loop oveer all treatments and save the results in a data frame
      pairwise_results <-rbind(pairwise_results,df2)
      
      # Save the results in RStudio
      pairwise_results <<- pairwise_results
    }  
  

# This is for the 12 samples (12AM-12PM)
  }else if(time_point == 12){
    for(var in treatment){
      
      ## Same as before but for the 12AM-12PM
      df <- data %>% filter(treatment == var) %>% 
  filter(time_string == "12:00:00" | time_string == "00:00:00") %>% 
  mutate(time_US = ifelse(time_string == "12:00:00", "T12", "T00")) %>% 
  select(tank_number, time_US, areal_flux) %>% 
  pivot_wider(names_from = time_US, values_from = areal_flux) 
      
      paired <- t.test(df$T12, df$T00, paired = TRUE, na.action = na.pass)
      
      df2 <- tibble(
        time_point = "12AM-12PM",
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      pairwise_results <- rbind(pairwise_results,df2)
      pairwise_results <<- pairwise_results
    }
}

# Display the results of the code 
print(pairwise_results)
}

```


```{r}
nunu_pairwise(data = DOC_24h, time_point = 12) 
```


## Test to deal with 12AM samples 

```{r}
DOC_flagged_outlier %>% 
  
  mutate(time_string = as.character(time)) %>% 
  filter(time_string == "00:00:00") %>% 
  mutate(npoc_u_m_diff = npoc_u_m - 119.84561, 
         resid_time = tank_vol/(avg_flow_rate*0.001/60),
         areal_flux = npoc_u_m_diff*tank_vol/ (resid_time*tank_area)) %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) %>% 
  filter(npoc_u_m<300) %>% 
  group_by(treatment) %>% 
  ggplot(aes(treatment, areal_flux, fill = treatment))+
  geom_violin()+
  geom_jitter(color = "black")+
  stat_summary(color = "yellow", alpha =0.5)+
  labs(title = "00:00:00 DOC Flux")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  geom_hline(yintercept=0, linetype='dashed')+
  nunu_theme
```


```{r}
DOC_flagged_outlier %>% 
  filter(npoc_u_m<300) %>% 
  group_by(treatment) %>% 
  reframe(mean = mean(npoc_u_m, na.rm = T))
```







