---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: MOBEAST DOC V2
output: 
  html_document:
     toc: true
     toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
#library(GGally)
library(emmeans)
library(performance)


```

## Load Data 
```{r}
DOC <- read_csv(here("data", "DOC", "DOC_clean.csv"))
table_id <- read_csv(here("data", "Laurel", "TableID.csv"))
```

## Data visualization theme
```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        panel.background = element_rect(fill = "azure1"))
```

## Data Clean Up

```{r}
table_id <- clean_names(table_id)

table_id <- table_id %>% rename(tank_number=tankid) %>% select(-treatment) %>% 
  mutate(tank_number = paste0("T", tank_number)) %>% 
  mutate(inflow_table = paste("Table", inflow_table))
DOC <- left_join(DOC, table_id, by= "tank_number")
```

```{r}
DOC <- DOC %>% mutate(inflow_table = ifelse(is.na(inflow_table), tank_number, inflow_table)) %>% 
   mutate(inflow_table = ifelse(inflow_table == "F1", "Table 1", 
                                ifelse(inflow_table == "F2", "Table 2", inflow_table))) %>% 
  mutate(inflow_table = factor(inflow_table),
         tank_number = factor(tank_number)) %>% 
  relocate(inflow_table, .after = "treatment")
```


```{r}

DOC_norm <- DOC %>% filter(date > as.Date("2024-05-31")) %>% # Removes end members
  filter(treatment != "Rain") %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
   group_by(date_time, inflow_table) %>%
  mutate(inflow= npoc_u_m[treatment == "Inflow"]) %>%  # Isolate inflow for each date_time and table
  ungroup() %>%
  group_by(date_time) %>% 
  mutate(inflow = ifelse(is.na(inflow),
      inflow[!is.na(inflow)][1],  # get the other table's value for that date
      inflow)) %>%
  ungroup() %>% 
  mutate(npoc_u_m_norm = npoc_u_m - inflow)  %>% # Normalize for treatments 
  relocate(inflow, .after = "npoc_u_m_norm") %>% 
  relocate(date_time, .before = "date") %>%
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
  
```


## Hypothesis 0: endmembers and inflow comparison

```{r}
DOC %>% filter(treatment == "Endmember" | treatment == "Rain"| treatment == "Inflow") %>% 
  filter(time != hms("00:00:00")) %>% 
  mutate(tank_number = as.character(tank_number)) %>% 
  mutate(tank_number = ifelse(tank_number == "Lagoon Tiahura", "Lagoon", tank_number),
         tank_number = factor(tank_number)) %>% 
  group_by(tank_number) %>% 
  summarize(avg = mean(npoc_u_m, na.rm =T),
            min = min(npoc_u_m, na.rm =T),
            max = max(npoc_u_m, na.rm =T))
```



```{r}
DOC_verif <- DOC %>% filter(treatment == "Endmember" | treatment == "Rain"| treatment == "Inflow") %>% 
  filter(time != hms("00:00:00")) %>% 
  mutate(tank_number = as.character(tank_number)) %>% 
  mutate(tank_number = ifelse(tank_number == "Lagoon Tiahura", "Lagoon", tank_number),
         tank_number = factor(tank_number, levels = c("Lagoon", "Ocean", "Bay", "F1", "F2")))
```

```{r}
mod <- lmer(npoc_u_m ~ tank_number + (1|date)+ (1|time)+  (1|treatment), data = DOC_verif)

summary(mod)
```

```{r, fig_height = 10, fig.width = 10}
plot_model(mod, type = "re")[[1]] +
  plot_model(mod, type = "re")[[2]] +
  plot_model(mod, type = "eff", terms = c('treatment')) +
  plot_model(mod, type = "diag")[[1]]
```
```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ tank_number)
```

```{r}
ggplot(DOC_verif, aes(x= tank_number, npoc_u_m, fill = treatment)) +
  geom_boxplot()+
  geom_jitter(aes(color = tank_number), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "All samples variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  facet_wrap(~date) + 
  #scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       #values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
```



## Hypothesis  1: Different communities will release difference amounts of DOC (12PM)

```{r}
DOC_norm_12PM <- DOC_norm %>% 
  filter(time == hms("12:00:00")) %>% 
  filter(treatment != "Inflow") 
```

```{r}
mod <- lmer(npoc_u_m_norm ~ treatment + (1|date) +(1|inflow_table/tank_number), data = DOC_norm_12PM)

summary(mod)
```

```{r, fig_height = 14, fig.width = 10}
plot_model(mod, type = "re")[[3]] +
  plot_model(mod, type = "re")[[2]] +
  plot_model(mod, type = "eff", terms = c('treatment')) +
  plot_model(mod, type = "diag")[[1]]
```

```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
#emmeans(mod, pairwise ~ treatment)

```

### Interpretation

- Random Effects: Sample date can explain 8 std of variation, Inflow table can explain 3.5 std of variation. Te residuals indicate that 22.6 std still remains unexplained by the model.
- Random effect for (1|tank_number) is 0 variance or std dev.
- Random effect for (1|inflow_table/tank_number) is 0 variance or std dev.
- Fixed effects: DOC values increased in algae tanks and stayed lower in coral tanks. There is a significant difference between algae tank and control (p = 0.025). Corals and CCA are not different from the control.
- Residual points are are close to the regression line, except for one point
- Emmeans does a paired comparison between control and treatments. The p-values are not significant. But algae tank is much smaller p-value than the other treatments. 

```{r}
ggplot(DOC_norm_12PM, aes(x= treatment, npoc_u_m_norm, fill = treatment)) +
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "12PM variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  facet_wrap(~date) + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
```


## Hypothesis 2: There are variation in exudation on a diel cycle 24h 

```{r}
DOC_norm_24h <- DOC_norm %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>% 
  filter(time != hms("00:00:00")) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!(tank_number %in% c("T8", "T11") & time == hms("18:00:00")))
```

```{r}
mod <- lmer(npoc_u_m_norm ~ treatment + (1|time)+ (1|date)+ (1|inflow_table/tank_number), data = DOC_norm_24h)

summary(mod)
```


```{r, fig_height = 10, fig.width = 10}
plot_model(mod, type = "re")[[1]] +
  plot_model(mod, type = "re")[[2]] +
  plot_model(mod, type = "eff", terms = c('treatment')) +
  plot_model(mod, type = "diag")[[1]]
```
```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```


### Interpretation
- Removed 00:00:00 because it was drastically different than the other 3 time points. Need to figure out what happened.
- Removed 18:00:00 T8 and T11 (control tanks) because they looked like outliers on the boxplots and also they skewed the model a lot!
- Random effect : inflow table does not explain the variance (variance = 0), nested effect helps to explain some of the variation. 
- Fixed effects: the p-values for the t-test are not significant but algae still has a smaller p-value than the other treatments. 
- Emmeans contrasts also do not show significant difference between the treatments. p>0.05

```{r}
ggplot(DOC_norm_24h, aes(x= treatment, npoc_u_m_norm, fill = treatment)) +
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "24h variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  facet_wrap(~time) + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
```

## Hypothesis 3: We'll see differences if we compare opposite times
```{r}
DOC_norm_6_18 <- DOC_norm %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>% 
  filter(time != hms("00:00:00")) %>% 
  filter(time != hms("12:00:00")) %>%
  filter(treatment != "Inflow") %>% 
  filter(!(tank_number %in% c("T8", "T11") & time == hms("18:00:00")))
```


```{r}
mod <- lmer(npoc_u_m_norm ~ treatment+ (1|time)+ (1|date)+ (1|inflow_table/tank_number), 
            data = DOC_norm_6_18)

summary(mod)
```
```{r, fig_height = 10, fig.width = 10}
plot_model(mod, type = "re")[[1]] +
  plot_model(mod, type = "re")[[2]] +
  plot_model(mod, type = "eff", terms = c('treatment')) +
  plot_model(mod, type = "diag")[[1]]
```

```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)

```


### Interpretion

- 06:00 vs 18:00 did not show significant differences in DOC concentrations between treatments either

```{r}
ggplot(DOC_norm_6_18, aes(x= treatment, npoc_u_m_norm, fill = treatment)) +
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "06:00-18:00 variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  facet_wrap(~time) + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
```

## Hypothesis 4: All samples

```{r}
DOC_norm_all <- DOC_norm %>% 
  filter(time != hms("00:00:00")) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!(tank_number %in% c("T8", "T11") & time == hms("18:00:00")))
```

```{r}
mod <- lmer(npoc_u_m_norm ~ treatment + (1|time)+ (1|date), 
            data = DOC_norm_all)

summary(mod)
```

```{r, fig_height = 10, fig.width = 10}
plot_model(mod, type = "re")[[1]] +
  plot_model(mod, type = "re")[[2]] +
  plot_model(mod, type = "eff", terms = c('treatment')) +
  plot_model(mod, type = "diag")[[1]]
```

```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
plot(emmeans(mod, ~ treatment), comparisons = TRUE)
```

### Interpretation

- Looking at all the samples together, lmer seems to indicate that Algae tanks are significantly different from the control
- Emmeans show that the difference between pairs is not significant. But the algae tanks have much low p-values than the other treatments



```{r}
ggplot(DOC_norm_all, aes(x= treatment, npoc_u_m_norm, fill = treatment)) +
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "All samples variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  facet_wrap(~date_time) + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
```

