---
title: "MOBEAST- DNA Ordination 24h"
format: html
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

# Load Libraries

```{r}
library(tidyverse)
#library(lubridate)
library(here)
library(janitor)
#library(ggrepel)
#library(ghibli)
#library(gt)
library(patchwork)
#library(RColorBrewer)
library(vegan)
library(phyloseq)
#library(GUniFrac)
```

# Data visualization

```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=32, face = "bold"), 
        plot.subtitle = element_text(size=30),
        axis.title = element_text(size = 28, face = "bold"),
        axis.text = element_text(size = 28),
        strip.text.x = element_text(size = 28, face = "bold"),
        legend.title=element_text(size=28, face = "bold"),
        legend.text=element_text(size=28))
```

# Load Data

```{r}
otu <- read_csv(here("data", "DNA", "MOBEAST_OTU_culled.csv"))
meta <- read_csv(here("data","MOBEAST_metadata_V2.csv"))
#tax <- read_csv(here("data","DNA", "MOBEAST_tax_culled.csv"))
OTU_ord <- read_csv(here("data","OTU_heatmap_ready.csv"))
cluster <- read_csv(here("data","DNA", "24h_clusters.csv"))
```

# Data Clean Up

```{, meta data clean up} 
meta\<- meta %\>% mutate(time = str_pad(time, width = 4, side = "left", pad = "0"), time = str_replace(time, "(\\d{2})(\\d{2})", "\\1:\\2")) %\>% \# "(\\d{2})(\\d{2})" separates the string in 2 groups of 2 characters \# "\\1:\\2", group1 + ":" + group2 mutate(time = paste0(time, ":00"), date_time = paste(date, time), date = mdy(date), date_time = mdy_hms(date_time), time = factor(time))
```         

## Preparing the metadata to annotate the heatmap

```{}
# renaming some of treatment so that environmental data have a label: endmember or rain
# Same for inflow table, environmental samples now have a label
## Convert tibble to df for later
meta <- meta %>% mutate(treatment = as.character(treatment)) %>% 
  mutate(treatment = ifelse(id_number == "MOBEAST115", "Rain", 
                            ifelse(is.na(treatment), "Endmember", treatment))) %>% 
  mutate(inflow_table = ifelse(is.na(inflow_table), "Environmental data", inflow_table)) %>% 
  mutate(date_string = as.character(date),
                time_string = as.character(time),
                date_time_string = paste(date_string, time_string)) %>% 
  as.data.frame()
```

```{r}
cluster<- cluster %>% 
  #mutate(cluster = factor(cluster)) %>% 
  mutate(cluster2 = ifelse(cluster == 1, "Algae", 
                           ifelse(cluster == 2, "CCA", 
                                  ifelse(cluster == 3, "Coral", "Control")))) %>% 
  select(-cluster) %>% 
  rename(cluster = cluster2)

meta<- meta %>% left_join(cluster)
```

## Transform the dataframes into matrices

```{r}
## Make an OTU matrix
otu<- otu %>% mutate(ASV = toupper(ASV))
otu_list<- otu$ASV
otu<- otu %>% select(-(117:122), -MOBEAST017b)
otu_mat <- as.matrix(otu)
rownames(otu_mat) <- otu_list

```

```{r}
OTU <- otu_table(otu_mat, taxa_are_rows = TRUE)
```

# Ordinations with long format df

## 24h Samples

```{r}
otu_ord <- OTU %>% 
  t() %>% 
  as.data.frame() 

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select 12PM samples 
id_number_24h <- meta %>% 
  filter(date_string == "2024-06-02"| date_string == "2024-06-03") %>% 
  pull(id_number)

otu_ord_24h <- otu_ord %>% filter(id_number %in% id_number_24h)


```

```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_24h <- otu_ord_24h %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances

```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_24h <- vegdist(otu_ord_mat_24h, method = "bray")
```

```{r}
#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_24h, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_24h <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_24h <- left_join(MOBEAST_nmds_24h, meta)
```



### Plotting NDMS 24h
#### Plotting by treatment

```{r, fig.width=16, fig.height= 6}
# Plot
MOBEAST_nmds_24h %>%   
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 4)+
  labs(title= "24h samples, ellipse by treatment")+
  scale_color_manual(values=treatment_colors )+
  scale_fill_manual(values= treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))

```



#### Identifying the 2 samples that cluster alone in the Heatmaps 
```{r}
MOBEAST_nmds_24h_bis <- MOBEAST_nmds_24h %>% 
  mutate(treatment_bis = treatment) %>% 
  mutate(treatment_bis = ifelse(id_number == "MOBEAST010" |id_number == "MOBEAST060", "weird", treatment_bis ))
```

```{r, fig.width=16, fig.height= 6}
# Plot
MOBEAST_nmds_24h_bis %>%  
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment_bis, fill= treatment_bis))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 4)+
  labs(title= "24h samples, ellipse by treatment")+
  scale_color_manual(values= c("Inflow"="#00838F", "Control"="grey", "Coral"="#03A9F4", "CCA"="#EF5350", "Algae"="#4CAF50","weird"= "yellow"))+
  scale_fill_manual(values= c("Inflow"="#00838F", "Control"="grey", "Coral"="#03A9F4", "CCA"="#EF5350", "Algae"="#4CAF50","weird"= "yellow")) +
  #nunu_theme+
  theme_bw()
#ggsave("24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))

```

#### Plotting by heatmap cluster

Note: cluster numbers are inaccurate and do not match what is seen in the heatmap. Problem with the distance matrix or clutering method or both...

```{r, fig.width=16, fig.height= 10}
# Plot
MOBEAST_nmds_24h %>%   
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  #stat_ellipse(aes(colour = cluster, fill = cluster),
               #alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 2)+
  labs(title= "24h samples, ellipse by treatment")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  ggforce::geom_mark_ellipse(aes(color = cluster, fill = cluster), alpha = 0.3)+
  #nunu_theme+
  theme_bw()
#ggsave("24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

```{r, fig.width=10, fig.height= 6}
# raw NMDS plot with all the samples
MOBEAST_nmds_24h %>%  
  mutate(time_string = factor(time_string, 
                              levels = c("12:00:00","18:00:00","00:00:00","06:00:00" ))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment))+
  stat_ellipse(aes(colour = time_string, fill = time_string), 
               alpha = 0.2, geom = "polygon", level = 0.95)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  labs(title= "24h samples, ellipse by time")+
  scale_color_manual(values=treatment_colors)+
  #scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("24h_time.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

### Stats on treatment only

```{r}
perma_24h <- adonis2(MOBEAST_dist_24h ~ MOBEAST_nmds_24h$treatment)#, 
        # formula to calculate p-value for treatments
        #strata = MOBEAST_nmds_24h$tank_number) # accounts for repeated measures
        # should I add the stratum here ? What should I put if I keep it? 
perma_24h
```

# Ordination of individual time points

## 12:00:00

```{r}
otu_ord <- OTU %>% t() %>% as.data.frame() 

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select 12PM samples 
samples_12PM <- meta %>% mutate(time_string = as.character(time_string),) %>% 
  filter(time_string == "12:00:00" & date == ymd("2024-06-02")) %>% pull(id_number)

otu_ord_12PM <- otu_ord %>% filter(id_number %in% samples_12PM)


```

```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_12PM <- otu_ord_12PM %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances

```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_12PM <- vegdist(otu_ord_mat_12PM, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_12PM, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_12PM <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

### Plotting NMDS

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_12PM <- left_join(MOBEAST_nmds_12PM, meta)
```

```{r, fig.width=16, fig.height= 12}
# Plot
MOBEAST_nmds_12PM %>%   
  # mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      # "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  #facet_wrap(~time_string, ncol = 2)+
  labs(title= "12:00 samples")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("12PM_24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

### Stats on treatment only

```{r}
perma_12PM <- adonis2(MOBEAST_dist_12PM ~ MOBEAST_nmds_12PM$treatment) #, 
        # formula to calculate p-value for treatments
        #strata = MOBEAST_nmds_12PM$tank_number) # accounts for repeated measures

perma_12PM
```

## 06:00:00

```{r}
otu_ord <- OTU %>% 
  t() %>% 
  as.data.frame() 

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select 12PM samples 
samples_6AM <- meta %>% mutate(time_string = as.character(time_string)) %>% 
  filter(time_string == "06:00:00") %>% pull(id_number)

otu_ord_6AM <- otu_ord %>% filter(id_number %in% samples_6AM)


```

```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_6AM <- otu_ord_6AM %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances

```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_6AM <- vegdist(otu_ord_mat_6AM, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_6AM, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_6AM <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

### Plotting NMDS

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_6AM <- left_join(MOBEAST_nmds_6AM, meta)
```

```{r, fig.width=16, fig.height= 12}
# Plot
MOBEAST_nmds_6AM %>%   
  # mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      # "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  #facet_wrap(~time_string, ncol = 2)+
  labs(title= "12:00 samples")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("12PM_24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

### Stats on treatment only

```{r}
perma_6AM <- adonis2(MOBEAST_dist_6AM ~ MOBEAST_nmds_6AM$treatment)#, 
        # formula to calculate p-value for treatments
        #strata = MOBEAST_nmds_6AM$tank_number) no need strata here because it restricts the permutations
perma_6AM
```

#### Accounting for uneven variation in the data.

If the variation is uneven without the groups, adonis may come up with a significant result, although it's not. We can control for that using betdisper() and permutest()

```         
# beta dispertion
bd_12 <- betadisper(MOBEAST_dist_12, MOBEAST_nmds_12$treatment)

anova(bd_12) 
permutest(bd_12) 
bd_12
```