---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: MOBEAST DOC verification
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE , 
                      warning = FALSE)
```

<!-- ## Load Libraries -->
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
```

<!--## Load Data -->
```{r}
meta <- read_delim(here("data", "MOBEAST_metadata.csv"))
DOC <- read_delim(here("data", "raw_data", "DOC", "MOBEAST_NPOC_TN_241010.csv"))
```

## Codes of the treatment  
```{r}
treat_code <- tibble(
  treatment = c("A", "K", "C", "R"),
  full_name = c("Algae", "Coral", "Control", "Rubble")
)
treat_table <- treat_code %>% gt() %>% 
  cols_label(treatment = "Treatment Code",
             full_name = "Treatment Name")
treat_table
```


<!-- ## Data clean up -->
```{r}
meta <- meta %>% clean_names() %>% drop_na() %>%
  mutate(long_name = str_trim(long_name)) %>% 
  filter(long_name != "Long Name") %>% 
  mutate(date = mdy(date), 
         time = format(strptime(time, format = "%H%M"), "%H:%M"),
         id_number = str_trim(id_number))  
```

```{r}
DOC <- DOC %>% clean_names() 
```

<!-- ## Joining the data -->
```{r}
DOC_join <- left_join(meta, DOC, by = "id_number")
```

## Check data for NPOC (mg/L)

The following points have a value different than the other data points: 

- 2024-06-02 18:00: **T8** and **T11**. These are the same points as the ones Laurel identified.  
- 2024-06-03 00:00: **F1, F2, T2, T5, T3** and **T13**. However, that's a lot of points so they might not be outliers.  

```{r}
DOC_join %>% drop_na(npoc_mg_l) %>% filter(treatment != "Endmember") %>% 
  ggplot(aes(x= treatment, y = npoc_mg_l, color = as.factor(time))) +
  geom_point() + 
  geom_text_repel(aes(label = tank_number)) +
  geom_point(data=DOC_join %>% filter(npoc_mg_l > 3.5), 
           pch=21, 
           size=4, 
           colour="red") +
  facet_wrap(.~date)+
  theme_bw()+
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle = 45, vjust = 0.5),
        panel.background = element_rect(fill = "azure1"))+
   scale_color_ghibli_d("SpiritedMedium", direction=-1)
  
  
  
```


## Check data for TN (mg/L)

For TN, **T11** at 2024-06-03 6:00 seems to be a clear outlier. The tanks that were identified as outliers in the NPOC plot above do not appear on the TN plot.  
```{r}
DOC_join %>% drop_na(tn_mg_l) %>% filter(treatment != "Endmember") %>% 
  ggplot(aes(x= treatment, y = tn_mg_l, color = as.factor(time))) +
  geom_point() + 
  geom_text_repel(aes(label = tank_number)) +
  geom_point(data=DOC_join %>% filter(tn_mg_l > 0.9), 
           pch=21, 
           size=4, 
           colour="red") +
  facet_wrap(.~date)+
  theme_bw()+
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle = 45, vjust = 0.5),
        panel.background = element_rect(fill = "azure1"))+
   scale_color_ghibli_d("SpiritedMedium", direction=-1)
```

<!--
## Overall Stats
```{r}
DOC_join %>% filter(treatment != "Endmember", treatment != "Rain" ) %>%
  group_by(date, time, treatment) %>% 
  summarise(avg = mean(npoc_mg_l, na.rm = T),
            std_dev = sd(npoc_mg_l, na.rm = T))
```
-->


