---
title: "MOBEAST fDOM V3"
format: 
  html:
    toc: true
    toc-location: left
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)
library(ggridges)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
#library(GGally)
library(emmeans)
library(performance)


```

## Data visualization theme

```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=32, face = "bold"), 
        plot.subtitle = element_text(size=30),
        axis.title = element_text(size = 28, face = "bold"),
        axis.text = element_text(size = 28),
        strip.text.x = element_text(size = 28, face = "bold"),
        legend.title=element_text(size=28, face = "bold"),
        legend.text=element_text(size=28))
```

```{r}
#| output: false
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```
## Load Data

```{r}
fDOM <- read_csv(here("data", "fDOM", "fDOM_sorted.csv"))

fDOM <- fDOM %>% mutate(treatment = ifelse(treatment == "Rubble", "CCA", treatment)) %>%
  select(-sample_type)

table_id <- read_csv(here("data", "Laurel", "TableID.csv"))
```

## Data Clean-Up
### Sub-catergorization of fDOM types
```{r}
fDOM_desc <- tibble(fdom = c("coble_a","coble_b","coble_c","coble_m","coble_t",
  "m_to_c","bix","hix","fpeak","fi",
  "lignin"),
  fdom_alt_name = c("UV-Humic","Tyrosine","Visible-Humic","Marine Humic","Tryptophan",
  "m_to_c","bix","hix","Phenylalanine","Fluvic Acid",
  "lignin"),
  fdom_cat= c("Humic","Proteinaceous","Humic","Humic","Proteinaceous",
  "Humic","Humic","Humic","Proteinaceous","Humic",
  "Other"))
```

### Assign Table ID to samples
```{r}
table_id <- clean_names(table_id)

table_id <- table_id %>% rename(tank_number=tankid) %>% select(-treatment) %>% 
  mutate(tank_number = paste0("T", tank_number)) %>% 
  mutate(inflow_table = paste("Table", inflow_table))
fDOM <- left_join(fDOM, table_id, by= "tank_number")
```

```{r}
fDOM <- fDOM %>% mutate(inflow_table = ifelse(is.na(inflow_table), tank_number, inflow_table)) %>% 
   mutate(inflow_table = ifelse(inflow_table == "F1", "Table 1", 
                                ifelse(inflow_table == "F2", "Table 2", inflow_table))) %>% 
  mutate(inflow_table = factor(inflow_table)) %>% 
  relocate(inflow_table, .after = "treatment") %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow", "Endmember")))
```

### Create long format df to make data analysis easier later
```{r}
fDOM_long <- fDOM %>% pivot_longer(cols = coble_a:lignin,
                      names_to = "fdom", 
                      values_to = "fdom_conc") %>% 
  select(-long_name)
```

Attaching the subcategories of fDOM to the long df
```{r}
fdom_long <- left_join(fDOM_long, fDOM_desc, by = "fdom")
```


## Outlier detection 

```{, log10 data for IQR}
# ATTENTION !!!!!!!!!!!!! Data was log transform to do the IQR method. 

## removing outliers from the data set
raw_flagged_outliers <- fdom_long%>%
  filter(treatment != "Endmember") %>% 
  # removing outliers
  ## grouping by date, time and treatment
  group_by(fdom, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  
# !!!!!!!!!!!!!  ATTENTION REQUIRE BELOW !!!!!!!!!!!!! 

  mutate(quant1 = quantile(log10(fdom_conc), probs = 0.25, na.rm = T),
         quant3 = quantile(log10(fdom_conc), probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( log10(fdom_conc) > quant3 + (IQR*1.5) | log10(fdom_conc) < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup()  
```


```{r}
## removing outliers from the data set
raw_flagged_outliers <- fdom_long%>%
  filter(treatment != "Endmember") %>% 
  # removing outliers
  ## grouping by date, time and treatment
  group_by(fdom, treatment) %>% 
  ## calculating the Interquantile range and of the raw data

  mutate(quant1 = quantile(fdom_conc, probs = 0.25, na.rm = T),
         quant3 = quantile(fdom_conc, probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse(fdom_conc > quant3 + (IQR*1.5) | fdom_conc < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup()  
```


```{r}
# Normalization version 2
## averaging the inflows for normalization 
## remove the outliers a right before the normalization 

## Normalization of the data to the inflow 
#nutrients_norm <- 
fDOM_norm <- raw_flagged_outliers %>% 
  mutate(time_string = as.character(time),
         date_string = as.character(date),
         date_time_string = paste(date_string, time_string),
         date_time = ymd_hms(date_time_string)) %>% 
  select(-IQR, -quant1, -quant3) %>% 
  
  filter(treatment != "Rain") %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table, fdom) %>%
  mutate(inflow = fdom_conc[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time, fdom) %>% 
  mutate(inflow = ifelse(outlier == "outlier", inflow == NA, inflow)) %>% 
  mutate(inflow = mean(inflow, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(conc_norm = fdom_conc - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```

## 12PM data
```{r}
# Version 2
## removing outliers from the data set
fDOM_12PM <- fDOM_norm %>%
  filter(time_string == "12:00:00") %>% 
  filter(treatment != "Inflow") %>% 
  
  ## remove the outliers 
  filter(outlier == "not outlier") 

```

### Iterative statistical modeling

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

fDOM_loop <- fDOM_12PM %>% select(fdom) %>% distinct() %>% pull()

models_12PM <- vector(mode = "list", length = length(fDOM_loop))
names(models_12PM) <- fDOM_loop

for (var in fDOM_loop) {
  df <- fDOM_12PM %>% filter(fdom == var)
  models_12PM[[var]] <- lmer(conc_norm ~ treatment + (1|tank_number), data = df)
}

```

### Iterative graphing

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## --> In the for-loop, filter the df by nutrient and make a graph that will display nutrient data
## Save the outputs of the model in the list such that each nutrient has its stats

fDOM_loop <- fDOM_12PM %>% select(fdom) %>% distinct() %>% pull()

graphs_12PM <- vector(mode = "list", length = length(fDOM_loop))
names(graphs_12PM) <- fDOM_loop

for (var in fDOM_loop) {
  
  df <- fDOM_12PM %>% filter(fdom == var)
  
  graphs_12PM[[var]] <- df %>%  
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("12PM", var, "concentrations"),
       x= "Treatment",
       y = "Concentration (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none")
}

```

### Humics
#### CobleA

```{r}
summary(models_12PM[["coble_a"]])
anova(models_12PM[["coble_a"]], type = 2)
emmeans(models_12PM[["coble_a"]], specs = trt.vs.ctrl ~ treatment)
```

```{r}
resid_panel(models_12PM[["coble_a"]])
graphs_12PM[["coble_a"]]
```

#### CobleM

```{r}
summary(models_12PM[["coble_m"]])
anova(models_12PM[["coble_m"]], type = 2)
emmeans(models_12PM[["coble_m"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[[1]], ~ treatment)
```

```{r}
resid_panel(models_12PM[["coble_m"]])
graphs_12PM[["coble_m"]]
```

#### CobleC *
Sig ANOVA with log10 data outlier detection but not contrast
```{r}
summary(models_12PM[["coble_c"]])
anova(models_12PM[["coble_c"]], type = 2)
emmeans(models_12PM[["coble_c"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["coble_c"]])
graphs_12PM[["coble_c"]]
```


#### Sig! M to C

```{r}
summary(models_12PM[["m_to_c"]])
anova(models_12PM[["m_to_c"]], type = 2)
emmeans(models_12PM[["m_to_c"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["m_to_c"]])
graphs_12PM[["m_to_c"]]
```


#### FI

```{r}
summary(models_12PM[["fi"]])
anova(models_12PM[["fi"]], type = 2)
emmeans(models_12PM[["fi"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["fi"]])
graphs_12PM[["fi"]]
```


#### BIX

```{r}
summary(models_12PM[["bix"]])
anova(models_12PM[["bix"]], type = 2)
emmeans(models_12PM[["bix"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["bix"]])
graphs_12PM[["bix"]]
```

#### HIX

```{r}
summary(models_12PM[["hix"]])
anova(models_12PM[["hix"]], type = 2)
emmeans(models_12PM[["hix"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["hix"]])
graphs_12PM[["hix"]]
```

### Proteinaceous
#### CobleB

```{r}
summary(models_12PM[["coble_b"]])
anova(models_12PM[["coble_b"]], type = 2)
emmeans(models_12PM[["coble_b"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["coble_b"]])
graphs_12PM[["coble_b"]]
```


#### CobleT

```{r}
summary(models_12PM[["coble_t"]])
anova(models_12PM[["coble_t"]], type = 2)
emmeans(models_12PM[["coble_t"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["coble_t"]])
graphs_12PM[["coble_t"]]
```


#### Fpeak

```{r}
summary(models_12PM[["fpeak"]])
anova(models_12PM[["fpeak"]], type = 2)
emmeans(models_12PM[["fpeak"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["fpeak"]])
graphs_12PM[["fpeak"]]
```


## 24h data 
```{r}
# Version 2
## removing outliers from the data set
fDOM_24h <- fDOM_norm %>%
  filter(date_string == "2024-06-02" | date_string == "2024-06-03" ) %>% 
  filter(treatment != "Inflow") %>% 
  
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00")))

```

### Iterative statistical modeling

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

fDOM_loop <- fDOM_24h %>% select(fdom) %>% distinct() %>% pull()

models_24h <- vector(mode = "list", length = length(fDOM_loop))
names(models_24h) <- fDOM_loop

for (var in fDOM_loop) {
  df <- fDOM_24h %>% filter(fdom == var)
  models_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + (1|tank_number), data = df)
}

```

### Iterative graphing

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## --> In the for-loop, filter the df by nutrient and make a graph that will display nutrient data
## Save the outputs of the model in the list such that each nutrient has its stats

fDOM_loop <- fDOM_24h %>% select(fdom) %>% distinct() %>% pull()

graphs_24h <- vector(mode = "list", length = length(fDOM_loop))
names(graphs_24h) <- fDOM_loop

for (var in fDOM_loop) {
  
  df <- fDOM_24h %>% filter(fdom == var)
  
  graphs_24h[[var]] <- df %>%  
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("24h", var, "concentrations"),
       x= "Treatment",
       y = "Concentration (µmol/L)")+
    facet_wrap(~time_fac, ncol = 4 )+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none")
}

```

### Humics
#### CobleA

```{r}
summary(models_24h[["coble_a"]])
anova(models_24h[["coble_a"]], type = 2)
emmeans(models_24h[["coble_a"]], specs = trt.vs.ctrl ~ treatment)
```

```{r}
resid_panel(models_24h[["coble_a"]])
graphs_24h[["coble_a"]]
```

#### CobleM

```{r}
summary(models_24h[["coble_m"]])
anova(models_24h[["coble_m"]], type = 2)
emmeans(models_24h[["coble_m"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[[1]], ~ treatment)
```

```{r}
resid_panel(models_24h[["coble_m"]])
graphs_24h[["coble_m"]]
```

#### Sig! CobleC 
Sig ANOVA and Contrasts with and without log10 outlier detection method
```{r}
summary(models_24h[["coble_c"]])
anova(models_24h[["coble_c"]], type = 2)
emmeans(models_24h[["coble_c"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["coble_c"]])
graphs_24h[["coble_c"]]
```


#### Sig! M to C 
Sig ANOVA and Contrasts with and without log10 outlier detection method

```{r}
summary(models_24h[["m_to_c"]])
anova(models_24h[["m_to_c"]], type = 2)
emmeans(models_24h[["m_to_c"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["m_to_c"]])
graphs_24h[["m_to_c"]]
```


#### Sig! FI  
Sig ANOVA and Contrasts with and without log10 outlier detection method

```{r}
summary(models_24h[["fi"]])
anova(models_24h[["fi"]], type = 2)
emmeans(models_24h[["fi"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["fi"]])
graphs_24h[["fi"]]
```


#### BIX

```{r}
summary(models_24h[["bix"]])
anova(models_24h[["bix"]], type = 2)
emmeans(models_24h[["bix"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["bix"]])
graphs_24h[["bix"]]
```

#### HIX *Sig ANOVA
Sig ANOVA with log10 outlier detection method but not the contrasts
Not sig without log10

```{r}
summary(models_24h[["hix"]])
anova(models_24h[["hix"]], type = 2)
emmeans(models_24h[["hix"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["hix"]])
graphs_24h[["hix"]]
```

### Proteinaceous
#### CobleB

```{r}
summary(models_24h[["coble_b"]])
anova(models_24h[["coble_b"]], type = 2)
emmeans(models_24h[["coble_b"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["coble_b"]])
graphs_24h[["coble_b"]]
```


#### CobleT

```{r}
summary(models_24h[["coble_t"]])
anova(models_24h[["coble_t"]], type = 2)
emmeans(models_24h[["coble_t"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["coble_t"]])
graphs_24h[["coble_t"]]
```


#### Fpeak

```{r}
summary(models_24h[["fpeak"]])
anova(models_24h[["fpeak"]], type = 2)
emmeans(models_24h[["fpeak"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["fpeak"]])
graphs_24h[["fpeak"]]
```

