---
title: "MOBEAST_all_variables"
created: 5 Fev 2025
edited: 25 Fev 2025
format: html
editor: source

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(janitor)
library(here)

library(ggrepel)
library(patchwork)
library(RColorBrewer)
library(measurements)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
library(emmeans)
library(performance)
library(multcompView)
library(broom.mixed)

# PCA packages
library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(factoextra) 
```


## Data visualization theme

```{r}
nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=24, face = "bold"), 
        plot.subtitle = element_text(size=20),
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=18, face = "bold"),
        legend.text=element_text(size=16),
        strip.background = element_rect(fill = "white"))
```



```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

## Load Data

```{r}
mobeast <- read_csv(here("data", "MOBEAST_full_merged_data_long.csv"))
#mobeast <- read_csv(here("data", "MOBEAST_full_merged_data.csv"))
```


## Data Clean Up

### Outlier detection on bulk data
```{r}
## flagging outliers from the data set
 
mobeast_flagged <- mobeast %>%
  mutate(time_string = as.character(time_string),
              date_string = as.character(date_string),
              date_time_string = paste(date_string,time_string))%>% 
  filter(!is.na(id_number)) %>% 
  # remove the DNA data because they should be treated differently
  filter(variable_cat != "DNA" ) %>% 
  filter(treatment != "Rain") %>% 
  filter(treatment != "Endmember") %>% 
  
  ## remove DOC values greater than 208
  #mutate(value = ifelse(variable == "npoc_u_m" & value >208, NA , value)) %>%

  # flagging outliers
  ## grouping by treatment, variable 
  group_by(variable) %>% 
  ## get the mean and the sd to calculate mean +/- 2sd
  mutate(mean =mean(value, na.rm = T), 
         stand_dev = sd(value, na.rm = T)) %>% 
  ## Values outside of the mean +/- 2sd range are flagged as outliers
  ## Important to filter outliers later to understand why they are outliers. 
  mutate(outlier = ifelse(value > mean +2*stand_dev | value < mean -2*stand_dev,
                           "outlier", "not outlier")) %>% 
  ungroup() 
  
```

```{r}
mobeast_flagged %>% filter(variable_cat == "nutrients") %>% 
  select(variable) %>% distinct()
```


### Biogeochemistry 
#### Calculating Fluxes 
(([aquarium]-[inflow])*Vol) / (RT*SA ) in [umol/SA/time]. 
```{r}
tank_area = conv_unit(506.25, from = "cm2", to = "m2") # convert area cm2 to m2
tank_vol = 10 # liters
# DOC (npoc_u_m) in uM, i.e. umol/L
# flowrate in mL/s converted to L/sec
# residence_time in hours
# areal_flux in umol/s/m2
```

```{r}

mobeast_flux <- mobeast_flagged %>%  
  
# Isolate inflow for each date_time and table
  group_by(date_time, inflow_table, variable) %>%
  mutate(inflow = value[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time, variable) %>% 
  #mutate(inflow = ifelse(outlier == "outlier", inflow == NA, inflow)) %>% 
  mutate(inflow = mean(inflow, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(conc_norm = value - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  # convert res time from h to sec
  mutate(residence_time = residence_time*3600) %>%
  
  # There is no residence time in 
  mutate(residence_time = ifelse(date_time_string == "2024-06-04 12:00:00", (residence_time[date_time_string == "2024-06-02 12:00:00"] + residence_time[date_time_string == "2024-06-06 12:00:00"])*0.5, residence_time)) %>% 

  # Unsure why we still normalize to the inflow so calculated both 
  # fluxes with raw data and normalized data.
  mutate(areal_flux = conc_norm*tank_vol/ (residence_time*tank_area)) %>% 
  filter(treatment != "Inflow") %>% 
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) %>% 
  select(-mean, -stand_dev, -inflow)

```

```{r}
## Some variables will not be studied in this project
exclude_var <- c("carbonate", "residence time")
# we also won't focus on the other time points, they are mainly for the FCM data and not other data was collected at those times anyway
exclude_time <- c("03:00:00","09:00:00","15:00:00","21:00:00")
mobeast_flux <- mobeast_flux %>% filter(!variable_cat %in%  exclude_var,
                        !time_string %in% exclude_time) 
```


```{r}
mobeast %>% filter(inflow_table != "Endmember", variable == "syn") %>% 
  group_by(inflow_table, date_time_string, tank_number) %>% reframe(mean = mean(value)) %>% drop_na() %>% 
  ggplot(aes(x= date_time_string, y = mean, group = tank_number, color = tank_number))+
  geom_point(aes(size = tank_number))+
  geom_line(aes(linetype = tank_number, linewidth = tank_number))+ 
  facet_wrap(.~ inflow_table)+
  labs(title = "Synechococcus Variation", 
       x= "Date", y = "Average Cell Count")+
  scale_size_manual(
    values = c("F1" = 4, "F2" = 4, 
               "T1"= 2, 
               "T2"=2, 
               "T10"= 2,
               "T11"= 2,
               "T5" = 2)) +
  scale_linetype_manual(values = c("F1" = "solid", 
                                   "F2" = "solid",
                                   "T1"= "dashed", 
                                   "T2"= "dashed",
                                   "T10"= "dashed",
                                   "T11"= "dashed",
                                   "T5" = "dashed"))+
  scale_linewidth_manual(
    values = c("F1" = 1.5, "F2" = 1.5, 
               "T1"= 0.5, 
               "T2"=0.5, 
               "T10"= 0.5,
               "T11"= 0.5,
               "T5" = 0.5))+
  theme_bw()+
  theme(legend.position = "none")
```




```{r}
# checking the residence time variations 
resid_time_plot_12PM <- mobeast_flux %>% select(date_time, time_string, treatment, residence_time, flowrate) %>% distinct() %>% 
  group_by(date_time, treatment) %>% 
  reframe(time_string, avg_flow = mean(flowrate, na.rm = TRUE)) %>% 
  ungroup() %>% 
  #filter(time_string =="12:00:00") %>% 
  ggplot(aes(x= date_time, y = avg_flow, color = treatment))+
  geom_point()+
  geom_line(aes(group= treatment))+
  labs(title = "Avg Residence Time",
       x= "date time",
       y = "avg residence time (h)")+
  scale_color_manual(values = treatment_colors)+ theme_bw()
resid_time_plot_12PM
```

```{r}
# check the avg areal flux for a variable and how it changes over the study period
flux_12PM_plot <- mobeast_flux %>% 
  filter(variable == "npoc_u_m") %>% 
  group_by(date_time, treatment) %>% 
  reframe(areal_flux,time_string, avg_flux = mean(areal_flux, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(time_string =="12:00:00") %>% 
  ggplot(aes(x= date_time, y = avg_flux, color = treatment,group = treatment))+
  geom_point()+
  geom_line()+
  geom_point(aes(y = areal_flux), size = 2)+
  labs(title = "Avg Areal Flux",
       x= "date time",
       y = "avg areal flux")+
  scale_color_manual(values = treatment_colors)+
  theme_bw()
flux_12PM_plot
```

```{r}
resid_time_plot_12PM +flux_12PM_plot
```


```{r}
mobeast_flux %>%  select(treatment, inflow_table, tank_number, flowrate) %>% 
  mutate(mean_all = mean(flowrate, na.rm = T), 
         sd_all = sd(flowrate, na.rm = T)) %>% 
  group_by(tank_number) %>% 
  reframe(mean = mean(flowrate, na.rm = T),
          sd = sd(flowrate, na.rm = T)#, 
          #mean_all, sd_all
          ) %>% 
  ungroup()

```



### DNA data 
```{r}
# prepaing the data for later
mobeast_DNA  <- mobeast %>%
  mutate(time_string = as.character(time_string),
              date_string = as.character(date_string),
              date_time_string = paste(date_string,time_string))%>% 
  filter(variable_cat == "DNA" ) %>% 
  filter(treatment != "Rain") %>% 
  filter(treatment != "Endmember") %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
  
  

```


## 12PM data
```{r}
# removing outliers from the data set
mobeast_12PM <- mobeast_flux %>%
  filter(time_string == "12:00:00") %>% 
  
  ## remove the outliers 
  filter(outlier == "not outlier") 



```
### Summary stats

```{r}
mobeast_12PM %>% group_by(variable,treatment) %>% 
  reframe(max = max(areal_flux, na.rm = TRUE),
          min = min(areal_flux, na.rm = TRUE),
    mean = mean(areal_flux, na.rm = TRUE),
          sd = sd(areal_flux, na.rm = TRUE))
```

```{r}
## removing the calculated outliers from the areal flux numbers 
mobeast_12PM %>% group_by(treatment, variable) %>% 
  mutate(mean = mean(areal_flux, na.rm = TRUE),
    sd = sd(areal_flux, na.rm = TRUE)) %>% 
  mutate(flux_outlier = ifelse(areal_flux > mean +2*sd | areal_flux < mean -2*sd,"outlier", "not outlier")) %>% 
  filter(variable == "het_bact") %>% 
  filter(flux_outlier != "outlier") %>% 
  ggplot(aes(x = treatment, y = areal_flux, fill = treatment)) +
      geom_violin() +
      geom_point() +
      geom_hline(yintercept = 0, linetype = "dashed") +
      stat_summary(color = "yellow", alpha = 0.5) +
      labs(title = paste("12PM Heterotrophic Bacteria Production"),
           x = "Treatment",
          y = "Areal Flux (indiv/(s.m2)") +
      scale_fill_manual(
        breaks = c("Control", "Coral", "CCA", "Algae"),
        values = c("grey", "#03A9F4", "#EF5350", "#4CAF50")
      ) +
      nunu_theme +
      theme(legend.position = "none")
```

### Lmer models  - Iterative  coding 

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

mobeast_loop <- mobeast_12PM %>% select(variable) %>% distinct() %>% pull()

models_12PM <- vector(mode = "list", length = length(mobeast_loop))
names(models_12PM) <- mobeast_loop

for (var in mobeast_loop) {
  df <- mobeast_12PM %>% filter(variable == var)
  models_12PM[[var]] <- lmer(areal_flux~ treatment + (1|tank_number), data = df)

}

```

#### singularity test 
```{r}
## check if the model soliution is singular and/or converges
models <- mobeast_12PM %>% group_by(variable) %>%
  group_modify(~ {
    mod <- lmer(areal_flux ~ treatment + (1|tank_number), data = .x)

    tibble(
      model = list(mod),
      singular = isSingular(mod),
      converged = is.null(mod@optinfo$conv$lme4$messages)
    )
  }) %>%
  ungroup() %>% filter(singular == TRUE)
```

#### model results (fixed/random)
```{r}
# save the outputs of the model in a dataframe (fixed and random effects)
results <- imap_dfr(models_12PM, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")
```

#### ANOVA results
```{r, warning= FALSE}
# So many Warnings! 
## Warning: The column names NumDF and DenDF in ANOVA output were not recognized or transformed.

# run anova on the models 
anova_results <- imap_dfr(models_12PM, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

## which variables have significant anova results
anova_results %>% filter(p.value <0.05)
```


```{}
## making sure that the code produces accurate outputs
test_12PM<- mobeast_12PM %>%  filter(variable =="npoc_u_m")
test_12PM_mod <- lmer(areal_flux ~ treatment + (1|tank_number), data = test_12PM)

anova(test_12PM_mod, type = 2)
```

#### emmeans results 
```{r}
# "estimate" are negative somehow. Why? 
## in manual models for DOC, estimate >0, here it's <0
## t.ratio in "statistics", and negative instead of positive
## p.values not shown, only p.adjusted. 
emmeans_results <- imap_dfr(models_12PM, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 


emmeans_results_sig <- emmeans_results %>% 
  filter(adj.p.value <0.05)

emmeans_results
emmeans_results_sig
```

```{}
## making sure that the code produces accurate outputs
emmeans(test_12PM_mod,specs = trt.vs.ctrl ~ treatment)
```


```{r}
emmean_sig <- emmeans_results_sig %>% pull(variable)
```

```{, warning=FALSE}
mobeast_12PM %>% filter(variable %in% emmean_sig) %>% 
  #filter(variable == "SiO2") %>% 
  ggplot(aes(x= treatment, y= areal_flux, fill = treatment))+
    geom_violin()+
    geom_jitter(aes(color = outlier), alpha = 0.5)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_wrap(~variable, scales = "free")+
    labs(title = paste("12PM Biogeochem Changes"),
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Flux XXX")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18))
```

### Lmer flux outliers 
```{r}
## removing the calculated outliers from the areal flux numbers 
mobeast_12PM_flux <- mobeast_12PM %>% group_by(variable) %>% 
  mutate(mean = mean(areal_flux, na.rm = TRUE),
    sd = sd(areal_flux, na.rm = TRUE)) %>% 
  mutate(flux_outlier = ifelse(areal_flux > mean +2*sd | areal_flux < mean -2*sd,"outlier", "not outlier")) %>% 
  ungroup() %>% 
  filter(flux_outlier != "outlier") 
  
```


#### Lmer models  - Iterative  coding 

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

mobeast_loop <- mobeast_12PM_flux %>% select(variable) %>% distinct() %>% pull()

models_12PM_2 <- vector(mode = "list", length = length(mobeast_loop))
names(models_12PM_2) <- mobeast_loop

for (var in mobeast_loop) {
  df <- mobeast_12PM_flux %>% filter(variable == var)
  models_12PM_2[[var]] <- lmer(areal_flux~ treatment + (1|tank_number), data = df)

}

```

#### singularity test
```{}
## check if the model soliution is singular and/or converges
models <- mobeast_12PM %>% group_by(variable) %>%
  group_modify(~ {
    mod <- lmer(areal_flux ~ treatment + (1|tank_number), data = .x)

    tibble(
      model = list(mod),
      singular = isSingular(mod),
      converged = is.null(mod@optinfo$conv$lme4$messages)
    )
  }) %>%
  ungroup() %>% filter(singular == TRUE)
```

#### model results 
```{r}
# save the outputs of the model in a dataframe (fixed and random effects)
results <- imap_dfr(models_12PM, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")
```

##### model verification
```{r}
DOC_12PM_flux <- mobeast_12PM_flux %>%  filter(variable == "npoc_u_m")

mod_DOC_12_flux <- lmer(areal_flux ~ treatment + (1|tank_number), data = DOC_12PM_flux)

anova(mod_DOC_12_flux, type = 2)

emmeans(mod_DOC_12_flux, specs = pairwise ~ treatment)
```


#### ANOVA results 
```{r, warning= FALSE}
# So many Warnings! 
## Warning: The column names NumDF and DenDF in ANOVA output were not recognized or transformed.

# run anova on the models 
anova_results <- imap_dfr(models_12PM_2, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

## which variables have significant anova results
anova_results %>% filter(p.value <0.05)
# tot-bact, syn, fpeak and npoc are significant. 
```


#### Emmeans results 
```{r}
# "estimate" are negative somehow. Why? 
## in manual models for DOC, estimate >0, here it's <0
## t.ratio in "statistics", and negative instead of positive
## p.values not shown, only p.adjusted. 
emmeans_results <- imap_dfr(models_12PM_2, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 


emmeans_results_sig <- emmeans_results %>% 
  filter(adj.p.value <0.05)

emmeans_results
emmeans_results_sig # there are no significant variables....
```

```{}
## making sure that the code produces accurate outputs
emmeans(test_12PM_mod,specs = trt.vs.ctrl ~ treatment)
```

#### Graphing the significant variables
```{r}
emmean_sig <- emmeans_results_sig %>% pull(variable)
```

```{, warning=FALSE}
mobeast_12PM %>% filter(variable %in% emmean_sig) %>% 
  #filter(variable == "SiO2") %>% 
  ggplot(aes(x= treatment, y= areal_flux, fill = treatment))+
    geom_violin()+
    geom_jitter(aes(color = outlier), alpha = 0.5)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_wrap(~variable, scales = "free")+
    labs(title = paste("12PM Biogeochem Changes"),
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Flux XXX")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18))
```




###  T-test
```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
t_test_12PM <- mobeast_12PM %>%
  group_by(variable, treatment) %>%
  summarize(
    t_test = list(t.test(areal_flux, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  ) %>% ungroup()

sig_var_12PM <- t_test_12PM %>% 
  filter(p_value <0.05) %>%  
  select(variable) %>% distinct() %>% pull()

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```

#### Graphing variables with significant results
```{r, warning=FALSE}

graphs_12PM <- sig_var_12PM %>%
  set_names() %>%
  map(~ {
    df <- mobeast_12PM %>% filter(variable == .x)
    
    ggplot(df, aes(x = treatment, y = areal_flux, fill = treatment)) +
      geom_violin() +
      geom_point() +
      geom_hline(yintercept = 0, linetype = "dashed") +
      stat_summary(color = "yellow", alpha = 0.5) +
      labs(title = paste("12PM", .x, "Production"),
           x = "Treatment",
          y = "Areal Flux (µmol/(s.m2)") +
      scale_fill_manual(
        breaks = c("Control", "Coral", "CCA", "Algae"),
        values = c("grey", "#03A9F4", "#EF5350", "#4CAF50")
      ) +
      nunu_theme +
      theme(legend.position = "none")
  })

graphs_12PM[["m_to_c"]] / graphs_12PM[["bix"]] / graphs_12PM[["hix"]]/ graphs_12PM[["coble_b"]] + plot_layout(axes = "collect")

graphs_12PM[["syn"]]+ labs(y = "Areal Flux (indiv/(s.m2))")
graphs_12PM[["het_bact"]]+ labs(title = "12PM Heterotrophic bacteria production", y = "Areal Flux (indiv/(s.m2))")

```

##### Clean Plots for t-test results

```{r}
variable_meta <- mobeast_12PM %>% 
  select(variable, variable_cat) %>%  distinct()
t_test_12PM %>%  left_join(variable_meta) %>%  filter(variable_cat == "fDOM") %>% 
  mutate(sig = ifelse(p_value<0.05 & mean > 0, "Enrichment", 
                                         ifelse(p_value<0.05 & mean < 0, "Depletion", "No Difference")))%>% 
  mutate(sig = factor(sig, 
                      levels = c("Enrichment","No Difference", "Depletion"))) %>% 
  arrange(variable) %>%   # ensures ordering in the data frame
  mutate(variable = factor(variable, levels = c(
    "coble_a","coble_c","coble_m", "m_to_c","bix","hix","fi",
    "coble_b", "coble_t","fpeak","lignin")),
    variable = fct_rev(variable)) %>% 
  
  ggplot(aes(x= treatment, y= variable, fill = sig))+
  geom_tile(color = "black")+
  labs(title = "fDOM Enrichment and Depletion Pattern (p-value) ")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw()
  
```


###  T-test flux outliers 
```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
t_test_12PM <- mobeast_12PM_flux %>% filter(flux_outlier != "outlier") %>% 
  group_by(variable, treatment) %>%
  summarize(
    t_test = list(t.test(areal_flux, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  ) %>% ungroup()

sig_var_12PM <- t_test_12PM %>% 
  filter(p_value <0.05) %>%  
  select(variable) %>% distinct() %>% pull()

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```

```{r, warning=FALSE}

graphs_12PM <- sig_var_12PM %>%
  set_names() %>%
  map(~ {
    df <- mobeast_12PM_flux %>% filter(variable == .x)
    
    ggplot(df, aes(x = treatment, y = areal_flux, fill = treatment)) +
      geom_violin() +
      geom_point() +
      geom_hline(yintercept = 0, linetype = "dashed") +
      stat_summary(color = "yellow", alpha = 0.5) +
      labs(title = paste("12PM", .x, "Production"),
           x = "Treatment",
          y = "Areal Flux (µmol/(s.m2)") +
      scale_fill_manual(
        breaks = c("Control", "Coral", "CCA", "Algae"),
        values = c("grey", "#03A9F4", "#EF5350", "#4CAF50")
      ) +
      nunu_theme +
      theme(legend.position = "none")
  })

#graphs_12PM[["m_to_c"]] / graphs_12PM[["bix"]] / graphs_12PM[["hix"]]/ #graphs_12PM[["coble_b"]] + plot_layout(axes = "collect")

graphs_12PM[["syn"]] +graphs_12PM[["pro"]]+ graphs_12PM[["het_bact"]] + labs(y = "Areal Flux (indiv/(s.m2))")
#graphs_12PM[["het_bact"]]+ labs(title = "12PM Heterotrophic bacteria production", y = "Areal Flux (indiv/(s.m2))")

#graphs_12PM[["npoc_u_m"]]
```


##### Clean Plots for t-test results

```{r}
variable_meta <- mobeast_12PM_flux %>% 
  select(variable, variable_cat) %>%  distinct()

variable_meta %>% select(variable_cat) %>%  distinct()
t_test_12PM %>%  left_join(variable_meta) %>%  #filter(variable_cat == "fDOM") %>% 
  mutate(sig = ifelse(p_value<0.05 & mean > 0, "Enrichment", 
                                         ifelse(p_value<0.05 & mean < 0, "Depletion", "No Difference")))%>% 
  mutate(sig = factor(sig, 
                      levels = c("Enrichment","No Difference", "Depletion"))) %>%
  arrange(variable_cat, variable) %>%
  mutate(variable = factor(variable, levels = unique(variable))) %>% 
  
  ggplot(aes(x= treatment, y= variable, fill = sig))+
  geom_tile(color = "black")+
  facet_grid(variable_cat ~ ., scales = "free_y", space = "free_y")+
  labs(title = "fDOM Enrichment and Depletion Pattern (p-value) ")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw()
  
```



## 24h data

```{r}
# subset the data set to onl the 24h samples
## removing outliers from the data set
mobeast_24h <- mobeast_flux %>%
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  filter(!is.na(id_number)) %>% 
  # remove the the DOC data because something happened  at 00:00. 
  filter(variable != "npoc_u_m") %>%  
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00"))) 
DOC_24h <- mobeast_flux %>%
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  filter(!is.na(id_number)) %>% 
  # remove the the DOC data because something happened  at 00:00. 
  filter(variable == "npoc_u_m") %>%  
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00"))) %>% 
  filter(time_string != "00:00:00") 
```

### Summary Stats 
```{r}
mobeast_flux %>%
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  filter(!is.na(id_number)) %>% 
  # remove the the DOC data because something happened  at 00:00. 
  #filter(variable == "npoc_u_m") %>%  
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  group_by(variable,time_string, treatment ) %>% 
  reframe(max = max(areal_flux, na.rm = TRUE),
          min = min(areal_flux, na.rm = TRUE),
    mean = mean(areal_flux, na.rm = TRUE),
          sd = sd(areal_flux, na.rm = TRUE)) %>% 
  filter(variable == "npoc_u_m") %>%  
  arrange(mean)
```

### Data verifications
```{r}
# chekcing out the residence time variation over 24h 
resid_time_plot_24h <- mobeast_24h %>% select(date_time, time_string, treatment, residence_time, flowrate) %>% distinct() %>% 
  group_by(date_time, treatment) %>% 
  reframe(time_string, avg_flow = mean(flowrate, na.rm = TRUE)) %>% 
  ungroup() %>% 

  ggplot(aes(x= date_time, y = avg_flow, color = treatment))+
  geom_point()+
  geom_line(aes(group= treatment))+
  labs(title = "Avg Flow Rate",
       x= "date time",
       y = "avg flow rate")+
  scale_color_manual(values = treatment_colors)+ theme_bw()
resid_time_plot_24h
```

```{r}
flux_24h_plot <- mobeast_24h %>% 
  filter(variable == "coble_a") %>% 
  group_by(date_time, treatment) %>% 
  reframe(areal_flux,time_string, avg_flux = mean(areal_flux, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(x= date_time, y = avg_flux, color = treatment,group = treatment))+
  geom_point()+
  geom_line()+
  #geom_point(aes(y = areal_flux), size = 2)+
  labs(title = "Avg Areal Flux",
       x= "date time",
       y = "avg areal flux")+
  scale_color_manual(values = treatment_colors)+
  theme_bw()
flux_24h_plot

resid_time_plot_24h + flux_24h_plot
```


### LMER models

```{r}
# Loop Function:
## Take the 24h samples and isolate the types of nutrient  present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

mobeast_loop <- mobeast_24h %>% select(variable) %>% distinct() %>% pull()

models_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- mobeast_24h %>% filter(variable == var)
  models_24h[[var]] <- lmer(areal_flux~ treatment*time_fac + (1|tank_number), data = df)
}


## separate model for DOC because 00:00 removed 
DOC_24h_mod <- lmer(areal_flux~ treatment*time_fac + (1|tank_number), data = DOC_24h)

```

#### model results (fixed and random)
```{r}
results <- imap_dfr(models_24h, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")

DOC_24h_results <- tidy(DOC_24h_mod) %>% mutate(variable = "npoc_u_m") %>% relocate(variable, .before = "effect")

results <- rbind(results, DOC_24h_results)
```

#### ANOVA results
```{r, warning= FALSE}
# So many Warnings! 
## Warning: The column names NumDF and DenDF in ANOVA output were not recognized or transformed.

anova_results <- imap_dfr(models_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

DOC_anova <- anova(DOC_24h_mod, type = 2) %>%  tidy() %>% mutate(variable = "npoc_u_m") %>% relocate(variable, .before = "term")

anova_results <- rbind(anova_results, DOC_anova)

anova_results %>% filter(p.value <0.05) 
```

#### emmeans results
```{r}
# "estimate" are negative somehow. Why? 
## in manual models for DOC, estimate >0, here it's <0
## t.ratio in "statistics", and negative instead of positive
## p.values not shown, only p.adjusted. 
emmeans_results <- imap_dfr(models_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 

DOC_emmeans <- emmeans(DOC_24h_mod, spec= pairwise ~ treatment)$contrast %>% 
  tidy() %>% 
  mutate(variable = "npoc_u_m") %>% 
  relocate(variable, .before = "term") %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 

emmeans_results <- rbind(emmeans_results, DOC_emmeans)

```

```{r}
emmeans_results_sig <- emmeans_results %>% 
  filter(adj.p.value <0.05)
emmean_sig <- emmeans_results_sig %>% pull(variable)
```


##### Violin plots
```{r, warning=FALSE}
mobeast_24h %>% 
  filter(variable == "syn") %>% 
  ggplot(aes(x= treatment, y= areal_flux, fill = treatment))+
    geom_violin()+
    #geom_jitter(aes(color = outlier), alpha = 0.5)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_wrap( .~ time_fac, ncol = 4 )+
    labs(title = paste("24h Synechococcus Changes"),
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Areal Flux (indiv/(s.m2)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18),
          legend.position = "none")
```


### Lmer Flux outliers 

```{r}
mobeast_24h_flux <- mobeast_24h %>%  group_by( variable) %>% 
  mutate(mean = mean(areal_flux, na.rm = TRUE),
    sd = sd(areal_flux, na.rm = TRUE)) %>% 
  mutate(flux_outlier = ifelse(areal_flux > mean +2*sd | areal_flux < mean -2*sd,"outlier", "not outlier")) %>% 
  ungroup() %>% 
  filter(flux_outlier != "outlier")

DOC_24h_flux <- DOC_24h %>%  group_by( variable) %>% 
  mutate(mean = mean(areal_flux, na.rm = TRUE),
    sd = sd(areal_flux, na.rm = TRUE)) %>% 
  mutate(flux_outlier = ifelse(areal_flux > mean +2*sd | areal_flux < mean -2*sd,"outlier", "not outlier")) %>% 
  ungroup() %>% 
  filter(flux_outlier != "outlier")
```

```{r}
# Loop Function:
## Take the 24h samples and isolate the types of nutrient  present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

mobeast_loop <- mobeast_24h_flux %>% select(variable) %>% distinct() %>% pull()

models_24h_flux <- vector(mode = "list", length = length(mobeast_loop))
names(models_24h_flux) <- mobeast_loop


for (var in mobeast_loop) {
  df <- mobeast_24h_flux %>% filter(variable == var)
  models_24h_flux[[var]] <- lmer(areal_flux~ treatment*time_fac + (1|tank_number), data = df)
}


## separate model for DOC because 00:00 removed 
DOC_24h_mod_flux <- lmer(areal_flux~ treatment*time_fac + (1|tank_number), data = DOC_24h_flux)

```

#### model results (fixed and random)
```{r}
results <- imap_dfr(models_24h_flux, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")

DOC_24h_results <- tidy(DOC_24h_mod_flux) %>% mutate(variable = "npoc_u_m") %>% relocate(variable, .before = "effect")

results <- rbind(results, DOC_24h_results)
```

#### ANOVA results
```{r, warning= FALSE}
# So many Warnings! 
## Warning: The column names NumDF and DenDF in ANOVA output were not recognized or transformed.

anova_results <- imap_dfr(models_24h_flux, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

DOC_anova <- anova(DOC_24h_mod_flux, type = 2) %>%  tidy() %>% mutate(variable = "npoc_u_m") %>% relocate(variable, .before = "term")

anova_results <- rbind(anova_results, DOC_anova)

anova_results %>% filter(p.value <0.05) 

sig_anova_flux_var <- anova_results %>% filter(p.value <0.05) %>% 
  group_by(variable) %>% 
  filter(
    !(all(c("time_fac", "treatment:time_fac") %in% term))
  ) %>% 
  ungroup() %>% 
  filter(term == "treatment", variable != "tot_bact") %>% 
   select(variable) %>% distinct() %>% pull()
```

#### emmeans results
```{r}
# "estimate" are negative somehow. Why? 
## in manual models for DOC, estimate >0, here it's <0
## t.ratio in "statistics", and negative instead of positive
## p.values not shown, only p.adjusted. 
emmeans_results <- imap_dfr(models_24h_flux, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 

DOC_emmeans <- emmeans(DOC_24h_mod_flux, spec= pairwise ~ treatment)$contrast %>% 
  tidy() %>% 
  mutate(variable = "npoc_u_m") %>% 
  relocate(variable, .before = "term") %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 

emmeans_results <- rbind(emmeans_results, DOC_emmeans)

```

```{r}
emmeans_results_sig <- emmeans_results %>% filter(variable %in% sig_anova_flux_var) %>% 
  filter(adj.p.value <0.05) %>% 
  select(variable, term, contrast,estimate, adj.p.value)
emmean_sig <- emmeans_results_sig %>% select(variable) %>% 
  distinct() %>% pull(variable)
```
##### Violin plots
```{r, warning=FALSE}
mobeast_24h_flux %>% 
  filter(variable == "het_bact") %>% 
  ggplot(aes(x= treatment, y= areal_flux, fill = treatment))+
    geom_violin()+
    #geom_jitter(aes(color = outlier), alpha = 0.5)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_wrap( .~ time_fac, ncol = 4 )+
    labs(title = paste("24h Synechococcus Changes"),
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Areal Flux (indiv/(s.m2)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18),
          legend.position = "none")
```

### DNA 

```{r}
mobeast_DNA_24h <- mobeast_DNA %>% filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(id_number)) %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00")))

exclude_family <- mobeast_DNA_24h %>% group_by(time, variable) %>% 
  reframe(avg = mean(value)) %>% filter(avg == 0) %>%  select(variable) %>% distinct() %>% pull()
```


```{r}
# Loop Function:
## Take the 24h samples and isolate the types of nutrient  present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

mobeast_loop <- mobeast_DNA_24h %>% filter(!variable %in% exclude_family) %>% select(variable) %>% distinct() %>% pull()

models_DNA_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_DNA_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- mobeast_DNA_24h %>% filter(variable == var)
  models_DNA_24h[[var]] <- lmer(value~ treatment*time_fac + (1|tank_number), data = df)

}

```


```{r}
results_DNA <- imap_dfr(models_DNA_24h, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")


```


```{r, warning= FALSE}
# So many Warnings! 
## Warning: The column names NumDF and DenDF in ANOVA output were not recognized or transformed.

anova_results <- imap_dfr(models_DNA_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

DNA_anova_sig <- anova_results %>% filter(p.value <0.05) %>% select(variable, anova_term = term, anova_p_val = p.value)


```

```{r}
# "estimate" are negative somehow. Why? 
## in manual models for DOC, estimate >0, here it's <0
## t.ratio in "statistics", and negative instead of positive
## p.values not shown, only p.adjusted. 
emmeans_results <- imap_dfr(models_DNA_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 


```

```{r}
emmeans_results_sig <- emmeans_results %>% filter(adj.p.value <0.05) %>% 
  select(variable, emmean_term = term, contrast, emmeans_p_val = adj.p.value) 

## Families that have a time:treatment interactions in ANOVA
interaction_family <- c("Francisellales_Francisellaceae_Francisella",
"Fusobacteriales_Leptotrichiaceae_Hypnocyclicus",
"Oceanospirillales_Saccharospirillaceae_Oleibacter",
"Synechococcales_Cyanobiaceae_Prochlorococcus_MIT9313")

emmeans_results_sig %>% left_join(DNA_anova_sig) %>% filter(!variable %in% interaction_family) %>% select(variable) %>%  distinct() %>% pull()

```

```{r}
DNA_anova_sig %>% filter(!variable %in% interaction_family,
                         anova_p_val <0.05, 
                         anova_term == "treatment") %>% 
  select(variable) %>% distinct()

DNA_anova_sig %>% filter(!variable %in% interaction_family,
                         anova_p_val <0.05, 
                         anova_term == "treatment") %>% 
  select(variable) %>% distinct()
```



```{r, warning=FALSE}
DNA_sig_var <- emmeans_results %>% 
  filter(adj.p.value <0.05) %>% 
  select(variable) %>% distinct() %>% pull()

graphs_DNA_24h <- DNA_sig_var %>%
  set_names() %>%
  map(~ {
    df <- mobeast_DNA_24h %>% filter(variable == .x)
    
    ggplot(df, aes(x = treatment, y = value, fill = treatment)) +
      geom_violin() +
      geom_point() +
      geom_hline(yintercept = 0, linetype = "dashed") +
      stat_summary(color = "yellow", alpha = 0.5) +
      labs(title = paste("12PM", .x, "Production"),
           x = "Treatment",
          y = "Areal Flux (µmol/(s.m2)") +
      scale_fill_manual(
        breaks = c("Control", "Coral", "CCA", "Algae", "Inflow"),
        values = c("grey", "#03A9F4", "#EF5350", "#4CAF50", "orange")
      ) +
      nunu_theme +
      theme(legend.position = "none")
  })


```




```{r}
emmeans_results %>%   
  mutate(sig = ifelse(adj.p.value<0.05 & estimate < 0, "Enrichment", 
                                         ifelse(adj.p.value<0.05 & estimate > 0, "Depletion", "No Difference")))%>% 
  mutate(sig = factor(sig, 
                      levels = c("Enrichment","No Difference", "Depletion"))) %>% 
  filter(variable %in% DNA_sig_var)  %>%
  filter(variable %in% interaction_family) %>% 
  mutate(contrast = factor(contrast, levels = c("Control - Coral", "Control - CCA", "Control - Algae"))) %>% 
  ggplot()+
  geom_tile(aes(x= contrast, y= variable, fill = sig), color = "black")+
  #facet_wrap(.~variable_cat)+
  labs(title = "24h Contrast (Emmeans): Control vs. Treatment",
       x= "Treatment",
       y = " Variables")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw() +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))# + 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")
  
```



### Pairwise comparison 

#### Pairwise comparison function

```{r}
nunu_pairwise <- function(data, type = "coble_a", time_point = 6){

  ## Get a list of the treatments for the for loop   
  treatment <- data %>% mutate(treatment = as.character(treatment)) %>%
  select(treatment) %>%  distinct() %>% pull()
  
  ## Create un empty data frame
pairwise_results <- tibble(
  time_point = vector(),
  treatment= vector(),
  mean = vector(),
  p_value = vector())

# this is the pairwise comparison for the 6 pairs (6AM-6PM)
  if(time_point == 6){
    for(var in treatment){
      
      ## only keep the 6AM and 6PM
      df <- data %>% 
        filter(variable == type) %>% 
        filter(treatment == var) %>% 
  filter(time_string == "18:00:00" | time_string == "06:00:00") %>% 
        ## Rename the times so they can become column names
  mutate(time_US = ifelse(time_string == "18:00:00", "T18", "T06")) %>% 
        ## Only keep the column of interest for the stats
  select(tank_number, time_US, areal_flux) %>% 
  pivot_wider(names_from = time_US, values_from = areal_flux) 
      
      ## Pairwise function 
      paired <- t.test(df$T18, df$T06, paired = TRUE, na.action = na.pass)
      
      ## Save the result in at data frame
      df2 <- tibble(
        time_point = "06AM-06PM",
        variable = type,
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      # Code will loop oveer all treatments and save the results in a data frame
      pairwise_results <-rbind(pairwise_results,df2)
      
      # Save the results in RStudio
      pairwise_results <<- pairwise_results
    }  
  

# This is for the 12 samples (12AM-12PM)
  }else if(time_point == 12){
    for(var in treatment){
      
      ## Same as before but for the 12AM-12PM
      df <- data %>% 
        filter(variable == type) %>% 
        filter(treatment == var) %>% 
  filter(time_string == "12:00:00" | time_string == "00:00:00") %>% 
  mutate(time_US = ifelse(time_string == "12:00:00", "T12", "T00")) %>% 
  select(tank_number, time_US, areal_flux) %>% 
  pivot_wider(names_from = time_US, values_from = areal_flux) 
      
      paired <- t.test(df$T12, df$T00, paired = TRUE, na.action = na.pass)
      
      df2 <- tibble(
        time_point = "12AM-12PM",
        variable = type,
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      pairwise_results <- rbind(pairwise_results, df2)
      pairwise_results <<- pairwise_results
    }
}

# Display the results of the code 
return(pairwise_results)
}

```

#### Results just fluxes

```{r}
# Loop Function:

mobeast_loop <- mobeast_24h %>% select(variable) %>% distinct() %>% pull()

pairwise_summary <- map_dfr(mobeast_loop, ~ {
  
  nunu_pairwise(
    data = mobeast_24h,
    time_point = 12,
    type = .x
  ) %>%
    dplyr::mutate(variable = .x)
  
})

```

```{r}
pairwise_summary %>% filter(p_value <0.05)
```

```{r}
nunu_pairwise(DOC_24h, type = "npoc_u_m",time_point = 6)

```



```{r}
variable_meta <- mobeast_24h %>% 
  select(variable, variable_cat) %>%  distinct()

sig_var <- pairwise_summary %>%  filter(p_value <0.05) %>% select(variable) %>% distinct() %>% pull()

pairwise_summary %>% 
  right_join(variable_meta) %>% 
  mutate(sig = ifelse(p_value<0.05 & mean > 0, "Enrichment", 
                                         ifelse(p_value<0.05 & mean < 0, "Depletion", "No Difference")))%>% 
  mutate(sig = factor(sig, 
                      levels = c("Enrichment","No Difference", "Depletion"))) %>% 
  arrange(variable_cat, variable) %>%
  mutate(variable = factor(variable, levels = unique(variable))) %>%
  filter(variable %in% sig_var) %>% filter(variable != "NO2", 
                                           variable != "tn_u_m") %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  ggplot()+
  geom_tile(aes(x= treatment, y= variable, fill = sig), color = "black")+
  #facet_wrap(.~variable_cat)+
  facet_grid(variable_cat ~ ., scales = "free_y", space = "free_y")+
  labs(title = "Pairwise comparison 12AM-12PM",
       x= "Treatment",
       y = " Variables")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw()+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")
```


```{r}
sig_var <- pairwise_summary %>%  filter(p_value <0.05) %>% select(variable) %>% distinct() %>% pull()

pairwise_summary %>% mutate(sig = ifelse(p_value<0.05 & mean > 0, "Enrichment", 
                                         ifelse(p_value<0.05 & mean < 0, "Depletion", "No Difference")))%>% 
  mutate(sig = factor(sig, levels = c("Enrichment","No Difference", "Depletion"))) %>% 
  filter(variable %in% sig_var) %>% 
  filter(variable != "pro", 
         variable != "pico_euk",
         variable != "NO2", 
         variable != "tn_u_m") %>% 
  mutate(variable = factor(variable, levels = c(
    "coble_a","coble_c","coble_m", "m_to_c","bix","hix","fi",
    "coble_b", "coble_t","fpeak","lignin")),
    variable = fct_rev(variable)) %>%
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  ggplot()+
  geom_tile(aes(x= treatment, y= variable, fill = sig), color = "black")+
  #facet_wrap(.~variable_cat)+
  labs(title = "Pairwise comparison 12PM vs. 12AM",
       x= "Treatment",
       y = " Variables")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw()+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")
```

```{r}
sig_fdom <-  pairwise_summary %>%  filter(p_value <0.05) %>% 
  filter(variable %in% sig_var) %>% 
  filter(variable != "pro", 
         variable != "pico_euk",
         variable != "NO2", 
         variable != "tn_u_m") %>% select(variable) %>% distinct() %>% pull()
 
```

```{r}
mobeast_24h %>% 
  filter(variable %in% sig_fdom) %>% 
  filter(time_string == "00:00:00"|time_string == "12:00:00") %>% 
  mutate(variable = factor(variable, levels = c(
    "coble_a","coble_c","coble_m", "m_to_c","bix","hix","fi",
    "coble_b", "coble_t","fpeak","lignin"))) %>% 
  ggplot(aes(x= treatment, y= areal_flux, fill = treatment))+
    geom_violin()+
    #geom_jitter(aes(color = outlier), alpha = 0.5)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_grid(variable ~ time_fac, scales = "free")+
    labs(title = paste("24h fDOM Changes"),
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Areal Flux (raman unit/(s.m2)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18), 
          legend.position = "none")
  
```



#### Results outlier fluxes


```{r}
# Loop Function:

mobeast_loop <- mobeast_24h_flux %>% select(variable) %>% distinct() %>% pull()

pairwise_summary <- map_dfr(mobeast_loop, ~ {
  
  nunu_pairwise(
    data = mobeast_24h,
    time_point = 12,
    type = .x
  ) %>%
    dplyr::mutate(variable = .x)
  
})

```


```{r}
pairwise_summary %>% filter(p_value <0.05)
```

```{r}
nunu_pairwise(DOC_24h_flux, type = "npoc_u_m",time_point = 6)

DOC_24h_flux %>% 
  ggplot(aes(x= treatment, y= areal_flux, fill = treatment))+
    geom_violin()+
    #geom_jitter(aes(color = outlier), alpha = 0.5)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_wrap( .~ time_fac, ncol = 4 )+
    labs(title = paste("24h DOC Changes"),
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Areal Flux (indiv/(s.m2)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18),
          legend.position = "none")

```



```{r}
## Note: There are no differences in the pattern of the heatmap with/without the calculated outliers 

variable_meta <- mobeast_24h_flux %>% 
  select(variable, variable_cat) %>%  distinct()

sig_var <- pairwise_summary %>%  filter(p_value <0.05) %>% select(variable) %>% distinct() %>% pull()

pairwise_summary %>% 
  right_join(variable_meta) %>% 
  mutate(sig = ifelse(p_value<0.05 & mean > 0, "Enrichment", 
                                         ifelse(p_value<0.05 & mean < 0, "Depletion", "No Difference")))%>% 
  mutate(sig = factor(sig, 
                      levels = c("Enrichment","No Difference", "Depletion"))) %>% 
  arrange(variable_cat, variable) %>%
  mutate(variable = factor(variable, levels = unique(variable))) %>%
  filter(variable %in% sig_var) %>% filter(variable != "NO2", 
                                           variable != "tn_u_m") %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  ggplot()+
  geom_tile(aes(x= treatment, y= variable, fill = sig), color = "black")+
  #facet_wrap(.~variable_cat)+
  facet_grid(variable_cat ~ ., scales = "free_y", space = "free_y")+
  labs(title = "Pairwise comparison 12AM-12PM",
       x= "Treatment",
       y = " Variables")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw()+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")



```


#### DNA Pairwise

```{r}

## modified function for DNA 
nunu_pairwise <- function(data, type = "coble_a", time_point = 6, value= areal_flux){

  ## Get a list of the treatments for the for loop   
  treatment <- data %>% mutate(treatment = as.character(treatment)) %>%
  select(treatment) %>%  distinct() %>% pull()
  
  ## Create un empty data frame
pairwise_results <- tibble(
  time_point = vector(),
  treatment= vector(),
  mean = vector(),
  p_value = vector())

# this is the pairwise comparison for the 6 pairs (6AM-6PM)
  if(time_point == 6){
    for(var in treatment){
      
      ## only keep the 6AM and 6PM
      df <- data %>% 
        filter(variable == type) %>% 
        filter(treatment == var) %>% 
  filter(time_string == "18:00:00" | time_string == "06:00:00") %>% 
        ## Rename the times so they can become column names
  mutate(time_US = ifelse(time_string == "18:00:00", "T18", "T06")) %>% 
        ## Only keep the column of interest for the stats
  select(tank_number, time_US, value) %>% 
  pivot_wider(names_from = time_US, values_from = value) 
      
      ## Pairwise function 
      paired <- t.test(df$T18, df$T06, paired = TRUE, na.action = na.pass)
      
      ## Save the result in at data frame
      df2 <- tibble(
        time_point = "06AM-06PM",
        variable = type,
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      # Code will loop oveer all treatments and save the results in a data frame
      pairwise_results <-rbind(pairwise_results,df2)
      
      # Save the results in RStudio
      pairwise_results <<- pairwise_results
    }  
  

# This is for the 12 samples (12AM-12PM)
  }else if(time_point == 12){
    for(var in treatment){
      
      ## Same as before but for the 12AM-12PM
      df <- data %>% 
        filter(variable == type) %>% 
        filter(treatment == var) %>% 
  filter(time_string == "12:00:00" | time_string == "00:00:00") %>% 
  mutate(time_US = ifelse(time_string == "12:00:00", "T12", "T00")) %>% 
  select(tank_number, time_US, value) %>% 
  pivot_wider(names_from = time_US, values_from = value) 
      
      paired <- t.test(df$T12, df$T00, paired = TRUE, na.action = na.pass)
      
      df2 <- tibble(
        time_point = "12AM-12PM",
        variable = type,
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      pairwise_results <- rbind(pairwise_results, df2)
      pairwise_results <<- pairwise_results
    }
}

# Display the results of the code 
return(pairwise_results)
}

```


```{r}
# Note: relative aboundance was logged to be able to compute the pairwise comparison 
# Loop Function:

mobeast_loop <- mobeast_DNA_24h %>% select(variable) %>% distinct() %>% pull()

mobeast_DNA_24h <- mobeast_DNA_24h %>% mutate(abun_log = log10(value +1))

pairwise_summary <- map_dfr(mobeast_loop, ~ {
  
  nunu_pairwise(
    data = mobeast_DNA_24h,
    time_point = 12,
    value = abun_log,
    type = .x
  ) %>%
    dplyr::mutate(variable = .x)
  
})

```

```{r}
sig_var <- pairwise_summary %>% filter(p_value <0.05) %>% 
  select(variable) %>% distinct() %>% pull()
```

```{r}
pairwise_summary %>% 
  mutate(sig = ifelse(p_value<0.05 & mean > 0, "Enrichment", 
                                         ifelse(p_value<0.05 & mean < 0, "Depletion", "No Difference")))%>% 
  mutate(sig = factor(sig, 
                      levels = c("Enrichment","No Difference", "Depletion"))) %>% 
  #mutate(variable = factor(variable, levels = unique(variable))) %>%
  filter(variable %in% sig_var) %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  ggplot()+
  geom_tile(aes(x= treatment, y= variable, fill = sig), color = "black")+
  #facet_grid(variable_cat ~ ., scales = "free_y", space = "free_y")+
  labs(title = "Pairwise comparison 12AM-12PM",
       x= "Treatment",
       y = " Variables")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw()#+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")

```

## PCA 

```{r}
## used the areal flux to calculate the PCA here
## normalization with scale(), but don't actually know what it noarmalizes to...
pca_12PM_df <- mobeast_12PM_flux %>% select(id_number, tank_number, treatment, inflow_table,
                             date_string, time_string, variable, areal_flux) %>%
  filter(variable != "cyano", variable != "tot_bact") %>% 
  pivot_wider(names_from = variable, values_from = areal_flux) %>% 
   na.omit()

pca_matrix_12PM_norm <- pca_12PM_df %>% select(pico_euk:hix) %>% scale() 

pca_matrix_12PM_norm %>% cor() %>% 
  ggcorrplot(
           hc.order = TRUE,
           #type = "lower",
           lab = TRUE)
```


```{r}

as.data.frame(pca_matrix_12PM_norm)
pca_matrix_12PM_norm<- pca_matrix_12PM_norm %>% as.data.frame() %>% 
  select(pico_euk, syn, pro, chla_conc,
                                coble_t, bix, fpeak, fi, PO4, NO3, 
                                het_bact, npoc_u_m, coble_a, 
                                coble_c, coble_m, hix) %>% as.matrix()
```


```{r}
data.pca <- princomp(pca_matrix_12PM_norm, cor = TRUE)
summary(data.pca)

data.pca$loadings[,1:4]

loading_results <- as.data.frame(data.pca$loadings[,1:4])

fviz_eig(data.pca, addlabels = TRUE)

fviz_pca_var(data.pca, col.var = "black")

fviz_pca_var(data.pca, col.var = "cos2",
            gradient.cols = c("blue", "lightgrey", "red"),
            repel = TRUE)
```

```{r}
pca_scores <- as.data.frame(data.pca$scores)

ggplot(pca_scores, aes(Comp.1,Comp.2))+
  geom_point(aes(color = pca_12PM_df$treatment, size= 3))+
  geom_segment(data = loading_results, 
               aes(x=0, y=0, xend = Comp.1, yend= Comp.2),
               arrow = arrow(length = unit(0.5, "cm")),
               alpha = 0.75, color = "black")+
  geom_text(data = loading_results, 
            aes(x = Comp.1, y = Comp.2, 
                label = rownames(loading_results)),
            vjust=1, hjust = 1)+
  theme_bw() +
      scale_color_manual(
        breaks = c("Control", "Coral", "CCA", "Algae"),
        values = c("grey", "#03A9F4", "#EF5350", "#4CAF50")
      )  
```




