---
title: "MOBEAST fDOM V2"
format: 
  html:
    toc: true
    toc-location: left
editor: visual
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)
library(ggridges)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
#library(GGally)
library(emmeans)
library(performance)


```

## Data visualization theme

```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=32, face = "bold"), 
        plot.subtitle = element_text(size=30),
        axis.title = element_text(size = 28, face = "bold"),
        axis.text = element_text(size = 28),
        strip.text.x = element_text(size = 28, face = "bold"),
        legend.title=element_text(size=28, face = "bold"),
        legend.text=element_text(size=28))
```

## Load Data

```{r}
fDOM <- read_csv(here("data", "fDOM", "fDOM_sorted.csv"))

fDOM <- fDOM %>% mutate(treatment = ifelse(treatment == "Rubble", "CCA", treatment)) %>%
  select(-sample_type)

table_id <- read_csv(here("data", "Laurel", "TableID.csv"))

```

## Data Clean-Up

```{r}
fDOM_desc <- tibble(fdom = c("coble_a","coble_b","coble_c","coble_m","coble_t",
  "m_to_c","bix","hix","fpeak","fi",
  "lignin"),
  fdom_alt_name = c("UV-Humic","Tyrosine","Visible-Humic","Marine Humic","Tryptophan",
  "m_to_c","bix","hix","Phenylalanine","Fluvic Acid",
  "lignin"),
  fdom_cat= c("Humic","Proteinaceous","Humic","Humic","Proteinaceous",
  "Humic","Humic","Humic","Proteinaceous","Humic",
  "Other"))

fDOM_desc %>% gt()
```

```{r}
table_id <- clean_names(table_id)

table_id <- table_id %>% rename(tank_number=tankid) %>% select(-treatment) %>% 
  mutate(tank_number = paste0("T", tank_number)) %>% 
  mutate(inflow_table = paste("Table", inflow_table))
fDOM <- left_join(fDOM, table_id, by= "tank_number")
```

```{r}
fDOM <- fDOM %>% mutate(inflow_table = ifelse(is.na(inflow_table), tank_number, inflow_table)) %>% 
   mutate(inflow_table = ifelse(inflow_table == "F1", "Table 1", 
                                ifelse(inflow_table == "F2", "Table 2", inflow_table))) %>% 
  mutate(inflow_table = factor(inflow_table),
         tank_number = factor(tank_number)) %>% 
  relocate(inflow_table, .after = "treatment")
```

```{r}
fDOM_long <- fDOM %>% pivot_longer(cols = coble_a:lignin,
                      names_to = "fdom", 
                      values_to = "fdom_conc")
```

```{r}
fdom_long <- left_join(fDOM_long, fDOM_desc, by = "fdom")

```

```{r}
fDOM_norm <- fdom_long %>% filter(date > as.Date("2024-05-31")) %>% # Removes end members
  filter(treatment != "Rain") %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
   group_by(date_time, inflow_table, fdom) %>%
  mutate(inflow= fdom_conc[treatment == "Inflow"]) %>%  # Isolate inflow for each date_time and table
  ungroup() %>% 
  #group_by(date_time) %>% 
  #mutate(inflow = ifelse(is.na(inflow),
   #   inflow[!is.na(inflow)][1],  # get the other table's value for that date
   #   inflow)) %>% 
  #ungroup() %>% 
  mutate(fdom_conc_norm = fdom_conc - inflow)  %>% # Normalize for treatments 
  relocate(inflow, .after = "fdom_conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
  

head(fDOM_norm, 15)

#write_csv(fDOM_norm, here("data", "fDOM", "fDOM_norm.csv"))
```

## Quick visulitation

```{r}
ggplot(fDOM_norm, aes(x= treatment, y = fdom_conc, fill = fdom))+
  geom_bar(position="stack", stat="identity")+
  facet_grid(fdom_cat~date, scales = "free")+
  nunu_theme
```

```{r}
fDOM_norm %>% filter(fdom_cat != "Other") %>% 
  ggplot(aes(x= treatment, y = fdom_conc, fill = fdom))+
  geom_boxplot()+
  facet_grid(fdom_cat~date, scales = "free")+
  nunu_theme
```

```{r}
fDOM_norm %>% mutate(fdom_cat = factor(fdom_cat, levels = c("Humic", "Proteinaceous", "Other"))) %>% 
  group_by(fdom_cat,fdom) %>% 
  summarize(min = round(min(fdom_conc, na.rm = T),3), 
            avg = round(mean(fdom_conc, na.rm = T),3),
            max = round(max(fdom_conc, na.rm = T),3)) %>% 
  ungroup() %>% 
  gt()
```

## Hypothesis 1: fDOM 12PM

```{r}
fDOM_12PM <- fDOM_norm %>% filter(treatment != "Inflow") %>% 
  filter(time == hms("12:00:00")) %>% 
  filter(fdom_cat != "Other")
```

### Humic

#### CobleA

```{r}
fDOM_12PM_cobA <- fDOM_12PM %>% filter(fdom == "coble_a") 
```

```{r}
## Outlier detection

fDOM_12PM_cobA <- fDOM_12PM_cobA %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(fdom_conc_norm, probs = 0.25),
         quant3 = quantile(fdom_conc_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( fdom_conc_norm > quant3 + (IQR*1.5) | fdom_conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

```{r}
mod <- lmer(fdom_conc_norm ~ treatment + (1|date), data = fDOM_12PM_cobA)
summary(mod)
```

```{r}
anova(mod, type= "2")
```

`{, fig_height = 10, fig.width = 10} plot_model(mod, type = "re")[[1]] +   plot_model(mod, type = "re")[[2]] +   plot_model(mod, type = "eff", terms = c('treatment')) +   plot_model(mod, type = "diag")[[1]]`

```{r}
resid_panel(mod)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, specs = ~ treatment)
```

##### Interpretation

```{r}
#sort(boxplot.stats(fDOM_12PM_cobA$fdom_conc)$out)
ggplot(fDOM_12PM_cobA, aes(x= treatment, y = fdom_conc_norm, fill = fdom))+
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  facet_grid(fdom_cat~date, scales = "free")+
  nunu_theme
```

#### CobleM

```{r}
fDOM_12PM_cobM <- fDOM_12PM %>% filter(fdom == "coble_m") 
```

```{r}
## Outlier detection

fDOM_12PM_cobM <- fDOM_12PM_cobM %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(fdom_conc_norm, probs = 0.25),
         quant3 = quantile(fdom_conc_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( fdom_conc_norm > quant3 + (IQR*1.5) | fdom_conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

###### Normality and Homogeneity

```{r}
bartlett.test(fdom_conc_norm ~ treatment,data = fDOM_12PM_cobM) # homogeneity
shapiro.test(fDOM_12PM_cobM$fdom_conc_norm) # normality
```

```{r}
mod <- lmer(fdom_conc_norm ~ treatment + (1|tank_number), data = fDOM_12PM_cobM)
summary(mod)
```

```{r}
resid_panel(mod)
```

```{r}
anova(mod)
kruskal.test(fdom_conc_norm ~ treatment, data = fDOM_12PM_cobM )
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, specs = ~ treatment)
```

```{r}
emm <- emmeans(mod, ~ treatment)
cld_letters <- multcomp::cld(emm, Letters = letters)
```

```{r, fig.width = 18, fig.height=8}
ggplot(fDOM_12PM_cobM, aes(x= treatment, fdom_conc, fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2,linewidth = 0.8)+
  geom_violin(linewidth = 1.5)+ 
  geom_point(color = "black", alpha = 0.5, size = 4)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  stat_summary(color = "yellow", alpha =0.5, size = 1.5)+
  labs(#title = "DOC Enrichment",
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Marine Humics Production (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  #ylim(-20,120)+
  nunu_theme +
  theme(legend.position = "none") 

#ggsave("CobM_12PM.png", width = 18,height = 8,path = here("outputs", "OOSC_2025"))
```

#### CobleC

```{r}
fDOM_12PM_cobC <- fDOM_12PM %>% filter(fdom == "coble_c") 
```

```{r}
## Outlier detection

fDOM_12PM_cobC <- fDOM_12PM_cobC %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(fdom_conc_norm, probs = 0.25),
         quant3 = quantile(fdom_conc_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( fdom_conc_norm > quant3 + (IQR*1.5) | fdom_conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

##### No data transformation

###### Normality and Homogeneity

```{r}
bartlett.test(fdom_conc_norm ~ treatment,data = fDOM_12PM_cobC) # homogeneity
shapiro.test(fDOM_12PM_cobC$fdom_conc_norm) # normality
```

```{r}
mod <- lmer(fdom_conc_norm ~ treatment + (1|tank_number), data = fDOM_12PM_cobC)
summary(mod)
```

```{r}
anova(mod)
```

```{r}
resid_panel(mod)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```

```{r}
ggplot(fDOM_12PM_cobC, aes(x= treatment, y = fdom_conc, fill = fdom))+
  geom_violin()+
  geom_point(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  #facet_grid(fdom_cat~date, scales = "free")+
  nunu_theme
```

##### With data transformation

###### Normality and Homogeneity

```{r}
power_model <- powerTransform(fDOM_12PM_cobC$fdom_conc_norm ~ 1, family = "yjPower")  # no predictors, just transforming y
summary(power_model)
## lambda = -186.9304 according to the above 

```

```{r}
bartlett.test(yjPower(fdom_conc_norm, lambda = -186.9304) ~ treatment,data = fDOM_12PM_cobC) # homogeneity
shapiro.test(yjPower(fDOM_12PM_cobC$fdom_conc_norm, lambda = -186.9304)) # normality
```

```{r}
fDOM_12PM_cobC <- fDOM_12PM_cobC %>% 
  mutate(fdom_conc_trans= yjPower(fdom_conc_norm, lambda = -186.9304)) %>% 
  relocate(fdom_conc_trans, .after = fdom_conc_norm)
```

```{r}
mod <- lmer(fdom_conc_trans ~ treatment + (1|tank_number), data = fDOM_12PM_cobC)
summary(mod)
```

```{r}
resid_panel(mod)
```

```{r}
anova(mod)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```

```{r}
emm <- emmeans(mod, ~ treatment)
cld_letters <- multcomp::cld(emm, Letters = letters)
```

##### Data Visualization

```{r, fig.width= 18, fig.height = 8}
ggplot(fDOM_12PM_cobC, aes(x= treatment, fdom_conc_trans, fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2,linewidth = 1)+
  geom_violin(linewidth = 1.5)+
  geom_point(color = "black", alpha = 0.5, size = 4)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  stat_summary(color = "yellow", alpha =0.5, size = 1.5)+
  labs(#title = "DOC Enrichment",
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Colored Humic Production (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  ylim(min(fDOM_12PM_cobC$fdom_conc_trans),max(fDOM_12PM_cobC$fdom_conc_trans)+0.001)+
  nunu_theme +
  theme(legend.position = "none") 

#ggsave("CobC_12PM.png", width = 18,height = 8,path = here("outputs", "OOSC_2025"))
```

#### M:C SIGNIFICANT

```{r}
fDOM_12PM_mtoc <- fDOM_12PM %>% filter(fdom == "m_to_c") 
```

```{r}
## Outlier detection

fDOM_12PM_mtoc <- fDOM_12PM_mtoc %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(fdom_conc_norm, probs = 0.25),
         quant3 = quantile(fdom_conc_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( fdom_conc_norm > quant3 + (IQR*1.5) | fdom_conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

##### No data transformation

###### Normality and Homogeneity

```{r}
bartlett.test(fdom_conc_norm ~ treatment,data = fDOM_12PM_mtoc) # homogeneity
shapiro.test(fDOM_12PM_mtoc$fdom_conc_norm) # normality
```

```{r}
mod <- lmer(fdom_conc_norm ~ treatment + (1|tank_number), data = fDOM_12PM_mtoc)
summary(mod)
```

```         
## Transforming the data with yjPower did not change the Q-Q plot 

power_model <- powerTransform(fDOM_12PM_mtoc$fdom_conc_norm ~ 1, family = "yjPower")  # no predictors, just transforming y
summary(power_model)

## lambda = 0.1449 according to the above 
mod <- lmer(yjPower(fdom_conc_norm, lambda = 0.1449) ~ treatment + (1|tank_number), data = fDOM_12PM_mtoc)
summary(mod)
```

```{r}
resid_panel(mod)
```

```{r}
anova(mod)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```

```{r}
emm <- emmeans(mod, ~ treatment)
cld_letters <- multcomp::cld(emm, Letters = letters)
```

##### Data Visualization

```{r,fig.width = 18, fig.height = 8}

ggplot(fDOM_12PM_mtoc, aes(x= treatment, fdom_conc, fill = treatment)) +
  #geom_hline(yintercept = 0, lty = 2,linewidth = 0.8)+
  geom_violin(linewidth = 1.5)+
  geom_point(color = "black", alpha = 0.5, size = 4)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  stat_summary(color = "yellow", alpha =0.5, size = 1.5)+
  labs(#title = "DOC Enrichment",
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Marine:Colored (M:C) Ratio ")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  ylim(0.3,0.9)+
  nunu_theme +
  theme(legend.position = "none") 

ggsave("MtoC_12PM.png", width = 18,height = 8,path = here("outputs", "OOSC_2025"))
```

```{r, fig.width=18, fig.height=8}
cobM_boxp <- ggplot(fDOM_12PM_cobM, aes(x= treatment, y = fdom_conc))+
  geom_boxplot(aes(fill = treatment), linewidth = 1.5)+
  geom_hline(yintercept = 0, lty = 2,linewidth = 1)+
  #geom_jitter(aes(color = treatment), alpha = 0.5)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  stat_summary(color = "yellow", alpha =0.5, size = 1.5)+
   ylim(0,0.016)+
   labs(title = "Marine Humics (M)",
        x= "Treatment",
        y= "Concentration ()")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  #facet_grid(~date)+
  nunu_theme+ theme(plot.title = element_text(hjust = 0.5),
                    legend.position = "none")
  
cobC_boxp <- ggplot(fDOM_12PM_cobC, aes(x= treatment, y = fdom_conc_trans))+
  geom_boxplot(aes(fill = treatment), linewidth = 1.5)+
  geom_hline(yintercept = 0, lty = 2,linewidth = 1)+
  stat_summary(color = "yellow", alpha =0.5, size = 1.5)+
  #geom_jitter(aes(color = treatment), alpha = 0.5)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
    ylim(0,0.016)+
    labs(title = "Colored Humics (C)",
         x= "Treatment",
         y= "Concentration ()")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  #facet_grid(~date)+
  nunu_theme + theme(plot.title = element_text(hjust = 0.5),
                     legend.position = "none")
  
mtoc_boxp <- ggplot(fDOM_12PM_mtoc, aes(x= treatment, y = fdom_conc))+
  geom_boxplot( fill = "yellow", linewidth = 1.5)+
  geom_hline(yintercept = 0, lty = 2,linewidth = 1)+
  #geom_jitter(aes(color = treatment), alpha = 0.5)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  #facet_grid(~date)+
    labs(title = "M:C Ratio",
         x= " Treatment",
         y= "Ratio")+
  nunu_theme + theme(plot.title = element_text(hjust = 0.5))

cobM_boxp + cobC_boxp + plot_layout(axis_titles = "collect")

ggsave("CobM_C_12PM.png", width = 18,height = 8,path = here("outputs", "OOSC_2025"))
```

#### FI

```{r}
fDOM_12PM_fi <- fDOM_12PM %>% filter(fdom == "fi") 
```

```{r}
## Outlier detection

fDOM_12PM_fi <- fDOM_12PM_fi %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(fdom_conc_norm, probs = 0.25),
         quant3 = quantile(fdom_conc_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( fdom_conc_norm > quant3 + (IQR*1.5) | fdom_conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

```{r}
mod <- lmer(fdom_conc_norm ~ treatment + (1|tank_number), data = fDOM_12PM_fi)
summary(mod)
```

```{r}
anova(mod, type= "2")
```

`{, fig_height = 10, fig.width = 10} plot_model(mod, type = "re")[[1]] +   plot_model(mod, type = "re")[[2]] +   plot_model(mod, type = "eff", terms = c('treatment')) +   plot_model(mod, type = "diag")[[1]]`

```{r}
resid_panel(mod)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```

```{r}
#sort(boxplot.stats(fDOM_12PM_cobC$fdom_conc)$out)
ggplot(fDOM_12PM_fi, aes(x= treatment, y = fdom_conc, fill = fdom))+
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  facet_grid(fdom_cat~date, scales = "free")+
  nunu_theme
```

#### BIX

```{r}
fDOM_12PM_bix <- fDOM_12PM %>% filter(fdom == "bix") 
```

```{r}
## Outlier detection

fDOM_12PM_bix <- fDOM_12PM_bix %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(fdom_conc_norm, probs = 0.25),
         quant3 = quantile(fdom_conc_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( fdom_conc_norm > quant3 + (IQR*1.5) | fdom_conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

```{r}
mod <- lmer(fdom_conc_norm ~ treatment + (1|date)+ (1|inflow_table), data = fDOM_12PM_bix)
summary(mod)
```

```{r}
anova(mod, type= "2")
```

`{, fig_height = 10, fig.width = 10} plot_model(mod, type = "re")[[1]] +   plot_model(mod, type = "re")[[2]] +   plot_model(mod, type = "eff", terms = c('treatment')) +   plot_model(mod, type = "diag")[[1]]`

```{r}
resid_panel(mod)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```

```{r}
#sort(boxplot.stats(fDOM_12PM_cobC$fdom_conc)$out)
ggplot(fDOM_12PM_bix, aes(x= treatment, y = fdom_conc, fill = fdom))+
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  facet_grid(fdom_cat~date, scales = "free")+
  nunu_theme
```

#### HIX

```{r}
fDOM_12PM_hix <- fDOM_12PM %>% filter(fdom == "hix") 
```

```{r}
## Outlier detection

fDOM_12PM_hix <- fDOM_12PM_hix %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(fdom_conc_norm, probs = 0.25),
         quant3 = quantile(fdom_conc_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( fdom_conc_norm > quant3 + (IQR*1.5) | fdom_conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

```{r}
mod <- lmer(fdom_conc_norm ~ treatment + (1|date)+ (1|inflow_table), data = fDOM_12PM_hix)
summary(mod)
```

```{r}
anova(mod, type= "2")
```

```{r}
resid_panel(mod)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```

```{r}
#sort(boxplot.stats(fDOM_12PM_cobC$fdom_conc)$out)
ggplot(fDOM_12PM_hix, aes(x= treatment, y = fdom_conc, fill = fdom))+
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  facet_grid(fdom_cat~date, scales = "free")+
  nunu_theme
```

### Proteinaceous

```{r}
fDOM_12PM %>% filter(fdom_cat == "Proteinaceous") %>% 
  select(fdom) %>% distinct()
```

#### CobleB

```{r}
fDOM_12PM_cobB <- fDOM_12PM %>% filter(fdom == "coble_b") 
```

```{r}
## Outlier detection

fDOM_12PM_cobB <- fDOM_12PM_cobB %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(fdom_conc_norm, probs = 0.25),
         quant3 = quantile(fdom_conc_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( fdom_conc_norm > quant3 + (IQR*1.5) | fdom_conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

```{r}
mod <- lmer(fdom_conc_norm ~ treatment + (1|date)+ (1|inflow_table), data = fDOM_12PM_cobB)
summary(mod)
```

```{r}
anova(mod, type= "2")
```

```{r}
resid_panel(mod)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, specs =  pairwise~ treatment)
```

```{r}
#sort(boxplot.stats(fDOM_12PM_cobC$fdom_conc)$out)
ggplot(fDOM_12PM_cobB, aes(x= treatment, y = fdom_conc, fill = fdom))+
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  facet_grid(fdom_cat~date, scales = "free")+
  nunu_theme
```

#### CobleT

```{r}
fDOM_12PM_cobT <- fDOM_12PM %>% filter(fdom == "coble_t") 
```

```{r}
## Outlier detection

fDOM_12PM_cobT <- fDOM_12PM_cobT %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(fdom_conc_norm, probs = 0.25),
         quant3 = quantile(fdom_conc_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( fdom_conc_norm > quant3 + (IQR*1.5) | fdom_conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

```{r}
mod <- lmer(fdom_conc_norm ~ treatment + (1|date), data = fDOM_12PM_cobT)
summary(mod)
```

```{r}
anova(mod, type= "2")
```

```{r}
resid_panel(mod)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, specs =  pairwise~ treatment)
```

```{r}
#sort(boxplot.stats(fDOM_12PM_cobC$fdom_conc)$out)
ggplot(fDOM_12PM_cobT, aes(x= treatment, y = fdom_conc, fill = fdom))+
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  facet_grid(fdom_cat~date, scales = "free")+
  nunu_theme
```

#### Fpeak

```{r}
fDOM_12PM_fpeak <- fDOM_12PM %>% filter(fdom == "fpeak") 
```

```{r}
## Outlier detection

fDOM_12PM_fpeak <- fDOM_12PM_fpeak %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(fdom_conc_norm, probs = 0.25),
         quant3 = quantile(fdom_conc_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( fdom_conc_norm > quant3 + (IQR*1.5) | fdom_conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

```{r}
mod <- lmer(fdom_conc ~ treatment + (1|date)+ (1|inflow_table), data = fDOM_12PM_fpeak)
summary(mod)
```

```{r}
anova(mod, type= "2")
```

```{r}
resid_panel(mod)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, specs =  pairwise~ treatment)
```

```{r}
#sort(boxplot.stats(fDOM_12PM_cobC$fdom_conc)$out)
ggplot(fDOM_12PM_fpeak, aes(x= treatment, y = fdom_conc, fill = fdom))+
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  facet_grid(fdom_cat~date, scales = "free")+
  nunu_theme
```

## 24h

```{r}
fDOM_24h <- fDOM_norm %>% filter(treatment != "Inflow") %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>% 
  mutate(time_fac = factor(as.character(time),levels = c("12:00:00","18:00:00","00:00:00",
                                            "06:00:00"))) %>% 
  mutate(date_fac = factor(date)) %>% 
  filter(fdom_cat != "Other")
```

### Humics

#### M:C

```{r}
fDOM_24h_mtoc <- fDOM_24h %>% filter(fdom == "m_to_c") 
```

```{r}
## Outlier detection

fDOM_24h_mtoc <- fDOM_24h_mtoc %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(fdom_conc_norm, probs = 0.25),
         quant3 = quantile(fdom_conc_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( fdom_conc_norm > quant3 + (IQR*1.5) | fdom_conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier")

```

##### No data transformation

###### Normality and Homogeneity

```{r}
bartlett.test(fdom_conc_norm ~ treatment,data = fDOM_24h_mtoc) # homogeneity
shapiro.test(fDOM_24h_mtoc$fdom_conc_norm) # normality
```

```{r}
mod <- lmer(fdom_conc_norm ~ treatment*time_fac + (1|tank_number), data = fDOM_24h_mtoc)
summary(mod)
```

```{r}
resid_panel(mod)
```

```{r}
anova(mod)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```

```{r}
emm <- emmeans(mod, ~ treatment)
cld_letters <- multcomp::cld(emm, Letters = letters)
```

##### Data Visualization

```{r,fig.width = 18, fig.height = 8}

fDOM_24h_mtoc %>% filter(time_fac != as.character("00:00:00")) %>% 
  ggplot(aes(x= treatment, fdom_conc, fill = treatment)) +
  #geom_hline(yintercept = 0, lty = 2,linewidth = 0.8)+
  geom_violin(linewidth = 1)+
  geom_point(color = "black", alpha = 0.5, size = 4)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  stat_summary(color = "yellow", alpha =0.5, size = 1.5)+
  labs(#title = "DOC Enrichment",
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Marine:Colored (M:C) Ratio ")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  facet_wrap(~time_fac)+
  ylim(0.3,0.9)+
  nunu_theme +
  theme(legend.position = "none",
        strip.background = element_rect(fill = "white")) 

ggsave("MtoC_24h.png", width = 18,height = 8,path = here("outputs", "OOSC_2025"))
```
