---
title: "MOBEAST Ordination"
format: html
editor: visual
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

# Load Libraries
```{r}
library(tidyverse)
#library(lubridate)
library(here)
library(janitor)
#library(ggrepel)
#library(ghibli)
#library(gt)
library(patchwork)
library(RColorBrewer)
library(vegan)
library(phyloseq)
library(GUniFrac)
```

# Data visualization

```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=32, face = "bold"), 
        plot.subtitle = element_text(size=30),
        axis.title = element_text(size = 28, face = "bold"),
        axis.text = element_text(size = 28),
        strip.text.x = element_text(size = 28, face = "bold"),
        legend.title=element_text(size=28, face = "bold"),
        legend.text=element_text(size=28))
```

# Load Data
```{r}
otu <- read_csv(here("data", "DNA", "MOBEAST_OTU_culled.csv"))
meta <- read_csv(here("data","MOBEAST_metadata_V2.csv"))
tax <- read_csv(here("data","DNA", "MOBEAST_tax_culled.csv"))
OTU_ord <- read_csv(here("data","OTU_heatmap_ready.csv"))
```

# Making a phyloseq object 
## Transform the dataframes into matrices 
```{r}
## Make an OTU matrix
otu<- otu %>% mutate(ASV = toupper(ASV))
otu_list<- otu$ASV
otu<- otu %>% select(-(117:122), -MOBEAST017b)
otu_mat <- as.matrix(otu)
rownames(otu_mat) <- otu_list

## Make a taxonomy matrix
tax_mat <- tax %>% mutate(OTUNumber = toupper(OTUNumber)) %>% 
   tibble::column_to_rownames("OTUNumber") %>% 
   as.matrix()

```

```{r}
## Clean the metadata 
meta_mat <-meta %>% select(-long_name) %>% 
  mutate(time = str_pad(time, width = 4, side = "left", pad = "0"),
               time = str_replace(time, "(\\d{2})(\\d{2})", "\\1:\\2")) %>% 
  # "(\\d{2})(\\d{2})" separates the string in 2 groups of 2 characters
  # "\\1:\\2", group1 + ":" + group2
  mutate(time = paste0(time, ":00"),
               date_time = paste(date, time),
               date = mdy(date), 
               date_time = mdy_hms(date_time),
               time = factor(time)) %>% 
    tibble::column_to_rownames("id_number") 
```

```{r, meta data clean up}
meta<- meta %>% mutate(time = str_pad(time, width = 4, side = "left", pad = "0"),
               time = str_replace(time, "(\\d{2})(\\d{2})", "\\1:\\2")) %>% 
  # "(\\d{2})(\\d{2})" separates the string in 2 groups of 2 characters
  # "\\1:\\2", group1 + ":" + group2
  mutate(time = paste0(time, ":00"),
               date_time = paste(date, time),
               date = mdy(date), 
               date_time = mdy_hms(date_time),
               time = factor(time))
  
  
```
## Preparing the metadata to annotate the heatmap

```{r}
# renaming some of treatment so that environmental data have a label: endmember or rain
# Same for inflow table, environmental samples now have a label
## Convert tibble to df for later
meta <- meta %>% mutate(treatment = as.character(treatment)) %>% 
  mutate(treatment = ifelse(id_number == "MOBEAST115", "Rain", 
                            ifelse(is.na(treatment), "Endmember", treatment))) %>% 
  mutate(inflow_table = ifelse(is.na(inflow_table), "Environmental data", inflow_table)) %>% 
  mutate(date_string = as.character(date),
                time_string = as.character(time),
                date_time_string = paste(date_string, time_string)) %>% 
  as.data.frame()
```

## Transform the data to phyloseq object
```{r}
OTU <- otu_table(otu_mat, taxa_are_rows = TRUE)
TAX <- tax_table(tax_mat)
samples <- sample_data(meta_mat)
```

```{r}
phy <- phyloseq(OTU, TAX, samples)
```

## Unsure what this is for but relabeling my OTU IDs for ordinations? 
```{r}
tax_concatonate <- data.frame(tax_mat)
tax_concatonate<- tax_concatonate %>% mutate(OTU = rownames(tax_concatonate)) 
rownames(tax_concatonate) <- NULL

tax_concatonate<- tax_concatonate %>% 
  mutate(tax_id = paste0(Phylum, "_", Class, "_", Order, "_", Family, "_", Genus,  "_",OTU)) %>% 
  relocate(tax_id, .before = Kingdom)

tax_id_df <- tax_concatonate %>% select(OTU,tax_id)
```

```{r}
otu_w_tax <- otu_mat
rownames(otu_w_tax) <- tax_id_df$tax_id

```

# Ordinations with long format df

## Using all the samples 
```{r}
# make a data frame from the OTU matrix wth OTU numbers in the columns and sample names in rows 
otu_ord <- OTU %>% t() %>% as.data.frame() 

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

```


```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat <- otu_ord %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances 
```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist <- vegdist(otu_ord_mat, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds <- left_join(MOBEAST_nmds, meta)
```

### Plot NDMS all samples
#### By treatments
```{r, fig.width=10, fig.height= 6}
# raw NMDS plot with all the samples
MOBEAST_nmds %>%  
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = treatment))+
  labs(title= "All samples, ellipse by treatments")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("NMDS_all_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

#### By date time 
```{r, fig.width=16, fig.height= 10}
# raw NMDS plot with all the samples
MOBEAST_nmds %>%  group_by(date_time_string) %>% 
  mutate(occurance = n()) %>% ungroup() %>% 
  filter(occurance > 3) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment))+
  stat_ellipse(aes(color = date_time_string, fill = date_time_string),alpha = 0.3, geom= "polygon", level = 0.95)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = treatment))+
  labs(title= "All samples, ellipse by date_time, endmembers removed")+
  scale_color_manual(values=treatment_colors)+
  #scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
ggsave("NMDS_all_date_time.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

#### By "Inflow Table"
```{r, fig.width=16, fig.height= 10}

MOBEAST_nmds %>%   
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment))+
  stat_ellipse(aes(color = inflow_table, fill = inflow_table),alpha = 0.3, geom= "polygon", level = 0.95)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = treatment))+
  labs(title= "All samples, ellipse: environmental data vs tanks")+
  scale_color_manual(values=treatment_colors)+
  #scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
ggsave("NMDS_all_inflow_table.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

```{r, fig.width=16, fig.height= 10}

MOBEAST_nmds %>%   
  mutate(location = ifelse(tank_number == "Bay"| tank_number == "Ocean", "Bay and Ocean", 
                           ifelse(tank_number == "Lagoon" | tank_number == "Lagoon Tiahura", "Lagoon",
                                  "Tanks"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment))+
  ggforce::geom_mark_ellipse(aes(color = location, fill = location), alpha = 0.3)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = treatment))+
  labs(title= "All samples, ellipse: environmental data vs tanks")+
  scale_color_manual(values=treatment_colors)+
  #scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
ggsave("NMDS_all_location.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

### Stats on all data 

```{r}
adonis2(MOBEAST_dist~MOBEAST_nmds$treatment, 
        # formula to calculate p-value for treatments
        strata = MOBEAST_nmds$tank_number) # accounts for repreated measures

```

#### Accounting for uneven variation in the data. 
If the variation is uneven without the groups, adonis may come up with a significant result, although it's not. 
We can control for that using betdisper() and permutest()

```{r, betadipser}
bd_all <- betadisper(MOBEAST_dist, MOBEAST_nmds$treatment)

```

```{r}
anova(bd_all)
permutest(bd_all)
bd_all
```



## 12PM samples
```{r}
# make a data frame from the OTU matrix wth OTU numbers in the columns and sample names in rows 
otu_ord <- OTU %>% 
  t() %>% 
  as.data.frame() %>% 
  select(-OTU0006) # Removing Nitrincolaceae as it's probably there due to a contamination

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select 12PM samples 
id_number_12PM <- meta %>% filter(time == "12:00:00") %>% pull(id_number)

otu_ord_12PM <- otu_ord %>% filter(id_number %in% id_number_12PM)


```


```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_12 <- otu_ord_12PM %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances 
```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_12 <- vegdist(otu_ord_mat_12, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_12, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_12 <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_12 <- left_join(MOBEAST_nmds_12, meta)
```

### Plotting NDMS 12PM

```{r, fig.width=10, fig.height= 6}
# Ellipse by treatments
MOBEAST_nmds_12 %>%  
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = treatment))+
  #facet_wrap(~date_string)+
  labs(title= "12PM samples, ellipse by treatments")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("NMDS_12PM_treatments.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

```{r, fig.width=10, fig.height= 6}
# Ellipse by dates
MOBEAST_nmds_12 %>%  
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment))+
  stat_ellipse(aes(colour = date_string, fill = date_string), 
               alpha = 0.2, geom= "polygon", level = 0.95)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = treatment))+
  labs(title= "12PM samples, ellipse by day")+
  scale_color_manual(values=treatment_colors)+
  #scale_fill_viridis_d() +
  #nunu_theme+
  theme_bw()
#ggsave("NMDS_12PM_day.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

```{r, fig.width=10, fig.height= 6}
# Ellipse by dates
MOBEAST_nmds_12 %>%  
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment))+
  stat_ellipse(aes(fill = treatment),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  stat_ellipse(aes(colour = date_string), 
               alpha = 0.2, geom= "polygon", level = 0.95, linetype = 2)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  labs(title= "12PM samples, ellipse by treatment, dashed ellipse by date")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
ggsave("NMDS_12PM_day_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

## 12PM Only using the treatment samples 

```{r}
# make a data frame from the OTU matrix wth OTU numbers in the columns and sample names in rows 
otu_ord <- OTU %>% 
  t() %>% 
  as.data.frame() %>% 
  select(-OTU0006) # Removing Nitrincolaceae as it's probably there due to a contamination

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select 12PM samples 
id_number_12PM <- meta %>% filter(time == "12:00:00") %>%
  filter(treatment != "Endmember"| treatment != "Rain") %>% 
  filter(treatment != "Inflow") %>% 
  pull(id_number)

otu_ord_12PM <- otu_ord %>% filter(id_number %in% id_number_12PM)


```


```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_12 <- otu_ord_12PM %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances 
```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_12 <- vegdist(otu_ord_mat_12, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_12, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_12 <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_12 <- left_join(MOBEAST_nmds_12, meta)
```


### Plotting NDMS 12PM with treatment only 
```{r}
centroid_12 <- MOBEAST_nmds_12 %>% 
  group_by(treatment) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))
```

```{r, fig.width=16, fig.height= 10}
# Ellipse by treatments
MOBEAST_nmds_12 %>%  
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  #stat_ellipse(alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  ggforce::geom_mark_ellipse(aes(color = treatment, fill = treatment), alpha = 0.3,
                             show.legend = FALSE)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  geom_point(data = centroid_12, aes(NMDS1, NMDS2, fill = treatment), 
             color = "yellow", shape= 23, size = 8, show.legend = FALSE)+
  #geom_label_repel(aes(label = treatment))+
  #facet_wrap(~date_string)+
  labs(title= "12PM treatment samples only, ellipse by treatments, OTU006 removed")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("NMDS_12PM_treatments.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

### Stats on treatment only 

```{r}
adonis2(MOBEAST_dist_12~MOBEAST_nmds_12$treatment, 
        # formula to calculate p-value for treatments
        strata = MOBEAST_nmds_12$tank_number) # accounts for repeated measures

```

#### Accounting for uneven variation in the data. 
If the variation is uneven without the groups, adonis may come up with a significant result, although it's not. 
We can control for that using betdisper() and permutest()

```{r, betadipser}
bd_12 <- betadisper(MOBEAST_dist_12, MOBEAST_nmds_12$treatment)

```

```{r}
anova(bd_12)
permutest(bd_12)
bd_12
```


## June 4 and June 6
```{r}
otu_ord <- OTU %>% 
  t() %>% 
  as.data.frame() %>% 
  select(-OTU0006) # Removing Nitrincolaceae as it's probably there due to a contamination

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select dates of interest
id_number_June4_6 <- meta %>% 
  filter(date_string == "2024-06-04"| date_string == "2024-06-06") %>% 
  pull(id_number)

otu_ord_June4_6 <- otu_ord %>% filter(id_number %in% id_number_June4_6)


```


```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_June4_6 <- otu_ord_June4_6 %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances 
```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_June4_6 <- vegdist(otu_ord_mat_June4_6, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_June4_6, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_June4_6 <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_June4_6 <- left_join(MOBEAST_nmds_June4_6, meta)
```

### Plotting NDMS 12PM

```{r, fig.width=10, fig.height= 6}
# raw NMDS plot with all the samples
MOBEAST_nmds_June4_6 %>%  
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(aes(colour = date_string, fill = date_string), 
               alpha = 0.2, geom= "polygon", level = 0.95)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  labs(title= "June 4 and June 6 samples, ellipse by day")+
  scale_color_manual(values=treatment_colors)+
  #scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("June4_6.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

```{r, fig.width=10, fig.height= 6}
# raw NMDS plot with all the samples
MOBEAST_nmds_June4_6 %>%  
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = treat, fill = date_string), 
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  labs(title= "June 4 and June 6 samples, ellipse by day")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("June4_6_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

```{r, fig.width=10, fig.height= 6}
# raw NMDS plot with all the samples
MOBEAST_nmds_June4_6 %>%  
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = treat, fill = date_string), 
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  stat_ellipse(aes(colour = date_string, fill = date_string), 
               alpha = 0.2, geom= "polygon", level = 0.95, linetype = 2)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  labs(title= "June 4 and June 6 samples, ellipse by treatment, dashed ellipse by day")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
ggsave("June4_6_treatment_day.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

## 24h Samples
```{r}
otu_ord <- OTU %>% 
  t() %>% 
  as.data.frame() 

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select 12PM samples 
id_number_24h <- meta %>% 
  filter(date_string == "2024-06-02"| date_string == "2024-06-03") %>% 
  pull(id_number)

otu_ord_24h <- otu_ord %>% filter(id_number %in% id_number_24h)


```


```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_24h <- otu_ord_24h %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances 
```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_24h <- vegdist(otu_ord_mat_24h, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_24h, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_24h <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_24h <- left_join(MOBEAST_nmds_24h, meta)
```

### Plotting NDMS 24h

```{r, fig.width=16, fig.height= 10}
# Plo
MOBEAST_nmds_24h %>%   
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string)+
  labs(title= "24h samples, ellipse by treatment")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```


```{r, fig.width=10, fig.height= 6}
# raw NMDS plot with all the samples
MOBEAST_nmds_24h %>%  
  mutate(time_string = factor(time_string, levels = c("12:00:00","18:00:00","00:00:00","06:00:00" ))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment))+
  stat_ellipse(aes(colour = time_string, fill = time_string), 
               alpha = 0.2, geom = "polygon", level = 0.95)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  labs(title= "24h samples, ellipse by time")+
  scale_color_manual(values=treatment_colors)+
  #scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("24h_time.png", width = 16, height = 10, path = here("outputs", "DNA"))
```


