---
author: 'Keanu Rochette'
title: "MOBEAST DOC V3"
created: '2025-11-10'
format: 
  html:
    toc: true
    toc-location: left
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
library(emmeans)
library(performance)
library(multcompView)



```

## Load Data

```{r}
mobeast <- read_csv(here("data", "MOBEAST_full_merged_data.csv"))
```

## Data visualization theme

```{r}
nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=24, face = "bold"), 
        plot.subtitle = element_text(size=20),
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=18, face = "bold"),
        legend.text=element_text(size=16),
        strip.background = element_rect(fill = "white"))
```



```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

## Data Clean Up

```{r}
mobeast %>% select(id_number:date_time, npoc_u_m, avg_res_time, avg_flow_rate)
```


### Outlier detection on bulk data
```{r}
# this data is log transformed
## flagging outliers from the data set
raw_flagged_outliers <- DOC %>%
  select(-npoc_mg_l, -tn_mg_l, -long_name, -sample) %>% 
  pivot_longer(cols = npoc_u_m:tn_u_m,
               names_to = "nutrient",
               values_to = "conc") %>% 
  filter(treatment != "Endmember") %>% 
  filter(nutrient != "TN") %>% 
  # flagging outliers
  ## grouping by date, time and treatment
  group_by(nutrient, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(log10(conc), probs = 0.25, na.rm = T),
         quant3 = quantile(log10(conc), probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( log10(conc) > quant3 + (IQR*1.5) | log10(conc) < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() #%>% 
  
  ## remove the outliers 
  #filter(outlier != "outlier") 

```


### Normalizing the data

- Avg inflow
- Norm = [Tank] - [Avg Inflow]

```{r}
# Normalization version 2
## averaging the inflows for normalization 
## remove the outliers a right before the normalization 

## Normalization of the data to the inflow 
#nutrients_norm <- 
DOC_norm <- raw_flagged_outliers %>% 
  mutate(time_string = as.character(time),
         date_string = as.character(date),
         date_time_string = paste(date_string, time_string),
         date_time = ymd_hms(date_time_string)) %>% 
  select(-IQR, -quant1, -quant3) %>% 
  
  filter(treatment != "Rain") %>% 
  
  ## testing the log10 data transformation
  #mutate(conc = log10(conc)) %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table, nutrient) %>%
  mutate(inflow = conc[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time, nutrient) %>% 
  mutate(inflow = ifelse(outlier == "outlier", inflow == NA, inflow)) %>% 
  mutate(inflow = mean(inflow, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(conc_norm = conc - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```

```{r}
DOC_norm <- DOC_norm %>% mutate(nutrient = ifelse(nutrient == "npoc_u_m", "DOC", "TN"))
```

```{}
# Dataset sent to Craig 
DOC_craig <- DOC_norm %>% filter(outlier != "outlier") %>% 
  filter(nutrient == "DOC") %>% select(-inflow)

#write_csv(DOC_craig, here("data","craig_data", "DOC_Craig.csv"))
```

```{r}
#Verifying that the outliers have been removed 
DOC_norm %>% filter(nutrient != "TN") %>% 
  #filter(outlier == "outlier") %>% 
  filter(time_string == "00:00:00") %>% 
  select(id_number, tank_number, treatment, date_time_string, conc, conc_norm) 
```


## 12PM data
```{r}
# Version 2
## removing outliers from the data set
DOC_12PM <- DOC_norm %>%
  filter(time_string == "12:00:00") %>% 
  filter(treatment != "Inflow") %>% 
  
  ## remove the outliers 
  filter(outlier == "not outlier") 

```


```{r}
DOC_12PM %>% 
  filter(nutrient == "DOC") %>% 
  select(tank_number, treatment, date_time, conc, inflow, conc_norm)
```

### Iterative statistical modeling

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

DOC_loop <- DOC_12PM %>% select(nutrient) %>% distinct() %>% pull()

models_12PM <- vector(mode = "list", length = length(DOC_loop))
names(models_12PM) <- DOC_loop

for (var in DOC_loop) {
  df <- DOC_12PM %>% filter(nutrient == var)
  models_12PM[[var]] <- lmer(conc_norm~ treatment + (1|tank_number), data = df)
}

```

### Iterative graphing

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## --> In the for-loop, filter the df by nutrient and make a graph that will display nutrient data
## Save the outputs of the model in the list such that each nutrient has its stats

DOC_loop <- DOC_12PM %>% select(nutrient) %>% distinct() %>% pull()

graphs_12PM <- vector(mode = "list", length = length(DOC_loop))
names(graphs_12PM) <- DOC_loop

for (var in DOC_loop) {
  
  df <- DOC_12PM %>% filter(nutrient == var)
  
  graphs_12PM[[var]] <- df %>%  
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("12PM", var, "Production"),
       x= "Treatment",
       y = "Production (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none")
}

```


#### 12PM DOC concentrations

```{r}
summary(models_12PM[["DOC"]])
anova(models_12PM[["DOC"]], type = 2)
emmeans(models_12PM[["DOC"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[[1]], ~ treatment)
```

```{r}
resid_panel(models_12PM[["DOC"]])

```

#### Data Visualization 

```{r, fig.width= 12, fig.height = 6}
graphs_12PM[["DOC"]]#+ 
  #geom_jitter(aes(color = inflow_table), size = 2)+
  #scale_color_manual(values = c("red", "blue"))+
  #facet_wrap(date~.)+
  #theme(legend.position = "right")

#ggsave(DOC_12PM.png, path = here("outputs", "Nov_com_meeting", "DOC_12PM.png"))
```

#### one-sample t-test

```{r}
qqnorm(DOC_12PM$conc_norm)
qqline(DOC_12PM$conc_norm)
```

```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
DOC_12PM %>%
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```


## Hypothesis 2: There are variation in exudation on a diel cycle 24h

```{r}
## removing outliers from the data set
DOC_24h <- DOC_norm %>%
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  filter(treatment != "Inflow") %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00"))) %>% 
  filter(time_string != "00:00:00") %>% 
  filter(nutrient == "DOC")
```



#### 24h Model setup and results
```{r}
mod <- lmer(conc_norm ~ treatment*time_fac + (1|tank_number), data = DOC_24h)
```

```{r}
summary(mod)
anova(mod, type = 2)
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
#emmeans(models_24h[["DOC"]], ~ treatment)
```

```{r}
resid_panel(mod)
```

#### Data Visualization 
```{r, fig.width= 12, fig.height = 6}
DOC_24h %>% 
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_wrap(~time_fac, ncol = 4 )+
    labs(title = paste("24h DOC Production"),
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Production (µmol/L)")+
    scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18),
        legend.position = "none") #+
  #geom_jitter(aes(color = inflow_table), size = 2)+
  #scale_color_manual(values = c("red", "blue"))+
  #theme(legend.position = "right")
```



### Pairwise comparison

```{r}
t_test_test <- DOC_12PM %>% 
  filter(time_string == "12:00:00",
         treatment == "Coral") 

t.test(t_test_test$conc_norm , mu=0)
```

```{r}
DOC_12PM %>% 
  filter(time_string == "12:00:00") %>% 
  #mutate(time_string = a)
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  ) %>% 
  select(-t_test) %>% 
  ungroup() 

```


```{}
control_24h <- DOC_24h %>% filter(treatment == "Coral") %>% 
  filter(time_string != "12:00:00") %>% 
  mutate(time_US = ifelse(time_string == "18:00:00", "six_pm", "six_am")) %>% 
  select(tank_number, time_US, conc_norm) %>% 
  pivot_wider(names_from = time_US, values_from = conc_norm) 

t.test(control_24h$six_pm, control_24h$six_am, paired = TRUE)
t.test(control_24h$six_pm, control_24h$six_am, paired = TRUE)$estimate
t.test(control_24h$six_pm, control_24h$six_am, paired = TRUE)$p.value
```



```{r}
nunu_pairwise <- function(data, time_point = 6){

  ## Get a list of the treatments for the for loop   
  treatment <- data %>% mutate(treatment = as.character(treatment)) %>%
  select(treatment) %>% distinct() %>% pull()
  
  ## Create un empty data frame
pairwise_results <- tibble(
  time_point = vector(),
  treatment= vector(),
  mean = vector(),
  p_value = vector())

# this is the pairwise comparison for the 6 pairs (6AM-6PM)
  if(time_point == 6){
    for(var in treatment){
      
      ## only keep the 6AM and 6PM
      df <- data %>% filter(treatment == var) %>% 
  filter(time_string == "18:00:00" | time_string == "06:00:00") %>% 
        ## Rename the times so they can become column names
  mutate(time_US = ifelse(time_string == "18:00:00", "T18", "T06")) %>% 
        ## Only keep the column of interest for the stats
  select(tank_number, time_US, conc_norm) %>% 
  pivot_wider(names_from = time_US, values_from = conc_norm) 
      
      ## Pairwise function 
      paired <- t.test(df$T18, df$T06, paired = TRUE, na.action = na.pass)
      
      ## Save the result in at data frame
      df2 <- tibble(
        time_point = "06AM-06PM",
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      # Code will loop oveer all treatments and save the results in a data frame
      pairwise_results <-rbind(pairwise_results,df2)
      
      # Save the results in RStudio
      pairwise_results <<- pairwise_results
    }  
  

# This is for the 12 samples (12AM-12PM)
  }else if(time_point == 12){
    for(var in treatment){
      
      ## Same as before but for the 12AM-12PM
      df <- data %>% filter(treatment == var) %>% 
  filter(time_string == "12:00:00" | time_string == "00:00:00") %>% 
  mutate(time_US = ifelse(time_string == "12:00:00", "T12", "T00")) %>% 
  select(tank_number, time_US, conc_norm) %>% 
  pivot_wider(names_from = time_US, values_from = conc_norm) 
      
      paired <- t.test(df$T12, df$T00, paired = TRUE, na.action = na.pass)
      
      df2 <- tibble(
        time_point = "12AM-12PM",
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      pairwise_results <- rbind(pairwise_results,df2)
      pairwise_results <<- pairwise_results
    }
}

# Display the results of the code 
print(pairwise_results)
}

```


```{r}
nunu_pairwise(data = DOC_24h, time_point = 6) 
```









