---
title: "MOBEAST - 24h Heatmap"
format: html
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)

library(ComplexHeatmap)
library(data.table)
library(dendextend)
library(grid)

# Rayna library
library(dendsort)
library(ggsci)
library(randomForest)
library(vegan)
library(ape)
library(devtools)
library(ggpubr)
library(viridis)
library(colorspace)
library(readr)
library(circlize)


library(cowplot)
library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
library(emmeans)
library(performance)
library(multcompView)
library(broom.mixed)
library(effectsize)

```

Heatmap info and tutorial: https://jokergoo.github.io/ComplexHeatmap-reference/book/a-single-heatmap.html

## Functions

```{r}
source(here("scripts", "reference_scripts", "heatmap_zscore_function.R"))
```

## Data visualization theme

```{r}

nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=24, face = "bold"), 
        plot.subtitle = element_text(size=20),
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=18, face = "bold"),
        legend.text=element_text(size=16),
        axis.text.x = element_text(angle= 45, hjust= 1),
        strip.background = element_rect(fill = "white"))
```

### Color palette

```{r}
treatment_colors = c(
  #Treatments
  "Control" = "grey",
  "Coral" = "#03A9F4",
  "CCA" = "#EF5350",
  "Algae" = "#4CAF50"
)

```

## Load Data

```{r}
meta <- read_csv(here("data", "MOBEAST_metadata_V2.csv"))
OTU_long <- read_csv(here("data","MOBEAST_DNA_long.csv"))
mobeast <- read_csv(here("data", "MOBEAST_full_merged_data_long.csv"))

```

```{r}
flow_meta <- mobeast %>%  
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>% 
  select(id_number, flowrate) %>% 
  distinct() %>% 
  filter(!is.na(id_number))
```


## Data Clean Up

```{r}
OTU_long <- OTU_long %>% clean_names() %>% select(-long_name, -tank_number, -treatment, -date,-time) %>%  rename(id_number= sample_id)

DNA <- meta %>% left_join(OTU_long)
```

### Preparing the OTU matrix for heatmap

```{r}
exclude_family <- DNA %>% group_by(family) %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>%
  reframe(sum = sum(abund)) %>% 
  filter(sum ==0) %>% 
  select(family) %>% 
  pull()

unclassified_families <- c("Alteromonadales_unclassified", "Alphaproteobacteria_unclassified", "Gammaproteobacteria_unclassified", "Campylobacteria_unclassified", "Bacteroidia_unclassified", "Bacteria_unclassified", 
                           "Proteobacteria_unclassified ")
```

```{r}
OTU_htm <- DNA %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>%
  filter(!family %in% exclude_family) %>% 
  mutate(family_tax = paste(family, "X", order, "X", class, "X", phylum),
         genus_tax = paste(genus, "X", family, "X", order, "X", class, "X", phylum))
```

```{r}
OTU_htm_24h <- OTU_htm %>% filter(treatment != "Inflow")
  
```



## Data with only the most abundant taxa 


```{r}
# genus
## per sample, min abundance 0.5% and more than 3x occurence. 

taxa_filtered <- OTU_htm_24h %>%
  
  mutate(genus_tax = paste0(genus, ", ", family, ", ", class)) %>% 
  filter(!family %in% unclassified_families) %>%
  
  #group_by(genus_tax) %>% 
  #mutate(sum_abund= sum(abund)) %>% 
  #mutate(max = max(sum_abund)) %>%
  #filter(max >0.005) %>%
  #ungroup() %>%

  ## remove taxa that don't have an abundance >0.5% in at least 3 samples
  group_by(id_number, genus_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(genus_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.005)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  mutate(taxa = genus_tax) %>% 
  select(-long_name, -kingdom, -family_tax, #-max, 
         -n_above_threshold) %>% distinct()

taxa_filtered
  
taxa_filtered %>% select(taxa) %>% distinct()
```




### SQRT-transformed Z-score 

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- taxa_filtered %>%
  group_by(id_number, taxa) %>%
  summarize(abund = sum(abund, na.rm = TRUE)) %>% 
  ungroup()

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = taxa, values_from = abund)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% tibble() %>%  select(-id_number)

```

```{r}
sqrt.matrix <- apply(culled.abund.OTU, 2, FUN=sqrt)
sqrt.zscore <- as.data.frame(apply(sqrt.matrix, 2, FUN = zscore.calculation))
```

```{r}
# bring back the sample ID number in the df 
sqrt.zscore$id_number = family.culled.abund$id_number


# Set the ID numbers as the rownames (to make a matrix)
rownames(sqrt.zscore) = sqrt.zscore$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
sqrt.zscore.matrix <- sqrt.zscore %>% select(-id_number) %>% 
  as.matrix() %>% t()


```

### Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(sqrt.zscore.matrix),
                           meta$id_number), 
                 "treatment"]

date_time <- meta[match(colnames(sqrt.zscore.matrix),
                           meta$id_number),
                  "date_time_string"]

time <- meta[match(colnames(sqrt.zscore.matrix),
                           meta$id_number),
                  "time_string"]

```


#### Organize the samples (ChatGPT)
```{r}

## organize the treatment in order they will be display on the heatmap
treatment2 <- factor(treatment$treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))

## organize the time in order they will be display on the heatmap
time2 <- factor(time$time_string, levels = c("12:00:00", "18:00:00", "00:00:00", "06:00:00"))

#store that info in a df
col_info <- data.frame(
  Treatment = treatment2,
  Time = time2) %>% drop_na()

# create the order in which the samples with be organize in the heatmap
sample_order <- order(col_info$Treatment,col_info$Time)

# tell the matrix with my z-scores how the samples should be arranged
mat_ordered <- sqrt.zscore.matrix[, sample_order]

# have vectors specifying the order of the treatment for the heatmap
treatment_ordered <- col_info$Treatment[sample_order]
time_ordered <- col_info$Time[sample_order]


```


```{r}
col_ha <- HeatmapAnnotation(
  Treatment = anno_block(
    gp = gpar(fill = c("grey", "#03A9F4", "#EF5350", "#4CAF50")),
    labels = c("Control", "Coral", "CCA", "Algae"), 
    labels_gp = gpar(col = "white", fontsize = 14, fontface = "bold")),
  
  Time = time_ordered, 
  
  col = list(Treatment = c("Control" = "grey", 
                           "Coral"   = "#03A9F4", 
                           "CCA"     = "#EF5350",
                           "Algae"   = "#4CAF50"),
             Time = c("12:00:00" = "black",
                      "18:00:00" = "green", 
                      "00:00:00" = "blue",
                      "06:00:00" = "grey")),
  
   # Controls the "Time" label on the left side of the ribbon *
   annotation_name_gp = gpar(fontsize = 16, fontface = "bold"), 
  
   # Controls the height of the Time ribbon bar itself 
   simple_anno_size = unit(6, "mm"), 
  
  annotation_legend_param = list(
    
     Time = list( 
       title_gp    = gpar(fontsize = 16, fontface = "bold"), 
       labels_gp   = gpar(fontsize = 16)
    )
  )
)
```




#### Printing heatmap

```{r, fig.height = 16, fig.width = 24}
# Version 2
#png(width=3000, height=2200, file= here("outputs", "zscore_24hdata.png"))
#png(width=2000, height=1200, file= here("outputs", "zscore_24hdata.png"))

ht = Heatmap(mat_ordered, 
             name = "Z-scored Abundance", 
             heatmap_legend_param = list(#legend_direction = "horizontal",
                                          title_gp= gpar(fontsize = 16, 
                                                          fontface = "bold"), 
                labels_gp       = gpar(fontsize = 16), 
                grid_height     = unit(6, "mm"), 
                grid_width      = unit(6, "mm"), 
                legend_width    = unit(8, "cm")),   # controls horizontal legend length),
             show_row_names = TRUE,
        row_names_max_width = unit(22, "cm"),
        clustering_method_rows = "complete",
        clustering_distance_rows = "euclidean",
             #rect_gp = gpar(col = "white", lwd = 0.1),
        
        column_title = "Bacterial Community Shift Over 24h", 
        column_title_gp = gpar(fontsize = 25, fontface = "bold"),
        #column_title_side = "bottom",
        cluster_rows = TRUE,
        cluster_columns =  FALSE,
        row_title = c("A", "B", "C") ,
        row_title_rot = 0,
        row_title_gp = gpar(fontsize = 30, fontface = "bold"),
        row_names_gp = gpar(fontsize = 16),
        row_dend_width = unit(2, "cm"),
        #column_names_gp = gpar(fontsize = 10),
        show_column_names = FALSE,
        row_km = 3, # 5? number of forced segregation in rows 
        #column_km = 4,#number of forced segregation in columns 
        top_annotation = col_ha,
        #bottom_annotation = col_ha,
        column_split = treatment_ordered
        )
set.seed(1999)
ht_drawn <- draw(ht, heatmap_legend_side = "right",
                 annotation_legend_side = "right", 
  merge_legends = TRUE)

#dev.off()

```

```{r, fig.height = 16, fig.width = 30}
# --- Step 1: draw once as before to get the k-means groupings ---
set.seed(1999)
ht_initial <- Heatmap(mat_ordered, 
             name = "Z-scored Abundance", 
             heatmap_legend_param = list(
                title_gp     = gpar(fontsize = 16, fontface = "bold"), 
                labels_gp    = gpar(fontsize = 16), 
                grid_height  = unit(6, "mm"), 
                grid_width   = unit(6, "mm"), 
                legend_width = unit(8, "cm")),
             show_row_names = TRUE,
             row_names_max_width = unit(22, "cm"),
             clustering_method_rows = "complete",
             clustering_distance_rows = "euclidean",
             column_title = "Bacterial Community Shift Over 24h", 
             column_title_gp = gpar(fontsize = 25, fontface = "bold"),
             cluster_rows = TRUE,
             cluster_columns = FALSE,
             row_title = c("A", "B", "C"),
             row_title_rot = 0,
             row_title_gp = gpar(fontsize = 30, fontface = "bold"),
             row_names_gp = gpar(fontsize = 16),
             row_dend_width = unit(2, "cm"),
             show_column_names = FALSE,
             row_km = 3,
             top_annotation = col_ha,
             column_split = treatment_ordered
             )

ht_drawn_initial <- draw(ht_initial, heatmap_legend_side = "right",
                          annotation_legend_side = "right", 
                          merge_legends = TRUE)

# --- Step 2: extract which rows belong to which displayed slice (1 = A, 2 = B, 3 = C) ---
row_order_list <- row_order(ht_drawn_initial)

split_vec <- integer(nrow(mat_ordered))
for (i in seq_along(row_order_list)) {
  split_vec[row_order_list[[i]]] <- i
}

# --- Step 3: build a factor that swaps A and B while keeping C in place ---
# original position 1 = "A", 2 = "B", 3 = "C"
# levels = c(2,1,3) means: whatever was position 2 now displays first, position 1 displays second
row_split_swapped <- factor(split_vec, levels = c(2, 1, 3), labels = c("A", "B", "C"))

# --- Step 4: redraw the heatmap using the manual split, order fixed (no re-clustering of slices) ---
ht = Heatmap(mat_ordered, 
             name = "Z-scored Abundance", 
             show_heatmap_legend = FALSE,
             #show_annotation_legend = FALSE,
             heatmap_legend_param = list(
                title_gp     = gpar(fontsize = 16, fontface = "bold"), 
                labels_gp    = gpar(fontsize = 16), 
                grid_height  = unit(6, "mm"), 
                grid_width   = unit(6, "mm"), 
                legend_width = unit(8, "cm")),
             show_row_names = TRUE,
             row_names_max_width = max_text_width(rownames(mat_ordered), 
                                                  gp = gpar(fontsize = 22)),
             #row_names_max_width = unit(22, "cm"),
             clustering_method_rows = "complete",
             clustering_distance_rows = "euclidean",
             column_title = "Bacterial Community Shift Over 24h", 
             column_title_gp = gpar(fontsize = 25, fontface = "bold"),
             cluster_rows = TRUE,
             cluster_columns = FALSE,
             row_split = row_split_swapped,
             cluster_row_slices = FALSE,   # <- keeps our custom order instead of auto-reordering
             row_title = c("A", "B", "C"),
             row_title_rot = 0,
             row_title_gp = gpar(fontsize = 30, fontface = "bold"),
             row_names_gp = gpar(fontsize = 16),
             row_dend_width = unit(2, "cm"),
             show_column_names = FALSE,
             top_annotation = col_ha,
             column_split = treatment_ordered
             )

set.seed(1999)
ht_drawn_vis <- draw(ht, heatmap_legend_side = "right",
                 annotation_legend_side = "right",
                 merge_legends = TRUE)
```

# Modules
## Data clean up 

```{r}

# 4. Extract row positions grouped by cluster slice
cluster_list <- row_order(ht_drawn)

# 5. Convert the list into a clean, unified data frame matching gene names to cluster IDs
cluster_membership <- do.call(rbind, lapply(names(cluster_list), function(cl_name) {
  row_indices <- cluster_list[[cl_name]]
  data.frame(
    taxa = rownames(mat_ordered)[row_indices],
    cluster = paste("Cluster", cl_name)
  )
}))

cluster_membership
```


```{r}
modules <- taxa_filtered %>%  select(id_number, tank_number,treatment,date, time, date_time,
                                     inflow_table,taxa, abund) %>% 
  left_join(cluster_membership) 

modules <- modules %>% left_join(flow_meta)


```

```{r}
modules <- modules %>%  
  mutate(cluster2 = cluster, 
         cluster = ifelse(cluster2 == "Cluster 1", "Module B",
                          ifelse(cluster2 == "Cluster 3", "Module A", "Module C")),
         cluster = factor(cluster, levels = c("Module A", "Module B", "Module C")))
```


```{r}
modules_24h <- modules %>% 
  mutate(time_fac = factor(time, levels = c("12:00:00", "18:00:00", "00:00:00", "06:00:00")),
         treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae")))
```


### Module averages and stuff 

```{r}
module_A_bact <- modules_24h %>%  
  group_by(taxa, cluster) %>% 
  reframe(avg = mean(abund)*100) %>%  
  filter(cluster == "Module A") %>% 
  arrange(desc(avg)) %>%  
  pull(taxa)
```



### Model of the modules

```{r}
modules_df <- modules_24h %>% 
  group_by(id_number, cluster) %>% 
  mutate(conc_norm = sum(abund)#,
         #conc_norm = sqrt(sum_ra_module)
         ) %>% 
  ungroup() %>% 
  mutate(log_flow = log10(flowrate)) %>%
  select(id_number, tank_number, treatment, date, time, date_time, time_fac, inflow_table, 
         conc_norm, cluster, log_flow) %>% distinct()

mobeast_loop <- modules_df %>% 
  select(cluster) %>% distinct() %>% pull()

models_modules <- vector(mode = "list", length = length(mobeast_loop))
names(models_modules) <- mobeast_loop


for (var in mobeast_loop) {
  df <- modules_df %>% filter(cluster == var)
  models_modules[[var]] <- lmer(conc_norm ~ treatment*time_fac + 
                                  log_flow + (1|tank_number), data = df)
  
}

```


### Module ANOVA

```{r, warning= FALSE}

anova_results <- imap_dfr(models_modules, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

anova_results %>% filter(p.value <0.05) %>% 
  select(variable, term, statistic, p.value)

```

### Module Emmeans 

```{r, message= FALSE}

emmeans_results_dna <- imap_dfr(models_modules, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae")

emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% 
  select(variable, term, estimate, contrast, adj.p.value)
```


### Plot of the clusters/module
```{r Family clustering, fig.width= 16, fig.height= 6}
modules_df %>% drop_na(cluster) %>%
  group_by(treatment, time_fac, cluster) %>%
  reframe(mean_abund = mean(conc_norm),
          sd_abund   = sd(conc_norm),
          conc_norm) %>%
  ungroup() %>% 
  ggplot(aes(x = time_fac, y = mean_abund, group = treatment, color = treatment)) +
  geom_line(linewidth = 1) +
  geom_point( size = 2)+
  geom_ribbon(aes(ymin = mean_abund - sd_abund,
                  ymax = mean_abund + sd_abund,
                  fill = treatment),
                  alpha = 0.15, color = NA) +
  geom_point(aes(x = time_fac, y = conc_norm), shape = 1, alpha = 0.7,
             size = 2.5) +
  labs(#title = "Modules from Genus Clustering", 
       x= "Time", 
       y= "Relative Abundance")+
  scale_color_manual(values = treatment_colors) +
  scale_fill_manual(values  = treatment_colors) +
  facet_grid(. ~ cluster)+
  nunu_theme



#ggsave(filename = "modules_ICRS.png", plot = last_plot(), path = here("outputs"))
```





# Taxa models OTU and Genus

## Data clean up 



```{r}
DNA <- OTU_htm %>% 
  left_join(flow_meta) %>% 
  mutate(date_string = as.character(date_string),
         time_string = as.character(time_string), 
         date_time_string = as.character(date_time_string)) %>% 
  mutate(OTU_tax = paste0(otu_number, ", ", family, ", ", order,", ", class),
         log_flow = log10(flowrate),
         time_fac = factor(time_string, 
                           levels = c("12:00:00", "18:00:00", "00:00:00","06:00:00"))) %>% 
  select(-long_name)

DNA_meta <- DNA %>% select(id_number, tank_number, treatment, time_fac, log_flow ) %>% distinct()
```


## Sqrt OTU Models : sqrt(RA)

```{r}
# Sqrt transform 
OTU_24h <- DNA %>%  
  filter(!family %in% unclassified_families) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, OTU_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  select(id_number, tank_number, treatment, date, time, date_time, time_fac, inflow_table, 
         sum_abund, OTU_tax, log_flow) %>% 
  
  # sqrt() transform
  mutate(conc_norm = sqrt(sum_abund)) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae")))
  
  
```


```{r}
exclude_tax <- OTU_24h %>% 
  group_by(time, OTU_tax) %>% 
  reframe(avg = mean(conc_norm)) %>% 
  filter(avg == 0) 
exclude_tax
```

```{r}
mobeast_loop <- OTU_24h %>% 
  filter(!OTU_tax %in% exclude_tax$OTU_tax & !OTU_tax %in% exclude_tax$time) %>%
  select(OTU_tax) %>% distinct() %>% pull()

models_DNA_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_DNA_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- OTU_24h %>% filter(OTU_tax == var)
  models_DNA_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + 
                                  log_flow + (1|tank_number), data = df)

}
```


```{}
resid_panel(models_DNA_24h[["Otu0024 X Rhodobacteraceae X Rhodobacterales X Alphaproteobacteria"]], plots = c("qq"))
```



```{r}
results_DNA <- imap_dfr(models_DNA_24h, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")

results_DNA %>%  filter(p.value <0.05)
```

### OTU ANOVA

```{r, warning= FALSE}

anova_results <- imap_dfr(models_DNA_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

anova_results %>% filter(term == "treatment", p.value <0.05) %>% 
  select(variable, term, statistic, p.value )

anova_sig <- anova_results %>% 
  filter(term == "treatment", p.value <0.05) %>%  
  select(variable) %>%  distinct() %>% pull()

anova_table <- anova_results %>% 
  select(OTU = variable, term, DF = NumDF, F_value = statistic, p.value) %>% 
  mutate(F_value = round(F_value, digits = 2),
         p.value = round(p.value, digits = 2))

write_csv(anova_table, path = here("outputs","otu_sqrt_anova.csv"))

```

### OTU Emmeans 

```{r, message= FALSE}

emmeans_results_dna <- imap_dfr(models_DNA_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae")

emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% 
  filter(variable %in% anova_sig, adj.p.value < 0.05) %>%
  select(variable, contrast, estimate, adj.p.value)
```


```{}

# Get a sorted, unique list of taxa
taxa_list <- OTU_24h %>% distinct(OTU_tax) %>% pull(OTU_tax) %>% sort()

# Function to build one QQ plot for a given taxon
make_qq <- function(taxon) {
  df <- OTU_24h %>% filter(OTU_tax == taxon)
  
  ggplot(df, aes(sample = conc_norm )) +
    stat_qq() +
    stat_qq_line(color = "red") +
    labs(title = taxon) +
    theme_minimal(base_size = 8) +
    theme(
      plot.title = element_text(size = 7, hjust = 0.5),
      axis.title = element_blank(),
      axis.text = element_text(size = 5)
    )
}

# Build the list of plots
qq_plots <- map(taxa_list, make_qq)

# Split into chunks of 16 and assemble each chunk into a 4x4 grid
chunks <- split(qq_plots, ceiling(seq_along(qq_plots) / 16))

grid_figures <- map(chunks, ~ wrap_plots(.x, ncol = 4, nrow = 4))

# Save all grids to the outputs folder, one file per page
walk2(grid_figures, seq_along(grid_figures), function(fig, i) {
  ggsave(
    filename = here("outputs", paste0("qq_grid_", i, ".png")),
    plot = fig,
    width = 14, height = 14, dpi = 300
  )
})
```


### OTU Tiles enrichment 
```{r}
sig_taxa <- emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% select(variable) %>%
  distinct() %>% pull()

emmeans_results_dna %>% filter(variable %in% sig_taxa) %>% 
  mutate(contrast = factor(contrast, levels = c("Control - Coral", "Control - CCA", "Control - Algae"))) %>% 
  mutate(sig = ifelse(estimate <0 & adj.p.value < 0.05, "Enrichment", 
                      ifelse(estimate > 0 & adj.p.value < 0.05, 
                             "Depletion", "No change"))) %>% 
  ggplot(aes(x= contrast, y = variable, fill = sig)) + 
  geom_tile(color = "black") +
  scale_fill_manual(breaks = c("Enrichment", "Depletion", "No change"),
                    values = c("lightgreen", "#EF5350", "#03A9F4")) +
  theme(legend.position = "bottom")+
  theme_bw()
```




## Delta OTU models

```{r}
# Delta no transformation, no z-scoring 
OTU_24h <- DNA %>%  
  filter(!family %in% unclassified_families) %>% 
  
  group_by(id_number, OTU_tax) %>% 
  mutate(sum_abund = sum(abund, na.rm = T)) %>% 
  ungroup() %>% 
  select(id_number, tank_number, treatment, date, time, date_time, time_fac, inflow_table, 
         sum_abund, OTU_tax, log_flow) %>%
  
  # calculating delta values
  mutate(inflow_isolated = ifelse(treatment == "Inflow", sum_abund, NA)) %>% 
  group_by(time, OTU_tax) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = T )) %>% 
  ungroup() %>%
  mutate(conc_norm  = sum_abund - inflow ) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae")))
   
```


```{r}
exclude_tax <- OTU_24h %>% 
  group_by(time, OTU_tax) %>% 
  reframe(avg = mean(conc_norm)) %>% 
  filter(avg == 0) 
exclude_tax
```

### OTU models
```{r}
mobeast_loop <- OTU_24h %>% 
  filter(!OTU_tax %in% exclude_tax$OTU_tax & !OTU_tax %in% exclude_tax$time) %>%
  select(OTU_tax) %>% distinct() %>% pull()

models_DNA_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_DNA_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- OTU_24h %>% filter(OTU_tax == var)
  models_DNA_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + 
                                  log_flow + (1|tank_number), data = df)

}
```


```{}
resid_panel(models_DNA_24h[["Otu0024 X Rhodobacteraceae X Rhodobacterales X Alphaproteobacteria"]], plots = c("qq"))
```



```{r}
results_DNA <- imap_dfr(models_DNA_24h, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")

results_DNA %>%  filter(p.value <0.05)
```

### OTU ANOVA

```{r, warning= FALSE}

anova_results <- imap_dfr(models_DNA_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

anova_results %>% filter(term == "treatment", p.value <0.05) %>%  select(variable, term, statistic, p.value)


anova_sig <- anova_results %>% 
  filter(term == "treatment", p.value <0.05) %>%  
  select(variable) %>%  distinct() %>% pull()
```


```{r}
anova_table <- anova_results %>% 
  select(OTU = variable, term, DF = NumDF, F_value = statistic, p.value) %>% 
  mutate(F_value = round(F_value, digits = 2),
         p.value = round(p.value, digits = 2))

write_csv(anova_table, path = here("outputs","otu_delta_anova.csv"))
```



### OTU Emmeans 

```{r, message= FALSE}

emmeans_results_dna <- imap_dfr(models_DNA_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae")

emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% 
  filter(variable %in% anova_sig, adj.p.value < 0.05) %>%
  select(variable, contrast, estimate, adj.p.value)
```


### OTU tiles 
```{r}
sig_taxa <- emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% select(variable) %>%
  distinct() %>% pull()

emmeans_results_dna %>% filter(variable %in% sig_taxa) %>% 
  mutate(contrast = factor(contrast, levels = c("Control - Coral", "Control - CCA", "Control - Algae"))) %>% 
  mutate(sig = ifelse(estimate <0 & adj.p.value < 0.05, "Enrichment", 
                      ifelse(estimate > 0 & adj.p.value < 0.05, 
                             "Depletion", "No change"))) %>% 
  ggplot(aes(x= contrast, y = variable, fill = sig)) + 
  geom_tile(color = "black") +
  scale_fill_manual(breaks = c("Enrichment", "Depletion", "No change"),
                    values = c("lightgreen", "#EF5350", "#03A9F4")) +
  theme(legend.position = "bottom")+
  theme_bw()
```





# Genus models

## Sqrt Genus Models

```{r}
genus_24h <- DNA %>%
  
  filter(!family %in% unclassified_families)  %>% 
  
  mutate(genus_tax = paste0(genus, ", ", family,", ", class),
         time_fac = factor(time, 
                           levels = c("12:00:00", "18:00:00", "00:00:00", "06:00:00")),
         log_flow = log10(flowrate)) %>% 

  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, genus_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  select(id_number, tank_number, treatment, date, time, date_time, time_fac, inflow_table, 
         sum_abund, genus_tax, log_flow) %>% 
  
  # outlier detection
  mutate(conc_norm = sqrt(sum_abund)) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) 

  
```


```{r}
exclude_tax <- genus_24h %>% 
  group_by(time, genus_tax) %>% 
  reframe(avg = mean(conc_norm)) %>% 
  filter(avg == 0) 
exclude_tax
```

### Genus models
```{r}
mobeast_loop <- genus_24h %>% 
  filter(!genus_tax %in% exclude_tax$genus_tax & !genus_tax %in% exclude_tax$time) %>%
  select(genus_tax) %>% distinct() %>% pull()

models_DNA_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_DNA_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- genus_24h %>% filter(genus_tax == var)
  models_DNA_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + 
                                  log_flow + (1|tank_number), data = df)

}
```


```{}
resid_panel(models_DNA_24h[["Vibrio X Vibrionaceae X Vibrionales X Gammaproteobacteria X Proteobacteria"]], plots = c("qq"))
```


### Genus ANOVA

```{r, warning= FALSE}  

anova_results <- imap_dfr(models_DNA_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 


anova_results %>% filter(term == "treatment", p.value <0.05) %>% 
  select(variable, term, p.value) 

anova_sig <- anova_results %>% 
  filter(term == "treatment", p.value <0.05) %>% 
  select(variable, term, p.value) %>% 
  select(variable) %>% 
  distinct() %>% 
  pull()

anova_table <- anova_results %>% 
  select(OTU = variable, term, DF = NumDF, F_value = statistic, p.value) %>% 
  mutate(F_value = round(F_value, digits = 2),
         p.value = round(p.value, digits = 7))

write_csv(anova_table, path = here("outputs","genus_sqrt_anova.csv"))
```

### Genus Emmeans 

```{r, message= FALSE}

emmeans_results_dna <- imap_dfr(models_DNA_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae")


emmeans_results_dna %>% 
  filter(adj.p.value < 0.05,
         variable %in% anova_sig) %>% 
  select(variable, contrast, estimate, adj.p.value) %>% 
  filter(variable %in% module_A_bact)


module_A_sig_emm <- emmeans_results_dna %>% 
  filter(adj.p.value < 0.05,
         variable %in% anova_sig) %>% 
  select(variable, contrast, estimate, adj.p.value) %>% 
  filter(variable %in% module_A_bact) %>% 
  pull(variable)


emmeans_table_genus_sqrt <- emmeans_results_dna %>% 
  filter(adj.p.value < 0.05,
         variable %in% anova_sig) %>% 
  select(variable, contrast, estimate, adj.p.value) %>%
  mutate(estimate = round(estimate, digits = 3),
         adj.p.value = round(adj.p.value, digits = 3))

write_csv(emmeans_table_genus_sqrt, path = here("outputs","genus_sqrt_contrasts.csv"))
```


### Percent decrease ? 

```{r}
genus_24h %>%  filter(treatment == "Algae"|treatment == "Control", 
                      genus_tax %in% module_A_sig_emm) %>% 
  group_by(genus_tax, time_fac, treatment) %>% 
  reframe(avg = mean(sum_abund)) %>% 
  mutate(time_fac = as.character(time_fac)) %>% 
  filter(#time_fac == "12:00:00" | 
           time_fac== "06:00:00") %>% 
  #mutate(time_fac = ifelse(time_fac == "12:00:00", "noon", "six")) %>% 
  pivot_wider(names_from = treatment, values_from = avg) %>% 
  mutate(diff = (Algae -Control)*100) %>% 
  select(genus_tax, diff)
    
```

### Genus tiles

```{r, fig.width= 15, fig.height= 12}
sig_taxa <- emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% select(variable) %>%
  distinct() %>% pull()

emmeans_results_dna %>% 
  filter(variable %in% anova_sig,
         variable %in% sig_taxa) %>% 
  mutate(contrast = factor(contrast, levels = c("Control - Coral", "Control - CCA", "Control - Algae"))) %>% 
  mutate(sig = ifelse(estimate <0 & adj.p.value < 0.05, "Enrichment", 
                      ifelse(estimate > 0 & adj.p.value < 0.05, 
                             "Depletion", "No change"))) %>% 
  ggplot(aes(x= contrast, y = variable, fill = sig)) + 
  geom_tile(color = "black") +
  labs(x= "Contrast", y = "Bacterial Taxa",
       fill = "Significance")+
  scale_fill_manual(breaks = c("Enrichment", "Depletion", "No change"),
                    values = c("#4CAF50", "#EF5350", "#90CAF9")) +
  theme(legend.position = "bottom")+
  theme_bw()+
  nunu_theme

ggsave(filename = "genus_enrichment.png", plot = last_plot(), path = here("outputs"))
```


## Delta Genus models

```{r}
# Delta no transformation, no z-scoring 
genus_24h <- DNA %>%  
  
  filter(!family %in% unclassified_families) %>%
  
  mutate(genus_tax = paste0(genus, ", ", family,", ", class)) %>% 
  
  group_by(id_number, genus_tax) %>% 
  mutate(sum_abund = sum(abund, na.rm = T)) %>% 
  ungroup() %>% 
  select(id_number, tank_number, treatment, date, time, date_time, time_fac, inflow_table, 
         sum_abund, genus_tax, log_flow) %>%
  
  # calculating delta values
  mutate(inflow_isolated = ifelse(treatment == "Inflow", sum_abund, NA)) %>% 
  group_by(time, genus_tax) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = T )) %>% 
  ungroup() %>%
  
  mutate(conc_norm  = sum_abund - inflow ) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) 
   
```


```{r}
exclude_tax <- genus_24h %>% 
  group_by(time, genus_tax) %>% 
  reframe(avg = mean(conc_norm)) %>% 
  filter(avg == 0) 
exclude_tax
```

### Genus Models
```{r}
mobeast_loop <- genus_24h %>% 
  filter(!genus_tax %in% exclude_tax$genus_tax & !genus_tax %in% exclude_tax$time) %>%
  select(genus_tax) %>% distinct() %>% pull()

models_DNA_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_DNA_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- genus_24h %>% filter(genus_tax == var)
  models_DNA_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + 
                                  log_flow + (1|tank_number), data = df)

}
```


```{}
resid_panel(models_DNA_24h[["Otu0024 X Rhodobacteraceae X Rhodobacterales X Alphaproteobacteria"]], plots = c("qq"))
```



```{r}
results_DNA <- imap_dfr(models_DNA_24h, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")

results_DNA %>%  filter(p.value <0.05)
```

### Genus ANOVA

```{r, warning= FALSE}

anova_results <- imap_dfr(models_DNA_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

anova_results %>% filter(term == "treatment", p.value <0.05) %>% 
  select(variable, term, statistic, p.value)

anova_sig <- anova_results %>% 
  filter(term == "treatment", p.value <0.05) %>%  
  select(variable) %>%  distinct() %>% pull()

anova_table <- anova_results %>% 
  select(OTU = variable, term, DF = NumDF, F_value = statistic, p.value) %>% 
  mutate(F_value = round(F_value, digits = 2),
         p.value = round(p.value, digits = 2))

write_csv(anova_table, path = here("outputs","genus_delta_anova.csv"))
```

### Genus Emmeans 

```{r, message= FALSE}

emmeans_results_dna <- imap_dfr(models_DNA_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae")

emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% 
  filter(variable %in% anova_sig) %>% 
  select(variable, contrast, estimate, adj.p.value)
```

```{r}
otu_contrast_tab <- emmeans_results_dna %>% 
  filter(variable %in% anova_sig, adj.p.value < 0.05) %>% 
  select(variable, contrast, estimate, adj.p.value) %>% 
  gt()
gtsave(otu_contrast_tab, "delta_genus_contrast_tab.png", path = here("outputs"))
```

### Genus tiles 

```{r}
sig_taxa <- emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% select(variable) %>%
  distinct() %>% pull()

emmeans_results_dna %>% filter(variable %in% sig_taxa) %>% 
  mutate(contrast = factor(contrast, levels = c("Control - Coral", "Control - CCA", "Control - Algae"))) %>% 
  mutate(sig = ifelse(estimate <0 & adj.p.value < 0.05, "Enrichment", 
                      ifelse(estimate > 0 & adj.p.value < 0.05, 
                             "Depletion", "No change"))) %>% 
  ggplot(aes(x= contrast, y = variable, fill = sig)) + 
  geom_tile(color = "black") +
  scale_fill_manual(breaks = c("Enrichment", "Depletion", "No change"),
                    values = c("lightgreen", "#EF5350", "#03A9F4")) +
  theme(legend.position = "bottom")+
  theme_bw()
```


## Z-score Delta Heatmap - Do not use, it does clsuter well with kmeans

```{}
# OTU 

# Delta no transformation, no z-scoring but **summed otu** used to make the heatmap
OTU_24h <- DNA %>%  
  
  filter(!family %in% unclassified_families) %>%
  
  # filtering out rare taxa
  ## removing taxa that have a max abundance >0.5%
  group_by(OTU_tax) %>% 
  mutate(max = max(abund)) %>%
  ungroup() %>% 
  filter(max >0.005) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, OTU_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(OTU_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.001)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  
  
  # outlier detection
  group_by(OTU_tax,treatment) %>%
  mutate(mean = mean(abund),
         sd = sd(abund),
         outlier = ifelse(abund > mean + 2*sd | abund < mean - 2*sd, "outlier", "not outlier")) %>%
  ungroup() %>% 
  filter(outlier != "outlier") %>%
  select(-mean, -sd) %>% 
  
  # calculating delta values
  mutate(inflow_isolated = ifelse(treatment == "Inflow", abund, NA)) %>% 
  group_by(time, OTU_tax) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = T )) %>% 
  ungroup() %>%
  mutate(conc_norm  = abund - inflow ) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate( z_score_conc = zscore.calculation(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae")))
```


```{}
# Family 

# Delta no transformation, no z-scoring but **summed otu** used to make the heatmap
OTU_24h_filtered <- DNA %>%  
  
  # filtering out rare taxa
  ## removing taxa that have a max abundance >0.5%
  group_by(family_tax) %>% 
  mutate(max = max(abund)) %>%
  ungroup() %>% 
  filter(max >0.005) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, family_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(family_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.001)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  
  
  # outlier detection
  group_by(family_tax,treatment) %>%
  mutate(mean = mean(abund),
         sd = sd(abund),
         outlier = ifelse(abund > mean + 2*sd | abund < mean - 2*sd, "outlier", "not outlier")) %>%
  ungroup() %>% 
  filter(outlier != "outlier") %>%
  select(-mean, -sd) 


  # calculating delta values
OTU_24h <- OTU_24h_filtered %>% 
  
  filter(!family %in% unclassified_families) %>% 
  
  mutate(inflow_isolated = ifelse(treatment == "Inflow", abund, NA)) %>% 
  group_by(time, family_tax) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = T )) %>% 
  ungroup() %>%
  
   group_by(id_number, treatment, time, family_tax) %>% 
  mutate(sum_abund = sum(abund, na.rm = T),
    conc_norm  = sum_abund - inflow ) %>% 
    ungroup() %>% 
  
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate(z_score_conc = zscore.calculation(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  select(id_number, tank_number, treatment, date,time,time_fac,date_time, inflow_table,
         z_score_conc,family_tax) %>% distinct()
```

```{r}
# Genus 

# Delta no transformation, no z-scoring but **summed otu** used to make the heatmap
OTU_24h_filtered <- DNA %>%  
  
  # filtering out rare taxa
  ## removing taxa that have a max abundance >0.5%
  group_by(genus_tax) %>% 
  mutate(sum_abund= sum(abund)) %>%
  mutate(max = max(sum_abund)) %>%
  ungroup() %>% 
  filter(max >0.005) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, genus_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(genus_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.001)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() 


  # calculating delta values
OTU_24h <- OTU_24h_filtered %>% 
  
  filter(!family %in% unclassified_families) %>% 
  mutate(genus_tax = paste0(genus, ", ", family, ", ", class)) %>% 
  
  mutate(inflow_isolated = ifelse(treatment == "Inflow", abund, NA)) %>% 
  group_by(time, genus_tax) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = T )) %>% 
  ungroup() %>%
  
   group_by(id_number, genus_tax, time_fac) %>% 
  mutate(sum_abund = sum(abund, na.rm = T)) %>% 
         ungroup() %>% 
  mutate(conc_norm  = sum_abund - inflow ) %>% 
  
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  group_by(genus_tax) %>% 
  mutate(z_score_conc = zscore.calculation(conc_norm)) %>% 
  ungroup() %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  select(id_number, tank_number, treatment, date,time,time_fac,date_time, inflow_table,
         z_score_conc,genus_tax) %>% distinct()
```

```{}
OTU_24h %>% filter(genus_tax == "Vibrio X Vibrionaceae X Vibrionales X Gammaproteobacteria X Proteobacteria")
```

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_24h %>%
  group_by(id_number, genus_tax) %>%
  summarize(z_score_conc) %>% 
  ungroup()

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = genus_tax, values_from = z_score_conc)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% as.data.frame() 

```


```{}
sqrt.matrix <- apply(culled.abund.OTU, 2, FUN=sqrt)
sqrt.zscore <- as.data.frame(apply(sqrt.matrix, 2, FUN = zscore.calculation))
```

```{r}
# Set the ID numbers as the rownames (to make a matrix)
rownames(culled.abund.OTU) = culled.abund.OTU$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
delta.matrix <- culled.abund.OTU %>% select(-id_number) %>% 
  as.matrix() %>% t()


```

### Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(delta.matrix),
                           meta$id_number), 
                 "treatment"]

date_time <- meta[match(colnames(delta.matrix),
                           meta$id_number),
                  "date_time_string"]

time <- meta[match(colnames(delta.matrix),
                           meta$id_number),
                  "time_string"]

```


#### Organize the samples (ChatGPT)
```{r}

## organize the treatment in order they will be display on the heatmap
treatment2 <- factor(treatment$treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))

## organize the time in order they will be display on the heatmap
time2 <- factor(time$time_string, levels = c("12:00:00", "18:00:00", "00:00:00", "06:00:00"))

#store that info in a df
col_info <- data.frame(
  Treatment = treatment2,
  Time = time2) %>% drop_na()

# create the order in which the samples with be organize in the heatmap
sample_order <- order(col_info$Treatment,col_info$Time)

# tell the matrix with my z-scores how the samples should be arranged
mat_ordered <- delta.matrix[, sample_order]

# have vectors specifying the order of the treatment for the heatmap
treatment_ordered <- col_info$Treatment[sample_order]
time_ordered <- col_info$Time[sample_order]


```

```{r}
col_ha <- HeatmapAnnotation(
  Treatment = anno_block(
    gp = gpar(fill = c("grey", "#03A9F4", "#EF5350", "#4CAF50")),
    labels = c("Control", "Coral", "CCA", "Algae"), 
    labels_gp = gpar(col = "white", fontsize = 14, fontface = "bold")),
  
  Time = time_ordered, 
  
  col = list(Treatment = c("Control" = "grey", 
                           "Coral"   = "#03A9F4", 
                           "CCA"     = "#EF5350",
                           "Algae"   = "#4CAF50"),
             Time = c("12:00:00" = "black",
                      "18:00:00" = "green", 
                      "00:00:00" = "blue",
                      "06:00:00" = "grey")),
  
   # Controls the "Time" label on the left side of the ribbon *
   annotation_name_gp = gpar(fontsize = 16, fontface = "bold"), 
  
   # Controls the height of the Time ribbon bar itself 
   simple_anno_size = unit(6, "mm"), 
  
  annotation_legend_param = list(
    
     Time = list( 
       title_gp    = gpar(fontsize = 16, fontface = "bold"), 
       labels_gp   = gpar(fontsize = 16)
    )
  )
)

```


#### Printing heatmap

```{r, fig.height = 16, fig.width = 24}
# Version 2
#png(width=3000, height=1200, file= here("outputs", "zscore_24hdata.png"))
#png(width=1800, height=1000, file= here("outputs", "zscore_delta_genus.png"))

ht = Heatmap(mat_ordered, 
             name = "Z-scored Abundance", 
             heatmap_legend_param = list(#legend_direction = "horizontal",
                                          title_gp= gpar(fontsize = 16, 
                                                          fontface = "bold"), 
                labels_gp       = gpar(fontsize = 16), 
                grid_height     = unit(6, "mm"), 
                grid_width      = unit(6, "mm"), 
                legend_width    = unit(8, "cm")),   # controls horizontal legend length),
             show_row_names = TRUE,
        row_names_max_width = unit(25, "cm"),
        clustering_method_rows = "complete",
        clustering_distance_rows = "euclidean",
             #rect_gp = gpar(col = "white", lwd = 0.1),
        
        column_title = "Bacterial Community Shift Over 24h", 
        column_title_gp = gpar(fontsize = 25, fontface = "bold"),
        #column_title_side = "bottom",
        cluster_rows = TRUE,
        cluster_columns =  FALSE,
        row_title = "Bacterial Taxa",
        row_title_gp = gpar(fontsize = 18, fontface = "bold"),
        row_names_gp = gpar(fontsize = 16),
        row_dend_width = unit(1.5, "cm"),
        #column_names_gp = gpar(fontsize = 10),
        show_column_names = FALSE,
        row_km = 3, # 5? number of forced segregation in rows 
        #column_km = 4,#number of forced segregation in columns 
        top_annotation = col_ha,
        #bottom_annotation = col_ha,
        column_split = treatment_ordered
        )
set.seed(1999)
ht_drawn <- draw(ht, heatmap_legend_side = "right",
                 annotation_legend_side = "right", 
  merge_legends = TRUE)

#dev.off()

```





```{, fig.height= 16, fig.width=12}
library(RColorBrewer)

OTU_24h %>% 
  ggplot(aes(x = time_fac, y = OTU_tax, fill = conc_norm)) + 
  geom_tile(color = "black") +
  facet_grid(. ~ treatment, scales = "free_x", space = "free_x") +
  scale_fill_distiller(palette = "RdBu") +
  labs(x = "Time point", y = NULL, fill = "Concentration") +
  theme_bw() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold"),
    panel.spacing = unit(0.3, "lines")
  )
```



```{r}
DNA %>% 
  filter(!family %in% unclassified_families) %>%  select(genus, family, order, class) %>%  distinct()
```

