---
title: "MOBEAST FCM"
author: "Keanu Rochette"
created: '22 May 2025'
updated: '`r format(Sys.Date())`'
format: 
  html:
    toc: true
    toc-location: left
editor: visual
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```


## Load Libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)
#library(ggpmisc)
library(hms)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
#library(GGally)
library(emmeans)
library(lme4)
library(lmerTest)
library(sjPlot)
library(performance)

```

## Data visualization theme

```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=32, face = "bold"), 
        plot.subtitle = element_text(size=30),
        axis.title = element_text(size = 28, face = "bold"),
        axis.text = element_text(size = 28),
        strip.text.x = element_text(size = 28, face = "bold"),
        legend.title=element_text(size=28, face = "bold"),
        legend.text=element_text(size=28))
```

## functions
```{}
lm_eqn <- function(df){
    m <- lm(y ~ x, df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}
```


# Load Data
```{r}
FCM <- read_csv(here("data", "MOBEAST_FCM.csv"))
meta <- read_csv(here("data", "MOBEAST_metadata.csv"))

FCM<- FCM %>% clean_names()
meta <-meta %>% clean_names()
```

```{r}
treat_code <- tibble(
  treatment = c("A", "K", "C", "R", "Inflow", "Endmember", "Rain"),
  full_name = c("Algae", "Coral", "Control", "CCA", "Inflow", "Endmember", "Rain")
)

```

```{r}
table_id <- read_csv(here("data", "Laurel", "TableID.csv"))
table_id <- clean_names(table_id)

table_id <- table_id %>% rename(tank_number=tankid) %>% select(-treatment) %>% 
  mutate(tank_number = paste0("T", tank_number)) %>% 
  mutate(inflow_table = paste("Table", inflow_table))
```

# Note
**Mislabeling**:  
- There were 2 batches of MOBEAST 2024-06-02 18:00 containing a total of 3 T4 samples. 
- MOBEAST 2024-06-02 21:00 is missing, one of the 18:00 is actually the 21:00 time point.
- One tube was missing a label (marked "X"). Assume it's the bay sample (2024-05-25).

For the preliminary results, I removed the duplicates for T4, only keeping 1 random T4 sample.  
I will be averaging the values of MOBEAST 2024-06-02 18:00. 

```{r}
FCM_pre <- FCM %>% filter(str_detect(sample, pattern = "MOBEAST")) %>% 
    mutate(sample = str_remove(sample, pattern = "X_")) %>% 
  filter(sample!= "2_MOBEAST_T4_A_2024-06-02_18:00", 
         sample!= "3_MOBEAST_T4_A_2024-06-02_18:00") %>% 
  separate_wider_delim(sample, names= c("project", "tank", "treatment", "date","time"), 
                             delim = "_") 


```

```{r}
FCM_pre <- left_join(FCM_pre, treat_code, by= "treatment")

FCM_pre <- FCM_pre %>% relocate(full_name,.before = treatment) %>%
  select(-treatment) %>% 
  rename(treatment = full_name)
```

```{r}
FCM_pre <- FCM_pre %>% mutate(date_time = ymd_hms(paste0(date," ", time,":00"))) %>% 
  select(tank, treatment, date_time, date, time, 
                   include_vol, 
                   het_bact = het_bact_events_u_l,
                   pico_euk = pico_euk_events_u_l,
                   syn = syn_events_u_l,
                   pro = pro_events_u_l,
                   tot_bact = bacteria_events_u_l) %>% 
  group_by(date_time, treatment, tank) %>% 
  mutate(cyano = sum(syn, pro, het_bact) - het_bact) %>%  
  ungroup() %>% 
  rename(tank_number = tank) %>% 
  mutate(date =  ymd(date),
         time = paste0(time,":00"),
         time = as_hms(time))
```


```{r}
FCM_norm <- FCM_pre %>% 
  select(-pro, -tot_bact, -cyano, -include_vol) %>% 
  pivot_longer(cols = het_bact:syn,
               names_to = "org", 
               values_to = "conc_ul") %>% 
   group_by(date_time, org) %>% 
  mutate(inflow_avg = mean(conc_ul[treatment == "Inflow"], na.rm = T)) %>%
  #calculating inflow average to normalize the data
  ungroup() %>%
   mutate(conc_ul_norm = ifelse(treatment %in% c("Coral", "Algae", "CCA", "Control"),
                                conc_ul-inflow_avg, NA)) %>%  # Normalize for coral & algae
  #filter(treatment != "Inflow") %>% # remove the inflow  
  mutate(treatment = factor(treatment,
                          levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) %>% 
  select(-conc_ul, -inflow_avg) 
```




```{}
DOC_norm <- read_csv(here("data", "DOC", "DOC_norm.csv"))
DOC_norm <- DOC_norm %>% select(-npoc_mg_l, -tn_mg_l,
                                -tn_u_m,-inflow,
                                -sample) 
```

```{}
DOC_FCM_norm <- left_join(FCM_norm, DOC_norm, 
                          by= c("date","time","date_time", "tank_number"))

DOC_FCM_norm <- DOC_FCM_norm %>% select(-treatment.y) %>% 
  relocate(date_time, .before=tank_number) %>% 
  relocate(date, .after= date_time) %>% 
  relocate(time, .after= date) %>% 
  relocate(id_number, .before=date_time) %>% 
  relocate(long_name, .after= id_number) %>% 
  relocate(inflow_table, .after= long_name) %>% 
  rename(treatment = treatment.x)
```

```{}
DOC_FCM_norm<- left_join(DOC_FCM_norm, table_id, by= "tank_number")

DOC_FCM_norm <-DOC_FCM_norm %>% 
  relocate(inflow_table.y, .after= inflow_table.x) %>% 
  select(-inflow_table.x) %>% 
  mutate(treatment = factor(treatment, levels= c("Control", "Coral", "CCA", "Algae")))
```

```{}
## Outlier detection

DOC_FCM <- DOC_FCM_norm %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m_norm, probs = 0.25, na.rm= TRUE),
         quant3 = quantile(npoc_u_m_norm, probs = 0.75, na.rm= TRUE),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m_norm > quant3 + (IQR*1.5) | npoc_u_m_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```


## 12PM 
```{r}
FCM_12PM_norm <- FCM_norm %>% filter(time == as_hms("12:00:00"))
```

```{r}
## Outlier detection

FCM_12PM_norm <- FCM_12PM_norm %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment, org) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(conc_ul_norm, probs = 0.25, na.rm= TRUE),
         quant3 = quantile(conc_ul_norm, probs = 0.75, na.rm= TRUE),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( conc_ul_norm > quant3 + (IQR*1.5) | conc_ul_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```


### Hypothesis 1: there will be more heterotrophic bacteria in algae tanks

```{r}
hbact_12PM <- FCM_12PM_norm %>% filter(org == "het_bact")
```

#### Normality and Homogenity 

```{r}
bartlett.test(conc_ul_norm ~ treatment,data = hbact_12PM) # homogeneity
shapiro.test(hbact_12PM$conc_ul_norm) # normality
```

```{r}
power_model <- powerTransform(hbact_12PM$conc_ul_norm ~ 1, family = "yjPower")  # no predictors, just transforming y
summary(power_model)
## lambda = 0.9345 according to the above 
```


```{}
hbact_12PM <- hbact_12PM %>% mutate(conc_ul_trans = yjPower(conc_ul_norm, lambda = 0.9345)) %>% 
  relocate(conc_ul_trans, .after =conc_ul_norm )
```

```{}
bartlett.test(conc_ul_trans ~ treatment,data = hbact_12PM) # homogeneity
shapiro.test(hbact_12PM$conc_ul_norm) # normality
```

```{r}
mod <- lmer(conc_ul_norm ~ treatment +(1|tank_number), data = hbact_12PM)

summary(mod)
```

```{r}
resid_panel(mod)
anova(mod)
kruskal.test(conc_ul_norm ~ treatment, data = hbact_12PM)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```

```{r}
emm <- emmeans(mod, ~ treatment)
cld_letters <- multcomp::cld(emm, Letters = letters)
```




```{r, fig.width= 18, fig.height = 8}
ggplot(hbact_12PM, aes(x= treatment, conc_ul_norm, fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2,linewidth = 1)+
  geom_violin(linewidth = 1.5)+
  geom_point(color = "black", alpha = 0.5, size = 3)+ #aes(color = treatment)
  stat_summary(color = "yellow", alpha =0.5, size = 1.5)+
  #stat_summary(color = "orange", alpha = 0.5)+
  labs(#title = "12PM variation in normalized DOC",
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Abundance (ind/µL)")+
  #facet_wrap(~date) + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_y_continuous(limits = c(-200,380), breaks = seq(-200, 400, by = 100))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
#ggsave("hbact_12PM.png", width = 18, height = 8,path = here("outputs", "OOSC_2025"))
```


### Hypothesis 2: Synechococcus

```{r}
syn_12PM <- FCM_12PM_norm %>% filter(org == "syn")
```

```{r}

ggplot(syn_12PM, aes(x= treatment, conc_ul_norm, fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2)+
  #stat_summary()+
  geom_violin()+
  geom_point(aes(color = treatment), alpha = 0.5)+
#  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "12PM variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  #facet_wrap(~date) + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))

```

### Hypothesis 3: Picoeukaryotes

```{r}
picoeuk_12PM <- FCM_12PM_norm %>% filter(org == "pico_euk")
```

```{r}

ggplot(picoeuk_12PM, aes(x= treatment, conc_ul_norm, fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2)+
  #stat_summary()+
  geom_violin()+
  geom_point(aes(color = treatment), alpha = 0.5)+
#  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "12PM variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  #facet_wrap(~date) + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))

```

## 24h 
WARNING !!! 18:00 IS DUPLICATED SOMEHOW! 

```{r}
FCM_24h_norm <- FCM_norm %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) 

```

```{}
FCM_18PM_norm <- FCM_24h_norm %>% filter(time == as_hms("18:00:00")) %>% 
  group_by(tank_number, org) %>% 
  reframe(tank_number, treatment,date_time, date, time, org, conc_ul_norm = mean(conc_ul_norm)) %>% 
  ungroup() %>% distinct()
```

```{}
FCM_24h_norm <- FCM_24h_norm %>% filter(time != as_hms("18:00:00"))
```

```{}
FCM_24h_norm <- rbind(FCM_24h_norm, FCM_18PM_norm)
```

```{}
FCM_24h_norm %>% mutate(time_fac = factor(time_fac, 
                                          levels=c("12:00:00", "15:00:00","18:00:00",
                                                   "21:00:00","00:00:00","03:00:00",
                                                   "06:00:00")))
```



```{r}
## Outlier detection

FCM_24h_norm <- FCM_24h_norm %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment, org) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(conc_ul_norm, probs = 0.25, na.rm= TRUE),
         quant3 = quantile(conc_ul_norm, probs = 0.75, na.rm= TRUE),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( conc_ul_norm > quant3 + (IQR*1.5) | conc_ul_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

### Hypothesis 1: there will be more heterotrophic bacteria in algae tanks on a diel cycle 

```{r}
hbact_24h <- FCM_24h_norm %>% filter(org == "het_bact") %>% 
  mutate(time_fac = factor(as.character(time),levels = c("12:00:00","15:00:00",
                                                         "18:00:00","21:00:00",
                                            "00:00:00","03:00:00","06:00:00",
                                            "09:00:00"))) %>% 
  mutate(date_fac = factor(date)) %>% 
  filter(time == as_hms("12:00:00") | 
         time == as_hms("18:00:00")|
         time == as_hms("06:00:00"))

```

#### Temporary correction of 18:00 mess up

```{r}
hbact_24h_temp <- hbact_24h %>% select(-quant1, -quant3, -IQR, - outlier)
```

```{r}
hbact_18h <- hbact_24h_temp %>% filter(time == as_hms("18:00:00")) %>% 
  group_by(tank_number, treatment, date_time, date, time, org) %>% 
  reframe(conc_ul_norm = mean(conc_ul_norm)) %>% 
  mutate(time_fac = factor(as.character(time),levels = c("12:00:00","15:00:00",
                                                         "18:00:00","21:00:00",
                                            "00:00:00","03:00:00","06:00:00",
                                            "09:00:00"))) %>% 
  mutate(date_fac = factor(date))


```

```{r}
hbact_24h_no_18 <- hbact_24h_temp %>% filter(time != as_hms("18:00:00"))
```


```{r}
hbact_24h_temp2 <- rbind(hbact_24h_no_18, hbact_18h)
```



#### Normality and Homogenity 

```{r}
bartlett.test(conc_ul_norm ~ treatment,data = hbact_24h_temp2) # homogeneity
shapiro.test(hbact_12PM$conc_ul_norm) # normality
```

```{r}
power_model <- powerTransform(hbact_24h_temp2$conc_ul_norm ~ 1, family = "yjPower")  # no predictors, just transforming y
summary(power_model)
## lambda = 0.9753 according to the above 
```

```{r}
hbact_24h_temp2 <- hbact_24h_temp2 %>% mutate(conc_ul_trans = yjPower(conc_ul_norm, lambda = 0.9753)) %>% 
  relocate(conc_ul_trans, .after =conc_ul_norm )
```

```{}
bartlett.test(conc_ul_trans ~ treatment,data = hbact_24h_temp2) # homogeneity
shapiro.test(hbact_24h$conc_ul_trans) # normality
```

```{r}
mod <- lmer(conc_ul_norm ~ treatment*time_fac +(1|tank_number), data = hbact_24h_temp2)

#summary(mod)
```

```{r}
resid_panel(mod)
anova(mod)
kruskal.test(conc_ul_norm ~ treatment, data = hbact_24h_temp2)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```


```{r, fig.width = 18, fig.height = 8}

ggplot(hbact_24h_temp2, aes(x= treatment, conc_ul_norm, fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2)+
  geom_violin(linewidth = 1)+
  geom_point(color = "black", alpha = 0.5, size = 4)+
  stat_summary(color = "yellow", alpha =0.5, size = 1.5)+
#  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(
       x= "Treatment",
       y = "Abundance (ind/µL)")+
  facet_wrap(~time_fac) + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        strip.background = element_rect(fill = "white"))

ggsave("hbact_24_TEMPORARY.png", width = 18, height = 8, path = here("outputs", "OOSC_2025"))
```


#### Reference section 
```{r}
power_model <- powerTransform(hbact_24h$conc_ul_norm ~ 1, family = "yjPower")  # no predictors, just transforming y
summary(power_model)
## lambda = 0.9779 according to the above 
```


```{r}
hbact_24h <- hbact_24h %>% mutate(conc_ul_trans = yjPower(conc_ul_norm, lambda = 0.9779)) %>% 
  relocate(conc_ul_trans, .after =conc_ul_norm )
```

```{r}
bartlett.test(conc_ul_trans ~ treatment,data = hbact_24h) # homogeneity
shapiro.test(hbact_24h$conc_ul_trans) # normality
```

```{r}
mod <- lmer(conc_ul_norm ~ treatment*time_fac +(1|tank_number), data = hbact_24h)

#summary(mod)
```

```{r}
resid_panel(mod)
anova(mod)
kruskal.test(conc_ul_norm ~ treatment, data = hbact_24h)
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```



```{r}
hbact_24h %>% #filter(time == as_hms("18:00:00")) %>%
  ggplot(aes(x= tank_number, conc_ul_norm, fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2)+
  #stat_summary()+
  #geom_violin()+
  geom_point(color = "black", alpha = 0.5)+ #aes(color = treatment)
#  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "12PM variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  #facet_wrap(~time_fac) + 
  scale_color_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
```


```{r, fig.width = 18, fig.height = 11}

ggplot(hbact_24h, aes(x= treatment, conc_ul_norm, fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2)+
  geom_violin(linewidth = 0.8)+
  geom_point(color = "black", alpha = 0.5)+ #aes(color = treatment)
  stat_summary(color = "yellow", alpha = 0.5)+
#  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(
       x= "Treatment",
       y = "Abundance (ind/µL)")+
  facet_wrap(~time_fac) + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none")
#ggsave("hbact_24.png", width = 18, height = 8, path = here("outputs", "OOSC_2025"))
```