---
author: 'Keanu Rochette '
created: '18 April 2025'
updated: '`r format(Sys.Date())`'
title: MOBEAST DOC V2
output: 
  html_document:
     toc: true
     toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
#library(GGally)
library(emmeans)
library(performance)
library(multcompView)



```

## Load Data 
```{r}
DOC <- read_csv(here("data", "DOC", "DOC_clean.csv"))
table_id <- read_csv(here("data", "Laurel", "TableID.csv"))
```

## Data visualization theme
```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=32, face = "bold"), 
        plot.subtitle = element_text(size=30),
        axis.title = element_text(size = 28, face = "bold"),
        axis.text = element_text(size = 28),
        strip.text.x = element_text(size = 28, face = "bold"),
        legend.title=element_text(size=28, face = "bold"),
        legend.text=element_text(size=28))
```

## Data Clean Up

```{r}
table_id <- clean_names(table_id)

table_id <- table_id %>% rename(tank_number=tankid) %>% select(-treatment) %>% 
  mutate(tank_number = paste0("T", tank_number)) %>% 
  mutate(inflow_table = paste("Table", inflow_table))
DOC <- left_join(DOC, table_id, by= "tank_number")
```


```{r}
DOC <- DOC %>% mutate(inflow_table = ifelse(is.na(inflow_table), tank_number, inflow_table)) %>% 
   mutate(inflow_table = ifelse(inflow_table == "F1", "Table 1", 
                                ifelse(inflow_table == "F2", "Table 2", inflow_table))) %>% 
  mutate(inflow_table = factor(inflow_table),
         tank_number = factor(tank_number)) %>% 
  relocate(inflow_table, .after = "treatment")
```


```{r}

DOC_norm <- DOC %>% filter(date > as.Date("2024-05-31")) %>% # Removes endmembers
  filter(treatment != "Rain") %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table) %>%
  mutate(inflow= npoc_u_m[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # get the other table's value for that date
  group_by(date_time) %>% 
  mutate(inflow = ifelse(is.na(inflow),
      inflow[!is.na(inflow)][1],  
      inflow)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(npoc_u_m_norm = npoc_u_m - inflow)  %>%  
  relocate(inflow, .after = "npoc_u_m_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
  

#write_csv(DOC_norm, here("data", "DOC", "DOC_norm.csv"))
```

<!--
## Hypothesis 0: endmembers and inflow comparison

```{r}
DOC %>% filter(treatment == "Endmember" | treatment == "Rain"| treatment == "Inflow") %>% 
  dplyr::filter(time != lubridate::hms("00:00:00")) %>% 
  mutate(tank_number = as.character(tank_number)) %>% 
  mutate(tank_number = ifelse(tank_number == "Lagoon Tiahura", "Lagoon", tank_number),
         tank_number = factor(tank_number)) %>% 
  group_by(tank_number) %>% 
  summarize(avg = mean(npoc_u_m, na.rm =T),
            min = min(npoc_u_m, na.rm =T),
            max = max(npoc_u_m, na.rm =T))
```



```{r}
DOC_verif <- DOC %>% filter(treatment == "Endmember" | treatment == "Rain"| treatment == "Inflow") %>% 
  filter(time != hms("00:00:00")) %>% 
  mutate(tank_number = as.character(tank_number)) %>% 
  mutate(tank_number = ifelse(tank_number == "Lagoon Tiahura", "Lagoon", tank_number),
         tank_number = factor(tank_number, levels = c("Lagoon", "Ocean", "Bay", "F1", "F2")))
```

```{r}
mod <- lmer(npoc_u_m ~ tank_number + (1|date)+ (1|time)+  (1|treatment), data = DOC_verif)

summary(mod)
```


```{r}
anova(mod, type = "2")
```



```{, fig_height = 10, fig.width = 10}
plot_model(mod, type = "re")[[1]] +
  plot_model(mod, type = "re")[[2]] +
  plot_model(mod, type = "eff", terms = c('treatment')) +
  plot_model(mod, type = "diag")[[1]]
```

```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ tank_number)
```

```{r}
ggplot(DOC_verif, aes(x= tank_number, npoc_u_m, fill = treatment)) +
  geom_boxplot()+
  geom_jitter(aes(color = tank_number), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "All samples variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  facet_wrap(~date) + 
  #scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       #values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
```







-->


## Hypothesis  1: Different communities will release difference amounts of DOC (12PM)
<!--
### Testing removing outliers first, then normalizing the data 
```{r}
DOC_12_pre_norm <- DOC %>% filter(date > as.Date("2024-05-31")) %>% # Removes end members
  filter(treatment != "Rain") %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
  ## Only keep the 12PM data points
  filter(time == hms("12:00:00")) %>% 
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m, probs = 0.25),
         quant3 = quantile(npoc_u_m, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m > quant3 + (IQR*1.5) | npoc_u_m < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  
  # Normalizing DOC data in tanks based on the inflow measurements
  group_by(date_time, inflow_table) %>%
  mutate(inflow= npoc_u_m[treatment == "Inflow"]) %>%  # Isolate inflow for each date_time and table
  ungroup() %>%
  group_by(date_time) %>% 
  mutate(inflow = ifelse(is.na(inflow),
      inflow[!is.na(inflow)][1],  # get the other table's value for that date
      inflow)) %>%
  ungroup() %>% 
  mutate(npoc_u_m_norm = npoc_u_m - inflow)  %>% # Normalize for treatments 
  relocate(inflow, .after = "npoc_u_m_norm") %>% 
  relocate(date_time, .before = "date") %>%
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) %>% 
  
  # removing the inflow data for modeling 
  filter(treatment != "Inflow")
  
```

```{r}
mod <- lmer(npoc_u_m_norm ~ treatment + (1|date) +(1|tank_number) , data = DOC_12_pre_norm)
mod <- lmer(npoc_u_m_norm ~ treatment + (1|date) , data = DOC_12_pre_norm)
summary(mod)
```

```{r}
Anova(mod, type= "2")
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```




-->

### No transformation 
#### Normality and Homogenity 
Data is not normal and variances are not homogenous
```{r}
bartlett.test(npoc_u_m_norm ~ treatment,data = DOC_12PM) # homogeneity
shapiro.test(DOC_12PM$npoc_u_m_norm) # normality
```


#### Modeling with data pre-normalized 

```{r}
DOC_norm_12PM <- DOC_norm %>% 
  filter(time == hms("12:00:00")) %>% 
  filter(treatment != "Inflow") 
```


```{r}
## Outlier detection

DOC_12PM <- DOC_norm_12PM %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m_norm, probs = 0.25),
         quant3 = quantile(npoc_u_m_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m_norm > quant3 + (IQR*1.5) | npoc_u_m_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```



```{r}
mod <- lmer(npoc_u_m_norm ~ treatment +(1|tank_number), data = DOC_12PM)
#mod <- lm(npoc_u_m_norm ~ treatment, data = DOC_std_12PM)

summary(mod)
```


```{r}
anova(mod)
kruskal.test(npoc_u_m_norm ~ treatment, data = DOC_12PM )
```


```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, ~ treatment)
```

```{r}
emm <- emmeans(mod, ~ treatment)
cld_letters <- cld(emm, Letters = letters)
```


### Data visualization 
```{r, fig.width= 16, fig.height = 8}
ggplot(DOC_12PM, aes(x= treatment, npoc_u_m_norm, fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2,linewidth = 0.8)+
  geom_violin(linewidth = 0.8)+
  geom_point(color = "black", alpha = 0.5)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  stat_summary(color = "yellow", alpha =0.5)+
  labs(#title = "DOC Enrichment",
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "DOC Production (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  ylim(-20,120)+
  nunu_theme +
  theme(legend.position = "none") 

#ggsave("DOC_12PM.png", width = 18,height = 8,path = here("outputs", "OOSC_2025"))
```


### With data transformation 
#### Transforming the data

```{r}
power_model <- powerTransform(DOC_12PM$npoc_u_m_norm ~ 1, family = "yjPower")  # no predictors, just transforming y
summary(power_model)
## lambda = 0.643 according to the above 
```

#### Normality and Homogenity 
Data is not normal and variances are not homogenous
```{r}
bartlett.test(yjPower(npoc_u_m_norm, lambda = 0.6425) ~ treatment,data = DOC_12PM) # homogeneity
shapiro.test(yjPower(DOC_12PM$npoc_u_m_norm, lambda = 0.6425)) # normality
```

#### Model Setup
```{r}
mod <- lmer(yjPower(npoc_u_m_norm, lambda = 0.6425) ~ treatment +(1|tank_number), data = DOC_12PM)

summary(mod)
```


```{r}
anova(mod)
kruskal.test(yjPower(npoc_u_m_norm, lambda = 0.6425) ~ treatment, data = DOC_12PM )
```

```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, ~ treatment)
```
```{r}
emm <- emmeans(mod, ~ treatment)
cld_letters <- multcomp::cld(emm, Letters = letters)
```

### data visualization 
```{r, fig.width= 18, fig.height = 8}

ggplot(DOC_12PM, aes(x= treatment, yjPower(npoc_u_m_norm, lambda = 0.6425), fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2,linewidth = 0.8)+
  geom_violin(linewidth = 0.8)+
  geom_point(color = "black", alpha = 0.5, size = 3)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  stat_summary(color = "yellow", alpha =0.5, size = 1.5)+
  labs(#title = "DOC Enrichment",
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "DOC Production (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_y_continuous(limits = c(-19,33), breaks = seq(-20, 33, by = 10))+
  nunu_theme +
  theme(legend.position = "none") 

#ggsave("DOC_12PM.png", width = 18,height = 8,path = here("outputs", "OOSC_2025"))
```


## Hypothesis 2: There are variation in exudation on a diel cycle 24h 

<!--
### Testing removing outliers first, then normalizing the data 
```{r}
DOC_24_pre_norm <- DOC %>% filter(date > as.Date("2024-05-31")) %>% # Removes end members
  filter(treatment != "Rain") %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
  
  # Only keep the 24h data points
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>% 
  filter(time != hms("00:00:00")) %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m, probs = 0.25, na.rm = TRUE),
         quant3 = quantile(npoc_u_m, probs = 0.75, na.rm = TRUE),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m > quant3 + (IQR*1.5) | npoc_u_m < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  
  # Normalizing DOC data in tanks based on the inflow measurements
  group_by(date, time, inflow_table) %>%
  mutate(inflow = npoc_u_m[treatment == "Inflow"]) %>%  # Isolate inflow for each date_time and table
  ungroup() %>%
  
  group_by(date_time) %>% 
  mutate(inflow = ifelse(is.na(inflow),
      inflow[!is.na(inflow)][1],  # get the other table's value for that date
      inflow)) %>%
  ungroup() %>% 
    
    ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  
  # Normalize for treatment
  mutate(npoc_u_m_norm = npoc_u_m - inflow)  %>% 
  relocate(inflow, .after = "npoc_u_m_norm") %>% 
  relocate(date_time, .before = "date") %>%
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) %>% 
  
  # removing the inflow data for modeling 
  filter(treatment != "Inflow") %>% 
  mutate(time = factor(time))
  
```


```{r}
mod <- lmer(npoc_u_m_norm ~ treatment*time + (1|date), data = DOC_24_pre_norm)

summary(mod)
```

```{r}
Anova(mod, type = "2")
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```



-->

### Detecting outliers after normalization 
```{r}
DOC_norm_24h <- DOC_norm %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>% 
  filter(treatment != "Inflow") %>% 
  mutate(time_fac = factor(as.character(time),levels = c("12:00:00","18:00:00","00:00:00",
                                            "06:00:00"))) %>% 
  mutate(date_fac = factor(date)) %>% 
  filter(time != hms("00:00:00")) %>%
  filter(!(tank_number %in% c("T8", "T11") & time == hms("18:00:00")))
```


```{r}
## Outlier detection

DOC_24h <- DOC_norm_24h %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m_norm, probs = 0.25, na.rm = TRUE),
         quant3 = quantile(npoc_u_m_norm, probs = 0.75, na.rm = TRUE),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m_norm > quant3 + (IQR*1.5) | npoc_u_m_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") #%>% 
  
  #filter(!(tank_number %in% c("T8", "T11") & time == hms("18:00:00")))
 
```


### No transformation 

#### Normality and Homogenity 
Data is not normal and variances are not homogenous
```{r}
bartlett.test(npoc_u_m_norm ~ treatment,data = DOC_24h) # homogeneity
shapiro.test(DOC_12PM$npoc_u_m_norm) # normality
```

#### Model Setup
```{r}
mod <- lmer(npoc_u_m_norm ~ treatment*time_fac + (1|tank_number), data = DOC_24h)

summary(mod)
anova(mod)
```


```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, pairwise ~ treatment)
```




#### Justfications
- Removed 18:00 T8 and T11 because they are very high and they are control tanks. Suspicion of contamination because there is no way that a control looks so different from the inflow. 
- Remove 00:00 because there is a inflow values are very high and it makes controls and corals have negative values while the control should be around the same value as the inflows. CCA and algae have large positive values indicating a large amount of enrichment, big enough to surpass the inflow values that are already high. Need further attention before being able to rule anything here. 



```{r, fig.width=10}
ggplot(DOC_24h, aes(x= treatment, npoc_u_m_norm, fill = treatment)) +
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "24h variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  facet_wrap(~time_fac, scales= "free") + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
```




### With data transformation 
#### Transforming the data

```{r}
power_model <- powerTransform(DOC_24h$npoc_u_m_norm ~ 1, family = "yjPower")  # no predictors, just transforming y
summary(power_model)
## lambda = 1.2012 according to the above with 00:00:00 time points
## lambda = 0.6717 according to the above without 00:00:00 time points
```

#### Normality and Homogenity 
Data is not normal and variances are not homogenous
```{r}
bartlett.test(yjPower(npoc_u_m_norm, lambda = 0.6717) ~ treatment,data = DOC_24h) # homogeneity
## bartlett test show that the variance are not homogenous so we cant do an anova here. 

shapiro.test(yjPower(DOC_24h$npoc_u_m_norm, lambda = 0.6717)) # normality
#shapiro test show that the data is normally distributed now. 

```

```{r}
mod <- lmer(yjPower(npoc_u_m_norm, lambda = 0.6717) ~ treatment*time_fac + (1|tank_number), data = DOC_24h)

summary(mod)
```


```{r}
anova(mod)
kruskal.test(yjPower(npoc_u_m_norm, lambda = 0.6717) ~ treatment, data = DOC_24h )
```


```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, ~ treatment)
```




#### Data visualization 

```{r, fig.width= 18, fig.height = 8}

ggplot(DOC_12PM, aes(x= treatment, yjPower(npoc_u_m_norm, lambda = 0.6425), fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2,linewidth = 0.8)+
  geom_violin(linewidth = 1.5)+
  geom_point(color = "black", alpha = 0.5, size = 4)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  stat_summary(color = "yellow", alpha =0.5, size = 1.5)+
  labs(#title = "DOC Enrichment",
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "DOC Production (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_y_continuous(limits = c(-19,33), breaks = seq(-20, 33, by = 10))+
  nunu_theme +
  theme(legend.position = "none") 

#ggsave("DOC_12PM.png", width = 18,height = 8,path = here("outputs", "OOSC_2025"))
```

```{r, fig.width= 18, fig.height = 8}
ggplot(DOC_24h, aes(x= treatment, yjPower(npoc_u_m_norm, lambda = 0.6717), fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2,linewidth = 1)+
  geom_violin(linewidth = 0.8)+
  geom_point(color = "black", alpha = 0.5, size = 3)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  stat_summary(color = "yellow", alpha =0.5, size = 1.5)+
  labs(#title = "24h variation in normalized DOC",
       #subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "DOC Production (µmol/L)")+
  facet_wrap(~time_fac) + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        strip.background = element_rect(fill = "white"))
ggsave("DOC_24h_freescale.png", width = 18, height = 8,path = here("outputs", "OOSC_2025"))
```


## Hypothesis 3: All samples

<!--
### Testing removing outliers first, then normalizing the data 

```{r}
DOC_all_pre_norm <- DOC %>% filter(date > as.Date("2024-05-31")) %>% # Removes end members
  filter(treatment != "Rain") %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m, probs = 0.25, na.rm = TRUE),
         quant3 = quantile(npoc_u_m, probs = 0.75, na.rm = TRUE),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m > quant3 + (IQR*1.5) | npoc_u_m < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  
  # Normalizing DOC data in tanks based on the inflow measurements
  group_by(date, time, inflow_table) %>%
  mutate(inflow = npoc_u_m[treatment == "Inflow"]) %>%  # Isolate inflow for each date_time and table
  ungroup() %>%
  
  group_by(date_time) %>% 
  mutate(inflow = ifelse(is.na(inflow),
      inflow[!is.na(inflow)][1],  # get the other table's value for that date
      inflow)) %>%
  ungroup() %>% 
    
    ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  
  # Normalize for treatment
  mutate(npoc_u_m_norm = npoc_u_m - inflow)  %>% 
  relocate(inflow, .after = "npoc_u_m_norm") %>% 
  relocate(date_time, .before = "date") %>%
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) %>% 
  
  # removing the inflow data for modeling 
  filter(treatment != "Inflow")  
  
  
```

```{r}
mod <- lmer(npoc_u_m_norm ~ treatment + (1|time) , data = DOC_all_pre_norm)

summary(mod)
```

```{r}
anova(mod, type = "2")
```

-->

### analysis post norm
```{r}
DOC_norm_all <- DOC_norm %>% 
  filter(time != hms("00:00:00")) %>% 
  filter(treatment != "Inflow") 
```

```{r}
## Outlier detection

DOC_all <- DOC_norm_all %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m_norm, probs = 0.25, na.rm = TRUE),
         quant3 = quantile(npoc_u_m_norm, probs = 0.75, na.rm = TRUE),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m_norm > quant3 + (IQR*1.5) | npoc_u_m_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))
```

### XXXX

```{r}
mod <- lmer(npoc_u_m_norm ~ treatment +(1|tank_number), data = DOC_all)

summary(mod)
```


```{r}
anova(mod, type= "2")
```



```{r, fig_height = 10, fig.width = 10}
plot_model(mod, type = "re")[[1]] +
  plot_model(mod, type = "re")[[2]] +
  plot_model(mod, type = "eff", terms = c('treatment')) +
  plot_model(mod, type = "diag")[[1]]
```

```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
plot(emmeans(mod, ~ treatment), comparisons = TRUE)
```

#### Interpretation

- Looking at all the samples together, lmer seems to indicate that Algae tanks are significantly different from the control
- Emmeans show that the difference between pairs is not significant. But the algae tanks have much low p-values than the other treatments




### Data visualization 

```{r}
ggplot(DOC_all, aes(x= treatment, npoc_u_m_norm, fill = treatment)) +
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "All samples variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  facet_wrap(~date_time, scales= "free") + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))+
  ylim(-15,80)
```


### With data transformation 
#### Transforming the data

```{r}
power_model <- powerTransform(DOC_all$npoc_u_m_norm ~ 1, family = "yjPower")  # no predictors, just transforming y
summary(power_model)
## lambda = 1.0694 according to the above with 00:00:00 time points
## lambda = 0.5511 according to the above with 00:00:00 time points
```

#### Normality and Homogenity 
Data is not normal and variances are not homogeneous
```{r}
bartlett.test(yjPower(npoc_u_m_norm, lambda = 0.5511) ~ treatment,data = DOC_all) # homogeneity
## bartlett test show that the variance are not homogenous so we cant do an anova here. 

shapiro.test(yjPower(DOC_all$npoc_u_m_norm, lambda = 0.5511)) # normality
#shapiro test show that the data is normally distributed now. 

```

```{r}
mod <- lmer(yjPower(npoc_u_m_norm, lambda = 0.5511) ~ treatment + (1|tank_number), data = DOC_all)

summary(mod)
```


```{r}
anova(mod)
kruskal.test(yjPower(npoc_u_m_norm, lambda = 0.5511) ~ treatment, data = DOC_all)
```


```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, ~ treatment)
```

#### Data visualization 
```{r, fig.width= 18, fig.height = 8}
ggplot(DOC_all, aes(x= treatment, yjPower(npoc_u_m_norm, lambda = 0.5511), fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2,linewidth = 0.8)+
  geom_violin(linewidth = 0.8)+
  geom_point(color = "black", alpha = 0.5)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  stat_summary(color = "yellow", alpha =0.5)+
  labs(#title = "24h variation in normalized DOC",
       #subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "DOC Production (µmol/L)")+
  facet_wrap(~date_time, scales= "free") + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
ggsave("DOC_24h.png", width = 18, height = 8,path = here("outputs", "OOSC_2025"))
```




