---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: MOBEAST fDOM Preliminary Results
output: 
  html_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE , 
                      warning = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(GGally)
library(emmeans)
```

## Load function
```{r}
source(here("functions", "function_ANOVA_nunu.R"))
```

```{r}
heatmap <- function(data , x, y, color) {
  #creating a function using the penguin dataset, 
  # x and y variables to plot and a color component 
  # geom_func allows to swap the style of the plot
  

  ggplot(data , aes(x= {{x}}, y= {{y}}, fill = {{color}})) +
    geom_tile()+
    theme_bw() +
    theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        panel.background = element_rect(fill = "azure1"))
}

```


## Load Data 
```{r}
fDOM <- read_csv(here("data", "fDOM", "fDOM_sorted.csv"))
```

## Data visualization theme
```{r}

nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        panel.background = element_rect(fill = "azure1"))
```

## Hierarchical hypothesis testing 

### Design function to automate the process 
```{r}
fDOM_hyp_test <- function(data, fDOM_var) {
  
  # look at the baseline of endmembers: lagoon, bay and ocean channel and rain
  ## table summary of average and stand dev with endmembers 
  ## the goal is to see is the endmember values are consistent
  
  endmem_summary <<- data %>% group_by(sample_type) %>% 
  summarise(avg= mean({{fDOM_var}}, rm.na = T), 
            std_dev = sd({{fDOM_var}}))
  
  ## plot the endmember values on a heat map for visualization 
  heat_map <<- ggplot(data , aes(x= sample_type, y = factor(date_time), 
                    fill = {{fDOM_var}})) +
    geom_tile()+
    scale_fill_distiller(palette = "RdYlBu")+
    nunu_theme
  
  
  # Is the inflow consistent ? 
  ## Remove the end members of the data set
  fDOM_inflow <<- data %>% select(tank_number, treatment, sample_type, date,
                                    time, date_time, {{fDOM_var}}) %>% 
    filter (treatment != "Endmember", treatment!= "Rain") 
  
  ## Boxplot to show the distribution of variables without endmembers
  boxgraph <<- fDOM_inflow %>% ggplot(aes(x= treatment, y= {{fDOM_var}}))+
      geom_boxplot(aes(fill = treatment))+
      nunu_theme
  
    ## calculate mean and std dev without the endmembers  
    inflow_summary <<- fDOM_inflow%>% 
      group_by(treatment) %>% 
      summarise(avg = mean({{fDOM_var}}, na.rm = T),
            std_dev = sd({{fDOM_var}}, na.rm = T))
    
  ## graph distribution of the std dev 
  point_graph <<- fDOM_inflow%>% 
    group_by(treatment) %>% 
    summarise(avg = mean({{fDOM_var}}, na.rm = T),
            std_dev = sd({{fDOM_var}}, na.rm = T)) %>% 
    ggplot(aes(x= treatment, y= std_dev))+
    geom_point(aes(color = treatment, size = 3))+
    scale_y_continuous(limits = c(0, NA))+
    nunu_theme
  
  # remove the inflow data point 
  fDOM_ctrl_trmt <- fDOM_inflow %>% 
    filter (treatment != "Inflow") %>% 
    mutate(treatment = factor(treatment, 
                            levels= c("Control", "Coral", "Rubble", "Algae")))
  fDOM_ctrl_trmt <<- fDOM_ctrl_trmt
      # treatment must be ordered factors for future linear model
  
  print("Outputs: endmem_summary, heat_map, boxgraph, fDOM_inflow, inflow_summary,
        point_graph, fDOM_ctrl_trmt")
}
```



### CobleA
```{r}
fDOM_hyp_test(data = fDOM, fDOM_var = coble_a)
```

#### Baseline assessment
Endmembers, i.e. Lagoon, Ocean channel, Bay and Rain, were measured to assess the "baseline" values of the environmental parameters measured, here cobleA which is a time of fDOM.  

We calculated the the average and standard deviation.  

From the heat maps, the values of cobleA in the endmembers seem close to one another and the standard deviations are low. We can consider that their values are consistent for now. 
```{r}
fDOM %>% group_by(sample_type) %>% 
  summarise(avg= mean(coble_a, rm.na = T), 
            std_dev = sd(coble_a))

```

```{r}
heatmap(data = fDOM, x= sample_type, y = factor(date_time), color = coble_a) +
  scale_fill_distiller(palette = "RdYlBu")


```


#### Variations in inflow
The inflow was measured as a baseline for water that entered the tank.  

It looks like the variation in flow are minimal and are there consistent.  
The box plot plot shows minimal spread of the data. Plotting the standard deviations shows a low standard deviation.
```{r}
fDOM %>% select(tank_number, treatment, sample_type, date, time, date_time, coble_a) %>% filter (treatment != "Endmember", treatment!= "Rain") %>% 
  group_by(treatment) %>% 
  ggplot(aes(x= treatment, y= coble_a))+
  geom_boxplot(aes(fill = treatment))


```

```{r}
fDOM %>% select(tank_number, treatment, sample_type, date, time, date_time, coble_a) %>% filter (treatment != "Endmember", treatment!= "Rain") %>% 
  group_by(treatment) %>% 
  summarise(avg = mean(coble_a, na.rm = T),
            std_dev = sd(coble_a, na.rm = T)) %>% 
  ggplot(aes(x= treatment, y= std_dev))+
  geom_point(aes(color = treatment, size = 3))+
  scale_y_continuous(limits = c(0, NA))


```

#### Comparing the treatments to the control tanks

We saw that the inflow was fairly consistent, we can now test the treatments vs. control. We are applying a Dunnett test to calculate the contrast between treatments and control. 
```{r}
cobA <- fDOM %>% select(tank_number, treatment, sample_type, date, time, date_time, coble_a) %>% filter (treatment != "Endmember", treatment!= "Rain", treatment != "Inflow") %>% 
  mutate(treatment = factor(treatment, levels= c("Control", "Coral", "Rubble", "Algae")))
```

Setting up the linear model to apply the Dunnett test via the emmeans() function. p-values of the contrast show not significant differences between the control and the treatments. 
```{r}
cobA_lm <-  lm(coble_a ~ treatment, data = cobA)

emmeans(cobA_lm, specs=(trt.vs.ctrl ~ treatment))
```

### Conclusion cobleA
There is no difference between control and treatments

```{r}
cobA %>% filter(treatment != "Control") %>% 
   ggplot(aes(x= treatment, y= coble_a))+
  geom_boxplot(aes(fill = treatment)) +
  geom_hline(yintercept = 0.009160615, color = "red", linewidth = 2, alpha = 0.5)+
  labs(caption = "red line is the average value of the inflow")+
  nunu_theme
```


### CobleM

```{r}
fDOM_hyp_test(data = fDOM, fDOM_var = coble_m)
```

```{r}
heat_map 
endmem_summary
boxgraph
inflow_summary
point_graph
```


```{r}

# run contrast between control and treatment
  ##  set up linear model to analyze contrasts
fDOM_lm <- lm(coble_m ~ treatment, data = fDOM_ctrl_trmt)
  
  
  ## run emmeans to see the difference btwn control and treatment
emmeans(fDOM_lm, specs=(trt.vs.ctrl ~ treatment))
```

```{r}
fDOM_ctrl_trmt %>% filter(treatment != "Control") %>% 
   ggplot(aes(x= treatment, y= coble_m))+
  geom_boxplot(aes(fill = treatment)) +
  geom_hline(yintercept = 0.01105936, color = "red", linewidth = 2, alpha = 0.5)+
  labs(caption = "red line is the average value of the inflow")+
  nunu_theme
```


### CobleB
```{r}
fDOM_hyp_test(data = fDOM, fDOM_var = coble_b)
```

```{r}
endmem_summary 
```

```{r}
heat_map
```

```{r}
boxgraph
```

```{r}
inflow_summary
```

```{r}
point_graph
```


```{r}
# run contrast between control and treatment
  ##  set up linear model to analyze contrasts
fDOM_lm <- lm(coble_b ~ treatment, data = fDOM_ctrl_trmt)
  
  
  ## run emmeans to see the difference btwn control and treatment
emmeans(fDOM_lm, specs=(trt.vs.ctrl ~ treatment))
```

```{r}
fDOM_ctrl_trmt %>% filter(treatment != "Control") %>% 
   ggplot(aes(x= treatment, y= coble_b))+
  geom_boxplot(aes(fill = treatment)) +
  geom_hline(yintercept = 0.05406242, color = "red", linewidth = 2, alpha = 0.5)+
  labs(caption = "red line is the average value of the inflow")+
  nunu_theme
```

### CobleC
```{r}
fDOM_hyp_test(data = fDOM, fDOM_var = coble_c)
```

```{r}
endmem_summary
```

```{r}
heat_map
```

```{r}
boxgraph
```

```{r}
inflow_summary
```

```{r}
point_graph
```

```{r}
# run contrast between control and treatment
  ##  set up linear model to analyze contrasts
fDOM_lm <- lm(coble_c~ treatment, data = fDOM_ctrl_trmt)
  
  
  ## run emmeans to see the difference btwn control and treatment
emmeans(fDOM_lm, specs=(trt.vs.ctrl ~ treatment))
```


```{r}
fDOM_ctrl_trmt %>% filter(treatment != "Control") %>% 
   ggplot(aes(x= treatment, y= coble_c))+
  geom_boxplot(aes(fill = treatment)) +
  geom_hline(yintercept = 0.005663924	, color = "red", linewidth = 2, alpha = 0.5)+
  labs(caption = "red line is the average value of the inflow")+
  nunu_theme
```



### CobleT
```{r}
fDOM_hyp_test(data = fDOM, fDOM_var = coble_t)
```

```{r}
endmem_summary 
```

```{r}
heat_map
```

```{r}
boxgraph
```

```{r}
inflow_summary
```

```{r}
point_graph
```


```{r}
# run contrast between control and treatment
  ##  set up linear model to analyze contrasts
fDOM_lm <- lm(coble_t ~ treatment, data = fDOM_ctrl_trmt)
  
  
  ## run emmeans to see the difference btwn control and treatment
emmeans(fDOM_lm, specs=(trt.vs.ctrl ~ treatment))
```

```{r}
fDOM_ctrl_trmt %>% filter(treatment != "Control") %>% 
   ggplot(aes(x= treatment, y= coble_t))+
  geom_boxplot(aes(fill = treatment)) +
  geom_hline(yintercept = 0.07117004, color = "red", linewidth = 2, alpha = 0.5)+
  labs(caption = "red line is the average value of the inflow")+
  nunu_theme
```

### M to C Ratio 
```{r}
fDOM_hyp_test(data = fDOM, fDOM_var = m_to_c)
```

```{r}
endmem_summary 
```

```{r}
heat_map
```

```{r}
boxgraph
```

```{r, fig.height= 8, fig.width=12}
fDOM_inflow %>% 
  ggplot(aes(x= as.factor(date_time), y = m_to_c, 
             group= date_time, fill = treatment)) +
  geom_boxplot()+
  facet_wrap(.~ treatment )+
  nunu_theme +
  theme(axis.text.x = element_text(angle= 90, hjust= 1))
```


```{r, fig.height= 8, fig.width=12}
# Are the fDOM concentration the same at 12PM  within treatments and different across treatments? 

## Creating a df with only the 12PM samples 
midday_mtoc <- fDOM_inflow %>% 
  mutate(time= as.character(time)) %>% 
  filter(time == "12:00:00") %>% # keep the 12 PM time
  filter(treatment != "Inflow") %>% # remove the inflow 
  mutate(treatment = factor(treatment, # reorder the factor treatments 
                            levels = c("Control","Coral",
                                       "Rubble", "Algae")))

## set a linear model 
midday_lm <- lm(m_to_c ~ treatment*date_time, data = midday_mtoc)

## anova to see differences between date_time
Anova(midday_lm)

## contrast between control and treatments
emmeans(midday_lm, specs = trt.vs.ctrl ~ treatment)


midday_mtoc %>% #mutate(time= as.character(time)) %>% 
  #filter(time == "12:00:00") %>% 
  ggplot(aes(x= as.factor(date_time), y = m_to_c, 
             group= date_time, fill = treatment)) +
  geom_boxplot()+
  geom_hline(yintercept = 0.5063610	, color = "red", linewidth = 2, alpha = 0.5)+
  facet_wrap(.~ treatment )+
  labs(title = "12 PM samples")+
  nunu_theme +
  theme(axis.text.x = element_text(angle= 30, hjust= 1))
```

```{r}


## Creating a df with only the 12PM samples 
H24_mtoc <- fDOM_inflow %>%  
  filter(date == "2024-06-02" | date == "2024-06-03") %>% 
  filter(treatment != "Inflow") %>% # remove the inflow 
  mutate(treatment = factor(treatment, # reorder the factor treatments 
                            levels = c("Control","Coral",
                                       "Rubble", "Algae"))) %>% 
  mutate(date_time = factor(date_time))

## set a linear model 
H24_mtoc_lm <- lm(m_to_c ~ date_time*treatment, data = H24_mtoc)

## anova to see differences between date_time
Anova(H24_mtoc_lm)

## contrast between control and treatments
pairs(emmeans(H24_mtoc_lm, specs = trt.vs.ctrl ~ date_time))


fDOM_inflow %>%  
  filter(date == "2024-06-02" | date == "2024-06-03") %>% 
  filter(treatment != "Inflow") %>% 
  ggplot(aes(x= as.factor(date_time), y = m_to_c, 
             group= date_time, fill = treatment)) +
  geom_boxplot()+
  geom_hline(yintercept = 0.5063610	, color = "red", linewidth = 2, alpha = 0.5)+
  facet_wrap(.~ treatment )+
  labs(title = "24h experiment")+
  nunu_theme +
  theme(axis.text.x = element_text(angle= 30, hjust= 0.9))
```


```{r}
inflow_summary
```

```{r}
point_graph
```


```{r}
# run contrast between control and treatment
  ##  set up linear model to analyze contrasts
fDOM_lm <- lm(m_to_c ~ treatment, data = fDOM_ctrl_trmt)
  
  
  ## run emmeans to see the difference btwn control and treatment
emmeans(fDOM_lm, specs=(trt.vs.ctrl ~ treatment))
```

```{r}
fDOM_ctrl_trmt %>% filter(treatment != "Control") %>% 
   ggplot(aes(x= treatment, y= m_to_c))+
  geom_boxplot(aes(fill = treatment)) +
  geom_hline(yintercept = 0.5063610	, color = "red", linewidth = 2, alpha = 0.5)+
  labs(caption = "red line is the average value of the inflow")+
  nunu_theme
```

### Fpeak
```{r}
fDOM_hyp_test(data = fDOM, fDOM_var = fpeak)
```

```{r}
endmem_summary 
```

```{r}
heat_map
```

```{r}
boxgraph
```

```{r}
inflow_summary
```

```{r}
point_graph
```


```{r}
# run contrast between control and treatment
  ##  set up linear model to analyze contrasts
fDOM_lm <- lm(fpeak ~ treatment, data = fDOM_ctrl_trmt)
  
  
  ## run emmeans to see the difference btwn control and treatment
emmeans(fDOM_lm, specs=(trt.vs.ctrl ~ treatment))
```

```{r}
fDOM_ctrl_trmt %>% filter(treatment != "Control") %>% 
   ggplot(aes(x= treatment, y= fpeak))+
  geom_boxplot(aes(fill = treatment)) +
  geom_hline(yintercept = 0.02720418		, color = "red", linewidth = 2, alpha = 0.5)+
  labs(caption = "red line is the average value of the inflow")+
  nunu_theme
```