---
title: "MOBEAST_all_variables_delta"
created: 5 Fev 2025
edited: 25 Fev 2025
format: html
editor: source

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(janitor)
library(here)

library(ggrepel)
library(patchwork)
library(RColorBrewer)
library(measurements)
library(gt)
library(webshot2)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
library(emmeans)
library(performance)
library(multcompView)
library(broom.mixed)
library(effectsize)


# PCA packages
library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(factoextra) 
```


## Data visualization theme

```{r}
nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=24, face = "bold"), 
        plot.subtitle = element_text(size=20),
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=18, face = "bold"),
        legend.text=element_text(size=16),
        strip.background = element_rect(fill = "white"))
```



```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

## Load Data

```{r}
mobeast_raw <- read_csv(here("data", "MOBEAST_full_merged_data_long.csv"))

```


## Data Clean Up
### Outlier detection on bulk data

```{r}
## flagging outliers from the data set
mobeast<-  mobeast_raw %>%
  mutate(time_string = as.character(time_string),
              date_string = as.character(date_string),
              date_time_string = paste(date_string,time_string))%>% 
  filter(!is.na(id_number)) %>% 
  
  # remove the DNA data because they should be treated differently
  filter(variable_cat != "DNA" ) %>% 
  filter(treatment != "Rain") %>% 
  filter(treatment != "Endmember")  %>% 
  mutate(units = ifelse(variable_cat == "FCM", "cell/µL", 
                        ifelse(variable_cat == "DOC", "µmol/L", 
                        ifelse(variable_cat == "fDOM", "R.U.",
                               ifelse(variable_cat == "nutrients", "µmol/L",
                                      ifelse(variable_cat == "chla", "µg/L", NA))))))
  
```

```{}
# first pass for outlier detection on DOC because extreme values detected
DOC <- mobeast %>% filter(variable == "npoc_u_m") %>%  
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>%
  mutate(mean = mean(value, na.rm = T),
         sd = sd(value, na.rm= T)) %>% 
  mutate(outlier = ifelse(value < mean - 2*sd | value > mean + 2*sd, 
                          "outlier", "not outlier")) %>% 
  filter(outlier != "outlier") %>% 
  select(-mean, -sd, - outlier)

mobeast_DOC <- mobeast %>% 
  # from the original data set, remove the old DOC data and replace with the new DOC data with inflow data that is not outliers
  filter(variable != "npoc_u_m") %>% 
  rbind(DOC)
```

```{r}
# first pass for outlier detection on DOC because extreme values detected
DOC <- mobeast %>% filter(variable == "npoc_u_m") %>%  
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>%
  filter(time_string != "00:00:00")

mobeast_DOC <- mobeast %>% 
  # from the original data set, remove the old DOC data and replace with the new DOC data with inflow data that is not outliers
  filter(variable != "npoc_u_m") %>% 
  rbind(DOC)
```


```{r}
mobeast_flagged <- mobeast_DOC %>% 
  
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  
  # flagging outliers
  ## grouping by variable 
  group_by(variable, treatment) %>% 
  
  
  ## get the mean and the sd to calculate mean +/- 2sd
  mutate(#log_value = log10(value),
    log_value = value, 
         mean = mean(log_value, na.rm = T), 
         stand_dev = sd(log_value, na.rm = T)) %>% 
  
  ## Values outside of the mean +/- 2sd range are flagged as outliers
  ## Important to filter outliers later to understand why they are outliers. 
  mutate(outlier = ifelse(log_value > mean + 2*stand_dev | log_value < mean - 2*stand_dev,
                           "outlier", "not outlier")) %>% 
  ungroup()
  

mobeast_flagged %>%  filter(variable == "NO3")
```



### Biogeochemistry 

```{r}

mobeast_delta <- mobeast_flagged %>%  
  
  #removing outlier values that may be in the inflows too
  filter(outlier != "outlier") %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, variable) %>% 
  #mutate(inflow_isolated = value[treatment == "Inflow"]) %>% 
  mutate(inflow_isolated = ifelse(treatment == "Inflow", value, NA)) %>% 
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time, variable) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  
  # Normalize for treatments
  mutate(conc_norm = value - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) %>% 
  select(-mean, -stand_dev, -inflow, -residence_time, - inflow_isolated)

```


```{r}
## Some variables will not be studied in this project
exclude_var <- c("tot_bact", "cyano", "NO2", "ph_diff", "delta_ta", "dic_mmol_kg",
                 "tn_u_m", "lignin")

# we also won't focus on the other time points, they are mainly for the FCM data and not other data was collected at those times anyway
exclude_time <- c("03:00:00","09:00:00","15:00:00","21:00:00")

mobeast_delta <- mobeast_delta %>% filter(!variable %in%  exclude_var,
                        !time_string %in% exclude_time) 
```



## 24h data

```{r}
# subset the data set to only the 24h samples
## removing outliers from the data set
mobeast_24h <- mobeast_delta %>%

 filter(treatment != "Inflow") %>% 

  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00"))) %>% 
  mutate(log_flow = log10(flowrate)) %>% 
  filter(!is.na(conc_norm))

```





```{r}
meta_var <- tibble(variable = c("het_bact", "pico_euk", "syn", "pro", "tot_bact", "cyano", 
                      "chla_conc", "tn_u_m", "coble_a", "coble_b", "coble_c",
                      "coble_m", "coble_t", "m_to_c", "bix", "hix", "fpeak", "fi",
                      "lignin", "N_N", "NO2", "PO4", "SiO2", "NH3", "NO3", "do_mg_l",
                      "ta", "dic_mmol_kg", "ph", "npoc_u_m"),
         variable_clean = c("Heterotrophic Bacteria", "Pico-Eukaryote", "Synechococcus",
                            "Prochlorococcus", "Total Bacteria", "Cyanobacteria", 
                            "Chlorophyll a", 
                            "TN", "Coble A", "Coble B", "Coble C", "Coble M",
                            "Coble T", "M:C", "BIX", "HIX", "Fpeak", "Fluorescence Index",
                            "Lignin", "N+N", "NO2", "PO4", "SiO2", "NH3", "NO3", 
                            "Dissolved Oxygen", "Total Alkalinity", "DIC", "pH", "DOC"))

mobeast_24h <- mobeast_24h %>% left_join(meta_var, by = "variable") %>% 
  relocate(variable_clean, .after = "variable")

```


### Summary Stats 
```{}
mobeast_delta %>%
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  filter(!is.na(id_number)) %>% 
  # remove the the DOC data because something happened  at 00:00. 
  #filter(variable == "npoc_u_m") %>%  
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  group_by(variable,time_string, treatment ) %>% 
  reframe(max = max(conc_norm, na.rm = TRUE),
          min = min(conc_norm, na.rm = TRUE),
    mean = mean(conc_norm, na.rm = TRUE),
          sd = sd(conc_norm, na.rm = TRUE)) %>% 
  filter(variable == "npoc_u_m") %>%  
  arrange(mean)
```





### LMER models

```{r}
# Loop Function:
## Take the 24h samples and isolate the types of nutrient  present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

mobeast_loop <- mobeast_24h %>% select(variable) %>% distinct() %>% pull()

models_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- mobeast_24h %>% filter(variable == var)
  models_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + log_flow  +  
                            (1|tank_number), data = df)
}

```

```{}
test_df <- mobeast_24h %>%  filter(variable == "lignin")
lmer(conc_norm ~ treatment*time_fac + log_flow  +  
                            (1|tank_number), data = test_df)
```



```{}
plot_grid(imap(models_24h, ~ {
  resid_panel(.x, plots = c("qq")) + plot_annotation(title = .y)}), 
margin = c(0.2, 0.2, 0.2, 0.2), tags = TRUE)

```

#### model results (fixed and random)
```{r}
results <- imap_dfr(models_24h, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")

results %>% filter(p.value < 0.05) 

```

#### ANOVA results

Your gatekeeper — if nothing is significant here, you stop
Reports F-statistic, df, and p-value for each term
Type II is appropriate because it accounts for the other terms in the model without assuming a particular order

```{r, warning= FALSE}
# So many Warnings! 
## Warning: The column names NumDF and DenDF in ANOVA output were not recognized or transformed.

anova_results <- imap_dfr(models_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

# all the variables that have significant treatment effects
sig_anova <- anova_results %>% 
  filter( p.value <0.05) %>%  filter(term == "treatment") %>%  
  select(variable) %>% pull()

anova_results %>% 
  filter( p.value <0.05,
    #term == "treatment:time_fac"
    ) %>%  
  #filter(term == "time_fac") %>% 
  filter(variable %in% c("PO4", "N_N", "NO3","NH3", "SiO2"))


# no signficant time interactions !
anova_results %>% 
  filter(p.value <0.05, 
         term == "treatment:time_fac") %>%  
  
  select(variable)

anova_results %>% 
  #filter( p.value <0.05) %>% 
  filter(variable %in% c("het_bact", "syn"))

# report results from the fdom
fdom<- mobeast_delta %>%  filter(variable_cat == "fDOM" | variable== "npoc_u_m") %>% 
  select("variable") %>%  distinct() %>% pull()

anova_results %>% 
  filter( p.value <0.05,
    #term == "treatment"
    ) %>%  
  #filter(term == "time_fac") %>% 
  filter(variable %in% fdom)

# report results from the FCM and chla 
fcm<- mobeast_delta %>%  filter(variable_cat == "FCM" | variable_cat== "chla") %>% 
  select("variable") %>%  distinct() %>% pull()

anova_results %>% 
  filter( p.value <0.05,
    term == "time_fac"
    ) %>%  
  #filter(term == "time_fac") %>% 
  filter(variable %in% fcm)
```


##### Nutrient ANOVA summary table 

```{r}
nutrient_stat_summary_table <- anova_results %>%
  filter(variable %in% c("NO3", "PO4", "NH3", "SiO2", "chla_conc")) %>%
  #filter(term != "log_flow") %>% 
  mutate(variable = recode_values(variable,
      "NO3"      ~"Nitrate",
      "PO4"       ~ "Phosphate",
      "NH3" ~ "Ammonia",
      "SiO2" ~ "Silicate",
      "chla_conc" ~ "Chlorophyll a"
    )) %>% 
  select(Variable = variable, Term = term, DF = NumDF, F_statistic = statistic, p.value) %>%
  mutate(
    p.value = round(p.value, digits = 3),
    F_statistic  = round(F_statistic, digits = 2)
  ) %>%
  mutate(Term = recode_values(Term,
      "treatment"      ~"Treatment",
      "time_fac"       ~ "Time",
      "treatment:time_fac" ~ "Treatment x Time", 
      "log_flow" ~ "Log10 Flow Rate"
    )) 


write_csv(nutrient_stat_summary_table, path = here("outputs","nutrients_anova.csv"))
```



#### emmeans results

Unpacks where the group differences are
Particularly important for your treatment * time interaction — you need to know at which time points treatments diverge
Controls for multiple comparisons


```{r}
# "estimate" are negative somehow. Why? 
## in manual models for DOC, estimate >0, here it's <0
## t.ratio in "statistics", and negative instead of positive
## p.values not shown, only p.adjusted. 
emmeans_results <- imap_dfr(models_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 

emmeans_results %>% filter(variable %in% sig_anova) %>%
  filter(adj.p.value <0.05) 
```


```{r}
emmeans_results %>% 
  #filter(variable %in% fcm) %>% 
  filter(adj.p.value <0.05)
```


```{r}
emmeans_results_sig <- emmeans_results %>% 
  filter(adj.p.value <0.05)
emmean_sig <- emmeans_results_sig %>% select(variable) %>% distinct() %>%  pull(variable)
```



##### Nutrient Tukey Time contrasts

```{r}
emmeans_time_results <- imap_dfr(models_24h, ~ {
  emmeans(.x, pairwise ~ time_fac | treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)
}) %>% 
  relocate(variable)

emmeans_time_results %>%  filter(adj.p.value<0.05)
```

##### FCM Time Contrasts

```{r}
tukey_FCM_results <- imap_dfr(models_24h, ~ {
  emmeans(.x, pairwise ~ time_fac)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)
}) %>% 
  relocate(variable)

tukey_FCM_results %>% filter(variable %in% fcm, 
                             adj.p.value <0.05) %>%
  mutate(abs_stat = abs(statistic)) %>% 
  group_by(variable) %>% 
  reframe(contrast, 
          statistic,
          max = max(abs_stat),
          adj.p.value) %>% 
  filter(abs(statistic) == max)
```


    
#### Contrast Effect Size 

Adds biological interpretability to each contrast
Tells you not just "these groups differ" but "by how much and in which direction"
Essential for comparing across nutrients since your residuals are on different scales


```{}
effect_size_results <- tibble()

for (var in sig_eta) {
  emm <- emmeans(models_24h[[var]], ~ treatment)
  contrasts <- contrast(emm, method = "trt.vs.ctrl")

## Effect size for each contrast
  df <- tidy(eff_size(emm, sigma = sigma(models_24h[[var]]), 
                edf = df.residual(models_24h[[var]])))
  
  df <- df %>% mutate(variable = var) %>% rename(effect.size = estimate) %>% 
    relocate(variable)
  
  effect_size_results <- rbind(effect_size_results, df)
  
}

effect_size_results %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") %>% filter(abs(effect.size) >=0.8)



```


#### Violin plots

```{r, warning=FALSE}

graph_loop <- mobeast_24h %>% #filter(variable %in% sig_eta) %>% 
  select(variable) %>% 
  distinct() %>% pull()
  
plots_24h <- vector(mode = "list", length = length(graph_loop))

for (var in graph_loop) {
  
  unit_plot <- mobeast_24h %>% filter(variable == var) %>% select(units) %>% distinct()
  
  variable_name <- mobeast_24h %>% filter(variable == var) %>% 
    select(variable_clean) %>% distinct()
  
  plots_24h[[var]] <- mobeast_24h %>% 
  filter(variable == var) %>% 
  ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    #geom_jitter(aes(color = outlier), alpha = 0.5)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #facet_wrap( .~ time_fac, ncol = 4 )+
    labs(
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = paste0("Delta ", variable_name, " (",unit_plot, ")"))+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18),
          legend.position = "none")
}

```

```{r, fig.width=14, fig.height=14}
# fDOM plots 
(plots_24h[["coble_c"]]+ labs(title = "A")) / (plots_24h[["m_to_c"]]+ labs(title = "B"))/
  (plots_24h[["hix"]]+ labs(title = "C")) / (plots_24h[["fi"]]+ labs(title = "D"))

#ggsave(filename = "24h_fDOM.png", plot = last_plot(), path = here("outputs"))
```

```{r, fig.width=14, fig.height=17}
# FCM plots 
(plots_24h[["het_bact"]]+ labs(title = "A")) /(plots_24h[["syn"]]+ labs(title = "B")) / (plots_24h[["pico_euk"]] + labs(title = "C")) / (plots_24h[["chla_conc"]] + labs(title = "D"))

ggsave(filename = "24h_FCM.png", plot = last_plot(), path = here("outputs"))
```


```{r, fig.width=14, fig.height=10}
# Nutrients plots 

(plots_24h[["N_N"]]+ labs(title = "A")) / (plots_24h[["SiO2"]]+ labs(title = "B"))


ggsave(filename = "24h_nutrients.png", plot = last_plot(), path = here("outputs"))
```

```{r, fig.width=14, fig.height=6}
plots_24h[["N_N"]]+
  labs(y = "Nitrate+Nitrite Difference from Inflow (µM)")

ggsave(filename = "24h_N+N.png", plot = last_plot(), path = here("outputs"))
```

```{r, fig.width=14, fig.height=12}
(plots_24h[["fi"]] + facet_wrap( .~ time_fac, ncol = 4 )) / (plots_24h[["m_to_c"]]+ facet_wrap( .~ time_fac, ncol = 4 )) + plot_layout(axes = "collect")

ggsave(filename = "24h_ICRS_FDOM.png", plot = last_plot(), path = here("outputs"))
```


```{r, fig.width=14, fig.height=6}
plots_24h[["het_bact"]]+
  labs(y = "Bacterial Abundance Difference (cell/µL)")#+
  #facet_wrap( .~ time_fac, ncol = 4 )

ggsave(filename = "24h_cobC.png", plot = last_plot(), path = here("outputs"))
```



```{r, fig.width=14, fig.height=10}
 plots_24h[["syn"]] / plots_24h[["het_bact"]]+ plot_layout(axes = "collect")

ggsave(filename = "24h_ICRS_FCM.png", plot = last_plot(), path = here("outputs"))
```


```{r, fig.width=14, fig.height=10}
# Carbonate plots 

(plots_24h[["ph"]]+ labs(title = "A", 
                         y = "Delta pH")) / 
  (plots_24h[["ta"]]+ labs(title = "B",
                           y = "Delta TA (μmol/kg)"))


ggsave(filename = "24h_carbonate.png", plot = last_plot(), path = here("outputs"))
```

```{r, fig.width=14, fig.height=30}
# Nutrients plots 

(plots_24h[["coble_c"]]+ labs(title = "Fluorescent DOM", subtitle = "A - Coble C")) /
  (plots_24h[["hix"]]+ labs(subtitle = "B - HIX")) / 
  
  #(plots_24h[["NO3"]]+ labs(title = "Nutrients",subtitle = "C - NO3"))/ 
  
  (plots_24h[["het_bact"]]+ labs(title = "Cell Count",subtitle = "D - Heterotrophic Bacteria"))/ 
  (plots_24h[["syn"]]+ labs(subtitle = "E - Synechococcus"))

#ggsave(filename = "24h_sig_var.png", plot = last_plot(), path = here("outputs"))
```

##### Nutrients 

```{r,message=FALSE, fig.width = 13, fig.height = 8}
mobeast_24h %>% 
   filter(variable %in% c("NO3", "PO4", "NH3", "SiO2", "chla_conc")) %>% 
  mutate(variable = recode_values(variable,
      "NO3"      ~"Nitrate",
      "PO4"       ~ "Phosphate",
      "NH3" ~ "Ammonia",
      "SiO2" ~ "Silicate",
      "chla_conc" ~ "Chlorophyll a"
    )) %>% 
  mutate(variable= factor(variable, levels = c("Nitrate",  "Ammonia", "Phosphate", "Silicate", "Chlorophyll a"))) %>% 
  ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    #geom_jitter(aes(color = outlier), alpha = 0.5)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_grid( variable ~ time_fac, scales = "free" )+
    labs(
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = paste0("Delta Nutrients (µM)"))+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18),
          legend.position = "none")

ggsave(filename = "nutrient_chla_plot.png" , plot = last_plot(), height = 8, width = 13,path = here("outputs"))
```


#### Summary Stats for Thesis

```{r}
mobeast_delta %>% filter(variable %in% sig_anova) %>% 
  group_by(variable, time, treatment) %>% 
  reframe(mean = mean(conc_norm, na.rm = TRUE),
          sd = sd(conc_norm, na.rm = TRUE))
```

## DOM x Het Bact
```{r}
mobeast_wide <- mobeast_24h %>% select(tank_number, treatment, inflow_table, 
                       time_fac,variable, conc_norm, log_flow) %>% 
  pivot_wider(names_from = variable, values_from = conc_norm)
```


### HIX 
```{r}
mod_hix <- lm(het_bact ~ hix + treatment + log_flow , 
                  data= mobeast_wide)
anova(mod_hix)
contrast(emmeans(mod_hix, ~treatment), methods= "trt.vs.ctrl")

ggplot(mobeast_wide, aes(hix, het_bact, color = treatment,group= treatment))+
  geom_point()+
  geom_smooth(method = "lm",se = F)+
  facet_wrap(.~treatment)+
  scale_color_manual(values = treatment_colors)+
  nunu_theme

ggplot(mobeast_wide, aes(hix, het_bact, fill = treatment,group= treatment))+
  geom_violin()+
  facet_wrap(.~treatment, nrow=1)+
  stat_summary(alpha = 0.5)+
  scale_fill_manual(values = treatment_colors)+
  nunu_theme
```

### Coble C

```{r}
mod_cobC <- lm(het_bact ~ coble_c + treatment + log_flow , 
                  data= mobeast_wide)
anova(mod_cobC)
contrast(emmeans(mod_cobC, ~treatment), methods= "trt.vs.ctrl")

ggplot(mobeast_wide, aes(coble_c, het_bact, color = treatment,group= treatment))+
  geom_point()+
  geom_smooth(method = "lm",se = F)+
  facet_wrap(.~treatment, nrow=1)+
  scale_color_manual(values = treatment_colors)+
  nunu_theme

ggplot(mobeast_wide, aes(coble_c, het_bact, fill = treatment,group= treatment))+
  geom_violin()+
  facet_wrap(.~treatment, nrow=1)+
  stat_summary(alpha = 0.5)+
  scale_fill_manual(values = treatment_colors)+
  nunu_theme
```



### Plots 


```{r}
# Define shading dataframe matching your exact factor levels
# Adjust period assignments to match your actual day/night schedule
shading <- data.frame(
  time_fac = levels(mobeast_24h$time_fac),
  period   = c("day", "night", "night", "day")  # adjust to your time points
)

ta <- mobeast_24h %>% 
  filter(variable == "ta") %>% 
  group_by(variable, time_fac, treatment) %>% 
  reframe(mean = mean(conc_norm, na.rm = T),
          sd   = sd(conc_norm, na.rm = T)) %>% 
  ggplot(aes(x = time_fac, y = mean, color = treatment, group = treatment)) +
  # Day/night shading — added before geom_point and geom_line
  geom_tile(
    data = shading,
    aes(x = time_fac, y = 0, fill = period),
    height = Inf, alpha = 0.3,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(
    values = c("day" = "lightyellow", "night" = "lightblue"),
    name   = "Period"
  ) +
  geom_point() +
  geom_line(aes(color = treatment)) +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  facet_grid(treatment ~ .) +
  labs(title = "Total Alkalinity ",
    x = "Time Point",
    y = paste0("Delta TA (µmol/kg)")
  ) +
  scale_color_manual(
    breaks = c("Control", "Coral", "CCA", "Algae"),
    values = c("grey", "#03A9F4", "#EF5350", "#4CAF50")
  ) +
  nunu_theme +
  theme(
    axis.text       = element_text(size = 18),
    legend.position = "none"
  )

ph <- mobeast_24h %>% 
  filter(variable == "ph") %>% 
  group_by(variable, time_fac, treatment) %>% 
  reframe(mean = mean(conc_norm, na.rm = T),
          sd   = sd(conc_norm, na.rm = T)) %>% 
  ggplot(aes(x = time_fac, y = mean, color = treatment, group = treatment)) +
  # Day/night shading — added before geom_point and geom_line
  geom_tile(
    data = shading,
    aes(x = time_fac, y = 0, fill = period),
    height = Inf, alpha = 0.3,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(
    values = c("day" = "lightyellow", "night" = "lightblue"),
    name   = "Period"
  ) +
  geom_point() +
  geom_line(aes(color = treatment)) +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  facet_grid(treatment ~ .) +
  labs(title = "pH",
    x = "Time Point",
    y = paste0("Delta pH")
  ) +
  scale_color_manual(
    breaks = c("Control", "Coral", "CCA", "Algae"),
    values = c("grey", "#03A9F4", "#EF5350", "#4CAF50")
  ) +
  nunu_theme +
  theme(
    axis.text       = element_text(size = 18),
    legend.position = "none"
  )

ta + ph
```


```{r}

# Define shading dataframe matching your exact factor levels
# Adjust period assignments to match your actual day/night schedule
shading <- data.frame(
  time_fac = levels(mobeast_24h$time_fac),
  period   = c("day", "night", "night", "day")  # adjust to your time points
)

ta <- mobeast_24h %>% 
  filter(variable == "ta") %>% 
  ggplot(aes(x = time_fac, y = conc_norm, fill = treatment)) +
  # Day/night shading - added before geom_point and geom_line
  #geom_point() +
  geom_tile(
    data = shading,
    aes(x = time_fac, y = 0, fill = period),
    height = Inf, alpha = 0.3,
    inherit.aes = FALSE
  ) +
  geom_violin() +
  stat_summary(color = "yellow", alpha =0.5)+
  geom_hline(yintercept = 0, linetype = 'dashed') +
  facet_grid(treatment ~ .) +
  labs(title = "Total Alkalinity ",
    x = "Time Point",
    y = paste0("Delta TA (µmol/kg)")
  ) +
  scale_fill_manual(
    breaks = c("Control", "Coral", "CCA", "Algae", "day", "night"),
    values = c("grey", "#03A9F4", "#EF5350", "#4CAF50", "lightyellow", "lightblue")
  ) +
  nunu_theme +
  theme(
    axis.text       = element_text(size = 18),
    legend.position = "none"
  )

ph <- mobeast_24h %>% 
  filter(variable == "ph") %>% 
  ggplot(aes(x = time_fac, y = conc_norm, fill = treatment)) +
  # Day/night shading - added before geom_point and geom_line
  #geom_point() +
  geom_tile(
    data = shading,
    aes(x = time_fac, y = 0, fill = period),
    height = Inf, alpha = 0.3,
    inherit.aes = FALSE
  ) +
  geom_violin() +
  stat_summary(color = "yellow", alpha =0.5)+
  geom_hline(yintercept = 0, linetype = 'dashed') +
  facet_grid(treatment ~ .) +
  labs(title = "pH",
    x = "Time Point",
    y = paste0("Delta pH")
  ) +
  scale_fill_manual(
    breaks = c("Control", "Coral", "CCA", "Algae", "day", "night"),
    values = c("grey", "#03A9F4", "#EF5350", "#4CAF50", "lightyellow", "lightblue")
  ) +
  nunu_theme +
  theme(
    axis.text       = element_text(size = 18),
    legend.position = "none"
  )

ta + ph
```



## PCA 

```{r}
## used the areal flux to calculate the PCA here
## normalization with scale(), but don't actually know what it noarmalizes to...
pca_24h_df <- mobeast_24h %>% select(id_number, tank_number, treatment, inflow_table,
                             date_string, time_string, variable, conc_norm) %>%
  pivot_wider(names_from = variable, values_from = conc_norm) %>% 
   na.omit()

pca_matrix_24h_norm <- pca_24h_df %>% select(het_bact:npoc_u_m) %>% scale() 

pca_matrix_24h_norm %>% cor() %>% 
  ggcorrplot(
           hc.order = TRUE,
           #type = "lower",
           lab = TRUE)
```


```{r}

pca_matrix_24h_norm<- pca_matrix_24h_norm %>% as.data.frame() %>% 
  select(het_bact, pico_euk, syn, pro, chla_conc, coble_b,SiO2,NH3,
                                coble_t, bix, fpeak, fi, PO4, NO3, 
                                het_bact, npoc_u_m, coble_a, 
                                coble_c, coble_m, hix,npoc_u_m ) %>% as.matrix()
```


```{r}
data.pca <- princomp(pca_matrix_24h_norm, cor = TRUE)
summary(data.pca)

data.pca$loadings[,1:4]

loading_results <- as.data.frame(data.pca$loadings[,1:4])

fviz_eig(data.pca, addlabels = TRUE)

fviz_pca_var(data.pca, col.var = "black")

fviz_pca_var(data.pca, col.var = "cos2",
            gradient.cols = c("blue", "lightgrey", "red"),
            repel = TRUE)
```

```{r}
pca_scores <- as.data.frame(data.pca$scores)

ggplot(pca_scores, aes(Comp.1,Comp.2))+
  geom_point(aes(color = pca_24h_df$treatment, size= 3))+
  geom_segment(data = loading_results, 
               aes(x=0, y=0, xend = Comp.1, yend= Comp.2),
               arrow = arrow(length = unit(0.5, "cm")),
               alpha = 0.75, color = "black")+
  geom_text(data = loading_results, 
            aes(x = Comp.1, y = Comp.2, 
                label = rownames(loading_results)),
            vjust=1, hjust = 1)+
  theme_bw() +
      scale_color_manual(
        breaks = c("Control", "Coral", "CCA", "Algae"),
        values = c("grey", "#03A9F4", "#EF5350", "#4CAF50")
      )  
```

