---
title: "MOBEAST- DNA Ordination 24h V2"
format: html
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

# Load Libraries

```{r}
library(tidyverse)
#library(lubridate)
library(here)
library(janitor)
#library(ggrepel)
#library(ghibli)
#library(gt)
library(patchwork)
#library(RColorBrewer)
library(vegan)
library(OmicFlow)
library(phyloseq)
#library(GUniFrac)

library(car) # ANOVA package
#library(ggeffects)
#library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
#library(sjPlot)
library(emmeans)
#library(performance)
#library(multcompView)
library(broom.mixed) # for tidy()
```

# Data visualization

```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

```{r}
nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=24, face = "bold"), 
        plot.subtitle = element_text(size=20),
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=18, face = "bold"),
        legend.text=element_text(size=16),
        strip.background = element_rect(fill = "white"))
```

# Load Data

```{r}
unifrac <- read.csv(here("data", "DNA", "unifrac", "unifrac_weighted_97.csv"), row.names = 1,
                    header = TRUE)

meta <- read_csv(here("data","MOBEAST_metadata_V2.csv"))

mobeast <- read_csv(here("data", "MOBEAST_full_merged_data_long.csv"))

DNA <- read.csv(here("data", "DNA", "MOBEAST_OTU_culled.csv"), header = 1)
tax <- read.csv(here("data", "DNA", "MOBEAST_tax_culled.csv"), header = 1)

```

# Data Clean Up

```{r}
endmember_list <- meta %>% 
  filter(treatment == "Endmember" |treatment == "Rain") %>% 
  pull(id_number)

inflow_list <- meta %>% 
  filter(treatment == "Inflow") %>% 
  pull(id_number)

diel_list <- meta %>% mutate(date_string = as.character(date_string)) %>% 
  filter(date_string == "2024-06-02" | date_string == "2024-06-03") %>% 
  pull(id_number)

sample_list <- meta %>% 
  #filter(!id_number %in% endmember_list) %>% 
  filter(!id_number %in% inflow_list) %>% 
  filter(id_number %in% diel_list) %>% 
  pull(id_number)
```

```{r}
unifrac_subset <- unifrac[rownames(unifrac) %in% sample_list, colnames(unifrac) %in% sample_list]

unifrac_dist <- as.dist(unifrac_subset)
```


# NMDS Calculation

```{r}

nmds <- metaMDS(unifrac_dist, k=2, trymax = 1000)
# put stress in the corner fo the nmds plot
nmds$stress
```

```{r}
nmds_scores <- as.data.frame(scores(nmds, display = "site"))
```

```{r}
nmds_scores$id_number= rownames(nmds_scores)
```

```{r}
nmds_merged <- merge(nmds_scores, meta, by.x ="id_number", by.y = "id_number")
```

```{r}
nmds_merged <- mobeast %>% select(id_number, flowrate) %>% 
  distinct() %>% right_join(nmds_merged) %>% mutate(log_flow = log10(flowrate))
```



```{r}
nmds_merged<- nmds_merged %>% mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```


# Plotting Ordination 


```{r, fig.width=16, fig.height= 6}
# Plot
nmds_merged %>%   
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.8, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 4)+
  labs(title= "Ordination of 24h samples")+
  scale_color_manual(values=treatment_colors )+
  scale_fill_manual(values= treatment_colors) +
  nunu_theme+
  theme(legend.position = "top")
#ggsave("24h_treatment.png", width = 16, height = 6, path = here("outputs", "DNA"))

```
## PERMANOVA 

```{r}
nmds_merged<- nmds_merged %>% mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00","00:00:00","06:00:00")))
```


```{r}
# Note : Version 2 of adonis 
set.seed(1999)
perma_24h <- adonis2(unifrac_dist ~ treatment+time_fac+ treatment:time_fac + log_flow, 
                     data = nmds_merged,
                     strata = nmds_merged$tank_number,by = "term")
print(perma_24h)

## Random effect (strata) should be applied on the tank because we sample repeatedly in the tank.
## We don't sample the same tank multiple times at a time point. 
## We may have a time interaction tho, so I added the date_time in the model formula. 
### Note: why is the output of the model not show fixed effects and random effects like lmer()? 


```

## Pairwise comparison 

- permnova : adonis2
- pairwise pairwise.adonis2(p.adjust.m = "BH"), make sure that it's a distance matrix (as.dist())

```{r}
set.seed(1999)
pairwise_adonis(x = unifrac_dist, 
                groups = as.vector(nmds_merged$treatment),
                  p.adjust.m = "BH")
```




# Ordination of individual time points

## 12:00:00

```{r}
samples_12PM <- meta %>% mutate(time_string = as.character(time_string),
                date_string = as.character(date_string)) %>% 
  filter(time_string == "12:00:00", 
         date_string == "2024-06-02", 
         treatment != "Inflow") %>% 
  select(id_number) %>%  pull()
```

```{r}
unifrac_12PM <- unifrac[rownames(unifrac) %in% samples_12PM, colnames(unifrac) %in% samples_12PM]

unifrac_dist_12PM <- as.dist(unifrac_12PM)
```


### NMDS Calculation

```{r}
nmds_12PM <- metaMDS(unifrac_dist_12PM, k=2, trymax = 1000)
# put stress in the corner fo the nmds plot
```

```{r}
nmds_scores_12PM <- as.data.frame(scores(nmds_12PM, display = "site"))
```

```{r}
nmds_scores_12PM$id_number= rownames(nmds_scores_12PM)
```

```{r}
nmds_merged_12PM <- merge(nmds_scores_12PM, meta, by.x ="id_number", by.y = "id_number")
```

```{r}
nmds_merged_12PM <- mobeast %>% select(id_number, flowrate) %>% 
  distinct() %>% right_join(nmds_merged_12PM) %>% mutate(log_flow = log10(flowrate))
```

```{r}
nmds_merged_12PM<- nmds_merged_12PM %>% mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```


### Plotting Ordination 


```{r, fig.width=12, fig.height= 8}
# Plot
nmds_merged_12PM %>%   
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.8, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  #facet_wrap(~time_string, ncol = 4)+
  labs(title= "Ordination 12:00 samples")+
  scale_color_manual(values=treatment_colors )+
  scale_fill_manual(values= treatment_colors) +
  nunu_theme+
  theme(legend.position = "top")
#ggsave("24h_treatment.png", width = 16, height = 6, path = here("outputs", "DNA"))

```

### PERMANOVA 

```{r}
nmds_merged_12PM<- nmds_merged_12PM %>% mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00","00:00:00","06:00:00")))
```


```{r}
# Note : Version 2 of adonis 
set.seed(1999)
perma_12PM <- adonis2(unifrac_dist_12PM ~ treatment + log_flow, 
                     data = nmds_merged_12PM, by = "term")
tidy(perma_12PM)

## Random effect (strata) should be applied on the tank because we sample repeatedly in the tank.
## We don't sample the same tank multiple times at a time point. 
## We may have a time interaction tho, so I added the date_time in the model formula. 
### Note: why is the output of the model not show fixed effects and random effects like lmer()? 


```

### Pairwise comparison
```{r}
set.seed(1999)
pairwise_adonis(x = unifrac_dist_12PM, 
                groups = as.vector(nmds_merged_12PM$treatment),
                  p.adjust.m = "BH")
```


## 06:00:00

```{r}
samples_6AM <- meta %>% mutate(time_string = as.character(time_string),
                date_string = as.character(date_string)) %>% 
  filter(time_string == "06:00:00", 
         date_string == "2024-06-03", 
         treatment != "Inflow") %>% 
  select(id_number) %>%  pull()
```

```{r}
unifrac_6AM <- unifrac[rownames(unifrac) %in% samples_6AM, colnames(unifrac) %in% samples_6AM]

unifrac_dist_6AM <- as.dist(unifrac_6AM)

```


### NMDS Calculation

```{r}
nmds_6AM <- metaMDS(unifrac_dist_6AM, k=2, trymax = 1000)
# put stress in the corner fo the nmds plot
```

```{r}
nmds_scores_6AM <- as.data.frame(scores(nmds_6AM, display = "site"))
```

```{r}
nmds_scores_6AM$id_number= rownames(nmds_scores_6AM)
```

```{r}
nmds_merged_6AM <- merge(nmds_scores_6AM, meta, by.x ="id_number", by.y = "id_number")
```

```{r}
nmds_merged_6AM <- mobeast %>% select(id_number, flowrate) %>% 
  distinct() %>% right_join(nmds_merged_6AM) %>% mutate(log_flow = log10(flowrate))
```

```{r}
nmds_merged_6AM <- nmds_merged_6AM %>% mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```


### Plotting Ordination 


```{r, fig.width=12, fig.height= 8}
# Plot
nmds_merged_6AM %>%   
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.8, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  #facet_wrap(~time_string, ncol = 4)+
  labs(title= "Ordination 06:00 samples")+
  scale_color_manual(values=treatment_colors )+
  scale_fill_manual(values= treatment_colors) +
  nunu_theme+
  theme(legend.position = "top")
#ggsave("24h_treatment.png", width = 16, height = 6, path = here("outputs", "DNA"))

```

### PERMANOVA 

```{r}
nmds_merged_6AM<- nmds_merged_6AM %>% mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00","00:00:00","06:00:00")))
```


```{r}
# Note : Version 2 of adonis 
set.seed(1999)
perma_6AM <- adonis2(unifrac_dist_6AM ~ treatment + log_flow, 
                     data = nmds_merged_6AM, by = "term")
tidy(perma_6AM)

## Random effect (strata) should be applied on the tank because we sample repeatedly in the tank.
## We don't sample the same tank multiple times at a time point. 
## We may have a time interaction tho, so I added the date_time in the model formula. 
### Note: why is the output of the model not show fixed effects and random effects like lmer()? 


```

### Pairwise comparison
```{r}

set.seed(1999)
pairwise_adonis(x = unifrac_dist_6AM, 
                groups = as.vector(nmds_merged_6AM$treatment),
                  p.adjust.m = "BH")
```




# Summary Ordindations
```{r}
meta_24h <- meta %>% mutate(time_string = as.character(time_string),
                date_string = as.character(date_string)) %>% 
  filter(date_string == "2024-06-02" | date_string == "2024-06-03", 
         treatment != "Inflow") 

mobeast_24h <- mobeast %>% 
  mutate(date_string = as.character(date_string),
         time_string = as.character(time_string)) %>% 
    filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  mutate(log_flow = log10(flowrate)) %>% 
  mutate(treatment = factor(treatment, 
                            levels = c("Control", "Coral", "CCA", "Algae"))) %>%
  filter(!tank_number %in% c("F1", "F2"), 
         !is.na(id_number)) %>% 
     select(id_number, treatment,tank_number, time_string, log_flow) %>% 
     distinct()

```

```{r}
time_point <- c("12:00:00", "18:00:00","00:00:00", "06:00:00")
perma_results <- tibble()
pairwise_results <- tibble()

for (time in time_point) {

  ## Isolating the samples that go in one time point
  samples <- mobeast_24h %>% filter(time_string == time) %>% 
  select(id_number) %>%  pull()
  
  ## Create a unifrac with a specific time point
  unifrac_time <- unifrac[rownames(unifrac) %in% samples, colnames(unifrac) %in% samples]

  unifrac_dist <- as.dist(unifrac_time)
  
  # mobeast filtering 
   df <- mobeast_24h %>% filter(time_string == time) 
  
   set.seed(1999)
  # Calculating the PERMANOVA 
  perma <- adonis2(unifrac_dist ~ treatment + log_flow, 
                     data = df, by = "term")
  
  # saving the results of the model in a clean df with the time point name
  results <- tidy(perma) %>% mutate(time_point = time) %>%  
    relocate(time_point) %>% 
    rename(Sum_Sq_PERMA = SumOfSqs, 
           R2_PERMA = R2,
           F_stat = statistic,
           p_val_PERMA = p.value)
  
  # bind to the final result df
  perma_results <- rbind(perma_results, results)
  
  set.seed(1999)
  # save the pairwise results in a df 
  pairwise <- pairwise_adonis(x = unifrac_dist, 
                groups = as.vector(df$treatment),
                  p.adjust.m = "BH")
  
  #bind to dinal result df 
  results_pair <- pairwise %>%  
    rename(Sum_Sq_pair = SumsOfSqs, 
           R2_pair = R2,
           F_stat_pair = F.Model,
           p_val_pair = p.value,
           p_adj_pair = p.adj) %>% 
    mutate(time_point = time) %>%  
    relocate(time_point) 
  
  pairwise_results <- rbind(pairwise_results, results_pair)
  
}

perma_results

pairwise_results %>% 
  filter(pairs %in% c("Algae vs Control", 
                      "CCA vs Control",
                      "Control vs Coral"))

```






# Microbiome diversity 

```{r}
div_meta <- mobeast %>% select(id_number, treatment, date_string, flowrate) %>% 
  distinct() %>% 
  mutate(log_flow = log10(flowrate))

# formatting the OTU matrix with sample names in the rows and OTU on the columns
otu <- DNA %>% 
  # remove the PCR controls from the df
  select(-P1_PCRPos,-P2_PCRPos, -P3_PCRPos, -P4_PCRPos, - P5_PCRPos) %>%
  # join with the taxonomic df 
  rename(OTUNumber = ASV) %>% 
  left_join(tax) %>% 
  # Here I want the relative abundance of microbes by family per samples
  select(MOBEAST001:MOBEAST115, Family) %>% 
  relocate(Family,.before = "MOBEAST001") %>% 
  # different OTU may result in bacteria that share a family but belong to different genus
  # we want to add the relative abundance of the bacteria in the same family
  group_by(Family) %>%
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) %>% 
  as.data.frame()

rownames(otu)<- otu$Family

#transpose so that the Families are the column names and the samples in the rows
otu_t <- otu %>%  select(-Family) %>% t()

#convert the matrix in a df that I can filter
otu_t <- otu_t %>% as.data.frame() %>% mutate(id_number = rownames(otu_t)) %>% 
  left_join(div_meta) %>% relocate(c("id_number", "treatment", "date_string"))

```

### 24h 
```{r}
# we only keep the 24h samples. we remove the inflow values. 
otu_24h <- otu_t %>% mutate(date_string = as.character(date_string)) %>% 
  filter(date_string == "2024-06-02" | date_string == "2024-06-03") %>% 
  filter(treatment != "Inflow") %>% 
  select(-treatment, -date_string)

rownames(otu_24h)<- otu_24h$id_number

# convert the df into a matrix
matrix_24h <- otu_24h %>% select(-id_number) %>% as.matrix() 

```

```{r}
# calculating different biodiversity metrics and indices. 
richness <- specnumber(matrix_24h)
shannon <- vegan::diversity(matrix_24h, index = "shannon")
simpson <- vegan::diversity(matrix_24h, index = "simpson")
evenness <- shannon/log(richness)
berger_parker <- apply(matrix_24h, 1, function(x) max(x)/sum(x))


# merge the results in one data frame that we can use for stat models
diversity_indices <- tibble(
  id_number = rownames(matrix_24h),
  richness = richness, 
  shannon = shannon, 
  simpson = simpson, 
  evenness = evenness, 
  berger_parker = berger_parker
)
```

```{r}
# merge the diversity metrics with the meta data
div_df_24h <- diversity_indices %>% left_join(meta) %>% 
  # only select the meta data that is useful for the models
  select(id_number, tank_number, treatment, inflow_table, time, date_string, time_string,
         richness,shannon,simpson, berger_parker, evenness) %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  mutate(time_fac = factor(time_string, 
                           levels = c("12:00:00", "18:00:00","00:00:00", "06:00:00"))) %>% 
  #focusing the shanon diversity index, we want to remove the outlier values
  ## we use the mean +/- 2SD method
  mutate(sd = sd(shannon), mean = mean(shannon)) %>% 
  filter(shannon < mean+ 2* sd | shannon > mean-2*sd) %>%
  select(-sd, -mean)  
  
```


```{r}
# statistical model to see if diversity changes by treatment and in time
# tank number used for repeated measure random effect
mod_div <- lmer(shannon ~ treatment*time+ (1|tank_number), data = div_df_24h)

summary(mod_div)
anova(mod_div, type= 2)
emmeans(mod_div, specs = trt.vs.ctrl ~ treatment) 
emmeans(mod_div, pairwise ~ treatment | time)
```

**Definition**: 
*Richness*: total number of different species. (discreet count)
*Evenness*: even population or 1 dominant group vs smaller groups. (1 is even,  0 is uneven)
*Shannon*: accounts for evenness and richness. (High value = higher diversity)
*Simpson*: probability that 2 indiv randomly picked are the same spp. (1 = higher diversity, 0= low/no diversity)
*Berger_parker*: numerical dominance (high = one single spp = low diversity)


```{r}
div_df_24h %>% 
  ggplot(aes(treatment, shannon, fill = treatment)) +
  geom_violin()+
  geom_point()+
  facet_wrap(.~time_fac, nrow = 1)+
  stat_summary(color = "yellow", alpha =0.5)+
  scale_fill_manual(values = treatment_colors)+
  nunu_theme
```
```{r}
div_df_24h %>% 
  
  ggplot(aes(richness, evenness, color = treatment)) +
  #geom_violin()+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+

  #facet_wrap(.~time_fac, nrow = 1)+
  #stat_summary(color = "yellow", alpha =0.5)+
  scale_color_manual(values = treatment_colors)+
  nunu_theme
```

```{r}
div_df_24h %>% 
  group_by(treatment) %>% 
  reframe(mean_rich = mean(richness), mean_even = mean(evenness)) %>% 
  ggplot(aes(mean_rich, mean_even, color = treatment)) +
  #geom_violin()+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  geom_label(aes(label = treatment), hjust = 0,vjust = 1.5)+

  #facet_wrap(.~time_fac, nrow = 1)+
  #stat_summary(color = "yellow", alpha =0.5)+
  scale_color_manual(values = treatment_colors)+
  nunu_theme
```
