---
title: "MOBEAST- DNA Ordination 24h"
format: html
editor: visual
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

# Load Libraries

```{r}
library(tidyverse)
#library(lubridate)
library(here)
library(janitor)
#library(ggrepel)
#library(ghibli)
#library(gt)
library(patchwork)
#library(RColorBrewer)
library(vegan)
library(phyloseq)
#library(GUniFrac)
```

# Data visualization

```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=32, face = "bold"), 
        plot.subtitle = element_text(size=30),
        axis.title = element_text(size = 28, face = "bold"),
        axis.text = element_text(size = 28),
        strip.text.x = element_text(size = 28, face = "bold"),
        legend.title=element_text(size=28, face = "bold"),
        legend.text=element_text(size=28))
```

# Load Data

```{r}
otu <- read_csv(here("data", "DNA", "MOBEAST_OTU_culled.csv"))
meta <- read_csv(here("data","MOBEAST_metadata_V2.csv"))
#tax <- read_csv(here("data","DNA", "MOBEAST_tax_culled.csv"))
OTU_ord <- read_csv(here("data","OTU_heatmap_ready.csv"))
cluster <- read_csv(here("data","DNA", "24h_clusters.csv"))
```

# Data Clean Up

```{r, meta data clean up}
meta<- meta %>% mutate(time = str_pad(time, width = 4, side = "left", pad = "0"),
               time = str_replace(time, "(\\d{2})(\\d{2})", "\\1:\\2")) %>% 
  # "(\\d{2})(\\d{2})" separates the string in 2 groups of 2 characters
  # "\\1:\\2", group1 + ":" + group2
  mutate(time = paste0(time, ":00"),
               date_time = paste(date, time),
               date = mdy(date), 
               date_time = mdy_hms(date_time),
               time = factor(time))
  
  
```

## Preparing the metadata to annotate the heatmap

```{r}
# renaming some of treatment so that environmental data have a label: endmember or rain
# Same for inflow table, environmental samples now have a label
## Convert tibble to df for later
meta <- meta %>% mutate(treatment = as.character(treatment)) %>% 
  mutate(treatment = ifelse(id_number == "MOBEAST115", "Rain", 
                            ifelse(is.na(treatment), "Endmember", treatment))) %>% 
  mutate(inflow_table = ifelse(is.na(inflow_table), "Environmental data", inflow_table)) %>% 
  mutate(date_string = as.character(date),
                time_string = as.character(time),
                date_time_string = paste(date_string, time_string)) %>% 
  as.data.frame()
```

```{r}
cluster<- cluster %>% 
  mutate(cluster = factor(cluster))

meta<- meta %>% left_join(cluster)
```

## Transform the dataframes into matrices

```{r}
## Make an OTU matrix
otu<- otu %>% mutate(ASV = toupper(ASV))
otu_list<- otu$ASV
otu<- otu %>% select(-(117:122), -MOBEAST017b)
otu_mat <- as.matrix(otu)
rownames(otu_mat) <- otu_list

```

```{r}
OTU <- otu_table(otu_mat, taxa_are_rows = TRUE)
```

# Ordinations with long format df

## 24h Samples

```{r}
otu_ord <- OTU %>% 
  t() %>% 
  as.data.frame() 

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select 12PM samples 
id_number_24h <- meta %>% 
  filter(date_string == "2024-06-02"| date_string == "2024-06-03") %>% 
  pull(id_number)

otu_ord_24h <- otu_ord %>% filter(id_number %in% id_number_24h)


```

```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_24h <- otu_ord_24h %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances

```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_24h <- vegdist(otu_ord_mat_24h, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_24h, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_24h <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_24h <- left_join(MOBEAST_nmds_24h, meta)
```

### Plotting NDMS 24h

#### Plotting by treatment

```{r, fig.width=16, fig.height= 12}
# Plot
MOBEAST_nmds_24h %>%   
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 2)+
  labs(title= "24h samples, ellipse by treatment")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

#### Plotting by heatmap cluster

Note: cluster numbers are inaccurate and do not match what is seen in the heatmap. Problem with the distance matrix or clutering method or both...

```{r, fig.width=16, fig.height= 10}
# Plot
MOBEAST_nmds_24h %>%   
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  #stat_ellipse(aes(colour = cluster, fill = cluster),
               #alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 2)+
  labs(title= "24h samples, ellipse by treatment")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  ggforce::geom_mark_ellipse(aes(color = cluster, fill = cluster), alpha = 0.3)+
  #nunu_theme+
  theme_bw()
#ggsave("24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

```{r, fig.width=10, fig.height= 6}
# raw NMDS plot with all the samples
MOBEAST_nmds_24h %>%  
  mutate(time_string = factor(time_string, 
                              levels = c("12:00:00","18:00:00","00:00:00","06:00:00" ))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment))+
  stat_ellipse(aes(colour = time_string, fill = time_string), 
               alpha = 0.2, geom = "polygon", level = 0.95)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  labs(title= "24h samples, ellipse by time")+
  scale_color_manual(values=treatment_colors)+
  #scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("24h_time.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

## Ordination of individual time points

### Stats on treatment only

```{r}
adonis2(MOBEAST_dist_12~MOBEAST_nmds_12$treatment, 
        # formula to calculate p-value for treatments
        strata = MOBEAST_nmds_12$tank_number) # accounts for repeated measures
```

#### Accounting for uneven variation in the data.

If the variation is uneven without the groups, adonis may come up with a significant result, although it's not. We can control for that using betdisper() and permutest()

```{r}
# beta dispertion
bd_12 <- betadisper(MOBEAST_dist_12, MOBEAST_nmds_12$treatment)

anova(bd_12) 
permutest(bd_12) 
bd_12

```
