---
title: "FCM Prelim Data"
author: "Keanu Rochette-Yu Tsuen"
date: "2025-02-16"
output: 
  html_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE , 
                      warning = FALSE)
```


## Load Libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)
library(ggpmisc)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(GGally)
library(emmeans)

```

## Data visualization theme
```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        panel.background = element_rect(fill = "azure1"))
```

## functions
```{r}
lm_eqn <- function(df){
    m <- lm(y ~ x, df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}
```


# Load Data
```{r}
FCM <- read_csv(here("data", "MOBEAST_FCM.csv"))
meta <- read_csv(here("data", "MOBEAST_metadata.csv"))

FCM<- FCM %>% clean_names()
meta <-meta %>% clean_names()
```

```{r}
treat_code <- tibble(
  treatment = c("A", "K", "C", "R", "Inflow", "Endmember", "Rain"),
  full_name = c("Algae", "Coral", "Control", "CCA", "Inflow", "Endmember", "Rain")
)

```


# Note
**Mislabeling**:  
- There were 2 batches of MOBEAST 2024-06-02 18:00 containing a total of 3 T4 samples. 
- MOBEAST 2024-06-02 21:00 is missing, one of the 18:00 is actually the 21:00 time point.
- One tube was missing a label (marked "X"). Assume it's the bay sample (2024-05-25).

For the preliminary results, I removed the duplicates for T4, only keeping 1 random T4 sample.  
I will be averaging the values of MOBEAST 2024-06-02 18:00. 
```{r}
FCM_pre <- FCM %>% filter(str_detect(sample, pattern = "MOBEAST")) %>% 
    mutate(sample = str_remove(sample, pattern = "X_")) %>% 
  filter(sample!= "2_MOBEAST_T4_A_2024-06-02_18:00", 
         sample!= "3_MOBEAST_T4_A_2024-06-02_18:00") %>% 
  separate_wider_delim(sample, names= c("project", "tank", "treatement", "date","time"), 
                             delim = "_") 


```

```{r}
FCM_pre<-FCM_pre %>% mutate(treatment = ifelse(treatement %in% treat_code$treatment, 
                                      treat_code$full_name, treatement)) %>% 
  relocate(treatment,.before = treatement) %>% 
  select(-treatement, -project)
```

```{r}
FCM_pre <- FCM_pre %>% mutate(date_time = ymd_hms(paste0(date," ", time,":00"))) %>% 
  select(tank, treatment, date_time, date, time, 
                   include_vol, 
                   het_bact = het_bact_events_u_l,
                   pico_euk = pico_euk_events_u_l,
                   syn = syn_events_u_l,
                   pro = pro_events_u_l,
                   tot_bact = bacteria_events_u_l) %>% 
  group_by(date_time, treatment, tank) %>% 
  mutate(cyano = sum(syn, pro, het_bact) - het_bact) %>%  
  ungroup()
```

```{r}
FCM_pre %>% pivot_longer(cols = het_bact:syn,
               names_to = "org", 
               values_to = "conc_ul") %>% 
  filter(str_detect(date, "06")) %>% 
  filter(treatment != "Endmember", treatment != "Rain" ) %>%
  ggplot(aes(date_time, conc_ul, color = treatment, group = org))+
  geom_point()+
  geom_line()+ 
  facet_wrap(.~treatment)
```


## BUOYS 

### 12PM 
```{r}
FCM_12PM <- FCM_pre %>% filter(time == "12:00")
```

```{r}
FCM_12PM_norm <- FCM_12PM %>% 
  select(-pro, -tot_bact, -cyano, -include_vol) %>% 
  pivot_longer(cols = het_bact:syn,
               names_to = "org", 
               values_to = "conc_ul") %>% 
   group_by(date_time, org) %>% 
  mutate(inflow_avg = mean(conc_ul[treatment == "Inflow"], na.rm = T)) %>%
  #calculating inflow average to normalize the data
  ungroup() %>%
   mutate(conc_ul_norm = ifelse(treatment %in% c("Coral", "Algae", "CCA", "Control"), conc_ul / inflow_avg, NA)) %>%  # Normalize for coral & algae
  filter(treatment != "Inflow") %>% # remove the inflow  
  mutate(treatment = factor(treatment,
                          levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) 

FCM_12PM_norm %>% 
  ggplot(aes(treatment, conc_ul_norm, group = treatment,fill= treatment))+
  geom_boxplot()+
  facet_grid(org~date_time, scales = "free")+
  nunu_theme
```

```{r}
FCM_12PM_norm %>% 
  ggplot(aes(date_time, conc_ul_norm, group = date_time,fill= treatment))+
  geom_boxplot()+
  facet_grid(org~treatment, scales = "free")+
  nunu_theme
```



#### Het bact 
```{r}
FCM_12PM_norm %>% filter(org == "het_bact") %>% 
  ggplot(aes(treatment, conc_ul_norm, group = treatment,fill= treatment))+
  geom_boxplot()+
  facet_grid(.~date_time)+
  nunu_theme
```

```{r}
FCM_12PM_norm_hb <- FCM_12PM_norm %>% filter(org == "het_bact") 
FCM_12_hb_lm<- lm(conc_ul_norm ~ date_time*treatment, data = FCM_12PM_norm_hb)

summary(FCM_12_hb_lm)

emmeans(FCM_12_hb_lm, specs = trt.vs.ctrl ~ treatment )
```


#### Synechococcus
```{r}
FCM_12PM_norm %>% filter(org == "syn") %>% 
  ggplot(aes(treatment, conc_ul_norm, group = treatment,fill= treatment))+
  geom_boxplot()+
  facet_grid(.~date_time)+
  nunu_theme
```

```{r}
FCM_12PM_norm_syn <- FCM_12PM_norm %>% filter(org == "syn") 
FCM_12_syn_lm<- lm(conc_ul_norm ~ date_time*treatment, 
                   data = FCM_12PM_norm_syn)

summary(FCM_12_syn_lm)

Anova(FCM_12_syn_lm)

emmeans(FCM_12_syn_lm, specs = trt.vs.ctrl ~ treatment )
```

#### Pico Euks
```{r}
FCM_12PM_norm %>% filter(org == "pico_euk") %>% 
  ggplot(aes(treatment, conc_ul_norm, group = treatment,fill= treatment))+
  geom_boxplot()+
  facet_grid(.~date_time)+
  nunu_theme
```

```{r}
FCM_12PM_norm_pe <- FCM_12PM_norm %>% filter(org == "pico_euk") 
FCM_12_pe_lm<- lm(conc_ul_norm ~ date_time*treatment, 
                   data = FCM_12PM_norm_pe)

summary(FCM_12_pe_lm)

Anova(FCM_12_pe_lm)

emmeans(FCM_12_pe_lm, specs = trt.vs.ctrl ~ treatment )
```



### 24h 
```{r}
FCM_24h <- FCM_pre %>%  filter(date != "2024-06-04", date != "2024-06-06") 
```

```{r}
FCM_24h_norm <- FCM_24h %>% 
  select(-pro, -tot_bact, -cyano, -include_vol) %>% 
  pivot_longer(cols = het_bact:syn,
               names_to = "org", 
               values_to = "conc_ul") %>% 
   group_by(date_time, org) %>% 
  mutate(inflow_avg = mean(conc_ul[treatment == "Inflow"], na.rm = T)) %>%
  #calculating inflow average to normalize the data
  ungroup() %>%
   mutate(conc_ul_norm = ifelse(treatment %in% c("Coral", "Algae", "CCA", "Control"), conc_ul / inflow_avg, NA)) %>%  # Normalize for coral & algae
  filter(treatment != "Inflow") %>% # remove the inflow  
  mutate(treatment = factor(treatment,
                          levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) %>%  
  filter(!str_detect(tank, "Lagoon")) %>% 
  filter(!str_detect(tank, "Ocean"))

FCM_24h_norm %>%  
  ggplot(aes(treatment, conc_ul_norm, group = treatment,fill= treatment))+
  geom_boxplot()+
  facet_wrap(.~date_time, scales = "free", nrow = 1)+
  nunu_theme
```

#### Het bact 
```{r}
FCM_24h_norm %>% filter(org == "het_bact") %>% 
  ggplot(aes(treatment, conc_ul_norm, group = treatment,fill= treatment))+
  geom_boxplot()+
  facet_grid(.~date_time)+
  nunu_theme
```
```{r}
FCM_24h_norm %>% filter(org == "het_bact") %>% 
  group_by(treatment, date_time) %>% 
  summarize(avg = mean(conc_ul_norm)) %>% 
  ggplot(aes(date_time, avg, group = treatment,color= treatment))+
  geom_point()+
  geom_line()+
 facet_grid(.~treatment)+
  nunu_theme
```

```{r}
FCM_24h_norm_hb <- FCM_24h_norm %>% filter(org == "het_bact") 
FCM_24_hb_lm<- lm(conc_ul_norm ~ date_time*treatment, data = FCM_24h_norm_hb)

summary(FCM_24_hb_lm)

emmeans(FCM_24_hb_lm, specs = trt.vs.ctrl ~ treatment )
```


#### Synechococcus
```{r}
FCM_24h_norm %>% filter(org == "syn") %>% 
  ggplot(aes(treatment, conc_ul_norm, group = treatment,fill= treatment))+
  geom_boxplot()+
  facet_grid(.~date_time)+
  nunu_theme
```

```{r}
FCM_24h_norm %>% filter(org == "syn") %>% 
  group_by(treatment, date_time) %>% 
  summarize(avg = mean(conc_ul_norm)) %>% 
  ggplot(aes(date_time, avg, group = treatment,color= treatment))+
  geom_point()+
  geom_line()+
 facet_grid(.~treatment)+
  nunu_theme
```

```{r}
FCM_24h_norm_syn <- FCM_24h_norm %>% filter(org == "syn") 
FCM_24_syn_lm<- lm(conc_ul_norm ~ date_time*treatment, 
                   data = FCM_24h_norm_syn)

summary(FCM_24_syn_lm)

emmeans(FCM_24_syn_lm, specs = trt.vs.ctrl ~ treatment )
```

#### Pico Euks
```{r}
FCM_24h_norm %>% filter(org == "pico_euk") %>% 
  ggplot(aes(treatment, conc_ul_norm, group = treatment,fill= treatment))+
  geom_boxplot()+
  facet_grid(.~date_time)+
  nunu_theme
```

```{r}
FCM_12PM_norm_pe <- FCM_12PM_norm %>% filter(org == "pico_euk") 
FCM_12_pe_lm<- lm(conc_ul_norm ~ date_time*treatment, 
                   data = FCM_12PM_norm_pe)

summary(FCM_12_pe_lm)

Anova(FCM_12_pe_lm)

emmeans(FCM_12_pe_lm, specs = trt.vs.ctrl ~ treatment )
```


```{r}
FCM_norm <- FCM_pre %>% 
  select(-pro, -tot_bact, -cyano, -include_vol) %>% 
  pivot_longer(cols = het_bact:syn,
               names_to = "org", 
               values_to = "conc_ul") %>% 
   group_by(date_time, org) %>% 
  mutate(inflow_avg = mean(conc_ul[treatment == "Inflow"], na.rm = T)) %>%
  #calculating inflow average to normalize the data
  ungroup() %>%
   mutate(conc_ul_norm = ifelse(treatment %in% c("Coral", "Algae", "CCA", "Control"), conc_ul / inflow_avg, NA)) %>%  # Normalize for coral & algae
  filter(treatment != "Inflow") %>% # remove the inflow  
  mutate(treatment = factor(treatment,
                          levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) %>% 
  select(-conc_ul, -inflow_avg) %>% 
  pivot_wider(names_from = org, 
              values_from = conc_ul_norm)
```

```{r}
DOC_norm<- DOC_norm %>% mutate(date_time=ymd_hms(date_time))
DOC_FCM_norm <- left_join(FCM_norm, DOC_norm, by= c("date_time", "treatment"))

```

```{r}
DOC_FCM_norm<- DOC_FCM_norm %>% select(-date.y, -time.y, -npoc_mg_l,-tn_mg_l,-tn_u_m) %>% 
  rename(date=date.x,time=time.x) %>% 
  mutate(treatment = factor(treatment, levels= c("Control", "Coral", "CCA", "Algae")))
```







```{r}
# syn
syn_doc_fcm <- DOC_FCM_norm %>% 
  mutate(treatment = factor(treatment,
                          levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment != "Control") %>% 
  ggplot(aes(x= npoc_norm, y = syn, color = treatment))+
  geom_point()+
  stat_poly_line() +
  stat_poly_eq(use_label(c("R2","p")),label.x = "left", label.y = "bottom") +
  #geom_smooth(method = "lm")+
  labs(title= "Synechococcus",
    x= "Normalized DOC",
       y= "Normalized Concentration")+
  facet_wrap(.~treatment)+
  nunu_theme+
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5)) +
  scale_color_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))
  
syn_doc_fcm 
ggsave(filename = "syn x doc.png",plot = last_plot(), path = here("outputs"), width = 12, height = 8)
```

```{r}
# het_bact
hb_doc_fcm <- DOC_FCM_norm %>% 
  mutate(treatment = factor(treatment,
                          levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment != "Control") %>% 
  ggplot(aes(x= npoc_norm, y = het_bact, color = treatment))+
  geom_point()+
  stat_poly_line() +
  stat_poly_eq(use_label(c("R2","p")), label.x = "right",
  label.y = "top") +
  #geom_smooth(method = "lm")+
  labs(title= "Heterotrophic Bacteria",
    x= "Normalized DOC",
       y= "Normalized Concentration")+
  facet_wrap(.~treatment)+
  nunu_theme+
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5)) +
  scale_color_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))
hb_doc_fcm 
#dggsave(filename = "hb x doc.png",plot = last_plot(), path = here("outputs"), width = 12, height = 8)
```

```{r}
# pico euk
pe_doc_fcm <- DOC_FCM_norm %>% 
  mutate(treatment = factor(treatment,
                          levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment != "Control") %>% 
  ggplot(aes(x= npoc_norm, y = pico_euk, color = treatment))+
  geom_point()+
  stat_poly_line() +
  stat_poly_eq(use_label(c("R2","p")),label.x = "right",
  label.y = "top") +
  #geom_smooth(method = "lm")+
  labs(title= "Picoeukaryotes",
    x= "Normalized DOC",
       y= "Normalized Concentration")+
  facet_wrap(.~treatment)+
  nunu_theme+
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5)) +
  scale_color_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))
pe_doc_fcm
#ggsave(filename = "pe x doc.png",plot = last_plot(), path = here("outputs"), width = 12, height = 8)
```

```{r}
syn_doc_fcm/ hb_doc_fcm / pe_doc_fcm

#ggsave(filename = "fcm x doc.png",plot = last_plot(), path = here("outputs"), width = 12, height = 8)
```
### Linear model FCM DOC
```{r}
syn_lm <- lm(syn~npoc_norm*treatment, data= DOC_FCM_norm)
emmeans(syn_lm, specs = trt.vs.ctrl ~ treatment)
aov(syn_lm)

#plot(allEffects(syn_lm))
#resid_panel(syn_lm, plots = c("resid", "qq", "lev", "hist"))
```

