---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: MOBEAST DOC verification
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE , 
                      warning = FALSE)
```

## Load Libraries 
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(GGally)
library(emmeans)
library(ggrepel)
```

## Load function
```{r}
source(here("functions", "function_ANOVA_nunu.R"))
```

## Load Data
```{r}
meta <- read_delim(here("data", "MOBEAST_metadata.csv"))
DOC <- read_delim(here("data", "raw_data", "DOC", "MOBEAST_NPOC_TN_241010.csv"))
```

## Codes of the treatment  
```{r}
treat_code <- tibble(
  treatment = c("A", "K", "C", "R", "Inflow", "Endmember", "Rain"),
  full_name = c("Algae", "Coral", "Control", "CCA", "Inflow", "Endmember", "Rain")
)
treat_code
#treat_table <- treat_code %>% gt() %>% 
#  cols_label(treatment = "Treatment Code",
#             full_name = "Treatment Name")
#treat_table
```

## Data visualization theme
```{r}

nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        panel.background = element_rect(fill = "azure1"))
```

## Data clean up 
```{r}
meta <- meta %>% clean_names() %>% drop_na() %>%
  mutate(long_name = str_trim(long_name)) %>% 
  filter(long_name != "Long Name") %>% 
  mutate(date = mdy(date), 
         time = format(strptime(time, format = "%H%M"), "%H:%M"),
         id_number = str_trim(id_number))  
```

```{r}
DOC <- DOC %>% clean_names() 
```

 ## Joining the data
```{r}
DOC_join <- left_join(meta, DOC, by = "id_number")
DOC_join <- left_join(DOC_join, treat_code, by = "treatment")
DOC_join <- DOC_join %>% select(-treatment) %>%  rename(treatment = full_name) %>%  relocate(treatment, .after = tank_number)
```

## Check data for NPOC (mg/L)

The following points have a value different than the other data points: 

- 2024-06-02 18:00: **T8** and **T11**. These are the same points as the ones Laurel identified.  
- 2024-06-03 00:00: **F1, F2, T2, T5, T3** and **T13**. However, that's a lot of points so they might not be outliers.  

```{r}
DOC_join %>% drop_na(npoc_mg_l) %>% filter(treatment != "Endmember") %>% 
  ggplot(aes(x= treatment, y = npoc_mg_l, color = as.factor(time))) +
  geom_point() + 
  geom_text_repel(aes(label = tank_number)) +
  geom_point(data=DOC_join %>% filter(npoc_mg_l > 3.5), 
           pch=21, 
           size=4, 
           colour="red") +
  facet_wrap(.~date)+
  theme_bw()+
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle = 45, vjust = 0.5),
        panel.background = element_rect(fill = "azure1"))+
   scale_color_ghibli_d("SpiritedMedium", direction=-1)
  
  
  
```


## Check data for TN (mg/L)

For TN, **T11** at 2024-06-03 6:00 seems to be a clear outlier. The tanks that were identified as outliers in the NPOC plot above do not appear on the TN plot.  
```{r}
DOC_join %>% drop_na(tn_mg_l) %>% filter(treatment != "Endmember") %>% 
  ggplot(aes(x= treatment, y = tn_mg_l, color = as.factor(time))) +
  geom_point() + 
  geom_text_repel(aes(label = tank_number)) +
  geom_point(data=DOC_join %>% filter(tn_mg_l > 0.9), 
           pch=21, 
           size=4, 
           colour="red") +
  facet_wrap(.~date)+
  theme_bw()+
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle = 45, vjust = 0.5),
        panel.background = element_rect(fill = "azure1"))+
   scale_color_ghibli_d("SpiritedMedium", direction=-1)
```


## Overall Stats
```{r}
DOC_join %>% filter(treatment != "Endmember", treatment != "Rain" ) %>%
  group_by(date, time, treatment) %>% 
  summarise(avg = mean(npoc_mg_l, na.rm = T),
            std_dev = sd(npoc_mg_l, na.rm = T))
```


## TN

```{r}
DOC_join
```


```{r}
TN_norm <- DOC_join %>% filter(date > as.Date("2024-05-31")) %>% 
  mutate(time= paste0(time,":00")) %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
   group_by(date_time) %>%
  mutate(inflow_avg = mean(tn_u_m[treatment == "Inflow"], na.rm = T)) %>%  # Calculate inflow avg
  ungroup() %>%
  mutate(tn_norm = ifelse(treatment %in% c("Coral", "Algae", "CCA", "Control"), tn_u_m / inflow_avg, NA))  # Normalize for coral & algae
  

```


```{r, fig.height=8, fig.width=12}
TN_norm %>% filter(time == "12:00:00") %>% 
mutate(treatment = factor(treatment, levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) %>% 
  ggplot(aes(treatment, tn_norm, fill = treatment)) +
  geom_boxplot() + facet_wrap(.~date_time)+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number))+
  labs(title = "12PM variation in normalized TN",
       x= "Treatment",
       y = "Normalized TN")+
    #ylim(0.8, 4)+
  #scale_y_continuous(sec.axis = sec_axis(~.*100, name="Average inflow")) +
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))

#ggsave(filename = "TN_norm_12PM.png",plot = last_plot(), path = here("outputs"), width = 12, height = 8)
```


## 12PM linear model 
```{r}
TN_treat<- TN_norm %>% filter(time == "12:00:00") %>% 
mutate(treatment = factor(treatment, levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) 


TN_norm_lm <- lm(tn_norm ~ date_time*treatment, data = TN_treat)

summary(TN_norm_lm)

resid_panel(TN_norm_lm)

emmeans(TN_norm_lm, specs = trt.vs.ctrl ~ treatment)
```

```{r}
TN_norm_lm <- lm(tn_norm ~ treatment, data = TN_treat)
Anova(TN_norm_lm)

```


```{r, fig.height=6, fig.width=12}
TN_norm %>% #filter(time != "00:00:00") %>% 
  filter(date != "2024-06-04", date != "2024-06-06") %>% 
mutate(treatment = factor(treatment, levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) %>% 
  #filter(tn_norm < 7) %>% 
  ggplot(aes(treatment, tn_norm, fill = treatment)) +
  geom_boxplot() + 
  facet_wrap(.~ date_time, nrow = 1)+
  geom_jitter(aes(color = treatment), alpha= 0.5, size = 3)+
  geom_text_repel(aes(label = tank_number))+
  labs(title = "24h variation in normalized TN",
       x= "Treatment",
       y = "Normalized TN")+
  #  ylim(0.8, 4)+
  #scale_y_continuous(sec.axis = sec_axis(~.*100, name="Average inflow")) +
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
  

#ggsave(filename = "TN_norm_24h.png",plot = last_plot(), path = here("outputs"), width = 12, height = 8)
```

## 24h linear model

```{r}
TN_24h <- TN_norm %>% #filter(time != "00:00:00") %>% 
  filter(date != "2024-06-04", date != "2024-06-06") %>% 
  mutate(treatment = factor(treatment, levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  #mutate(date_time = date_time) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) #%>% 
  #filter(tn_norm<7) 



TN_norm_lm <- lm(tn_norm ~ treatment*date_time, data = TN_24h)

summary(TN_norm_lm)
resid_panel(TN_norm_lm)

emmeans(TN_norm_lm, specs = trt.vs.ctrl ~ treatment)
```

```{r}
Anova(TN_norm_lm)
```








# BUOYS figures
## Preliminary visualization

Let's see how the DOC data behaves. From the graph below, we can see some intriging peaks in the different tanks. This may be off concern later. But also, as of 18 feb 2025, we are also stil waiting for some samples to be re-run by the S-LAB. 
```{r}
DOC_join %>% filter(date > as.Date("2024-05-31")) %>% 
  mutate(time= paste0(time,":00")) %>% 
  #filter(time != "00:00:00") %>% 
  mutate(date_time = ymd_hms(paste(date, time))) %>% 
  group_by(treatment, date_time) %>% 
  summarise(avg = mean(npoc_u_m, na.rm = T)) %>% 
  filter(avg <250) %>% 
  ggplot(aes(date_time, avg, color = treatment)) +
  geom_point(size= 3)+
  geom_line()+ nunu_theme
```

For now, we will exclude the data points that seem like outliers. From conversations with Craig, I will also just remove the 00:00 time point for BUOYS because it's created confusion due to the outliers. We'll simplify my analysis that way. 
```{r}
DOC_join %>% filter(date > as.Date("2024-05-31")) %>% 
  mutate(time= paste0(time,":00")) %>% 
  filter(time != "00:00:00") %>% 
  mutate(date_time = ymd_hms(paste(date, time))) %>% 
  group_by(treatment, date_time) %>% 
  summarise(avg = mean(npoc_u_m, na.rm = T)) %>% 
  filter(avg <250) %>% 
  ggplot(aes(date_time, avg, color = treatment)) +
  geom_point(size= 3)+
  geom_line()+ nunu_theme
```

```{r}
DOC_norm <- DOC_join %>% filter(date > as.Date("2024-05-31")) %>% 
  mutate(time= paste0(time,":00")) %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
   group_by(date_time) %>%
  mutate(inflow_avg = mean(npoc_u_m[treatment == "Inflow"], na.rm = T)) %>%  # Calculate inflow avg
  ungroup() %>%
  mutate(npoc_norm = ifelse(treatment %in% c("Coral", "Algae", "CCA", "Control"), npoc_u_m / inflow_avg, NA))  # Normalize for coral & algae
  

```


```{r, fig.height=8, fig.width=12}
DOC_norm %>% filter(time == "12:00:00") %>% 
mutate(treatment = factor(treatment, levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) %>% 
  ggplot(aes(treatment, npoc_norm, fill = treatment)) +
  geom_boxplot() + facet_wrap(.~date_time)+
  #geom_jitter(aes(color = treatment), alpha = 0.5)+
  labs(title = "12PM variation in normalized DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  #scale_y_continuous(sec.axis = sec_axis(~.*100, name="Average inflow")) +
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))

#ggsave(filename = "DOC_norm_12PM.png",plot = last_plot(), path = here("outputs"), width = 12, height = 8)
```



## 12PM linear model 
```{r}
DOC_treat<- DOC_norm %>% filter(time == "12:00:00") %>% 
mutate(treatment = factor(treatment, levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) 


DOC_norm_lm <- lm(npoc_norm ~ date_time*treatment, data = DOC_treat)

summary(DOC_norm_lm)

emmeans(DOC_norm_lm, specs = trt.vs.ctrl ~ treatment)
```

```{r}
DOC_norm_lm <- lm(npoc_norm ~ treatment, data = DOC_treat)
Anova(DOC_norm_lm)
```

```{r}
anova_nunu(data = DOC_treat, indep_var = treatment, dep_var = npoc_norm)
```


```{r}
#resid_panel(DOC_norm_lm, plots = c("resid", "qq", "lev", "hist"))
```

```{r}
plot(allEffects(DOC_norm_lm))
```

```{r, fig.height=6, fig.width=12}
DOC_norm %>% filter(time != "00:00:00") %>% 
  filter(date != "2024-06-04", date != "2024-06-06") %>% 
mutate(treatment = factor(treatment, levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) %>% 
  filter(npoc_norm<3.5) %>% 
  ggplot(aes(treatment, npoc_norm, fill = treatment)) +
  geom_boxplot() + 
  facet_wrap(.~ date_time, nrow = 1)+
  #geom_jitter(aes(color = treatment), alpha= 0.5, size = 3)+
  labs(title = "24h variation in normalized DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  #scale_y_continuous(sec.axis = sec_axis(~.*100, name="Average inflow")) +
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
  

#ggsave(filename = "DOC_norm_24h.png",plot = last_plot(), path = here("outputs"), width = 12, height = 8)
```

## 24h linear model

```{r}
DOC_24h <- DOC_norm %>% filter(time != "00:00:00") %>% 
  filter(date != "2024-06-04", date != "2024-06-06") %>% 
  mutate(treatment = factor(treatment, levels= c("Control", "Coral", "CCA", "Algae"))) %>% 
  mutate(date_time = date_time) %>% 
  filter(treatment %in% c("Coral", "Algae", "CCA", "Control")) %>% 
  filter(npoc_norm<3.5) 


DOC_norm_lm <- lm(npoc_norm ~ treatment*date_time, data = DOC_24h)

summary(DOC_norm_lm)

emmeans(DOC_norm_lm, specs = trt.vs.ctrl ~ treatment)
```

```{r}
Anova(DOC_norm_lm)
```

