---
title: "MOBEAST FCM V3"
author: "Keanu Rochette"
created: '11 November 2025'
updated: '`r format(Sys.Date())`'
format: 
  html:
    toc: true
    toc-location: left
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)
#library(ggpmisc)
library(hms)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
#library(GGally)
library(emmeans)
library(lme4)
library(lmerTest)
library(sjPlot)
library(performance)

```

## Data visualization theme

```{r}
nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"))
```

## Functions

```         
lm_eqn <- function(df){
    m <- lm(y ~ x, df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}
```

# Load Data
```{r}
FCM <- read_csv(here("data", "MOBEAST_FCM.csv"))
meta <- read_csv(here("data", "MOBEAST_metadata.csv"))
# can't use the good metadata V2 bcs we sampled FCM more frequently than the other variable
# we are missing some id numbers for the additional samples...

FCM<- FCM %>% clean_names()
meta <-meta %>% clean_names()
```

```{r}
treat_code <- tibble(
  treatment = c("A", "K", "C", "R", "Inflow", "Endmember", "Rain"),
  full_name = c("Algae", "Coral", "Control", "CCA", "Inflow", "Endmember", "Rain")
)

```

```{r}
table_id <- read_csv(here("data", "Laurel", "TableID.csv"))
table_id <- clean_names(table_id)

table_id <- table_id %>% rename(tank_number=tankid) %>% select(-treatment) %>% 
  mutate(tank_number = paste0("T", tank_number)) %>% 
  mutate(inflow_table = paste("Table", inflow_table))
```

# Data Cleaning 

Note

**Mislabeling**:\
- There were 2 batches of MOBEAST 2024-06-02 18:00 containing a total of 3 T4 samples. 
- MOBEAST 2024-06-02 21:00 is missing, one of the 18:00 is actually the 21:00 time point. 
- One tube was missing a label (marked "X"). Assume it's the bay sample (2024-05-25).

For the preliminary results, I removed the duplicates for T4, only keeping 1 random T4 sample.\
I will be averaging the values of MOBEAST 2024-06-02 18:00.

```{r}
FCM_pre <- FCM %>% filter(str_detect(sample, pattern = "MOBEAST")) %>% 
    mutate(sample = str_remove(sample, pattern = "X_")) %>% 
  filter(sample!= "2_MOBEAST_T4_A_2024-06-02_18:00", 
         sample!= "3_MOBEAST_T4_A_2024-06-02_18:00") %>% 
  separate_wider_delim(sample, names= c("project", "tank", "treatment", "date","time"), 
                             delim = "_") 


```

```{r}
FCM_pre <- left_join(FCM_pre, treat_code, by= "treatment")

FCM_pre <- FCM_pre %>% relocate(full_name,.before = treatment) %>%
  select(-treatment) %>% 
  rename(treatment = full_name) %>% 
  rename(tank_number = tank)
```

```{r}
FCM_pre <- FCM_pre %>% left_join(table_id, by = "tank_number") %>% 
  mutate(inflow_table = ifelse(tank_number == "F1", "Table 1",
                               ifelse(tank_number == "F2", "Table 2", inflow_table))) %>% 
  mutate(inflow_table = ifelse(is.na(inflow_table), "Endmember", inflow_table))
```


```{r}
FCM_pre <- FCM_pre %>% 
  
  ## 18:00 is duplicated because one batch is actually 21:00
  ## there is also multiple samples for 1 tank (T4?)
  ## removed for simplicity (Nov 11 2025)
  filter(time != "18:00") %>% 
  mutate(date_time = ymd_hms(paste0(date," ", time,":00"))) %>% 
  select(tank_number, treatment, inflow_table, date_time, date, time, 
                   het_bact = het_bact_events_u_l,
                   pico_euk = pico_euk_events_u_l,
                   syn = syn_events_u_l,
                   pro = pro_events_u_l,
                   tot_bact = bacteria_events_u_l) %>% 
  group_by(date_time, treatment, tank_number) %>% 
  mutate(cyano = sum(syn, pro, het_bact) - het_bact) %>%  
  ungroup() %>% 
  mutate(date =  ymd(date),
         time = paste0(time,":00"),
         time = as_hms(time)) 
```


## Outlier detection 

```{r}
raw_flagged_outliers <- FCM_pre %>% 
  select(-pro, -tot_bact, -cyano) %>% 
  pivot_longer(cols = het_bact:syn,
               names_to = "org", 
               values_to = "conc_ul") %>% 
  filter(treatment != "Endmember", treatment != "Rain") %>% 
  # flagging outliers
  ## grouping by date, time and treatment
  group_by(org, treatment) %>% 
  ## calculating the Interquantile range and of the raw data

  mutate(quant1 = quantile(conc_ul, probs = 0.25, na.rm = T),
         quant3 = quantile(conc_ul, probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse(conc_ul > quant3 + (IQR*1.5) | conc_ul < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup()  
```


```{r}
# Normalization version 2
## averaging the inflows for normalization 
## remove the outliers a right before the normalization 

## Normalization of the data to the inflow 
 
FCM_norm <- raw_flagged_outliers %>% 
  mutate(time_string = as.character(time),
         date_string = as.character(date),
         date_time_string = paste(date_string, time_string)) %>% 
  select(-IQR, -quant1, -quant3) %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table, org) %>%
  mutate(inflow = conc_ul[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time, org) %>% 
  mutate(inflow = ifelse(outlier == "outlier", inflow == NA, inflow)) %>% 
  mutate(inflow = mean(inflow, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(conc_norm = conc_ul - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```

```{r}
# Dataset sent to Craig 
FCM_craig <- FCM_norm %>% filter(outlier != "outlier") %>% 
  select(-inflow)

write_csv(FCM_craig, here("data","craig_data", "fcm_Craig.csv"))
```


## 12PM

```{r}
FCM_12PM <- FCM_norm %>% filter(time_string == "12:00:00") %>% 
  filter(outlier != "outlier") %>% 
  filter(treatment != "Inflow")
```

## Iterative statistical modeling

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

FCM_loop <- FCM_12PM %>% select(org) %>% distinct() %>% pull()

models_12PM <- vector(mode = "list", length = length(FCM_loop))
names(models_12PM) <- FCM_loop

for (var in FCM_loop) {
  df <- FCM_12PM %>% filter(org == var)
  models_12PM[[var]] <- lmer(conc_norm ~ treatment + (1|tank_number), data = df)
}

```

## Iterative graphing

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## --> In the for-loop, filter the df by nutrient and make a graph that will display nutrient data
## Save the outputs of the model in the list such that each nutrient has its stats

FCM_loop <- FCM_12PM %>% select(org) %>% distinct() %>% pull()

graphs_12PM <- vector(mode = "list", length = length(FCM_loop))
names(graphs_12PM) <- FCM_loop

for (var in FCM_loop) {
  
  df <- FCM_12PM %>% filter(org == var)
  
  graphs_12PM[[var]] <- df %>%  
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("12PM", var, "concentrations"),
       x= "Treatment",
       y = "Concentration (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none")
}

```


### Heterotrophic Bacteria

```{r}
summary(models_12PM[["het_bact"]])
anova(models_12PM[["het_bact"]], type = 2)
emmeans(models_12PM[["het_bact"]], specs = trt.vs.ctrl ~ treatment)
```

```{r, fig.height=6, fig.width=12}
resid_panel(models_12PM[["het_bact"]])
graphs_12PM[["het_bact"]]
  
ggsave("het_bact_12PM.png",width = 16, height = 8, path = here("outputs"))
```

### t-test 
```{r}
FCM_12PM %>%
  filter(org == "het_bact") %>% 
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )
```


### Synechococcus 

```{r}
summary(models_12PM[["syn"]])
anova(models_12PM[["syn"]], type = 2)
emmeans(models_12PM[["syn"]], specs = trt.vs.ctrl ~ treatment)
```

```{r}
resid_panel(models_12PM[["syn"]])
graphs_12PM[["syn"]]
```


### Pico Eukaryotes 

```{r}
summary(models_12PM[["pico_euk"]])
anova(models_12PM[["pico_euk"]], type = 2)
emmeans(models_12PM[["pico_euk"]], specs = trt.vs.ctrl ~ treatment)
```

```{r}
resid_panel(models_12PM[["pico_euk"]])
graphs_12PM[["pico_euk"]]
```



# 24h

```{r}
# Version 2
## removing outliers from the data set
FCM_24h <- FCM_norm %>%
  filter(date_string == "2024-06-02" | date_string == "2024-06-03" ) %>% 
  #filter(time_string == "12:00:00"|time_string == "00:00:00"| time_string == "06:00:00") %>%
  filter(treatment != "Inflow") %>% 
  
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "15:00:00","18:00:00",
                                                  "21:00:00", "00:00:00", "03:00:00",
                                                  "06:00:00", "09:00:00")))
  # no 18h and 21h due to mixed up samples

```





## Iterative statistical modeling

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

FCM_loop <- FCM_24h %>% select(org) %>% distinct() %>% pull()

models_24h <- vector(mode = "list", length = length(FCM_loop))
names(models_24h) <- FCM_loop

for (var in FCM_loop) {
  df <- FCM_24h %>% filter(org == var)
  models_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + (1|tank_number), data = df)
}

```

## Iterative graphing

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## --> In the for-loop, filter the df by nutrient and make a graph that will display nutrient data
## Save the outputs of the model in the list such that each nutrient has its stats

FCM_loop <- FCM_24h %>% select(org) %>% distinct() %>% pull()

graphs_24h <- vector(mode = "list", length = length(FCM_loop))
names(graphs_24h) <- FCM_loop

for (var in FCM_loop) {
  
  df <- FCM_24h %>% filter(org == var)
  
  graphs_24h[[var]] <- df %>%  
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("24h", var, "concentrations"),
       x= "Treatment",
       y = "Concentration (µmol/L)")+
    facet_wrap(~time_fac, ncol = 3 )+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none")
}

```

### Heterotrophic Bacteria
```{r}
summary(models_24h[["het_bact"]])
anova(models_24h[["het_bact"]], type = 2)
emmeans(models_24h[["het_bact"]], specs = trt.vs.ctrl ~ treatment)
```


```{r, fig.width=12, fig.height=6}
resid_panel(models_24h[["het_bact"]])
graphs_24h[["het_bact"]]
```

#### Selected time to display 

```{r}
df <- FCM_24h %>% filter(org == "het_bact")%>% 
  filter(time_string == "12:00:00"| time_string == "00:00:00"| time_string == "06:00:00")

mod <- lmer(conc_norm ~ treatment*time_fac + (1|tank_number), data = df)
```

```{r}
summary(mod)
anova(mod, type = 2)
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```

```{r}
FCM_24h %>% filter(org == "het_bact") %>% 
  filter(time_string == "12:00:00"| time_string == "00:00:00"| time_string == "06:00:00") %>% 
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("24h Bacterial Abundance"),
       x= "Treatment",
       y = "Concentration (ind/µL)")+
    facet_wrap(~time_fac, ncol = 3 )+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none")

ggsave("het_bact_24h_2.png",width = 16, height = 8, path = here("outputs"))
```


#### T-test
```{r}

FCM_24h %>%
  filter(org == "het_bact") %>% 
  filter(time_string == "12:00:00"| time_string == "00:00:00"| time_string == "06:00:00") %>%
  mutate(time_fac = fct_rev(time_fac)) %>% 
  #filter(time_string == "12:00:00"| time_string == "00:00:00"| time_string == "06:00:00") %>%
  group_by(time_fac, treatment ) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value) %>% 
   mutate(significance = ifelse(p_value <0.05, "Enrichment", "No Difference"),
         significance = ifelse(mean <0 & p_value <0.05, "Depletion", significance)#,
         #significance = as.character(significance)
         ) %>% 
    ungroup() %>% 
  
  ggplot(aes(x = treatment, y= time_fac, fill = significance))+
  geom_tile(color = "white", size = 1)+
  #facet_wrap(.~time_fac, nrow = 1)+
  labs(title = " 24h DOC Variations",
       x= "Treatment",
       y = "Sample Collection Time")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw()+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")

#ggsave("fdom_tiles_24h_t_test.png",width = 16, height = 8, path = here("outputs"))
```

#### Pairwise

```{r}
nunu_pairwise_FCM <- function(data, time_1 = "12:00:00", time_2 = "00:00:00"){

  ## Get a list of the treatments for the for loop   
  treatment <- data %>% mutate(treatment = as.character(treatment)) %>%
  select(treatment) %>% distinct() %>% pull()
  
  ## Create un empty data frame
pairwise_results <- tibble(
  time_point = vector(),
  treatment= vector(),
  mean = vector(),
  p_value = vector())

# this is the pairwise comparison for the 6 pairs (6AM-6PM)
    for(var in treatment){
      
      ## only keep the 6AM and 6PM
      df <- data %>% 
        mutate(treatment = factor(treatment, 
                                  levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
        filter(org == "het_bact") %>% 
        filter(treatment == var) %>% 
  filter(time_string == time_1 | time_string == time_2) %>% 
        ## Rename the times so they can become column names
  mutate(time_US = ifelse(time_string == time_1, "T01", "T02")) %>% 
        ## Only keep the column of interest for the stats
  select(tank_number, time_US, conc_norm) %>% 
  pivot_wider(names_from = time_US, values_from = conc_norm) 
      
      ## Pairwise function 
      paired <- t.test(df$T01, df$T02, paired = TRUE, na.action = na.pass)
      
      ## Save the result in at data frame
      df2 <- tibble(
        time_point = paste0(time_1, "-", time_2),
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      # Code will loop oveer all treatments and save the results in a data frame
      pairwise_results <-rbind(pairwise_results,df2)
      
      # Save the results in RStudio
      pairwise_results <<- pairwise_results
    }  
print(pairwise_results)
}
```

```{r}
pairwise_heatmap <- nunu_pairwise_FCM(data = FCM_24h, time_1 = "12:00:00", time_2 = "00:00:00")

pairwise_heatmap 
```


```{r}
nunu_pairwise_FCM(data = FCM_24h, time_1 = "03:00:00", time_2 = "15:00:00")

pairwise_heatmap <- rbind(pairwise_heatmap, nunu_pairwise_FCM(data = FCM_24h, time_1 = "03:00:00", time_2 = "15:00:00"))
```


```{r}
pairwise_heatmap %>%
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  mutate(significance = ifelse(p_value <0.05, "Diel Variations", "No Variations")#,
         #significance = as.character(significance)
         ) %>% 
    ungroup() %>% 
  
  ggplot(aes(x = treatment, y= time_point, fill = significance))+
  #geom_tile(aes(color = fdom_cat), size = 1)+
  geom_tile(color = "white", size = 1)+
  labs(title= "FCM Diel Variations", 
       x= "Treatment",
       y = "fDOM Type")+
  scale_fill_manual(breaks = c("No Variations", "Diel Variations"),
                    values = c( "#90CAF9", "#43A047"))+
  theme_bw()+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")

ggsave("fdom_tiles_24h_pairwise.png",width = 10, height = 8, path = here("outputs"))

```



### Synechococcus
```{r}
summary(models_24h[["syn"]])
anova(models_24h[["syn"]], type = 2)
emmeans(models_24h[["syn"]], specs = trt.vs.ctrl ~ treatment)
```

```{r}
resid_panel(models_24h[["syn"]])
graphs_24h[["syn"]]
```


### Pico Eukaryotes
```{r}
summary(models_24h[["pico_euk"]])
anova(models_24h[["pico_euk"]], type = 2)
emmeans(models_24h[["pico_euk"]], specs = trt.vs.ctrl ~ treatment)
```

```{r}
resid_panel(models_24h[["pico_euk"]])
graphs_24h[["pico_euk"]]
```

