---
title: "MOBEAST- DNA Ordination 24h V2"
format: html
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

# Load Libraries

```{r}
library(tidyverse)
#library(lubridate)
library(here)
library(janitor)
#library(ggrepel)
#library(ghibli)
#library(gt)
library(patchwork)
#library(RColorBrewer)
library(vegan)
library(OmicFlow)
library(phyloseq)
#library(GUniFrac)
```

# Data visualization

```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

```{r}
nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=24, face = "bold"), 
        plot.subtitle = element_text(size=20),
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=18, face = "bold"),
        legend.text=element_text(size=16),
        strip.background = element_rect(fill = "white"))
```

# Load Data

```{r}
unifrac <- read.csv(here("data", "DNA", "unifrac", "unifrac_weighted_97.csv"), row.names = 1,
                    header = TRUE)

meta <- read_csv(here("data","MOBEAST_metadata_V2.csv"))

```

# Data Clean Up

```{r}
endmember_list <- meta %>% 
  filter(treatment == "Endmember" |treatment == "Rain") %>% 
  pull(id_number)

inflow_list <- meta %>% 
  filter(treatment == "Inflow") %>% 
  pull(id_number)

sample_list <- meta %>% 
  filter(!id_number %in% endmember_list) %>% 
  filter(!id_number %in% inflow_list) %>% 
  pull(id_number)
```

```{r}
unifrac_subset <- unifrac[rownames(unifrac) %in% sample_list, colnames(unifrac) %in% sample_list]

unifrac_dist <- as.dist(unifrac_subset)
```



```{}
cluster<- cluster %>% 
  #mutate(cluster = factor(cluster)) %>% 
  mutate(cluster2 = ifelse(cluster == 1, "Algae", 
                           ifelse(cluster == 2, "CCA", 
                                  ifelse(cluster == 3, "Coral", "Control")))) %>% 
  select(-cluster) %>% 
  rename(cluster = cluster2)

meta<- meta %>% left_join(cluster)
```



# NMDS Calculation

```{r}
nmds <- metaMDS(unifrac_dist, k=2, trymax = 1000)
# put stress in the corner fo the nmds plot
```

```{r}
nmds_scores <- as.data.frame(scores(nmds, display = "site"))
```

```{r}
nmds_scores$id_number= rownames(nmds_scores)
```

```{r}
nmds_merged <- merge(nmds_scores, meta, by.x ="id_number", by.y = "id_number")
```

```{r}
nmds_merged<- nmds_merged %>% mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```


# Plotting Ordination 


```{r, fig.width=16, fig.height= 6}
# Plot
nmds_merged %>%   
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 4)+
  labs(title= "24h samples, ellipse by treatment")+
  scale_color_manual(values=treatment_colors )+
  scale_fill_manual(values= treatment_colors) +
  nunu_theme
#ggsave("24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))

```
## PERMANOVA 

```{r}
# Note : Version 2 of adonis 
perma_24h <- adonis2(unifrac_dist ~ treatment + date_time + treatment:date_time, 
                     data = nmds_merged,
                     #permutations = 1000, 
                     strata = nmds_merged$tank_number)
perma_24h

## Random effect (strata) should be applied on the tank because we sample repeatedly in the tank.
## We don't sample the same tank multiple times at a time point. 
## We may have a time interaction tho, so I added the date_time in the model formula. 
### Note: why is the output of the model not show fixed effects and random effects like lmer()? 


```

## Pairwise comparison 

- permnova : adonis2
- pairwise pairwise.adonis2(p.adjust.m = "BH"), make sure that it's a distance matrix (as.dist())

```{r}
pairwise_adonis(x = unifrac_dist, 
                groups = as.vector(nmds_merged$treatment),
                  p.adjust.m = "BH")
```



# Analyzing data with Heatmap 

## Identifying the 2 samples that cluster alone in the Heatmaps 
```{r}
MOBEAST_nmds_24h_bis <- MOBEAST_nmds_24h %>% 
  mutate(treatment_bis = treatment) %>% 
  mutate(treatment_bis = ifelse(id_number == "MOBEAST010" |id_number == "MOBEAST060", "weird", treatment_bis ))
```

```{r, fig.width=16, fig.height= 6}
# Plot
MOBEAST_nmds_24h_bis %>%  
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment_bis, fill= treatment_bis))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 4)+
  labs(title= "24h samples, ellipse by treatment")+
  scale_color_manual(values= c("Inflow"="#00838F", "Control"="grey", "Coral"="#03A9F4", "CCA"="#EF5350", "Algae"="#4CAF50","weird"= "yellow"))+
  scale_fill_manual(values= c("Inflow"="#00838F", "Control"="grey", "Coral"="#03A9F4", "CCA"="#EF5350", "Algae"="#4CAF50","weird"= "yellow")) +
  #nunu_theme+
  theme_bw()
#ggsave("24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))

```

#### Plotting by heatmap cluster

Note: cluster numbers are inaccurate and do not match what is seen in the heatmap. Problem with the distance matrix or clutering method or both...

```{r, fig.width=16, fig.height= 10}
# Plot
MOBEAST_nmds_24h %>%   
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  #stat_ellipse(aes(colour = cluster, fill = cluster),
               #alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 2)+
  labs(title= "24h samples, ellipse by treatment")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  ggforce::geom_mark_ellipse(aes(color = cluster, fill = cluster), alpha = 0.3)+
  #nunu_theme+
  theme_bw()
#ggsave("24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

```{r, fig.width=10, fig.height= 6}
# raw NMDS plot with all the samples
MOBEAST_nmds_24h %>%  
  mutate(time_string = factor(time_string, 
                              levels = c("12:00:00","18:00:00","00:00:00","06:00:00" ))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment))+
  stat_ellipse(aes(colour = time_string, fill = time_string), 
               alpha = 0.2, geom = "polygon", level = 0.95)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  labs(title= "24h samples, ellipse by time")+
  scale_color_manual(values=treatment_colors)+
  #scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("24h_time.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

### Stats on treatment only

```{r}
# Tranform the distance matrix into a df that I can use adonis on 
id_number_24h <- meta %>% filter(date_string == "2024-06-02" |date_string == "2024-06-03") %>% pull(id_number)

distance_24h <- MOBEAST_dist_24h %>% as.matrix() %>% as.data.frame() %>% 
  mutate(id_number = id_number_24h)%>% 
  relocate(id_number, .before = MOBEAST001)

rownames(distance_24h) <- NULL
```

```{r}
meta_distance_24h <- inner_join(meta, distance_24h, by = "id_number") %>% 
  select(-long_name, -date_string, -time_string, -date_time_string,-cluster)
meta_distance_24h 

all_dist <- meta_distance_24h %>% 
  select(all_of(meta_distance_24h[["id_number"]])) %>% 
  as.dist()
```

```{r}
# Note : Version 2 of adonis 
perma_24h <- adonis2(all_dist ~ treatment + date_time + treatment:date_time, 
                     data = meta_distance_24h,
                     #permutations = 1000, 
                     strata = meta_distance_24h$tank_number)
perma_24h

## Random effect (strata) should be applied on the tank because we sample repeatedly in the tank.
## We don't sample the same tank multiple times at a time point. 
## We may have a time interaction tho, so I added the date_time in the model formula. 
### Note: why is the output of the model not show fixed effects and random effects like lmer()? 


```


```{r}
# Note : Version 1 of adonis
## the results are the same between adonis stats V1 and V2
perma_24h <- adonis2(MOBEAST_dist_24h ~ treatment*date_time, 
                     data = MOBEAST_nmds_24h, 
        # formula to calculate p-value for treatments
        strata = MOBEAST_nmds_24h$date_time) # accounts for repeated measures

perma_24h
```

# Ordination of individual time points

## 12:00:00

```{r}
otu_ord <- OTU %>% t() %>% as.data.frame() 

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select 12PM samples 
samples_12PM <- meta %>% mutate(time_string = as.character(time_string),) %>% 
  filter(time_string == "12:00:00" & date == ymd("2024-06-02")) %>% pull(id_number)

otu_ord_12PM <- otu_ord %>% filter(id_number %in% samples_12PM)


```

```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_12PM <- otu_ord_12PM %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances

```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_12PM <- vegdist(otu_ord_mat_12PM, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_12PM, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_12PM <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

### Plotting NMDS

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_12PM <- left_join(MOBEAST_nmds_12PM, meta)
```

```{r, fig.width=16, fig.height= 12}
# Plot
MOBEAST_nmds_12PM %>%   
  # mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      # "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  #facet_wrap(~time_string, ncol = 2)+
  labs(title= "12:00 samples")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("12PM_24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

### Stats on treatment only

```{r}
perma_12PM <- adonis2(MOBEAST_dist_12PM ~ MOBEAST_nmds_12PM$treatment) #, 
        # formula to calculate p-value for treatments
        #strata = MOBEAST_nmds_12PM$tank_number) # accounts for repeated measures

perma_12PM
```

## 06:00:00

```{r}
otu_ord <- OTU %>% 
  t() %>% 
  as.data.frame() 

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select 12PM samples 
samples_6AM <- meta %>% mutate(time_string = as.character(time_string)) %>% 
  filter(time_string == "06:00:00") %>% pull(id_number)

otu_ord_6AM <- otu_ord %>% filter(id_number %in% samples_6AM)


```

```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_6AM <- otu_ord_6AM %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances

```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_6AM <- vegdist(otu_ord_mat_6AM, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_6AM, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_6AM <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

### Plotting NMDS

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_6AM <- left_join(MOBEAST_nmds_6AM, meta)
```

```{r, fig.width=16, fig.height= 12}
# Plot
MOBEAST_nmds_6AM %>%   
  # mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      # "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  #facet_wrap(~time_string, ncol = 2)+
  labs(title= "12:00 samples")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("12PM_24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

### Stats on treatment only

```{r}
perma_6AM <- adonis2(MOBEAST_dist_6AM ~ MOBEAST_nmds_6AM$treatment)#, 
        # formula to calculate p-value for treatments
        #strata = MOBEAST_nmds_6AM$tank_number) no need strata here because it restricts the permutations
perma_6AM
```

#### Accounting for uneven variation in the data.

If the variation is uneven without the groups, adonis may come up with a significant result, although it's not. We can control for that using betdisper() and permutest()

```         
# beta dispertion
bd_12 <- betadisper(MOBEAST_dist_12, MOBEAST_nmds_12$treatment)

anova(bd_12) 
permutest(bd_12) 
bd_12
```