---
author: 'Keanu Rochette '
created: '18 April 2025'
updated: '`r format(Sys.Date())`'
title: MOBEAST DOC V2
output: 
  html_document:
     toc: true
     toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
#library(GGally)
library(emmeans)
library(performance)
library(multcompView)



```

## Load Data 
```{r}
DOC <- read_csv(here("data", "DOC", "DOC_clean.csv"))
table_id <- read_csv(here("data", "Laurel", "TableID.csv"))
```

## Data visualization theme
```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=32, face = "bold"), 
        plot.subtitle = element_text(size=30),
        axis.title = element_text(size = 28, face = "bold"),
        axis.text = element_text(size = 28),
        strip.text.x = element_text(size = 28, face = "bold"),
        legend.title=element_text(size=28, face = "bold"),
        legend.text=element_text(size=28))
```

## Data Clean Up

```{r}
table_id <- clean_names(table_id)

table_id <- table_id %>% rename(tank_number=tankid) %>% select(-treatment) %>% 
  mutate(tank_number = paste0("T", tank_number)) %>% 
  mutate(inflow_table = paste("Table", inflow_table))
DOC <- left_join(DOC, table_id, by= "tank_number")
```


```{r}
DOC <- DOC %>% mutate(inflow_table = ifelse(is.na(inflow_table), tank_number, inflow_table)) %>% 
   mutate(inflow_table = ifelse(inflow_table == "F1", "Table 1", 
                                ifelse(inflow_table == "F2", "Table 2", inflow_table))) %>% 
  mutate(inflow_table = factor(inflow_table),
         tank_number = factor(tank_number)) %>% 
  relocate(inflow_table, .after = "treatment")
```


```{r}

DOC_norm <- DOC %>% filter(date > as.Date("2024-05-31")) %>% # Removes endmembers
  filter(treatment != "Rain") %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table) %>%
  mutate(inflow= npoc_u_m[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # get the other table's value for that date
  group_by(date_time) %>% 
  mutate(inflow = ifelse(is.na(inflow),
      inflow[!is.na(inflow)][1],  
      inflow)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(npoc_u_m_norm = npoc_u_m - inflow)  %>%  
  relocate(inflow, .after = "npoc_u_m_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
  

#write_csv(DOC_norm, here("data", "DOC", "DOC_norm.csv"))
```


## Hypothesis 0: endmembers and inflow comparison

```{r}
DOC %>% filter(treatment == "Endmember" | treatment == "Rain"| treatment == "Inflow") %>% 
  dplyr::filter(time != lubridate::hms("00:00:00")) %>% 
  mutate(tank_number = as.character(tank_number)) %>% 
  mutate(tank_number = ifelse(tank_number == "Lagoon Tiahura", "Lagoon", tank_number),
         tank_number = factor(tank_number)) %>% 
  group_by(tank_number) %>% 
  summarize(avg = mean(npoc_u_m, na.rm =T),
            min = min(npoc_u_m, na.rm =T),
            max = max(npoc_u_m, na.rm =T))
```



```{r}
DOC_verif <- DOC %>% filter(treatment == "Endmember" | treatment == "Rain"| treatment == "Inflow") %>% 
  filter(time != hms("00:00:00")) %>% 
  mutate(tank_number = as.character(tank_number)) %>% 
  mutate(tank_number = ifelse(tank_number == "Lagoon Tiahura", "Lagoon", tank_number),
         tank_number = factor(tank_number, levels = c("Lagoon", "Ocean", "Bay", "F1", "F2")))
```

```{r}
mod <- lmer(npoc_u_m ~ tank_number + (1|date)+ (1|time)+  (1|treatment), data = DOC_verif)

summary(mod)
```


```{r}
anova(mod, type = "2")
```



```{, fig_height = 10, fig.width = 10}
plot_model(mod, type = "re")[[1]] +
  plot_model(mod, type = "re")[[2]] +
  plot_model(mod, type = "eff", terms = c('treatment')) +
  plot_model(mod, type = "diag")[[1]]
```

```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ tank_number)
```

```{r}
ggplot(DOC_verif, aes(x= tank_number, npoc_u_m, fill = treatment)) +
  geom_boxplot()+
  geom_jitter(aes(color = tank_number), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "All samples variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  facet_wrap(~date) + 
  #scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       #values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
```







## Hypothesis  1: Different communities will release difference amounts of DOC (12PM)


<!--
### Testing removing outliers first, then normalizing the data 
```{r}
DOC_12_pre_norm <- DOC %>% filter(date > as.Date("2024-05-31")) %>% # Removes end members
  filter(treatment != "Rain") %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
  ## Only keep the 12PM data points
  filter(time == hms("12:00:00")) %>% 
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m, probs = 0.25),
         quant3 = quantile(npoc_u_m, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m > quant3 + (IQR*1.5) | npoc_u_m < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  
  # Normalizing DOC data in tanks based on the inflow measurements
  group_by(date_time, inflow_table) %>%
  mutate(inflow= npoc_u_m[treatment == "Inflow"]) %>%  # Isolate inflow for each date_time and table
  ungroup() %>%
  group_by(date_time) %>% 
  mutate(inflow = ifelse(is.na(inflow),
      inflow[!is.na(inflow)][1],  # get the other table's value for that date
      inflow)) %>%
  ungroup() %>% 
  mutate(npoc_u_m_norm = npoc_u_m - inflow)  %>% # Normalize for treatments 
  relocate(inflow, .after = "npoc_u_m_norm") %>% 
  relocate(date_time, .before = "date") %>%
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) %>% 
  
  # removing the inflow data for modeling 
  filter(treatment != "Inflow")
  
```

```{r}
mod <- lmer(npoc_u_m_norm ~ treatment + (1|date) +(1|tank_number) , data = DOC_12_pre_norm)
mod <- lmer(npoc_u_m_norm ~ treatment + (1|date) , data = DOC_12_pre_norm)
summary(mod)
```

```{r}
Anova(mod, type= "2")
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```




-->

### Normality and Homogenity 
Data is not normal and variances are not homogenous
```{r}
bartlett.test(yjPower(npoc_u_m_norm) ~ treatment,data = DOC_12PM) # homogeneity
shapiro.test(DOC_12PM$npoc_u_m_norm) # normality
```


### modeling with data pre-normalized 

```{r}
DOC_norm_12PM <- DOC_norm %>% 
  filter(time == hms("12:00:00")) %>% 
  filter(treatment != "Inflow") 
```


```{r}
## Outlier detection

DOC_12PM <- DOC_norm_12PM %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m_norm, probs = 0.25),
         quant3 = quantile(npoc_u_m_norm, probs = 0.75),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m_norm > quant3 + (IQR*1.5) | npoc_u_m_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))

```

```{r}
mod <- lmer(npoc_u_m_norm ~ treatment +(1|tank_number), data = DOC_12PM)
#mod <- lm(npoc_u_m_norm ~ treatment, data = DOC_std_12PM)

summary(mod)
```


```{r}
anova(mod)
kruskal.test(npoc_u_m_norm ~ treatment, data = DOC_12PM )
```


```{, fig_height = 14, fig.width = 10}
plot_model(mod, type = "re")[[1]] +
  plot_model(mod, type = "re")[[2]] +
  plot_model(mod, type = "eff", terms = c('treatment')) +
  plot_model(mod, type = "diag")[[1]]
```

```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, ~ treatment)
```

```{r}
emm <- emmeans(mod, ~ treatment)
cld_letters <- cld(emm, Letters = letters)
```


### data visualization 
```{r, fig.width= 10, fig.height = 6}
ggplot(DOC_12PM, aes(x= treatment, npoc_u_m_norm, fill = treatment)) +
  geom_hline(yintercept = 0, lty = 2)+
  geom_violin()+
  geom_point(color = "black", alpha = 0.5)+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  stat_summary(color = "yellow", alpha =0.5)+
  labs(#title = "DOC Enrichment",
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "DOC Production (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  ylim(-20,120)+
  nunu_theme +
  theme(legend.position = "none") 

ggsave("DOC_12PM.png", width = 10,height = 6,path = here("outputs", "OOSC_2025"))
```

## Hypothesis 2: There are variation in exudation on a diel cycle 24h 

<!--
### Testing removing outliers first, then normalizing the data 
```{r}
DOC_24_pre_norm <- DOC %>% filter(date > as.Date("2024-05-31")) %>% # Removes end members
  filter(treatment != "Rain") %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
  
  # Only keep the 24h data points
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>% 
  filter(time != hms("00:00:00")) %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m, probs = 0.25, na.rm = TRUE),
         quant3 = quantile(npoc_u_m, probs = 0.75, na.rm = TRUE),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m > quant3 + (IQR*1.5) | npoc_u_m < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  
  # Normalizing DOC data in tanks based on the inflow measurements
  group_by(date, time, inflow_table) %>%
  mutate(inflow = npoc_u_m[treatment == "Inflow"]) %>%  # Isolate inflow for each date_time and table
  ungroup() %>%
  
  group_by(date_time) %>% 
  mutate(inflow = ifelse(is.na(inflow),
      inflow[!is.na(inflow)][1],  # get the other table's value for that date
      inflow)) %>%
  ungroup() %>% 
    
    ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  
  # Normalize for treatment
  mutate(npoc_u_m_norm = npoc_u_m - inflow)  %>% 
  relocate(inflow, .after = "npoc_u_m_norm") %>% 
  relocate(date_time, .before = "date") %>%
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) %>% 
  
  # removing the inflow data for modeling 
  filter(treatment != "Inflow") %>% 
  mutate(time = factor(time))
  
```


```{r}
mod <- lmer(npoc_u_m_norm ~ treatment*time + (1|date), data = DOC_24_pre_norm)

summary(mod)
```

```{r}
Anova(mod, type = "2")
```

```{r}
emmeans(mod, specs = trt.vs.ctrl ~ treatment)
```



-->


### Detecting outliers after normalization 
```{r}
DOC_norm_24h <- DOC_norm %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>% 
  #filter(time != hms("00:00:00")) %>%
  filter(treatment != "Inflow") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))
```


```{r}
## Outlier detection

DOC_24h <- DOC_norm_24h %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m_norm, probs = 0.25, na.rm = TRUE),
         quant3 = quantile(npoc_u_m_norm, probs = 0.75, na.rm = TRUE),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m_norm > quant3 + (IQR*1.5) | npoc_u_m_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") #%>% 
  
  #filter(!(tank_number %in% c("T8", "T11") & time == hms("18:00:00")))
 
```



```{r}
mod <- lmer(npoc_u_m_norm ~ treatment*time_fac + (1|tank_number), data = DOC_24h)

summary(mod)
anova(mod)
```


```{, fig_height = 10, fig.width = 10}
plot_model(mod, type = "re")[[1]] +
  plot_model(mod, type = "re")[[2]] +
  plot_model(mod, type = "eff", terms = c('treatment')) +
  plot_model(mod, type = "diag")[[1]]
```


```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
emmeans(mod, pairwise ~ treatment)
```


#### Interpretation
- Removed 00:00:00 because it was drastically different than the other 3 time points. Need to figure out what happened.
- Removed 18:00:00 T8 and T11 (control tanks) because they looked like outliers on the boxplots and also they skewed the model a lot!
- Random effect : inflow table does not explain the variance (variance = 0), nested effect helps to explain some of the variation. 
- Fixed effects: the p-values for the t-test are not significant but algae still has a smaller p-value than the other treatments. 
- Emmeans contrasts also do not show significant difference between the treatments. p>0.05


### Data visualization 
```{r}
ggplot(DOC_24h, aes(x= treatment, npoc_u_m_norm, fill = treatment)) +
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "24h variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  facet_wrap(~time, scales= "free") + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))
```



## Hypothesis 3: All samples

<!--
### Testing removing outliers first, then normalizing the data 

```{r}
DOC_all_pre_norm <- DOC %>% filter(date > as.Date("2024-05-31")) %>% # Removes end members
  filter(treatment != "Rain") %>% 
  mutate(date_time = factor(ymd_hms(paste(date, time)))) %>% 
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m, probs = 0.25, na.rm = TRUE),
         quant3 = quantile(npoc_u_m, probs = 0.75, na.rm = TRUE),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m > quant3 + (IQR*1.5) | npoc_u_m < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  
  # Normalizing DOC data in tanks based on the inflow measurements
  group_by(date, time, inflow_table) %>%
  mutate(inflow = npoc_u_m[treatment == "Inflow"]) %>%  # Isolate inflow for each date_time and table
  ungroup() %>%
  
  group_by(date_time) %>% 
  mutate(inflow = ifelse(is.na(inflow),
      inflow[!is.na(inflow)][1],  # get the other table's value for that date
      inflow)) %>%
  ungroup() %>% 
    
    ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  
  # Normalize for treatment
  mutate(npoc_u_m_norm = npoc_u_m - inflow)  %>% 
  relocate(inflow, .after = "npoc_u_m_norm") %>% 
  relocate(date_time, .before = "date") %>%
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) %>% 
  
  # removing the inflow data for modeling 
  filter(treatment != "Inflow")  
  
  
```

```{r}
mod <- lmer(npoc_u_m_norm ~ treatment + (1|time) , data = DOC_all_pre_norm)

summary(mod)
```

```{r}
anova(mod, type = "2")
```

-->

### analysis post norm
```{r}
DOC_norm_all <- DOC_norm %>% 
  #filter(time != hms("00:00:00")) %>% 
  filter(treatment != "Inflow") 
```

```{r}
## Outlier detection

DOC_all <- DOC_norm_all %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(npoc_u_m_norm, probs = 0.25, na.rm = TRUE),
         quant3 = quantile(npoc_u_m_norm, probs = 0.75, na.rm = TRUE),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( npoc_u_m_norm > quant3 + (IQR*1.5) | npoc_u_m_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time)) %>% 
  mutate(date_fac = factor(date))
```


```{r}
mod <- lmer(npoc_u_m_norm ~ treatment + (1|time)+(1|date), data = DOC_all)

summary(mod)
```


```{r}
anova(mod, type= "2")
```



```{r, fig_height = 10, fig.width = 10}
plot_model(mod, type = "re")[[1]] +
  plot_model(mod, type = "re")[[2]] +
  plot_model(mod, type = "eff", terms = c('treatment')) +
  plot_model(mod, type = "diag")[[1]]
```

```{r}
resid_panel(mod)

emmeans(mod, specs = trt.vs.ctrl ~ treatment)
plot(emmeans(mod, ~ treatment), comparisons = TRUE)
```

#### Interpretation

- Looking at all the samples together, lmer seems to indicate that Algae tanks are significantly different from the control
- Emmeans show that the difference between pairs is not significant. But the algae tanks have much low p-values than the other treatments



### Data visualization 




```{r}
ggplot(DOC_all, aes(x= treatment, npoc_u_m_norm, fill = treatment)) +
  geom_boxplot()+
  geom_jitter(aes(color = treatment), alpha = 0.5)+
  geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  labs(title = "All samples variation in normalized DOC",
       subtitle = "Norm DOC = Tank DOC - Avg Inflow DOC",
       x= "Treatment",
       y = "Normalized DOC")+
  facet_wrap(~date_time, scales= "free") + 
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none",
        axis.text.x = element_text(angle= 0, hjust= 0.5))+
  ylim(-15,80)
```








