---
title: "MOBEAST- DNA Ordination 24h V2"
format: html
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

# Load Libraries

```{r}
library(tidyverse)
#library(lubridate)
library(here)
library(janitor)
#library(ggrepel)
#library(ghibli)
#library(gt)
library(patchwork)
#library(RColorBrewer)
library(vegan)
library(OmicFlow)
library(phyloseq)
#library(GUniFrac)

library(car) # ANOVA package
#library(ggeffects)
#library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
#library(sjPlot)
library(emmeans)
#library(performance)
#library(multcompView)
library(broom.mixed) # for tidy()
```

# Data visualization

```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

```{r}
nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=24, face = "bold"), 
        plot.subtitle = element_text(size=20),
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=18, face = "bold"),
        legend.text=element_text(size=16),
        strip.background = element_rect(fill = "white"))
```

# Load Data

```{r}
unifrac <- read.csv(here("data", "DNA", "unifrac", "unifrac_weighted_97.csv"), row.names = 1,
                    header = TRUE)

meta <- read_csv(here("data","MOBEAST_metadata_V2.csv"))

DNA <- read.csv(here("data", "DNA", "MOBEAST_OTU_culled.csv"), header = 1)
tax <- read.csv(here("data", "DNA", "MOBEAST_tax_culled.csv"), header = 1)

```

# Data Clean Up

```{r}
endmember_list <- meta %>% 
  filter(treatment == "Endmember" |treatment == "Rain") %>% 
  pull(id_number)

inflow_list <- meta %>% 
  filter(treatment == "Inflow") %>% 
  pull(id_number)

diel_list <- meta %>% mutate(date_string = as.character(date_string)) %>% 
  filter(date_string == "2024-06-02" | date_string == "2024-06-03") %>% 
  pull(id_number)

sample_list <- meta %>% 
  #filter(!id_number %in% endmember_list) %>% 
  filter(!id_number %in% inflow_list) %>% 
  filter(id_number %in% diel_list) %>% 
  pull(id_number)
```

```{r}
unifrac_subset <- unifrac[rownames(unifrac) %in% sample_list, colnames(unifrac) %in% sample_list]

unifrac_dist <- as.dist(unifrac_subset)
```


# NMDS Calculation

```{r}
nmds <- metaMDS(unifrac_dist, k=2, trymax = 1000)
# put stress in the corner fo the nmds plot
```

```{r}
nmds_scores <- as.data.frame(scores(nmds, display = "site"))
```

```{r}
nmds_scores$id_number= rownames(nmds_scores)
```

```{r}
nmds_merged <- merge(nmds_scores, meta, by.x ="id_number", by.y = "id_number")
```

```{r}
nmds_merged<- nmds_merged %>% mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```


# Plotting Ordination 


```{r, fig.width=16, fig.height= 6}
# Plot
nmds_merged %>%   
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.8, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 4)+
  labs(title= "Ordination of 24h samples")+
  scale_color_manual(values=treatment_colors )+
  scale_fill_manual(values= treatment_colors) +
  nunu_theme+
  theme(legend.position = "top")
#ggsave("24h_treatment.png", width = 16, height = 6, path = here("outputs", "DNA"))

```
## PERMANOVA 

```{r}
# Note : Version 2 of adonis 
perma_24h <- adonis2(unifrac_dist ~ treatment + date_time + treatment:date_time, 
                     data = nmds_merged,
                     #permutations = 1000, 
                     strata = nmds_merged$tank_number)
perma_24h

## Random effect (strata) should be applied on the tank because we sample repeatedly in the tank.
## We don't sample the same tank multiple times at a time point. 
## We may have a time interaction tho, so I added the date_time in the model formula. 
### Note: why is the output of the model not show fixed effects and random effects like lmer()? 


```

## Pairwise comparison 

- permnova : adonis2
- pairwise pairwise.adonis2(p.adjust.m = "BH"), make sure that it's a distance matrix (as.dist())

```{r}
pairwise_adonis(x = unifrac_dist, 
                groups = as.vector(nmds_merged$treatment),
                  p.adjust.m = "BH")
```



# Analyzing data with Heatmap 

## Identifying the 2 samples that cluster alone in the Heatmaps 
```{r}
MOBEAST_nmds_24h_bis <- MOBEAST_nmds_24h %>% 
  mutate(treatment_bis = treatment) %>% 
  mutate(treatment_bis = ifelse(id_number == "MOBEAST010" |id_number == "MOBEAST060", "weird", treatment_bis ))
```

```{r, fig.width=16, fig.height= 6}
# Plot
MOBEAST_nmds_24h_bis %>%  
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment_bis, fill= treatment_bis))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 4)+
  labs(title= "24h samples, ellipse by treatment")+
  scale_color_manual(values= c("Inflow"="#00838F", "Control"="grey", "Coral"="#03A9F4", "CCA"="#EF5350", "Algae"="#4CAF50","weird"= "yellow"))+
  scale_fill_manual(values= c("Inflow"="#00838F", "Control"="grey", "Coral"="#03A9F4", "CCA"="#EF5350", "Algae"="#4CAF50","weird"= "yellow")) +
  #nunu_theme+
  theme_bw()
#ggsave("24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))

```

#### Plotting by heatmap cluster

Note: cluster numbers are inaccurate and do not match what is seen in the heatmap. Problem with the distance matrix or clutering method or both...

```{r, fig.width=16, fig.height= 10}
# Plot
MOBEAST_nmds_24h %>%   
  mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  #stat_ellipse(aes(colour = cluster, fill = cluster),
               #alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  facet_wrap(~time_string, ncol = 2)+
  labs(title= "24h samples, ellipse by treatment")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  ggforce::geom_mark_ellipse(aes(color = cluster, fill = cluster), alpha = 0.3)+
  #nunu_theme+
  theme_bw()
#ggsave("24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

```{r, fig.width=10, fig.height= 6}
# raw NMDS plot with all the samples
MOBEAST_nmds_24h %>%  
  mutate(time_string = factor(time_string, 
                              levels = c("12:00:00","18:00:00","00:00:00","06:00:00" ))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment))+
  stat_ellipse(aes(colour = time_string, fill = time_string), 
               alpha = 0.2, geom = "polygon", level = 0.95)+
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  labs(title= "24h samples, ellipse by time")+
  scale_color_manual(values=treatment_colors)+
  #scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("24h_time.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

### Stats on treatment only

```{r}
# Tranform the distance matrix into a df that I can use adonis on 
id_number_24h <- meta %>% filter(date_string == "2024-06-02" |date_string == "2024-06-03") %>% pull(id_number)

distance_24h <- MOBEAST_dist_24h %>% as.matrix() %>% as.data.frame() %>% 
  mutate(id_number = id_number_24h)%>% 
  relocate(id_number, .before = MOBEAST001)

rownames(distance_24h) <- NULL
```

```{r}
meta_distance_24h <- inner_join(meta, distance_24h, by = "id_number") %>% 
  select(-long_name, -date_string, -time_string, -date_time_string,-cluster)
meta_distance_24h 

all_dist <- meta_distance_24h %>% 
  select(all_of(meta_distance_24h[["id_number"]])) %>% 
  as.dist()
```

```{r}
# Note : Version 2 of adonis 
perma_24h <- adonis2(all_dist ~ treatment + date_time + treatment:date_time, 
                     data = meta_distance_24h,
                     #permutations = 1000, 
                     strata = meta_distance_24h$tank_number)
perma_24h

## Random effect (strata) should be applied on the tank because we sample repeatedly in the tank.
## We don't sample the same tank multiple times at a time point. 
## We may have a time interaction tho, so I added the date_time in the model formula. 
### Note: why is the output of the model not show fixed effects and random effects like lmer()? 


```


```{r}
# Note : Version 1 of adonis
## the results are the same between adonis stats V1 and V2
perma_24h <- adonis2(MOBEAST_dist_24h ~ treatment*date_time, 
                     data = MOBEAST_nmds_24h, 
        # formula to calculate p-value for treatments
        strata = MOBEAST_nmds_24h$date_time) # accounts for repeated measures

perma_24h
```

# Ordination of individual time points

## 12:00:00

```{r}
otu_ord <- OTU %>% t() %>% as.data.frame() 

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select 12PM samples 
samples_12PM <- meta %>% mutate(time_string = as.character(time_string),) %>% 
  filter(time_string == "12:00:00" & date == ymd("2024-06-02")) %>% pull(id_number)

otu_ord_12PM <- otu_ord %>% filter(id_number %in% samples_12PM)


```

```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_12PM <- otu_ord_12PM %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances

```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_12PM <- vegdist(otu_ord_mat_12PM, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_12PM, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_12PM <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

### Plotting NMDS

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_12PM <- left_join(MOBEAST_nmds_12PM, meta)
```

```{r, fig.width=16, fig.height= 12}
# Plot
MOBEAST_nmds_12PM %>%   
  # mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      # "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  #facet_wrap(~time_string, ncol = 2)+
  labs(title= "12:00 samples")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("12PM_24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

### Stats on treatment only

```{r}
perma_12PM <- adonis2(MOBEAST_dist_12PM ~ MOBEAST_nmds_12PM$treatment) #, 
        # formula to calculate p-value for treatments
        #strata = MOBEAST_nmds_12PM$tank_number) # accounts for repeated measures

perma_12PM
```

## 06:00:00

```{r}
otu_ord <- OTU %>% 
  t() %>% 
  as.data.frame() 

# Make the sample ID the rownames of the data frame
otu_ord<- otu_ord %>% mutate(id_number = rownames(otu_ord)) 

# only select 12PM samples 
samples_6AM <- meta %>% mutate(time_string = as.character(time_string)) %>% 
  filter(time_string == "06:00:00") %>% pull(id_number)

otu_ord_6AM <- otu_ord %>% filter(id_number %in% samples_6AM)


```

```{r}
# only keep the OTU numbers in the matrix and remove the id_number column
## ID numbers are now rownames. then convert the df in a matrix.
otu_ord_mat_6AM <- otu_ord_6AM %>%  select(-id_number) %>%  as.matrix()

```

### Calculate distances

```{r}
# metaMDS uses a random number to calculate the distance matrix. Need to set a "seed" to get the consistent results. 
set.seed(2000)

# calcultat the distance matrix with bray-curtis method
MOBEAST_dist_6AM <- vegdist(otu_ord_mat_6AM, method = "bray")

#calculate the MDS scores such that it converges to a solutions. 
## trymax was randomly set to 10000 (based on the internet), because solutions did not converge on the initial trials.
nmds <- metaMDS(MOBEAST_dist_6AM, trymax = 10000)

# save the NMDS scores in a tibble with the ID numbers. 
MOBEAST_nmds_6AM <- scores(nmds) %>% as_tibble(rownames = "id_number")
  
```

### Plotting NMDS

```{r}
# add the metadata to the NMDS tibble to make plots 
MOBEAST_nmds_6AM <- left_join(MOBEAST_nmds_6AM, meta)
```

```{r, fig.width=16, fig.height= 12}
# Plot
MOBEAST_nmds_6AM %>%   
  # mutate(time_string = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                      # "00:00:00", "06:00:00"))) %>% 
  ggplot(aes(x= NMDS1, y = NMDS2, colour = treatment, fill= treatment))+
  stat_ellipse(#aes(colour = cluster, fill = cluster),
               alpha = 0.2, geom= "polygon", level = 0.95, show.legend = FALSE)+
  
  geom_point(size = 4) +
  geom_point(shape = 1, color = "black",size = 4) +
  #geom_label_repel(aes(label = date_string))+
  #facet_wrap(~time_string, ncol = 2)+
  labs(title= "12:00 samples")+
  scale_color_manual(values=treatment_colors)+
  scale_fill_manual(values=treatment_colors) +
  #nunu_theme+
  theme_bw()
#ggsave("12PM_24h_treatment.png", width = 16, height = 10, path = here("outputs", "DNA"))
```

### Stats on treatment only

```{r}
perma_6AM <- adonis2(MOBEAST_dist_6AM ~ MOBEAST_nmds_6AM$treatment)#, 
        # formula to calculate p-value for treatments
        #strata = MOBEAST_nmds_6AM$tank_number) no need strata here because it restricts the permutations
perma_6AM
```

#### Accounting for uneven variation in the data.

If the variation is uneven without the groups, adonis may come up with a significant result, although it's not. We can control for that using betdisper() and permutest()

```         
# beta dispertion
bd_12 <- betadisper(MOBEAST_dist_12, MOBEAST_nmds_12$treatment)

anova(bd_12) 
permutest(bd_12) 
bd_12
```

## Microbiome diversity 

```{r}
div_meta <- meta %>% select(id_number, treatment, date_string)

# formatting the OTU matrix with sample names in the rows and OTU on the columns
otu <- DNA %>% 
  # remove the PCR controls from the df
  select(-P1_PCRPos,-P2_PCRPos, -P3_PCRPos, -P4_PCRPos, - P5_PCRPos) %>%
  # join with the taxonomic df 
  rename(OTUNumber = ASV) %>% 
  left_join(tax) %>% 
  # Here I want the relative abundance of microbes by family per samples
  select(MOBEAST001:MOBEAST115, Family) %>% 
  relocate(Family,.before = "MOBEAST001") %>% 
  # different OTU may result in bacteria that share a family but belong to different genus
  # we want to add the relative abundance of the bacteria in the same family
  group_by(Family) %>%
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) %>% 
  as.data.frame()

rownames(otu)<- otu$Family

#transpose so that the Families are the column names and the samples in the rows
otu_t <- otu %>%  select(-Family) %>% t()

#convert the matrix in a df that I can filter
otu_t <- otu_t %>% as.data.frame() %>% mutate(id_number = rownames(otu_t)) %>% 
  left_join(div_meta) %>% relocate(c("id_number", "treatment", "date_string"))

```

### 24h 
```{r}
# we only keep the 24h samples. we remove the inflow values. 
otu_24h <- otu_t %>% mutate(date_string = as.character(date_string)) %>% 
  filter(date_string == "2024-06-02" | date_string == "2024-06-03") %>% 
  filter(treatment != "Inflow") %>% 
  select(-treatment, -date_string)

rownames(otu_24h)<- otu_24h$id_number

# convert the df into a matrix
matrix_24h <- otu_24h %>% select(-id_number) %>% as.matrix() 

```

```{r}
# calculating different biodiversity metrics and indices. 
richness <- specnumber(matrix_24h)
shannon <- vegan::diversity(matrix_24h, index = "shannon")
simpson <- vegan::diversity(matrix_24h, index = "simpson")
evenness <- shannon/log(richness)
berger_parker <- apply(matrix_24h, 1, function(x) max(x)/sum(x))


# merge the results in one data frame that we can use for stat models
diversity_indices <- tibble(
  id_number = rownames(matrix_24h),
  richness = richness, 
  shannon = shannon, 
  simpson = simpson, 
  evenness = evenness, 
  berger_parker = berger_parker
)
```

```{r}
# merge the diversity metrics with the meta data
div_df_24h <- diversity_indices %>% left_join(meta) %>% 
  # only select the meta data that is useful for the models
  select(id_number, tank_number, treatment, inflow_table, time, date_string, time_string,
         richness,shannon,simpson, berger_parker, evenness) %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  mutate(time_fac = factor(time_string, 
                           levels = c("12:00:00", "18:00:00","00:00:00", "06:00:00"))) %>% 
  #focusing the shanon diversity index, we want to remove the outlier values
  ## we use the mean +/- 2SD method
  mutate(sd = sd(shannon), mean = mean(shannon)) %>% 
  filter(shannon < mean+ 2* sd | shannon > mean-2*sd) %>%
  select(-sd, -mean)  
  
```


```{r}
# statistical model to see if diversity changes by treatment and in time
# tank number used for repeated measure random effect
mod_div <- lmer(shannon ~ treatment*time+ (1|tank_number), data = div_df_24h)

summary(mod_div)
anova(mod_div, type= 2)
emmeans(mod_div, specs = trt.vs.ctrl ~ treatment) 
emmeans(mod_div, pairwise ~ treatment | time)
```

**Definition**: 
*Richness*: total number of different species. (discreet count)
*Evenness*: even population or 1 dominant group vs smaller groups. (1 is even,  0 is uneven)
*Shannon*: accounts for evenness and richness. (High value = higher diversity)
*Simpson*: probability that 2 indiv randomly picked are the same spp. (1 = higher diversity, 0= low/no diversity)
*Berger_parker*: numerical dominance (high = one single spp = low diversity)


```{r}
div_df_24h %>% 
  ggplot(aes(treatment, shannon, fill = treatment)) +
  geom_violin()+
  geom_point()+
  facet_wrap(.~time_fac, nrow = 1)+
  stat_summary(color = "yellow", alpha =0.5)+
  scale_fill_manual(values = treatment_colors)+
  nunu_theme
```
```{r}
div_df_24h %>% 
  
  ggplot(aes(richness, evenness, color = treatment)) +
  #geom_violin()+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+

  #facet_wrap(.~time_fac, nrow = 1)+
  #stat_summary(color = "yellow", alpha =0.5)+
  scale_color_manual(values = treatment_colors)+
  nunu_theme
```

```{r}
div_df_24h %>% 
  group_by(treatment) %>% 
  reframe(mean_rich = mean(richness), mean_even = mean(evenness)) %>% 
  ggplot(aes(mean_rich, mean_even, color = treatment)) +
  #geom_violin()+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  geom_label(aes(label = treatment), hjust = 0,vjust = 1.5)+

  #facet_wrap(.~time_fac, nrow = 1)+
  #stat_summary(color = "yellow", alpha =0.5)+
  scale_color_manual(values = treatment_colors)+
  nunu_theme
```
