---
title: "MOBEAST DNA delta"
created: 21 June 2026
edited: 21 June 2026
format: html
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(janitor)
library(here)

library(ggrepel)
library(patchwork)
library(RColorBrewer)
library(measurements)
library(gt)
library(webshot2)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
library(emmeans)
library(performance)
library(multcompView)
library(broom.mixed)
library(effectsize)


# PCA packages
library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(factoextra) 
```


## Data visualization theme

```{r}
nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=24, face = "bold"), 
        plot.subtitle = element_text(size=20),
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=18, face = "bold"),
        legend.text=element_text(size=16),
        strip.background = element_rect(fill = "white"))
```



```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

## Load Data

```{r}
mobeast_raw <- read_csv(here("data", "MOBEAST_full_merged_data_long.csv"))

```


## DNA 

### DNA data clean up 
```{r}
# prepaing the data for later
mobeast_DNA  <- mobeast_raw %>%
  mutate(time_string = as.character(time_string),
              date_string = as.character(date_string),
              date_time_string = paste(date_string,time_string))%>% 
  filter(variable_cat == "DNA" ) %>% 
  filter(treatment != "Rain") %>% 
  filter(treatment != "Endmember") %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
  
  

```

```{r}
mobeast_DNA_24h_raw <- mobeast_DNA %>% filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  #filter(treatment != "Inflow") %>% 
  filter(!is.na(id_number)) %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00"))) %>% 
  mutate(log_flow = log10(flowrate))

```

```{r}
## outlier detection on the relative abundance??
mobeast_DNA_24h_flagged <- mobeast_DNA_24h_raw %>%  
  group_by(variable) %>% 
  mutate(mean = mean(value, na.rm= T), 
         sd = sd(value, na.rm= T)) %>% 
  mutate(outlier = ifelse(value > mean + 2*sd |value < mean - 2*sd, "outlier", "not outlier")) %>% 
  ungroup() %>% 
  filter(outlier != "outlier")
```


```{r}
## outlier detection on the relative abundance??
mobeast_DNA_24h <- mobeast_DNA_24h_flagged %>%  
  mutate(inflow_isolated = ifelse(treatment == "Inflow", value, NA)) %>% 
  group_by(time, variable) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = T )) %>% 
  ungroup() %>% 
  mutate(conc_norm  = value - inflow ) %>% 
  filter(treatment != "Inflow" )
```




### LMER Models
```{r}
# Loop Function:
## Take the 24h samples and isolate the types of nutrient  present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

exclude_family <- mobeast_DNA_24h %>% 
  group_by(time, variable) %>% 
  reframe(avg = mean(value)) %>% 
  filter(avg == 0) 

mobeast_loop <- mobeast_DNA_24h %>% filter(!variable %in% exclude_family$variable & !variable %in% exclude_family$time) %>%
  select(variable) %>% distinct() %>% pull()

models_DNA_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_DNA_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- mobeast_DNA_24h %>% filter(variable == var)
  models_DNA_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + 
                                  log_flow + (1|tank_number), data = df)

}

```


```{r}
results_DNA <- imap_dfr(models_DNA_24h, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")

results_DNA %>%  filter(p.value <0.05)
```

### DNA ANOVA

```{r, warning= FALSE}
# So many Warnings! 
## Warning: The column names NumDF and DenDF in ANOVA output were not recognized or transformed.

anova_results <- imap_dfr(models_DNA_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

DNA_anova_sig <- anova_results %>% filter(p.value <0.05) %>% select(variable, anova_term = term, anova_p_val = p.value)

sig_treat <- anova_results %>% filter(p.value <0.05) %>% 
  filter(term == "treatment") %>%
  select(variable) %>% 
  pull()

fam_time <- anova_results %>% filter(p.value <0.05) %>% 
  filter(term == "time_fac") %>% select(variable) %>% distinct() 

fam_treatment <- anova_results %>% filter(p.value <0.05) %>% 
  filter(term == "treatment") %>% select(variable) %>% distinct() %>% pull()

fam_time %>%  filter(variable %in% fam_treatment) %>%  pull()

fam_interaction <- anova_results %>% filter(p.value <0.05) %>% 
  filter(term == "treatment:time_fac")

fam_interaction %>% filter(variable %in% fam_treatment)
interaction_family <- fam_interaction %>% filter(variable %in% fam_treatment) %>% 
  select(variable) %>%  pull()
```



#### Partial Eta squared 

Partial η²

Answers "so what?" after a significant ANOVA
Tells you whether the significant effect is actually large enough to care about biologically
A p < 0.05 with η² = 0.01 is statistically real but possibly biologically trivial

```{r}
eta_results <- tibble()

for (var in sig_treat) {
  df <- tibble(eta_squared(anova(models_DNA_24h[[var]]), partial = TRUE))
  df <- df %>% mutate(variable = var ) %>% relocate(variable)
  
  eta_results <- rbind(eta_results, df)
  
  eta_results <- eta_results 
  
}

sig_eta <- eta_results %>% filter(Eta2_partial > 0.14) %>% select(variable) %>% distinct() %>% pull()

sig_eta 
```


### DNA Emmeans 

```{r, message= FALSE}
# "estimate" are negative somehow. Why? 
## in manual models for DOC, estimate >0, here it's <0
## t.ratio in "statistics", and negative instead of positive
## p.values not shown, only p.adjusted. 
emmeans_results_dna <- imap_dfr(models_DNA_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 


```

```{r}
emmeans_sig_dna <- emmeans_results_dna %>% filter(adj.p.value <0.05) %>% 
  select(variable, emmean_term = term, contrast, emmeans_p_val = adj.p.value) 

emmeans_sig_dna %>% left_join(DNA_anova_sig) %>% filter(!variable %in% interaction_family) %>% select(variable) %>%  distinct() %>% pull()

```



```{r, warning=FALSE}
DNA_sig_var <- emmeans_results_dna %>% 
  filter(adj.p.value <0.05) %>% 
  select(variable) %>% distinct() %>% pull()

graphs_DNA_24h <- DNA_sig_var %>%
  set_names() %>%
  map(~ {
    df <- mobeast_DNA_24h %>% filter(variable == .x)
    
    ggplot(df, aes(x = treatment, y = value, fill = treatment)) +
      geom_violin() +
      geom_point() +
      geom_hline(yintercept = 0, linetype = "dashed") +
      stat_summary(color = "yellow", alpha = 0.5) +
      labs(title = paste("24h", .x, "Production"),
           x = "Treatment",
          y = "Relative Abundance") +
      scale_fill_manual(
        breaks = c("Control", "Coral", "CCA", "Algae", "Inflow"),
        values = c("grey", "#03A9F4", "#EF5350", "#4CAF50", "orange")
      ) +
      nunu_theme +
      theme(legend.position = "none")
  })


```




```{r}
emmeans_results_dna %>%  
  mutate(sig = ifelse(adj.p.value<0.05 & estimate < 0, "Enrichment", 
                                         ifelse(adj.p.value<0.05 & estimate > 0, "Depletion", "No Difference")))%>% 
  mutate(sig = factor(sig, 
                      levels = c("Enrichment","No Difference", "Depletion"))) %>% 
  filter(variable %in% sig_treat)  %>%
  #filter(variable %in% interaction_family) %>% 
  mutate(contrast = factor(contrast, levels = c("Control - Coral", "Control - CCA", "Control - Algae"))) %>%
  
  # these families have sig ANOVA but not sig control vs treatment contrast
  ## probs have sig treatment vs treatment contrasts 
  filter(variable != "Oligoflexaceae", variable != "NS9_marine_group") %>% 
  
  ggplot()+
  geom_tile(aes(x= contrast, y= variable, fill = sig), color = "black")+
  #facet_wrap(.~variable_cat)+
  labs(
       x= "Treatment",
       y = " Variables")+
  guides(fill=guide_legend(title="p-value Interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw() +
  nunu_theme+
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
  
ggsave(filename = "dna_enrichment_tiles.png", path = here("outputs"), height = 6, width=12)
```

### Effect Size 

Adds biological interpretability to each contrast
Tells you not just "these groups differ" but "by how much and in which direction"
Essential for comparing across nutrients since your residuals are on different scales


```{r}
effect_size_results <- tibble()

for (var in sig_eta) {
  emm <- emmeans(models_DNA_24h[[var]], ~ treatment)
  contrasts <- contrast(emm, method = "trt.vs.ctrl")

## Effect size for each contrast
  df <- tidy(eff_size(emm, sigma = sigma(models_DNA_24h[[var]]), 
                edf = df.residual(models_DNA_24h[[var]])))
  
  df <- df %>% mutate(variable = var) %>% rename(effect.size = estimate) %>% 
    relocate(variable)
  
  effect_size_results <- rbind(effect_size_results, df)
  
}

effect_size_results %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") %>% filter(abs(effect.size) >=0.8) %>%  filter (variable %in%sig_treat)



```

### Summary DNA table 

```{r}

anova_sig_family <- anova_results %>% 
  select(variable, term, df = NumDF, F_stat = statistic, p.value) %>%
  filter(term != "log_flow") %>%
  group_by(variable) %>%
  filter(any(p.value < 0.05)) %>% 
  ungroup() %>%  
  select(variable) %>% distinct() %>% pull()
group1 <- anova_sig_family[1:12]
group2 <- anova_sig_family[13:24]
group3 <- anova_sig_family[25:36]
group4 <- anova_sig_family[37:48]
```

```{r}
### 1/4 table summary 
dna_stat_summary_table1 <- anova_results %>% 
  select(variable, term, df = NumDF, F_stat = statistic, p.value) %>%
  filter(term != "log_flow") %>%
  filter(variable %in% group1) %>% 
  group_by(variable) %>%
  filter(any(p.value < 0.05)) %>%
  mutate(term = recode_values(term,
      "treatment"      ~"Treatment",
      "time_fac"       ~ "Time",
      "treatment:time_fac" ~ "Treatment x Time"
    ),
    F_stat  = round(F_stat, 2),
    p.value = round(p.value, 3)
  ) %>%
  gt(groupname_col = "variable") %>%        # ← this merges the repeated variable cells
  cols_label(
    term    = "Term",
    df      = "df",
    F_stat  = "F",
    p.value = "p-value"
  ) %>%
  fmt_number(columns = F_stat,  decimals = 2) %>%
  fmt_number(columns = p.value, decimals = 3) %>%
  tab_style(
    style     = cell_fill(color = "#f0f0f0"),
    locations = cells_row_groups()            # shaded group header rows
  ) %>%
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_row_groups()            # bold group headers
  ) %>%
  tab_style(
    style     = cell_text(weight = "bold", color = "red"),
    locations = cells_body(
      columns = p.value,
      rows    = p.value < 0.05
    )
  ) %>% 
  tab_options(
  table.font.size      = 12,
  row_group.font.weight = "bold",
  table.width           = pct(75)  # or a fixed value like px(800)
)

dna_stat_summary_table1

gtsave(dna_stat_summary_table1, here("outputs", "dna_anova_table_summary1.png")) 
       #vwidth = 1200,   # width in pixels  ← increase this to make it wider
       #vheight = 1600) # height in pixels ← increase if table gets cut off
```


```{r}
### 2/4 table summary 
dna_stat_summary_table2 <- anova_results %>% 
  select(variable, term, df = NumDF, F_stat = statistic, p.value) %>%
  filter(term != "log_flow") %>%
  filter(variable %in% group2) %>% 
  group_by(variable) %>%
  filter(any(p.value < 0.05)) %>%
  mutate(term = recode_values(term,
      "treatment"      ~"Treatment",
      "time_fac"       ~ "Time",
      "treatment:time_fac" ~ "Treatment x Time"
    ),
    F_stat  = round(F_stat, 2),
    p.value = round(p.value, 3)
  ) %>%
  gt(groupname_col = "variable") %>%        # ← this merges the repeated variable cells
  cols_label(
    term    = "Term",
    df      = "df",
    F_stat  = "F",
    p.value = "p-value"
  ) %>%
  fmt_number(columns = F_stat,  decimals = 2) %>%
  fmt_number(columns = p.value, decimals = 3) %>%
  tab_style(
    style     = cell_fill(color = "#f0f0f0"),
    locations = cells_row_groups()            # shaded group header rows
  ) %>%
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_row_groups()            # bold group headers
  ) %>%
  tab_style(
    style     = cell_text(weight = "bold", color = "red"),
    locations = cells_body(
      columns = p.value,
      rows    = p.value < 0.05
    )
  )  %>% 
  tab_options(
  table.font.size      = 12,
  row_group.font.weight = "bold",
  table.width           = pct(75)  # or a fixed value like px(800)
)

gtsave(dna_stat_summary_table2, here("outputs", "dna_anova_table_summary2.png")) 
       #vwidth = 1200,   # width in pixels  ← increase this to make it wider
       #vheight = 1600) # height in pixels ← increase if table gets cut off
```


```{r}
### 3/4 table summary 
dna_stat_summary_table3 <- anova_results %>% 
  select(variable, term, df = NumDF, F_stat = statistic, p.value) %>%
  filter(term != "log_flow") %>%
  filter(variable %in% group3) %>% 
  group_by(variable) %>%
  filter(any(p.value < 0.05)) %>%
  mutate(term = recode_values(term,
      "treatment"      ~"Treatment",
      "time_fac"       ~ "Time",
      "treatment:time_fac" ~ "Treatment x Time"
    ),
    F_stat  = round(F_stat, 2),
    p.value = round(p.value, 3)
  ) %>%
  gt(groupname_col = "variable") %>%        # ← this merges the repeated variable cells
  cols_label(
    term    = "Term",
    df      = "df",
    F_stat  = "F",
    p.value = "p-value"
  ) %>%
  fmt_number(columns = F_stat,  decimals = 2) %>%
  fmt_number(columns = p.value, decimals = 3) %>%
  tab_style(
    style     = cell_fill(color = "#f0f0f0"),
    locations = cells_row_groups()            # shaded group header rows
  ) %>%
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_row_groups()            # bold group headers
  ) %>%
  tab_style(
    style     = cell_text(weight = "bold", color = "red"),
    locations = cells_body(
      columns = p.value,
      rows    = p.value < 0.05
    )
  )  %>% 
  tab_options(
  table.font.size      = 12,
  row_group.font.weight = "bold",
  table.width           = pct(75)  # or a fixed value like px(800)
)

gtsave(dna_stat_summary_table3, here("outputs", "dna_anova_table_summary3.png")) 
       #vwidth = 1200,   # width in pixels  ← increase this to make it wider
       #vheight = 1600) # height in pixels ← increase if table gets cut off
```



```{r}
### 4/4 table summary 
dna_stat_summary_table4 <- anova_results %>% 
  select(variable, term, df = NumDF, F_stat = statistic, p.value) %>%
  filter(term != "log_flow") %>%
  filter(variable %in% group4) %>% 
  group_by(variable) %>%
  filter(any(p.value < 0.05)) %>%
  mutate(term = recode_values(term,
      "treatment"      ~"Treatment",
      "time_fac"       ~ "Time",
      "treatment:time_fac" ~ "Treatment x Time"
    ),
    F_stat  = round(F_stat, 2),
    p.value = round(p.value, 3)
  ) %>%
  gt(groupname_col = "variable") %>%        # ← this merges the repeated variable cells
  cols_label(
    term    = "Term",
    df      = "df",
    F_stat  = "F",
    p.value = "p-value"
  ) %>%
  fmt_number(columns = F_stat,  decimals = 2) %>%
  fmt_number(columns = p.value, decimals = 3) %>%
  tab_style(
    style     = cell_fill(color = "#f0f0f0"),
    locations = cells_row_groups()            # shaded group header rows
  ) %>%
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_row_groups()            # bold group headers
  ) %>%
  tab_style(
    style     = cell_text(weight = "bold", color = "red"),
    locations = cells_body(
      columns = p.value,
      rows    = p.value < 0.05
    )
  )  %>% 
  tab_options(
  table.font.size      = 12,
  row_group.font.weight = "bold",
  table.width           = pct(75)  # or a fixed value like px(800)
)

gtsave(dna_stat_summary_table4, here("outputs", "dna_anova_table_summary4.png")) 
       #vwidth = 1200,   # width in pixels  ← increase this to make it wider
       #vheight = 1600) # height in pixels ← increase if table gets cut off
```



```{r}
sig_var <- pairwise_summary %>% filter(p_value <0.05) %>% 
  select(variable) %>% distinct() %>% pull()
```

```{r}
pairwise_summary %>% 
  mutate(sig = ifelse(p_value<0.05 & mean > 0, "Enrichment", 
                                         ifelse(p_value<0.05 & mean < 0, "Depletion", "No Difference")))%>% 
  mutate(sig = factor(sig, 
                      levels = c("Enrichment","No Difference", "Depletion"))) %>% 
  #mutate(variable = factor(variable, levels = unique(variable))) %>%
  filter(variable %in% sig_var) %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  ggplot()+
  geom_tile(aes(x= treatment, y= variable, fill = sig), color = "black")+
  #facet_grid(variable_cat ~ ., scales = "free_y", space = "free_y")+
  labs(title = "Pairwise comparison 12AM-12PM",
       x= "Treatment",
       y = " Variables")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw()#+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")

```
