---
title: "MOBEAST - 24h Heatmap"
format: html
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)

library(ComplexHeatmap)
library(data.table)
library(dendextend)
library(grid)

# Rayna library
library(dendsort)
library(ggsci)
library(randomForest)
library(vegan)
library(ape)
library(devtools)
library(ggpubr)
library(viridis)
library(colorspace)
library(readr)
library(circlize)


library(cowplot)
library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
library(emmeans)
library(performance)
library(multcompView)
library(broom.mixed)
library(effectsize)

```

Heatmap info and tutorial: https://jokergoo.github.io/ComplexHeatmap-reference/book/a-single-heatmap.html

## Functions

```{r}
source(here("scripts", "reference_scripts", "heatmap_zscore_function.R"))
```

## Data visualization theme

```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        panel.background = element_rect(fill = "white"))
```

### Color palette

```{r}
treatment_colors = c(
  #Treatments
  "Control" = "grey",
  "Coral" = "#03A9F4",
  "CCA" = "#EF5350",
  "Algae" = "#4CAF50"
)

```

## Load Data

```{r}
meta <- read_csv(here("data", "MOBEAST_metadata_V2.csv"))
OTU_long <- read_csv(here("data","MOBEAST_DNA_long.csv"))

```

## Data Clean Up

```{r}
OTU_long <- OTU_long %>% clean_names() %>% select(-long_name, -tank_number, -treatment, -date,-time) %>%  rename(id_number= sample_id)

DNA <- meta %>% left_join(OTU_long)
```

### Preparing the OTU matrix for heatmap

```{r}
exclude_family <- DNA %>% group_by(family) %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>%
  reframe(sum = sum(abund)) %>% 
  filter(sum ==0) %>% 
  select(family) %>% 
  pull()
```

```{r}
OTU_htm <- DNA %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>%
  filter(!family %in% exclude_family) %>% 
  mutate(family_tax = paste(family, "X", order, "X", class, "X", phylum),
         genus_tax = paste(genus, "X", family, "X", order, "X", class, "X", phylum))
```

```{r}
OTU_htm_24h <- OTU_htm %>% filter(treatment != "Inflow")
  
```

```{r}
OTU_htm %>% filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>% 
  select(id_number) %>% distinct()
```



## Data with only the most abundant taxa 
```{r}
rare_taxa_threshold = 0.005
```

```{}
# family
## per sample, min abundance "threshold" and more than 3x occurence. 
taxa_filtered <- OTU_htm_24h %>%
  group_by(id_number, family_tax) %>%
  mutate(sum_abund = sum(abund),
            time, treatment) %>% 
  ungroup() %>% 
  group_by(family_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= rare_taxa_threshold)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  mutate(taxa = family_tax)

taxa_filtered %>% select(taxa) %>% distinct()

```

```{r}
# genus
## per sample, min abundance 0.5% and more than 3x occurence. 

taxa_filtered <- OTU_htm_24h %>%
  filter(!str_detect(genus_tax, "unclassified")) %>%
  group_by(id_number, genus_tax) %>%
  mutate(sum_abund = sum(abund)) %>% 
  ungroup() %>% 
  group_by(genus_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= rare_taxa_threshold)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  mutate(taxa = genus_tax)
  
taxa_filtered %>% select(taxa) %>% distinct()
```

```{}
#family
## max abundance > threshold
taxa_filtered <- OTU_htm_24h %>%
  group_by(family_tax) %>%
  mutate(max = max(abund)) %>% 
  ungroup() %>% 
  filter(max >= rare_taxa_threshold) %>% 
  ungroup() %>% 
  mutate(taxa = family_tax)

taxa_filtered %>% select(taxa) %>% distinct()
```


```{}
# genus 
## max abundance > threshold

taxa_filtered <- OTU_htm_24h %>%
  group_by(genus_tax) %>%
  mutate(max = max(abund)) %>% 
  ungroup() %>% 
  filter(max >= rare_taxa_threshold) %>% 
  ungroup() %>% 
  mutate(taxa = genus_tax)

taxa_filtered %>% select(taxa) %>% distinct()
```


### SQRT-transformed Z-score 

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- taxa_filtered %>%
  group_by(id_number, taxa) %>%
  summarize(abund = sum(abund, na.rm = TRUE)) %>% 
  ungroup()

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = taxa, values_from = abund)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% tibble() %>%  select(-id_number)

```

```{r}
sqrt.matrix <- apply(culled.abund.OTU, 2, FUN=sqrt)
sqrt.zscore <- as.data.frame(apply(sqrt.matrix, 2, FUN = zscore.calculation))
```

```{r}
# bring back the sample ID number in the df 
sqrt.zscore$id_number = family.culled.abund$id_number


# Set the ID numbers as the rownames (to make a matrix)
rownames(sqrt.zscore) = sqrt.zscore$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
sqrt.zscore.matrix <- sqrt.zscore %>% select(-id_number) %>% 
  as.matrix() %>% t()


```

### Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(sqrt.zscore.matrix),
                           meta$id_number), 
                 "treatment"]

date_time <- meta[match(colnames(sqrt.zscore.matrix),
                           meta$id_number),
                  "date_time_string"]

time <- meta[match(colnames(sqrt.zscore.matrix),
                           meta$id_number),
                  "time_string"]

```


#### Organize the samples (ChatGPT)
```{r}

## organize the treatment in order they will be display on the heatmap
treatment2 <- factor(treatment$treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))

## organize the time in order they will be display on the heatmap
time2 <- factor(time$time_string, levels = c("12:00:00", "18:00:00", "00:00:00", "06:00:00"))

#store that info in a df
col_info <- data.frame(
  Treatment = treatment2,
  Time = time2) %>% drop_na()

# create the order in which the samples with be organize in the heatmap
sample_order <- order(col_info$Treatment,col_info$Time)

# tell the matrix with my z-scores how the samples should be arranged
mat_ordered <- sqrt.zscore.matrix[, sample_order]

# have vectors specifying the order of the treatment for the heatmap
treatment_ordered <- col_info$Treatment[sample_order]
time_ordered <- col_info$Time[sample_order]


```

```{r}
## Create Annotation for the heatmap 
col_ha <- HeatmapAnnotation(
  #Treatment = treatment_ordered,
  
  ## Creates annotations blocks from treatments
  Treatment = anno_block(
    
  gp = gpar(fill = c("grey", "#03A9F4", "#EF5350", "#4CAF50")), # choose label colors
  #choose the name of the blocks
        labels = c("Control", "Coral", "CCA", "Algae"), 
  # customize the font color, size and font type
        labels_gp = gpar(col = "white", fontsize = 14,fontface = "bold")),
  
  # Creates clustering for time
  Time = time_ordered, 
  
  #Specifying the labels and the colors of the cluster groups.
  ## here Treatment is not very useful bcs they are specified above already
  col = list(Treatment = c("Control"="grey", 
                           "Coral"="#03A9F4", "CCA"="#EF5350",
                           "Algae"="#4CAF50"),
             Time = c("12:00:00" = "black",
                          "18:00:00" = "green", 
                          "00:00:00" = "blue",
                          "06:00:00" = "grey"))
)

```





#### Printing heatmap

```{r, fig.height = 16, fig.width = 24}
# Version 2
#png(width=1600, height=1000, file= here("outputs", "zscore_24hdata.png"))

ht = Heatmap(mat_ordered, 
             name = "sqrt z-scored", 
             heatmap_legend_param = list(legend_direction = "horizontal"),
             show_row_names = TRUE,
        row_names_max_width = unit(18, "cm"),
        clustering_method_rows = "complete",
        clustering_distance_rows = "euclidean",
             #rect_gp = gpar(col = "white", lwd = 0.1),
        
        column_title = "Bacterial Community Shift Over 24h", 
        column_title_gp = gpar(fontsize = 18, fontface = "bold"),
        #column_title_side = "bottom",
        cluster_rows = TRUE,
        cluster_columns =  FALSE,
        row_title = "Bacterial Families",
        row_title_gp = gpar(fontsize = 14, fontface = "bold"),
        row_names_gp = gpar(fontsize = 11),
        row_dend_width = unit(1.5, "cm"),
        #column_names_gp = gpar(fontsize = 10),
        show_column_names = FALSE,
        row_km = 3, # 5? number of forced segregation in rows 
        #column_km = 4,#number of forced segregation in columns 
        top_annotation = col_ha,
        #bottom_annotation = col_ha,
        column_split = treatment_ordered
        )
set.seed(1999)
ht_drawn <- draw(ht, heatmap_legend_side = "bottom",annotation_legend_side = "bottom", 
  merge_legends = TRUE)

#dev.off()

```

### Family Clusters and modules 

```{r}

# 4. Extract row positions grouped by cluster slice
cluster_list <- row_order(ht_drawn)

# 5. Convert the list into a clean, unified data frame matching gene names to cluster IDs
cluster_membership <- do.call(rbind, lapply(names(cluster_list), function(cl_name) {
  row_indices <- cluster_list[[cl_name]]
  data.frame(
    taxa = rownames(mat_ordered)[row_indices],
    cluster = paste("Cluster", cl_name)
  )
}))

cluster_membership
```


```{r}
modules <- taxa_filtered %>%  
  left_join(cluster_membership) 
  
```

```{}
modules <- modules %>%  
  mutate(cluster2 = cluster, 
         cluster = ifelse(cluster2 == "Cluster 1", "Oligotroph",
                          ifelse(cluster2 == "Cluster 2", "Copiotroph", "Generalist")),
         cluster = factor(cluster, levels = c("Generalist", "Oligotroph", "Copiotroph")))
```




## Plotting 

```{r}
modules_24h <- modules %>% 
  mutate(time_fac = factor(time, levels = c("12:00:00", "18:00:00", "00:00:00", "06:00:00")),
         treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae")))
```


```{r Family clustering, fig.width= 10, fig.height= 5}
modules_24h %>% drop_na(cluster) %>%
  group_by(id_number, cluster) %>% 
  reframe(sum_abund = sum(abund), time_fac, treatment) %>% 
  ungroup() %>% 
  group_by(treatment, time_fac, cluster) %>%
  reframe(mean_abund = mean(sum_abund),
          sd_abund   = sd(sum_abund),
          sum_abund) %>%
  ungroup() %>% 
  ggplot(aes(x = time_fac, y = mean_abund, group = treatment, color = treatment)) +
  geom_line() +
  geom_point()+
  geom_ribbon(aes(ymin = mean_abund - sd_abund,
                  ymax = mean_abund + sd_abund,
                  fill = treatment),
              alpha = 0.15, color = NA) +
  geom_point(aes(x = time_fac, y = sum_abund), shape = 1, alpha = 0.1) +
  labs(title = "Modules from Family Clustering", 
       x= "Time", 
       y= "Relative Abundance")+
  scale_color_manual(values = treatment_colors) +
  scale_fill_manual(values  = treatment_colors) +
  facet_wrap(. ~ cluster)+
  nunu_theme
```

# Stats 

## Taxa abundance 

### Data clean up 

```{r}
mobeast <- read_csv(here("data", "MOBEAST_full_merged_data_long.csv"))

```

```{r}
flow_meta <- mobeast %>%  
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) %>% 
  select(id_number, flowrate) %>% 
  distinct() %>% 
  filter(!is.na(id_number))
```

```{r}
DNA <- OTU_htm %>% 
  left_join(flow_meta) %>% 
  mutate(date_string = as.character(date_string),
         time_string = as.character(time_string), 
         date_time_string = as.character(date_time_string)) %>% 
  mutate(OTU_tax = paste(otu_number, "X", family, "X", order,"X", class),
         log_flow = log10(flowrate),
         time_fac = factor(time_string, 
                           levels = c("12:00:00", "18:00:00", "00:00:00","06:00:00"))) %>% 
  select(-long_name)

DNA_meta <- DNA %>% select(id_number, tank_number, treatment, time_fac, log_flow ) %>% distinct()
```




## Models by OTU 

```{}
OTU_24h <- DNA %>%  
  
  # filtering out rare taxa
  ## removing taxa that have a max abundance >0.5%
  group_by(OTU_tax) %>% 
  mutate(max = max(abund)) %>%
  ungroup() %>% 
  filter(max >0.005) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, OTU_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(OTU_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.001)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  
  
  # outlier detection
  mutate(sqrt_abund = sqrt(abund)) %>% 
  group_by(OTU_tax,treatment) %>%
  mutate(mean = mean(sqrt_abund),
         sd = sd(sqrt_abund),
         outlier = ifelse(sqrt_abund > mean + 2*sd | sqrt_abund < mean - 2*sd, "outlier", "not outlier")) %>%
  ungroup() %>% 
  filter(outlier != "outlier") %>%
  select(-mean, -sd) %>% 
  
  # calculating delta values 
  mutate(inflow_isolated = ifelse(treatment == "Inflow", abund, NA)) %>% 
  group_by(time, OTU_tax) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = T )) %>% 
  ungroup() %>%
  mutate(conc_norm  = abund - inflow ) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm))
  

   
  
```


```{r}
# Sqrt transform and Z-scoring 
OTU_24h <- DNA %>%  
  
  # filtering out rare taxa
  ## removing taxa that have a max abundance >0.5%
  group_by(OTU_tax) %>% 
  mutate(max = max(abund)) %>%
  ungroup() %>% 
  filter(max >0.005) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, OTU_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(OTU_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.001)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  
  
  # outlier detection
  mutate(sqrt_abund = sqrt(abund)) %>% 
  group_by(OTU_tax,treatment) %>%
  mutate(mean = mean(sqrt_abund),
         sd = sd(sqrt_abund),
         outlier = ifelse(sqrt_abund > mean + 2*sd | sqrt_abund < mean - 2*sd, "outlier", "not outlier")) %>%
  ungroup() %>% 
  filter(outlier != "outlier") %>%
  select(-mean, -sd) %>% 
  
  # calculating delta values
  #group_by(id_number) %>% 
  mutate(conc_norm  = zscore.calculation(sqrt_abund)) %>% 
  ungroup() %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>%
  filter(!str_detect(OTU_tax, "unclassified")) 
  
```


```{}
# Delta no transformation, no z-scoring 
OTU_24h <- DNA %>%  
  
  # filtering out rare taxa
  ## removing taxa that have a max abundance >0.5%
  group_by(OTU_tax) %>% 
  mutate(max = max(abund)) %>%
  ungroup() %>% 
  filter(max >0.005) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, OTU_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(OTU_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.001)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  
  
  # outlier detection
  group_by(OTU_tax,treatment) %>%
  mutate(mean = mean(abund),
         sd = sd(abund),
         outlier = ifelse(abund > mean + 2*sd | abund < mean - 2*sd, "outlier", "not outlier")) %>%
  ungroup() %>% 
  filter(outlier != "outlier") %>%
  select(-mean, -sd) %>% 
  
  # calculating delta values
  mutate(inflow_isolated = ifelse(treatment == "Inflow", abund, NA)) %>% 
  group_by(time, OTU_tax) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = T )) %>% 
  ungroup() %>%
  mutate(conc_norm  = abund - inflow ) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae")))
```


```{r}
exclude_tax <- OTU_24h %>% 
  group_by(time, OTU_tax) %>% 
  reframe(avg = mean(conc_norm)) %>% 
  filter(avg == 0) 
exclude_tax
```

```{r}
mobeast_loop <- OTU_24h %>% 
  filter(!OTU_tax %in% exclude_tax$OTU_tax & !OTU_tax %in% exclude_tax$time) %>%
  select(OTU_tax) %>% distinct() %>% pull()

models_DNA_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_DNA_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- OTU_24h %>% filter(OTU_tax == var)
  models_DNA_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + 
                                  log_flow + (1|tank_number), data = df)

}
```


```{r}
resid_panel(models_DNA_24h[["Otu0024 X Rhodobacteraceae X Rhodobacterales X Alphaproteobacteria"]], plots = c("qq"))
```



```{r}
results_DNA <- imap_dfr(models_DNA_24h, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")

results_DNA %>%  filter(p.value <0.05)
```

#### OTU ANOVA

```{r, warning= FALSE}

anova_results <- imap_dfr(models_DNA_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

anova_results %>% filter(term == "treatment", p.value <0.05)

otu_anova_tab <- anova_results %>% filter(term == "treatment", p.value <0.05) %>% gt()
gtsave(otu_anova_tab, "otu_anova_tab.png", path = here("outputs"))
```

#### OTU Emmeans 

```{r, message= FALSE}

emmeans_results_dna <- imap_dfr(models_DNA_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae")

emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% select(variable, contrast, estimate, adj.p.value)
```

```{r}
otu_contrast_tab <- emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% 
  select(variable, contrast, estimate, adj.p.value) %>% 
  gt()
gtsave(otu_contrast_tab, "otu_contrast_tab.png", path = here("outputs"))
```

```{r}

# Get a sorted, unique list of taxa
taxa_list <- OTU_24h %>% distinct(OTU_tax) %>% pull(OTU_tax) %>% sort()

# Function to build one QQ plot for a given taxon
make_qq <- function(taxon) {
  df <- OTU_24h %>% filter(OTU_tax == taxon)
  
  ggplot(df, aes(sample = conc_norm )) +
    stat_qq() +
    stat_qq_line(color = "red") +
    labs(title = taxon) +
    theme_minimal(base_size = 8) +
    theme(
      plot.title = element_text(size = 7, hjust = 0.5),
      axis.title = element_blank(),
      axis.text = element_text(size = 5)
    )
}

# Build the list of plots
qq_plots <- map(taxa_list, make_qq)

# Split into chunks of 16 and assemble each chunk into a 4x4 grid
chunks <- split(qq_plots, ceiling(seq_along(qq_plots) / 16))

grid_figures <- map(chunks, ~ wrap_plots(.x, ncol = 4, nrow = 4))

# Save all grids to the outputs folder, one file per page
walk2(grid_figures, seq_along(grid_figures), function(fig, i) {
  ggsave(
    filename = here("outputs", paste0("qq_grid_", i, ".png")),
    plot = fig,
    width = 14, height = 14, dpi = 300
  )
})
```



```{r}
OTU_24h %>%  filter(OTU_tax == "Otu0023 X Flavobacteriaceae X Flavobacteriales X Bacteroidia") %>% 
  ggplot(aes(x= treatment, y = conc_norm, fill = treatment)) +
  geom_violin()
```


```{r}
sig_taxa <- emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% select(variable) %>%
  distinct() %>% pull()

emmeans_results_dna %>% filter(variable %in% sig_taxa) %>% 
  mutate(contrast = factor(contrast, levels = c("Control - Coral", "Control - CCA", "Control - Algae"))) %>% 
  mutate(sig = ifelse(estimate <0 & adj.p.value < 0.05, "Enrichment", 
                      ifelse(estimate > 0 & adj.p.value < 0.05, 
                             "Depletion", "No change"))) %>% 
  ggplot(aes(x= contrast, y = variable, fill = sig)) + 
  geom_tile(color = "black") +
  scale_fill_manual(breaks = c("Enrichment", "Depletion", "No change"),
                    values = c("lightgreen", "#EF5350", "#03A9F4")) +
  theme(legend.position = "bottom")+
  theme_bw()
```





## Models by family 

```{r}
family_24h <- DNA %>%  
  
  # filtering out rare taxa
  ## removing taxa that have a max abundance >0.5%
  group_by(family_tax) %>% 
  mutate(max = max(abund)) %>%
  ungroup() %>% 
  filter(max >0.005) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, family_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(family_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.001)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  
  
  # outlier detection
  mutate(sqrt_abund = sqrt(abund)) %>% 
  group_by(family_tax,treatment) %>%
  mutate(mean = mean(sqrt_abund),
         sd = sd(sqrt_abund),
         outlier = ifelse(sqrt_abund > mean + 2*sd | sqrt_abund < mean - 2*sd, "outlier", "not outlier")) %>%
  ungroup() %>% 
  filter(outlier != "outlier") %>%
  select(-mean, -sd) %>% 
  
  # calculating delta values
  mutate(conc_norm  = zscore.calculation(sqrt_abund)) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae")))
  

   
  
```


```{r}
exclude_tax <- family_24h %>% 
  group_by(time, family_tax) %>% 
  reframe(avg = mean(conc_norm)) %>% 
  filter(avg == 0) 
exclude_tax
```

```{r}
mobeast_loop <- family_24h %>% #filter(!OTU_tax %in% exclude_tax$OTU_tax & !OTU_tax %in% exclude_tax$time) %>%
  select(family_tax) %>% distinct() %>% pull()

models_DNA_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_DNA_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- family_24h %>% filter(family_tax == var)
  models_DNA_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + 
                                  log_flow + (1|tank_number), data = df)

}
```


```{r}
resid_panel(models_DNA_24h[["Arenicellaceae X Arenicellales X Gammaproteobacteria X Proteobacteria"]], plots = c("qq"))
```



```{r}
results_DNA <- imap_dfr(models_DNA_24h, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")

results_DNA %>%  filter(p.value <0.05)
```

#### Family ANOVA

```{r, warning= FALSE}

anova_results <- imap_dfr(models_DNA_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

anova_results %>% filter(term == "treatment", p.value <0.05) %>% select(variable, term, p.value)

```

#### Family Emmeans 

```{r, message= FALSE}

emmeans_results_dna <- imap_dfr(models_DNA_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae")

emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% 
  select(variable, term, contrast, adj.p.value)
```

```{r}
sig_taxa <- emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% select(variable) %>%
  distinct() %>% pull()

emmeans_results_dna %>% filter(variable %in% sig_taxa) %>% 
  mutate(contrast = factor(contrast, levels = c("Control - Coral", "Control - CCA", "Control - Algae"))) %>% 
  mutate(sig = ifelse(estimate <0 & adj.p.value < 0.05, "Enrichment", 
                      ifelse(estimate > 0 & adj.p.value < 0.05, 
                             "Depletion", "No change"))) %>% 
  ggplot(aes(x= contrast, y = variable, fill = sig)) + 
  geom_tile(color = "black") +
  scale_fill_manual(breaks = c("Enrichment", "Depletion", "No change"),
                    values = c("lightgreen", "#EF5350", "#03A9F4")) +
  theme(legend.position = "bottom")+
  theme_bw()
```



## Models by genus 

```{r}
genus_24h <- DNA %>%  
  
  # filtering out rare taxa
  ## removing taxa that have a max abundance >0.5%
  group_by(genus_tax) %>% 
  mutate(max = max(abund)) %>%
  ungroup() %>% 
  filter(max >0.005) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, genus_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(genus_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.001)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  
  
  # outlier detection
  mutate(sqrt_abund = sqrt(abund)) %>% 
  group_by(genus_tax, treatment) %>%
  mutate(mean = mean(sqrt_abund),
         sd = sd(sqrt_abund),
         outlier = ifelse(sqrt_abund > mean + 2*sd | sqrt_abund < mean - 2*sd, "outlier", "not outlier")) %>%
  ungroup() %>% 
  filter(outlier != "outlier") %>%
  select(-mean, -sd) %>% 
  
  # calculating delta values
  mutate(conc_norm  = zscore.calculation(sqrt_abund)) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>%
  filter(!str_detect(genus_tax, "unclassified")) 
  
```


```{r}
exclude_tax <- genus_24h %>% 
  group_by(time, genus_tax) %>% 
  reframe(avg = mean(conc_norm)) %>% 
  filter(avg == 0) 
exclude_tax
```

```{r}
mobeast_loop <- genus_24h %>% #filter(!OTU_tax %in% exclude_tax$OTU_tax & !OTU_tax %in% exclude_tax$time) %>%
  select(genus_tax) %>% distinct() %>% pull()

models_DNA_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_DNA_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- genus_24h %>% filter(genus_tax == var)
  models_DNA_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + 
                                  log_flow + (1|tank_number), data = df)

}
```


```{r}
resid_panel(models_DNA_24h[["Vibrio X Vibrionaceae X Vibrionales X Gammaproteobacteria X Proteobacteria"]], plots = c("qq"))
```


#### Genus ANOVA


```{r, warning= FALSE}

anova_results <- imap_dfr(models_DNA_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 


anova_results %>% filter(term == "treatment", p.value <0.05) %>% 
  select(variable, term, p.value) 

```

#### Genus Emmeans 

```{r, message= FALSE}

emmeans_results_dna <- imap_dfr(models_DNA_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae")


emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% 
  select(variable, contrast, adj.p.value) 


```



```{r}
sig_taxa <- emmeans_results_dna %>% filter(adj.p.value < 0.05) %>% select(variable) %>%
  distinct() %>% pull()

emmeans_results_dna %>% filter(variable %in% sig_taxa) %>% 
  mutate(contrast = factor(contrast, levels = c("Control - Coral", "Control - CCA", "Control - Algae"))) %>% 
  mutate(sig = ifelse(estimate <0 & adj.p.value < 0.05, "Enrichment", 
                      ifelse(estimate > 0 & adj.p.value < 0.05, 
                             "Depletion", "No change"))) %>% 
  ggplot(aes(x= contrast, y = variable, fill = sig)) + 
  geom_tile(color = "black") +
  scale_fill_manual(breaks = c("Enrichment", "Depletion", "No change"),
                    values = c("lightgreen", "#EF5350", "#03A9F4")) +
  theme(legend.position = "bottom")+
  theme_bw()
```



## Z-score Delta Heatmap 

```{r}
# OTU 

# Delta no transformation, no z-scoring but **summed otu** used to make the heatmap
OTU_24h <- DNA %>%  
  
  filter(!str_detect(family_tax, "unclassified")) %>%
  
  # filtering out rare taxa
  ## removing taxa that have a max abundance >0.5%
  group_by(OTU_tax) %>% 
  mutate(max = max(abund)) %>%
  ungroup() %>% 
  filter(max >0.005) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, OTU_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(OTU_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.001)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  
  
  # outlier detection
  group_by(OTU_tax,treatment) %>%
  mutate(mean = mean(abund),
         sd = sd(abund),
         outlier = ifelse(abund > mean + 2*sd | abund < mean - 2*sd, "outlier", "not outlier")) %>%
  ungroup() %>% 
  filter(outlier != "outlier") %>%
  select(-mean, -sd) %>% 
  
  # calculating delta values
  mutate(inflow_isolated = ifelse(treatment == "Inflow", abund, NA)) %>% 
  group_by(time, OTU_tax) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = T )) %>% 
  ungroup() %>%
  mutate(conc_norm  = abund - inflow ) %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate( z_score_conc = zscore.calculation(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae")))
```


```{r}
# Family 

# Delta no transformation, no z-scoring but **summed otu** used to make the heatmap
OTU_24h_filtered <- DNA %>%  
  
  # filtering out rare taxa
  ## removing taxa that have a max abundance >0.5%
  group_by(family_tax) %>% 
  mutate(max = max(abund)) %>%
  ungroup() %>% 
  filter(max >0.005) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, family_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(family_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.001)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  
  
  # outlier detection
  group_by(family_tax,treatment) %>%
  mutate(mean = mean(abund),
         sd = sd(abund),
         outlier = ifelse(abund > mean + 2*sd | abund < mean - 2*sd, "outlier", "not outlier")) %>%
  ungroup() %>% 
  filter(outlier != "outlier") %>%
  select(-mean, -sd) 


  # calculating delta values
OTU_24h <- OTU_24h_filtered %>% 
  
  filter(!str_detect(family_tax, "unclassified")) %>% 
  
  mutate(inflow_isolated = ifelse(treatment == "Inflow", abund, NA)) %>% 
  group_by(time, family_tax) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = T )) %>% 
  ungroup() %>%
  
   group_by(id_number, treatment, time, family_tax) %>% 
  mutate(sum_abund = sum(abund, na.rm = T),
    conc_norm  = sum_abund - inflow ) %>% 
    ungroup() %>% 
  
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate(z_score_conc = zscore.calculation(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  select(id_number, tank_number, treatment, date,time,time_fac,date_time, inflow_table,
         z_score_conc,family_tax) %>% distinct()
```

```{}
# Genus 

# Delta no transformation, no z-scoring but **summed otu** used to make the heatmap
OTU_24h_filtered <- DNA %>%  
  
  # filtering out rare taxa
  ## removing taxa that have a max abundance >0.5%
  group_by(genus_tax) %>% 
  mutate(max = max(abund)) %>%
  ungroup() %>% 
  filter(max >0.005) %>% 
  
  ## remove taxa that don't have an abundance >0.1% in at least 3 samples
  group_by(id_number, genus_tax) %>% 
  mutate(sum_abund= sum(abund)) %>% 
  ungroup() %>% 
  group_by(genus_tax) %>%
  mutate(n_above_threshold = sum(sum_abund >= 0.001)) %>%
  filter(n_above_threshold >= 3) %>% 
  ungroup() %>% 
  
  
  # outlier detection
  group_by(genus_tax,treatment) %>%
  mutate(mean = mean(abund),
         sd = sd(abund),
         outlier = ifelse(abund > mean + 2*sd | abund < mean - 2*sd, "outlier", "not outlier")) %>%
  ungroup() %>% 
  filter(outlier != "outlier") %>%
  select(-mean, -sd) 
  



  # calculating delta values
OTU_24h <- OTU_24h_filtered %>% 
  
  filter(!str_detect(genus_tax, "unclassified")) %>% 
  
  mutate(inflow_isolated = ifelse(treatment == "Inflow", abund, NA)) %>% 
  group_by(time, genus_tax) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = T )) %>% 
  ungroup() %>%
  
   group_by(id_number, treatment, time, genus_tax) %>% 
  mutate(sum_abund = sum(abund, na.rm = T),
    conc_norm  = sum_abund - inflow ) %>% 
    ungroup() %>% 
  
  filter(treatment != "Inflow") %>% 
  filter(!is.na(conc_norm)) %>% 
  mutate(z_score_conc = zscore.calculation(conc_norm)) %>% 
  mutate(treatment= factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  select(id_number, tank_number, treatment, date,time,time_fac,date_time, inflow_table,
         z_score_conc,genus_tax) %>% distinct()
```

```{}
OTU_24h %>% filter(genus_tax == "Vibrio X Vibrionaceae X Vibrionales X Gammaproteobacteria X Proteobacteria")
```

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_24h %>%
  group_by(id_number, family_tax) %>%
  summarize(z_score_conc) %>% 
  ungroup()

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = family_tax, values_from = z_score_conc)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% as.data.frame() 

```


```{}
sqrt.matrix <- apply(culled.abund.OTU, 2, FUN=sqrt)
sqrt.zscore <- as.data.frame(apply(sqrt.matrix, 2, FUN = zscore.calculation))
```

```{r}
# Set the ID numbers as the rownames (to make a matrix)
rownames(culled.abund.OTU) = culled.abund.OTU$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
delta.matrix <- culled.abund.OTU %>% select(-id_number) %>% 
  as.matrix() %>% t()


```

### Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(delta.matrix),
                           meta$id_number), 
                 "treatment"]

date_time <- meta[match(colnames(delta.matrix),
                           meta$id_number),
                  "date_time_string"]

time <- meta[match(colnames(delta.matrix),
                           meta$id_number),
                  "time_string"]

```


#### Organize the samples (ChatGPT)
```{r}

## organize the treatment in order they will be display on the heatmap
treatment2 <- factor(treatment$treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))

## organize the time in order they will be display on the heatmap
time2 <- factor(time$time_string, levels = c("12:00:00", "18:00:00", "00:00:00", "06:00:00"))

#store that info in a df
col_info <- data.frame(
  Treatment = treatment2,
  Time = time2) %>% drop_na()

# create the order in which the samples with be organize in the heatmap
sample_order <- order(col_info$Treatment,col_info$Time)

# tell the matrix with my z-scores how the samples should be arranged
mat_ordered <- delta.matrix[, sample_order]

# have vectors specifying the order of the treatment for the heatmap
treatment_ordered <- col_info$Treatment[sample_order]
time_ordered <- col_info$Time[sample_order]


```

```{r}
## Create Annotation for the heatmap 
col_ha <- HeatmapAnnotation(
  #Treatment = treatment_ordered,
  
  ## Creates annotations blocks from treatments
  Treatment = anno_block(
    
  gp = gpar(fill = c("grey", "#03A9F4", "#EF5350", "#4CAF50")), # choose label colors
  #choose the name of the blocks
        labels = c("Control", "Coral", "CCA", "Algae"), 
  # customize the font color, size and font type
        labels_gp = gpar(col = "white", fontsize = 14,fontface = "bold")),
  
  # Creates clustering for time
  Time = time_ordered, 
  
  #Specifying the labels and the colors of the cluster groups.
  ## here Treatment is not very useful bcs they are specified above already
  col = list(Treatment = c("Control"="grey", 
                           "Coral"="#03A9F4", "CCA"="#EF5350",
                           "Algae"="#4CAF50"),
             Time = c("12:00:00" = "black",
                          "18:00:00" = "green", 
                          "00:00:00" = "blue",
                          "06:00:00" = "grey"))
)

```


#### Printing heatmap

```{r, fig.height = 16, fig.width = 24}
# Version 2
#png(width=1600, height=1000, file= here("outputs", "zscore_24hdata.png"))

ht = Heatmap(mat_ordered, 
             name = "sqrt z-scored", 
             heatmap_legend_param = list(legend_direction = "horizontal"),
             show_row_names = TRUE,
             na_col = "black", 
        row_names_max_width = unit(18, "cm"),
        clustering_method_rows = "complete",
        clustering_distance_rows = "euclidean",
             #rect_gp = gpar(col = "white", lwd = 0.1),
        
        column_title = "Bacterial Community Shift Over 24h", 
        column_title_gp = gpar(fontsize = 18, fontface = "bold"),
        #column_title_side = "bottom",
        cluster_rows = TRUE,
        cluster_columns =  FALSE,
        row_title = "Bacterial Families",
        row_title_gp = gpar(fontsize = 14, fontface = "bold"),
        row_names_gp = gpar(fontsize = 11),
        row_dend_width = unit(1.5, "cm"),
        #column_names_gp = gpar(fontsize = 10),
        show_column_names = FALSE,
        #row_km = 3, # 5? number of forced segregation in rows 
        #column_km = 4,#number of forced segregation in columns 
        top_annotation = col_ha,
        #bottom_annotation = col_ha,
        column_split = treatment_ordered
        )
set.seed(1999)
ht_drawn <- draw(ht, heatmap_legend_side = "bottom",annotation_legend_side = "bottom", 
  merge_legends = TRUE)

#dev.off()

```





```{r, fig.height= 16, fig.width=12}
library(RColorBrewer)

OTU_24h %>% 
  ggplot(aes(x = time_fac, y = OTU_tax, fill = conc_norm)) + 
  geom_tile(color = "black") +
  facet_grid(. ~ treatment, scales = "free_x", space = "free_x") +
  scale_fill_distiller(palette = "RdBu") +
  labs(x = "Time point", y = NULL, fill = "Concentration") +
  theme_bw() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold"),
    panel.spacing = unit(0.3, "lines")
  )
```