---
title: "MOBEAST_all_variables_delta"
created: 5 Fev 2025
edited: 25 Fev 2025
format: html
editor: source

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(janitor)
library(here)

library(ggrepel)
library(patchwork)
library(RColorBrewer)
library(measurements)
library(gt)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
library(emmeans)
library(performance)
library(multcompView)
library(broom.mixed)
library(effectsize)


# PCA packages
library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(factoextra) 
```


## Data visualization theme

```{r}
nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=24, face = "bold"), 
        plot.subtitle = element_text(size=20),
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=18, face = "bold"),
        legend.text=element_text(size=16),
        strip.background = element_rect(fill = "white"))
```



```{r}
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

## Load Data

```{r}
mobeast_raw <- read_csv(here("data", "MOBEAST_full_merged_data_long.csv"))

```


## Data Clean Up
### Outlier detection on bulk data

```{r}
## flagging outliers from the data set
mobeast<-  mobeast_raw %>%
  mutate(time_string = as.character(time_string),
              date_string = as.character(date_string),
              date_time_string = paste(date_string,time_string))%>% 
  filter(!is.na(id_number)) %>% 
  
  # remove the DNA data because they should be treated differently
  filter(variable_cat != "DNA" ) %>% 
  filter(treatment != "Rain") %>% 
  filter(treatment != "Endmember")  %>% 
  mutate(units = ifelse(variable_cat == "FCM", "cell/µL", 
                        ifelse(variable_cat == "DOC", "µmol/L", 
                        ifelse(variable_cat == "fDOM", "R.U.",
                               ifelse(variable_cat == "nutrients", "µmol/L",
                                      ifelse(variable_cat == "chla", "µg/L", NA))))))
  
```

```{r}

DOC <- mobeast %>% filter(variable == "npoc_u_m") %>%  
  mutate(inflow = ifelse(treatment == "Inflow", value, NA)) %>% 
  mutate(inflow = ifelse(inflow > 208, NA, inflow)) %>% 
  mutate(value = ifelse(treatment == "Inflow" & value > 208, mean(inflow, na.rm= TRUE), value)) %>% 
  mutate(value = ifelse(value >208, NA, value)) %>% select(-inflow)

mobeast_DOC_avg <- mobeast %>% filter(variable != "npoc_u_m") %>% 
  rbind(DOC)
```


```{r}
mobeast_flagged <- mobeast_DOC_avg %>% 
  # flagging outliers
  ## grouping by treatment, variable 
  group_by(variable) %>% 
  
  ## get the mean and the sd to calculate mean +/- 2sd
  mutate(mean =mean(value, na.rm = T), 
         stand_dev = sd(value, na.rm = T)) %>% 
  
  ## Values outside of the mean +/- 2sd range are flagged as outliers
  ## Important to filter outliers later to understand why they are outliers. 
  mutate(outlier = ifelse(value > mean +2*stand_dev | value < mean -2*stand_dev,
                           "outlier", "not outlier")) %>% 
  ungroup() 
  
```



### Biogeochemistry 

```{r}

mobeast_delta <- mobeast_flagged %>%  
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table, variable) %>%
  #mutate(inflow = value[treatment == "Inflow"]) %>% 
  mutate(inflow_isolated = ifelse(treatment == "Inflow", value, NA)) %>% 
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time, variable) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  
  # Normalize for treatments
  mutate(conc_norm = value - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) %>% 
  select(-mean, -stand_dev, -inflow, -residence_time, - inflow_isolated)

```

```{r}
## Some variables will not be studied in this project
exclude_var <- c("carbonate", "residence time")

# we also won't focus on the other time points, they are mainly for the FCM data and not other data was collected at those times anyway
exclude_time <- c("03:00:00","09:00:00","15:00:00","21:00:00")

mobeast_delta <- mobeast_delta %>% filter(!variable_cat %in%  exclude_var,
                        !time_string %in% exclude_time) 
```


## 24h data

```{r}
# subset the data set to only the 24h samples
## removing outliers from the data set
mobeast_24h <- mobeast_delta %>%
  filter(!is.na(id_number)) %>%
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  # remove the the DOC data because something happened  at 00:00. 
 filter(treatment != "Inflow") %>% 
  ## remove the outliers 
  #filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00"))) %>% mutate(log_flow = log10(flowrate)) 

```

### Summary Stats 
```{r}
mobeast_delta %>%
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  filter(!is.na(id_number)) %>% 
  # remove the the DOC data because something happened  at 00:00. 
  #filter(variable == "npoc_u_m") %>%  
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  group_by(variable,time_string, treatment ) %>% 
  reframe(max = max(conc_norm, na.rm = TRUE),
          min = min(conc_norm, na.rm = TRUE),
    mean = mean(conc_norm, na.rm = TRUE),
          sd = sd(conc_norm, na.rm = TRUE)) %>% 
  filter(variable == "npoc_u_m") %>%  
  arrange(mean)
```



### LMER models

```{r}
# Loop Function:
## Take the 24h samples and isolate the types of nutrient  present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

mobeast_loop <- mobeast_24h %>% select(variable) %>% distinct() %>% pull()

models_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- mobeast_24h %>% filter(variable == var)
  models_24h[[var]] <- lmer(conc_norm~ treatment*time_fac + log_flow +  
                            (1|tank_number), data = df)
}


```


```{r}
plot_grid(imap(models_24h, ~ {
  resid_panel(.x, plots = c("qq")) + plot_annotation(title = .y)}), 
margin = c(0.2, 0.2, 0.2, 0.2), tags = TRUE)

```

#### model results (fixed and random)
```{r}
results <- imap_dfr(models_24h, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")

results %>% filter(p.value < 0.05) 

```

#### ANOVA results

Your gatekeeper — if nothing is significant here, you stop
Reports F-statistic, df, and p-value for each term
Type II is appropriate because it accounts for the other terms in the model without assuming a particular order

```{r, warning= FALSE}
# So many Warnings! 
## Warning: The column names NumDF and DenDF in ANOVA output were not recognized or transformed.

anova_results <- imap_dfr(models_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

# all the variables that have significant treatment effects
sig_anova <- anova_results %>% 
  filter( p.value <0.05) %>%  filter(term == "treatment") %>%  
  select(variable) %>% pull()

# no signficant time interactions !
anova_results %>% 
  filter(p.value <0.05, term == "treatment:time_fac") %>%  
  select(variable)

anova_results %>% 
  #filter( p.value <0.05) %>% 
  filter(variable %in% c("het_bact", "syn"))

```


##### Nutrient ANOVA summary table 

```{r}

anova_results %>%
  filter(variable %in% c("NO3", "PO4", "NH3", "SiO2")) %>%
  filter(term != "log_flow") %>%
  select(variable, term, df = NumDF, F_stat = statistic, p.value) %>%
  mutate(
    p.value = round(p.value, 3),
    F_stat  = round(F_stat, 2)
  ) %>%
  mutate(term = recode_values(term,
      "treatment"      ~"Treatment",
      "time_fac"       ~ "Time",
      "treatment:time_fac" ~ "Treatment x Time"
    )) %>% 
  gt(groupname_col = "variable") %>%
  tab_header(
    title    = "ANOVA Results",
    subtitle = "Linear mixed model — Type II tests"
  ) %>%
  cols_label(
    term    = "Term",
    df      = "df",
    F_stat  = "F",
    p.value = "p-value"
  ) %>%
  #tab_style(
   # style = cell_text(bold = TRUE),
   # locations = cells_row_groups()) %>%
  tab_style(
    style = cell_fill(color = "#f0f0f0"),
    locations = cells_row_groups()
  ) %>%
  fmt_number(columns = F_stat, decimals = 2) %>%
  fmt_number(columns = p.value, decimals = 3) %>%
  tab_options(
    table.font.size  = 12,
    row_group.font.weight = "bold"
  ) %>% 
    tab_style(
  style = cell_text(weight = "bold", color = "red"),
  locations = cells_body(
    columns = p.value,
    rows    = p.value < 0.05
  )
)
```



#### Partial Eta squared 

Partial η²

Answers "so what?" after a significant ANOVA
Tells you whether the significant effect is actually large enough to care about biologically
A p < 0.05 with η² = 0.01 is statistically real but possibly biologically trivial


```{r}
eta_results <- tibble()

for (var in sig_anova) {
  df <- tibble(eta_squared(anova(models_24h[[var]]), partial = TRUE))
  df <- df %>% mutate(variable = var ) %>% relocate(variable)
  
  eta_results <- rbind(eta_results, df)
  
  eta_results <- eta_results 
  
}

sig_eta <- eta_results %>% filter(Eta2_partial > 0.14, variable != "tot_bact", variable != "cyano", variable != "N_N") %>% select(variable) %>% distinct() %>% pull()

```

#### emmeans results

Unpacks where the group differences are
Particularly important for your treatment * time interaction — you need to know at which time points treatments diverge
Controls for multiple comparisons


```{r}
# "estimate" are negative somehow. Why? 
## in manual models for DOC, estimate >0, here it's <0
## t.ratio in "statistics", and negative instead of positive
## p.values not shown, only p.adjusted. 
emmeans_results <- imap_dfr(models_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 

emmeans_results %>% #filter(variable %in% sig_eta) %>%
  filter(adj.p.value <0.05)
```



```{r}
emmeans_results_sig <- emmeans_results %>% 
  filter(adj.p.value <0.05)
emmean_sig <- emmeans_results_sig %>% select(variable) %>% distinct() %>%  pull(variable)
```



##### Nutrient Time contrasts

```{r}
## Nutrient time contrast
time_contrast <- tibble()

for (var in c("NO3", "PO4", "NH3", "SiO2")) {
  emm <- emmeans(models_24h[[var]], ~ time_fac | treatment)
  
  df <- contrast(emm, method = "trt.vs.ctrl") %>% 
    tidy() %>% mutate(variable = var) %>%  relocate(variable)
  
  
  time_contrast <- rbind(time_contrast, df)
}

time_contrast %>% filter(adj.p.value < 0.05)
```

##### FCM Time Contrasts

```{r}
## FCM time contrast
time_contrast <- tibble()

for (var in c("het_bact", "syn")) {
  emm <- emmeans(models_24h[[var]], ~ time_fac | treatment)
  
  df <- contrast(emm, method = "trt.vs.ctrl") %>% 
    tidy() %>% mutate(variable = var) %>%  relocate(variable)
  
  
  time_contrast <- rbind(time_contrast, df)
}

time_contrast %>% filter(adj.p.value < 0.05)
```


#### Emmeans if time interactions ! 
3b. If interaction is significant → decompose emmeans by time point
    emmeans(model, pairwise ~ treatment | time)
    
    If interaction is NOT significant → look at main effects only
    emmeans(model, pairwise ~ treatment)

NOT APPLICABLE BCS NO TIME INTERACTION IN THE ANOVA 

    

#### Contrast Effect Size 

Adds biological interpretability to each contrast
Tells you not just "these groups differ" but "by how much and in which direction"
Essential for comparing across nutrients since your residuals are on different scales


```{r}
effect_size_results <- tibble()

for (var in sig_eta) {
  emm <- emmeans(models_24h[[var]], ~ treatment)
  contrasts <- contrast(emm, method = "trt.vs.ctrl")

## Effect size for each contrast
  df <- tidy(eff_size(emm, sigma = sigma(models_24h[[var]]), 
                edf = df.residual(models_24h[[var]])))
  
  df <- df %>% mutate(variable = var) %>% rename(effect.size = estimate) %>% 
    relocate(variable)
  
  effect_size_results <- rbind(effect_size_results, df)
  
}

effect_size_results %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") %>% filter(abs(effect.size) >=0.8)



```




#### Violin plots

```{r}
meta_var <- mobeast_24h %>% select(variable) %>% distinct() %>% 
  mutate(variable_clean = c("Heterotrophic Bacteria", "Eukaryote", "Synechococcus",
                            "Prochlorococcus", "Total Bacteria", "Cyanobacteria", 
                            "Chlorophyll a", 
                            "TN", "Coble A", "Coble B", "Coble C", "Coble M",
                            "Coble T", "M:C", "BIX", "HIX", "Fpeak", "Fulvic Acid",
                            "Lignin", "N+N", "NO2", "PO4", "SiO2", "NH3", "NO3", "DOC"))

mobeast_24h <- mobeast_24h %>% left_join(meta_var, by = "variable") %>% 
  relocate(variable_clean, .after = "variable")
```

```{r, warning=FALSE}

#graph_loop <- mobeast_24h %>% filter(variable %in% sig_eta) %>% select(variable_clean) %>% 
  #distinct()
  
plots_24h <- vector(mode = "list", length = length(sig_eta))

for (var in sig_eta) {
  
  unit_plot <- mobeast_24h %>% filter(variable == var) %>% select(units) %>% distinct()
  
  variable_name <- mobeast_24h %>% filter(variable == var) %>% 
    select(variable_clean) %>% distinct()
  
  plots_24h[[var]] <- mobeast_24h %>% 
  filter(variable == var) %>% 
  ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    #geom_jitter(aes(color = outlier), alpha = 0.5)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_wrap( .~ time_fac, ncol = 4 )+
    labs(
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = paste0("Delta ", variable_name, " (",unit_plot, ")"))+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18),
          legend.position = "none")
}

```

```{r, fig.width=14, fig.height=10}
# fDOM plots 
(plots_24h[["coble_c"]]+ labs(title = "A")) / (plots_24h[["hix"]]+ labs(title = "B"))

ggsave(filename = "24h_fDOM.png", plot = last_plot(), path = here("outputs"))
```

```{r, fig.width=14, fig.height=10}
# FCM plots 
(plots_24h[["het_bact"]]+ labs(title = "A")) /(plots_24h[["syn"]]+ labs(title = "B"))

ggsave(filename = "24h_FCM.png", plot = last_plot(), path = here("outputs"))
```


```{r, fig.width=14, fig.height=7}
# Nutrients plots 

plots_24h[["NO3"]]

ggsave(filename = "24h_N03.png", plot = last_plot(), path = here("outputs"))
```



```{r, fig.width=14, fig.height=30}
# Nutrients plots 

(plots_24h[["coble_c"]]+ labs(title = "Fluorescent DOM", subtitle = "A - Coble C")) /
  (plots_24h[["hix"]]+ labs(subtitle = "B - HIX")) / 
  
  #(plots_24h[["NO3"]]+ labs(title = "Nutrients",subtitle = "C - NO3"))/ 
  
  (plots_24h[["het_bact"]]+ labs(title = "Cell Count",subtitle = "D - Heterotrophic Bacteria"))/ 
  (plots_24h[["syn"]]+ labs(subtitle = "E - Synechococcus"))

#ggsave(filename = "24h_sig_var.png", plot = last_plot(), path = here("outputs"))
```

##### Nutrients 

```{r}
 mobeast_24h %>% 
   filter(variable %in% c("NO3", "PO4", "NH3", "SiO2")) %>% 
  ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    #geom_jitter(aes(color = outlier), alpha = 0.5)+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_grid( variable ~ time_fac, scales = "free" )+
    labs(
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = paste0("Delta Nutrients (µM)"))+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  scale_color_manual(breaks = c("outlier", "not outlier"), 
                       values=c("red", "black"))+
    nunu_theme + 
    theme(axis.text = element_text(size = 18),
          legend.position = "none")
```


#### Summary Stats for Thesis

```{r}
mobeast_delta %>% filter(variable %in% sig_anova) %>% 
  group_by(variable, time, treatment) %>% 
  reframe(mean = mean(conc_norm, na.rm = TRUE),
          sd = sd(conc_norm, na.rm = TRUE))
```

## DOM x Het Bact
```{r}
mobeast_wide <- mobeast_24h %>% select(tank_number, treatment, inflow_table, 
                       time_fac,variable, conc_norm, log_flow) %>% 
  pivot_wider(names_from = variable, values_from = conc_norm)
```


### HIX - No sig ANOVA
```{r}
mod_hix <- lm(het_bact ~ hix, 
                  data= mobeast_wide)
anova(mod_hix)
```

### Coble C

```{r}
mobeast_wide_algae <- filter(mobeast_wide, treatment == "Algae")
cobC_lm <- lm(het_bact ~ coble_c,
                data= mobeast_wide_algae)

summary(cobC_lm) 

resid_panel(cobC_lm, plots = "qq") 

anova(cobC_lm, type=2)
```


```{r}
mobeast_wide %>% filter(treatment == "Algae") %>% 
  ggplot(aes(hix, het_bact, color = treatment, group = treatment)) +
  geom_point()+
  geom_smooth(method = "lm")+
  #facet_wrap(.~treatment)+
  scale_color_manual(
        breaks = c("Control", "Coral", "CCA", "Algae"),
        values = c("grey", "#03A9F4", "#EF5350", "#4CAF50")
      )  +
  nunu_theme
    
```




## Pairwise comparison 

### Pairwise comparison function

```{r}
nunu_pairwise <- function(data, type = "coble_a", time_point = 6){

  ## Get a list of the treatments for the for loop   
  treatment <- data %>% mutate(treatment = as.character(treatment)) %>%
  select(treatment) %>%  distinct() %>% pull()
  
  ## Create un empty data frame
pairwise_results <- tibble(
  time_point = vector(),
  treatment= vector(),
  mean = vector(),
  p_value = vector())

# this is the pairwise comparison for the 6 pairs (6AM-6PM)
  if(time_point == 6){
    for(var in treatment){
      
      ## only keep the 6AM and 6PM
      df <- data %>% 
        filter(variable == type) %>% 
        filter(treatment == var) %>% 
  filter(time_string == "18:00:00" | time_string == "06:00:00") %>% 
        ## Rename the times so they can become column names
  mutate(time_US = ifelse(time_string == "18:00:00", "T18", "T06")) %>% 
        ## Only keep the column of interest for the stats
  select(tank_number, time_US, conc_norm) %>% 
  pivot_wider(names_from = time_US, values_from = conc_norm) 
      
      ## Pairwise function 
      paired <- t.test(df$T18, df$T06, paired = TRUE, na.action = na.pass)
      
      ## Save the result in at data frame
      df2 <- tibble(
        time_point = "06AM-06PM",
        variable = type,
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      # Code will loop oveer all treatments and save the results in a data frame
      pairwise_results <-rbind(pairwise_results,df2)
      
      # Save the results in RStudio
      pairwise_results <<- pairwise_results
    }  
  

# This is for the 12 samples (12AM-12PM)
  }else if(time_point == 12){
    for(var in treatment){
      
      ## Same as before but for the 12AM-12PM
      df <- data %>% 
        filter(variable == type) %>% 
        filter(treatment == var) %>% 
  filter(time_string == "12:00:00" | time_string == "00:00:00") %>% 
  mutate(time_US = ifelse(time_string == "12:00:00", "T12", "T00")) %>% 
  select(tank_number, time_US, conc_norm) %>% 
  pivot_wider(names_from = time_US, values_from = conc_norm) 
      
      paired <- t.test(df$T12, df$T00, paired = TRUE, na.action = na.pass)
      
      df2 <- tibble(
        time_point = "12AM-12PM",
        variable = type,
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      pairwise_results <- rbind(pairwise_results, df2)
      pairwise_results <<- pairwise_results
    }
}

# Display the results of the code 
return(pairwise_results)
}

```



```{r}
nunu_pairwise(mobeast_24h, type = "npoc_u_m", time_point = 6)
```


## pH and TA 

```{r}
mobeast_carb <- mobeast_flagged %>%  filter(variable_cat == "carbonate") %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table, variable) %>%
  #mutate(inflow = value[treatment == "Inflow"]) %>% 
  mutate(inflow_isolated = ifelse(treatment == "Inflow", value, NA)) %>% 
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time, variable) %>% 
  mutate(inflow = mean(inflow_isolated, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  
  # Normalize for treatments
  mutate(conc_norm = value - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow"))) %>% 
  select(-mean, -stand_dev, -inflow, -residence_time, - inflow_isolated) %>% 
  filter(variable %in% c("ta", "ph")) %>% 
  filter(!time_string %in% exclude_time) %>% 
  filter(!is.na(id_number)) %>%
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  # remove the the DOC data because something happened  at 00:00. 
 filter(treatment != "Inflow") %>% 
  ## remove the outliers 
  #filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00"))) %>% mutate(log_flow = log10(flowrate)) 

```


### Carbonate LMER models

```{r}
# Loop Function:
## Take the 24h samples and isolate the types of nutrient  present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

mobeast_loop <- mobeast_carb %>% select(variable) %>% distinct() %>% pull()

models_carb <- vector(mode = "list", length = length(mobeast_loop))
names(models_carb) <- mobeast_loop


for (var in mobeast_loop) {
  df <- mobeast_carb %>% filter(variable == var)
  models_carb[[var]] <- lmer(conc_norm~ treatment*time_fac + log_flow +  
                            (1|tank_number), data = df)
}


```


```{r}
plot_grid(imap(models_carb, ~ {
  resid_panel(.x, plots = c("qq")) + plot_annotation(title = .y)}), 
margin = c(0.2, 0.2, 0.2, 0.2), tags = TRUE)

```

### ANOVA results

Your gatekeeper — if nothing is significant here, you stop
Reports F-statistic, df, and p-value for each term
Type II is appropriate because it accounts for the other terms in the model without assuming a particular order

```{r, warning= FALSE}
# So many Warnings! 
## Warning: The column names NumDF and DenDF in ANOVA output were not recognized or transformed.

anova_carb <- imap_dfr(models_carb, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

# all the variables that have significant treatment effects
anova_carb %>% 
  filter( p.value <0.05) %>%  filter(term == "treatment")

# no signficant time interactions !
anova_carb %>% 
  filter(p.value <0.05, term == "treatment:time_fac")



```




### Partial Eta squared 

Partial η²

Answers "so what?" after a significant ANOVA
Tells you whether the significant effect is actually large enough to care about biologically
A p < 0.05 with η² = 0.01 is statistically real but possibly biologically trivial


```{r}
eta_carb <- tibble()

for (var in c("ph", "ta")) {
  df <- tibble(eta_squared(anova(models_carb[[var]]), partial = TRUE))
  df <- df %>% mutate(variable = var ) %>% relocate(variable)
  
  eta_carb <- rbind(eta_carb, df)
  
  eta_carb <- eta_carb 
  
}

eta_carb %>% filter(Eta2_partial > 0.14)

sig_eta_carb <- eta_carb %>% filter(Eta2_partial > 0.14) %>% select(variable) %>% distinct() %>% pull()

```


### emmeans results

Unpacks where the group differences are
Particularly important for your treatment * time interaction — you need to know at which time points treatments diverge
Controls for multiple comparisons


```{r}
# "estimate" are negative somehow. Why? 
## in manual models for DOC, estimate >0, here it's <0
## t.ratio in "statistics", and negative instead of positive
## p.values not shown, only p.adjusted. 
emmeans_carb <- imap_dfr(models_carb, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 

emmeans_carb %>% #filter(variable %in% sig_eta) %>%
  filter(adj.p.value <0.05)
```

#### Contrast Effect Size 

Adds biological interpretability to each contrast
Tells you not just "these groups differ" but "by how much and in which direction"
Essential for comparing across nutrients since your residuals are on different scales


```{r}
effect_size_carb <- tibble()

for (var in c("ph", "ta")) {
  emm <- emmeans(models_carb[[var]], ~ treatment)
  contrasts <- contrast(emm, method = "trt.vs.ctrl")

## Effect size for each contrast
  df <- tidy(eff_size(emm, sigma = sigma(models_carb[[var]]), 
                edf = df.residual(models_carb[[var]])))
  
  df <- df %>% mutate(variable = var) %>% rename(effect.size = estimate) %>% 
    relocate(variable)
  
  effect_size_carb <- rbind(effect_size_carb, df)
  
}

effect_size_carb %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") %>% filter(abs(effect.size) >=0.8)



```


### Plots 


```{r}
# Define shading dataframe matching your exact factor levels
# Adjust period assignments to match your actual day/night schedule
shading <- data.frame(
  time_fac = levels(mobeast_carb$time_fac),
  period   = c("day", "night", "night", "day")  # adjust to your time points
)

ta <- mobeast_carb %>% 
  filter(variable == "ta") %>% 
  group_by(variable, time_fac, treatment) %>% 
  reframe(mean = mean(conc_norm, na.rm = T),
          sd   = sd(conc_norm, na.rm = T)) %>% 
  ggplot(aes(x = time_fac, y = mean, color = treatment, group = treatment)) +
  # Day/night shading — added before geom_point and geom_line
  geom_tile(
    data = shading,
    aes(x = time_fac, y = 0, fill = period),
    height = Inf, alpha = 0.3,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(
    values = c("day" = "lightyellow", "night" = "lightblue"),
    name   = "Period"
  ) +
  geom_point() +
  geom_line(aes(color = treatment)) +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  facet_grid(treatment ~ ., scales = "free_y") +
  labs(title = "Total Alkalinity ",
    x = "Time Point",
    y = paste0("Delta TA (µmol/kg)")
  ) +
  scale_color_manual(
    breaks = c("Control", "Coral", "CCA", "Algae"),
    values = c("grey", "#03A9F4", "#EF5350", "#4CAF50")
  ) +
  nunu_theme +
  theme(
    axis.text       = element_text(size = 18),
    legend.position = "none"
  )

ph <- mobeast_carb %>% 
  filter(variable == "ph") %>% 
  group_by(variable, time_fac, treatment) %>% 
  reframe(mean = mean(conc_norm, na.rm = T),
          sd   = sd(conc_norm, na.rm = T)) %>% 
  ggplot(aes(x = time_fac, y = mean, color = treatment, group = treatment)) +
  # Day/night shading — added before geom_point and geom_line
  geom_tile(
    data = shading,
    aes(x = time_fac, y = 0, fill = period),
    height = Inf, alpha = 0.3,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(
    values = c("day" = "lightyellow", "night" = "lightblue"),
    name   = "Period"
  ) +
  geom_point() +
  geom_line(aes(color = treatment)) +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  facet_grid(treatment ~ ., scales = "free_y") +
  labs(title = "pH",
    x = "Time Point",
    y = paste0("Delta pH")
  ) +
  scale_color_manual(
    breaks = c("Control", "Coral", "CCA", "Algae"),
    values = c("grey", "#03A9F4", "#EF5350", "#4CAF50")
  ) +
  nunu_theme +
  theme(
    axis.text       = element_text(size = 18),
    legend.position = "none"
  )

ta + ph
```


```{r}

# Define shading dataframe matching your exact factor levels
# Adjust period assignments to match your actual day/night schedule
shading <- data.frame(
  time_fac = levels(mobeast_carb$time_fac),
  period   = c("day", "night", "night", "day")  # adjust to your time points
)

ta <- mobeast_carb %>% 
  filter(variable == "ta") %>% 
  ggplot(aes(x = time_fac, y = conc_norm, fill = treatment)) +
  # Day/night shading - added before geom_point and geom_line
  #geom_point() +
  geom_tile(
    data = shading,
    aes(x = time_fac, y = 0, fill = period),
    height = Inf, alpha = 0.3,
    inherit.aes = FALSE
  ) +
  geom_violin() +
  stat_summary(color = "yellow", alpha =0.5)+
  geom_hline(yintercept = 0, linetype = 'dashed') +
  facet_grid(treatment ~ ., scales = "free_y") +
  labs(title = "Total Alkalinity ",
    x = "Time Point",
    y = paste0("Delta TA (µmol/kg)")
  ) +
  scale_fill_manual(
    breaks = c("Control", "Coral", "CCA", "Algae", "day", "night"),
    values = c("grey", "#03A9F4", "#EF5350", "#4CAF50", "lightyellow", "lightblue")
  ) +
  nunu_theme +
  theme(
    axis.text       = element_text(size = 18),
    legend.position = "none"
  )

ph <- mobeast_carb %>% 
  filter(variable == "ph") %>% 
  ggplot(aes(x = time_fac, y = conc_norm, fill = treatment)) +
  # Day/night shading - added before geom_point and geom_line
  #geom_point() +
  geom_tile(
    data = shading,
    aes(x = time_fac, y = 0, fill = period),
    height = Inf, alpha = 0.3,
    inherit.aes = FALSE
  ) +
  geom_violin() +
  stat_summary(color = "yellow", alpha =0.5)+
  geom_hline(yintercept = 0, linetype = 'dashed') +
  facet_grid(treatment ~ ., scales = "free_y") +
  labs(title = "pH",
    x = "Time Point",
    y = paste0("Delta pH")
  ) +
  scale_fill_manual(
    breaks = c("Control", "Coral", "CCA", "Algae", "day", "night"),
    values = c("grey", "#03A9F4", "#EF5350", "#4CAF50", "lightyellow", "lightblue")
  ) +
  nunu_theme +
  theme(
    axis.text       = element_text(size = 18),
    legend.position = "none"
  )

ta + ph
```





## DNA 

### DNA data clean up 
```{r}
# prepaing the data for later
mobeast_DNA  <- mobeast_raw %>%
  mutate(time_string = as.character(time_string),
              date_string = as.character(date_string),
              date_time_string = paste(date_string,time_string))%>% 
  filter(variable_cat == "DNA" ) %>% 
  filter(treatment != "Rain") %>% 
  filter(treatment != "Endmember") %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
  
  

```

```{r}
mobeast_DNA_24h <- mobeast_DNA %>% filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  filter(treatment != "Inflow") %>% 
  filter(!is.na(id_number)) %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00"))) %>% mutate(log_flow = log10(flowrate))

exclude_family <- mobeast_DNA_24h %>% 
  group_by(time, variable) %>% 
  reframe(avg = mean(value)) %>% 
  filter(avg == 0) 
```

### LMER Models
```{r}
# Loop Function:
## Take the 24h samples and isolate the types of nutrient  present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

mobeast_loop <- mobeast_DNA_24h %>% filter(!variable %in% exclude_family$variable & !variable %in% exclude_family$time) %>%
  select(variable) %>% distinct() %>% pull()

models_DNA_24h <- vector(mode = "list", length = length(mobeast_loop))
names(models_DNA_24h) <- mobeast_loop


for (var in mobeast_loop) {
  df <- mobeast_DNA_24h %>% filter(variable == var)
  models_DNA_24h[[var]] <- lmer(value ~ treatment*time_fac + 
                                  log_flow + (1|tank_number), data = df)

}

```


```{r}
results_DNA <- imap_dfr(models_DNA_24h, ~ tidy(.x) %>%
                     dplyr::mutate(variable = .y)) %>% 
  relocate(variable, .before = "effect")

results_DNA %>%  filter(p.value <0.05)
```

### DNA ANOVA

```{r, warning= FALSE}
# So many Warnings! 
## Warning: The column names NumDF and DenDF in ANOVA output were not recognized or transformed.

anova_results <- imap_dfr(models_DNA_24h, ~ {
  stats::anova(.x, type = 2) %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>%
  relocate(variable) 

DNA_anova_sig <- anova_results %>% filter(p.value <0.05) %>% select(variable, anova_term = term, anova_p_val = p.value)

sig_treat <- anova_results %>% filter(p.value <0.05) %>% 
  filter(term == "treatment") %>%
  select(variable) %>% 
  pull()

fam_time <- anova_results %>% filter(p.value <0.05) %>% 
  filter(term == "time_fac") %>% select(variable) %>% distinct() 

fam_treatment <- anova_results %>% filter(p.value <0.05) %>% 
  filter(term == "treatment") %>% select(variable) %>% distinct() %>% pull()

fam_time %>%  filter(variable %in% fam_treatment) %>%  pull()

fam_interaction <- anova_results %>% filter(p.value <0.05) %>% 
  filter(term == "treatment:time_fac")

fam_interaction %>% filter(variable %in% fam_treatment)
interaction_family <- fam_interaction %>% filter(variable %in% fam_treatment) %>% 
  select(variable) %>%  pull()
```



#### Partial Eta squared 

Partial η²

Answers "so what?" after a significant ANOVA
Tells you whether the significant effect is actually large enough to care about biologically
A p < 0.05 with η² = 0.01 is statistically real but possibly biologically trivial

```{r}
eta_results <- tibble()

for (var in sig_treat) {
  df <- tibble(eta_squared(anova(models_DNA_24h[[var]]), partial = TRUE))
  df <- df %>% mutate(variable = var ) %>% relocate(variable)
  
  eta_results <- rbind(eta_results, df)
  
  eta_results <- eta_results 
  
}



sig_eta <- eta_results %>% filter(Eta2_partial > 0.14) %>% select(variable) %>% distinct() %>% pull()

sig_eta 
```


### DNA Emmeans 

```{r, message= FALSE}
# "estimate" are negative somehow. Why? 
## in manual models for DOC, estimate >0, here it's <0
## t.ratio in "statistics", and negative instead of positive
## p.values not shown, only p.adjusted. 
emmeans_results_dna <- imap_dfr(models_DNA_24h, ~ {
  emmeans(.x, pairwise ~ treatment)$contrasts %>%
    broom::tidy() %>%
    mutate(variable = .y)}) %>% 
  relocate(variable) %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") 


```

```{r}
emmeans_sig_dna <- emmeans_results_dna %>% filter(adj.p.value <0.05) %>% 
  select(variable, emmean_term = term, contrast, emmeans_p_val = adj.p.value) 

emmeans_sig_dna %>% left_join(DNA_anova_sig) %>% filter(!variable %in% interaction_family) %>% select(variable) %>%  distinct() %>% pull()

```



```{r, warning=FALSE}
DNA_sig_var <- emmeans_results_dna %>% 
  filter(adj.p.value <0.05) %>% 
  select(variable) %>% distinct() %>% pull()

graphs_DNA_24h <- DNA_sig_var %>%
  set_names() %>%
  map(~ {
    df <- mobeast_DNA_24h %>% filter(variable == .x)
    
    ggplot(df, aes(x = treatment, y = value, fill = treatment)) +
      geom_violin() +
      geom_point() +
      geom_hline(yintercept = 0, linetype = "dashed") +
      stat_summary(color = "yellow", alpha = 0.5) +
      labs(title = paste("24h", .x, "Production"),
           x = "Treatment",
          y = "Relative Abundance") +
      scale_fill_manual(
        breaks = c("Control", "Coral", "CCA", "Algae", "Inflow"),
        values = c("grey", "#03A9F4", "#EF5350", "#4CAF50", "orange")
      ) +
      nunu_theme +
      theme(legend.position = "none")
  })


```




```{r}
emmeans_results_dna %>%   
  mutate(sig = ifelse(adj.p.value<0.05 & estimate < 0, "Enrichment", 
                                         ifelse(adj.p.value<0.05 & estimate > 0, "Depletion", "No Difference")))%>% 
  mutate(sig = factor(sig, 
                      levels = c("Enrichment","No Difference", "Depletion"))) %>% 
  filter(variable %in% sig_treat)  %>%
  #filter(variable %in% interaction_family) %>% 
  mutate(contrast = factor(contrast, levels = c("Control - Coral", "Control - CCA", "Control - Algae"))) %>%
  ggplot()+
  geom_tile(aes(x= contrast, y= variable, fill = sig), color = "black")+
  #facet_wrap(.~variable_cat)+
  labs(title = "24h Contrast (Emmeans): Control vs. Treatment",
       x= "Treatment",
       y = " Variables")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw() +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))# + 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")
  
```

### Effect Size 

Adds biological interpretability to each contrast
Tells you not just "these groups differ" but "by how much and in which direction"
Essential for comparing across nutrients since your residuals are on different scales


```{r}
effect_size_results <- tibble()

for (var in sig_eta) {
  emm <- emmeans(models_DNA_24h[[var]], ~ treatment)
  contrasts <- contrast(emm, method = "trt.vs.ctrl")

## Effect size for each contrast
  df <- tidy(eff_size(emm, sigma = sigma(models_DNA_24h[[var]]), 
                edf = df.residual(models_DNA_24h[[var]])))
  
  df <- df %>% mutate(variable = var) %>% rename(effect.size = estimate) %>% 
    relocate(variable)
  
  effect_size_results <- rbind(effect_size_results, df)
  
}

effect_size_results %>% 
  filter(contrast == "Control - Coral"| contrast == "Control - CCA"|contrast == "Control - Algae") %>% filter(abs(effect.size) >=0.8) %>%  filter (variable %in%sig_treat)



```






#### DNA Pairwise

```{r}

## modified function for DNA 
nunu_pairwise <- function(data, type = "coble_a", time_point = 6, value= areal_flux){

  ## Get a list of the treatments for the for loop   
  treatment <- data %>% mutate(treatment = as.character(treatment)) %>%
  select(treatment) %>%  distinct() %>% pull()
  
  ## Create un empty data frame
pairwise_results <- tibble(
  time_point = vector(),
  treatment= vector(),
  mean = vector(),
  p_value = vector())

# this is the pairwise comparison for the 6 pairs (6AM-6PM)
  if(time_point == 6){
    for(var in treatment){
      
      ## only keep the 6AM and 6PM
      df <- data %>% 
        filter(variable == type) %>% 
        filter(treatment == var) %>% 
  filter(time_string == "18:00:00" | time_string == "06:00:00") %>% 
        ## Rename the times so they can become column names
  mutate(time_US = ifelse(time_string == "18:00:00", "T18", "T06")) %>% 
        ## Only keep the column of interest for the stats
  select(tank_number, time_US, value) %>% 
  pivot_wider(names_from = time_US, values_from = value) 
      
      ## Pairwise function 
      paired <- t.test(df$T18, df$T06, paired = TRUE, na.action = na.pass)
      
      ## Save the result in at data frame
      df2 <- tibble(
        time_point = "06AM-06PM",
        variable = type,
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      # Code will loop oveer all treatments and save the results in a data frame
      pairwise_results <-rbind(pairwise_results,df2)
      
      # Save the results in RStudio
      pairwise_results <<- pairwise_results
    }  
  

# This is for the 12 samples (12AM-12PM)
  }else if(time_point == 12){
    for(var in treatment){
      
      ## Same as before but for the 12AM-12PM
      df <- data %>% 
        filter(variable == type) %>% 
        filter(treatment == var) %>% 
  filter(time_string == "12:00:00" | time_string == "00:00:00") %>% 
  mutate(time_US = ifelse(time_string == "12:00:00", "T12", "T00")) %>% 
  select(tank_number, time_US, value) %>% 
  pivot_wider(names_from = time_US, values_from = value) 
      
      paired <- t.test(df$T12, df$T00, paired = TRUE, na.action = na.pass)
      
      df2 <- tibble(
        time_point = "12AM-12PM",
        variable = type,
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      pairwise_results <- rbind(pairwise_results, df2)
      pairwise_results <<- pairwise_results
    }
}

# Display the results of the code 
return(pairwise_results)
}

```


```{r}
# Note: relative aboundance was logged to be able to compute the pairwise comparison 
# Loop Function:

mobeast_loop <- mobeast_DNA_24h %>% select(variable) %>% distinct() %>% pull()

mobeast_DNA_24h <- mobeast_DNA_24h %>% mutate(abun_log = log10(value +1))

pairwise_summary <- map_dfr(mobeast_loop, ~ {
  
  nunu_pairwise(
    data = mobeast_DNA_24h,
    time_point = 12,
    value = abun_log,
    type = .x
  ) %>%
    dplyr::mutate(variable = .x)
  
})

```

```{r}
sig_var <- pairwise_summary %>% filter(p_value <0.05) %>% 
  select(variable) %>% distinct() %>% pull()
```

```{r}
pairwise_summary %>% 
  mutate(sig = ifelse(p_value<0.05 & mean > 0, "Enrichment", 
                                         ifelse(p_value<0.05 & mean < 0, "Depletion", "No Difference")))%>% 
  mutate(sig = factor(sig, 
                      levels = c("Enrichment","No Difference", "Depletion"))) %>% 
  #mutate(variable = factor(variable, levels = unique(variable))) %>%
  filter(variable %in% sig_var) %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  ggplot()+
  geom_tile(aes(x= treatment, y= variable, fill = sig), color = "black")+
  #facet_grid(variable_cat ~ ., scales = "free_y", space = "free_y")+
  labs(title = "Pairwise comparison 12AM-12PM",
       x= "Treatment",
       y = " Variables")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw()#+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")

```

## PCA 

```{r}
## used the areal flux to calculate the PCA here
## normalization with scale(), but don't actually know what it noarmalizes to...
pca_24h_df <- mobeast_24h %>% select(id_number, tank_number, treatment, inflow_table,
                             date_string, time_string, variable, conc_norm) %>%
  filter(variable != "cyano", variable != "tot_bact") %>% 
  pivot_wider(names_from = variable, values_from = conc_norm) %>% 
   na.omit()

pca_matrix_24h_norm <- pca_24h_df %>% select(het_bact:npoc_u_m) %>% scale() 

pca_matrix_24h_norm %>% cor() %>% 
  ggcorrplot(
           hc.order = TRUE,
           #type = "lower",
           lab = TRUE)
```


```{r}

pca_matrix_24h_norm<- pca_matrix_24h_norm %>% as.data.frame() %>% 
  select(het_bact, pico_euk, syn, pro, chla_conc, coble_b,SiO2,NH3,
                                coble_t, bix, fpeak, fi, PO4, NO3, 
                                het_bact, npoc_u_m, coble_a, 
                                coble_c, coble_m, hix,npoc_u_m ) %>% as.matrix()
```


```{r}
data.pca <- princomp(pca_matrix_24h_norm, cor = TRUE)
summary(data.pca)

data.pca$loadings[,1:4]

loading_results <- as.data.frame(data.pca$loadings[,1:4])

fviz_eig(data.pca, addlabels = TRUE)

fviz_pca_var(data.pca, col.var = "black")

fviz_pca_var(data.pca, col.var = "cos2",
            gradient.cols = c("blue", "lightgrey", "red"),
            repel = TRUE)
```

```{r}
pca_scores <- as.data.frame(data.pca$scores)

ggplot(pca_scores, aes(Comp.1,Comp.2))+
  geom_point(aes(color = pca_24h_df$treatment, size= 3))+
  geom_segment(data = loading_results, 
               aes(x=0, y=0, xend = Comp.1, yend= Comp.2),
               arrow = arrow(length = unit(0.5, "cm")),
               alpha = 0.75, color = "black")+
  geom_text(data = loading_results, 
            aes(x = Comp.1, y = Comp.2, 
                label = rownames(loading_results)),
            vjust=1, hjust = 1)+
  theme_bw() +
      scale_color_manual(
        breaks = c("Control", "Coral", "CCA", "Algae"),
        values = c("grey", "#03A9F4", "#EF5350", "#4CAF50")
      )  
```





## PCA 6AM only

```{r}
## used the areal flux to calculate the PCA here
## normalization with scale(), but don't actually know what it noarmalizes to...
pca_6AM_df <- mobeast_24h %>% select(id_number, tank_number, treatment, inflow_table,
                             date_string, time_string, variable, conc_norm) %>%
  filter(variable != "cyano", variable != "tot_bact",
         variable != "tn_u_m", variable != "m_to_c",
         variable != "N_N", variable != "lignin",
         variable != "chla_conc") %>% 
  pivot_wider(names_from = variable, values_from = conc_norm) %>% 
  drop_na()

pca_matrix_6AM_norm <- pca_6AM_df %>% select(het_bact:npoc_u_m) %>% scale() 

pca_matrix_6AM_norm %>% cor() %>% 
  ggcorrplot(
           hc.order = TRUE,
           #type = "lower",
           lab = TRUE)
```


```{r}

pca_matrix_6AM_norm <- pca_matrix_6AM_norm %>% as.data.frame() %>% 
  select(het_bact, pico_euk, syn, coble_a, coble_b,
         coble_c, coble_m, coble_t, hix, bix, fpeak, fi, 
         SiO2, NH3,PO4, NO3, npoc_u_m) %>% as.matrix()
dim(pca_matrix_6AM_norm)
```


```{r}
data.pca <- princomp(pca_matrix_6AM_norm, cor = TRUE)
#data.pca <- prcomp(pca_matrix_6AM_norm, scale. = FALSE)
summary(data.pca)

data.pca$loadings[,1:4]

loading_results <- as.data.frame(data.pca$loadings[,1:4])

fviz_eig(data.pca, addlabels = TRUE)

fviz_pca_var(data.pca, col.var = "black")

fviz_pca_var(data.pca, col.var = "cos2",
            gradient.cols = c("blue", "lightgrey", "red"),
            repel = TRUE)
```

