---
title: "MOBEAST - 24h Heatmap"
format: html
editor: visual
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)

library(ComplexHeatmap)
library(data.table)
library(dendextend)

# Rayna library
library(dendsort)
library(ggsci)
library(randomForest)
library(vegan)
library(ape)
library(devtools)
library(ggpubr)
library(viridis)
library(colorspace)
library(readr)
library(circlize)

```

Heatmap info and tutorial: https://jokergoo.github.io/ComplexHeatmap-reference/book/a-single-heatmap.html

## Functions

```{r}
source(here("scripts", "reference_scripts", "heatmap_zscore_function.R"))
```

## Data visualization theme

```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        panel.background = element_rect(fill = "azure1"))
```

### Color palette

```{r}
treatment_colors = c(
  #Treatments
  "Control" = "grey",
  "Coral" = "#03A9F4",
  "CCA" = "#EF5350",
  "Algae" = "#4CAF50"
)

```

## Load Data

```{r}
#OTU <- read_csv(here("data", "DNA","MOBEAST_OTU_culled.csv"))
meta <- read_csv(here("data","MOBEAST_metadata_heatmaps.csv"))
#tax <- read_csv(here("data","DNA", "MOBEAST_tax_culled.csv"))
OTU_long <- read_csv(here("data","MOBEAST_DNA_long.csv"))

# Note: DO NOT DELETE METADATA_HEATMAP.CSV !
# Somehow if we use a clean-up version of the metadata, 
# heatmap annotations don't work!

```

## Data Clean Up
Reformating the date, time and date-time in the metadata file 
```{r, meta data clean up}
meta<- meta %>% mutate(time = str_pad(time, width = 4, side = "left", pad = "0"),
               time = str_replace(time, "(\\d{2})(\\d{2})", "\\1:\\2")) %>% 
  # "(\\d{2})(\\d{2})" separates the string in 2 groups of 2 characters
  # "\\1:\\2", group1 + ":" + group2
  mutate(time = paste0(time, ":00"),
               date_time = paste(date, time),
               date = mdy(date), 
               date_time = mdy_hms(date_time),
               time = factor(time))
  
  
```
### Preparing the metadata to annotate the heatmap

```{r}
# renaming some of treatment so that environmental data have a label: endmember or rain
# Same for inflow table, environmental samples now have a label
## Convert tibble to df for later
meta <- meta %>% mutate(treatment = as.character(treatment)) %>% 
  mutate(treatment = ifelse(id_number == "MOBEAST115", "Rain", 
                            ifelse(is.na(treatment), "Endmember", treatment))) %>% 
  mutate(inflow_table = ifelse(is.na(inflow_table), "Environmental data", inflow_table)) %>% 
  mutate(date_string = as.character(date),
                time_string = as.character(time),
                date_time_string = paste(date_string, time_string)) %>% 
  as.data.frame()
```

Clean up the long format data to make it easier to merge with the meta data file

```{r}
OTU_long<- OTU_long %>% clean_names() %>% 
  mutate(id_number = sample_id) %>% 
  relocate(id_number, .before= otu_number) %>% 
  select(-sample_id, -date, -treatment, -time)
```

Fusion!

```{r}
OTU_htm <- left_join(OTU_long, meta)
#write_csv(OTU_htm, "OTU_heatmap_ready.csv", path = here("data", "OTU_heatmap_ready.csv"))
```

## 24h Z-Scored Heatmap

### Preparing the OTU matrix for heatmap

```{r}
OTU_htm_24h <- OTU_htm %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) 
OTU_htm_24h <- OTU_htm_24h %>% filter(family != "Bradymonadales_fa")
```

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_htm_24h %>%
  group_by(id_number, family) %>%
  summarize(abund = sum(abund, na.rm = TRUE))

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = family, values_from = abund)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% ungroup() %>% tibble() %>%  select(-id_number)

# Calculate Z-score on df/matrix
z.zscore.data.otu.subset <- as.data.frame(apply(culled.abund.OTU, 2, FUN=zscore.calculation))

# bring back the sample ID number in the df 
z.zscore.data.otu.subset$id_number = family.culled.abund$id_number


# Set the ID numbers as the rownames (to make a matrix)
rownames(z.zscore.data.otu.subset) = z.zscore.data.otu.subset$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
zscore.data.otu.matrix <- z.zscore.data.otu.subset %>% select(-id_number) %>% 
  as.matrix() %>% t()
```

### Generate Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number), 
                 "treatment"]

date_time <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number),
                  "date_time_string"]

```

```{r}
## 12PM Samples
ha = HeatmapAnnotation(
  treatment = treatment,
  date_time = date_time,
  col = list(treatment = c("Inflow"="#00838F","Control"="grey", "Coral"="#03A9F4", 
                           "CCA"="#EF5350", "Algae"="#4CAF50",
                           "Endmember" = "black", "Rain"= "yellow"),
             date_time = c("2024-06-02 12:00:00" = "red",
                           "2024-06-02 18:00:00" = "yellow", 
                           "2024-06-03 00:00:00" = "green",
                           "2024-06-03 06:00:00" = "blue"
                           ))
)


```

#### Printing heatmap

```{r, fig.height = 10, fig.width = 16}
#png(width=1800, height=1000, file= here("outputs", "zscore_24hdata.png"))

ht = Heatmap(zscore.data.otu.matrix, name = "z-scored", 
             #rect_gp = gpar(col = "white", lwd = 0.1),
        
        column_title = "24h Sampling - All samples", 
        column_title_gp = gpar(fontsize = 16, fontface = "bold"),
        #column_title_side = "bottom",
        row_title = "Family",
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10),
        row_km = 5, #number of forced segregation in rows 
        column_km = 4,#number of forced segregation in columns 
        top_annotation = ha
        )
set.seed(1999)
ht_drawn <- draw(ht)

#dev.off()

```

### Cluster membership

```{r}
col_orders<- column_order(ht_drawn)

cluster_df <- data.frame(
  id_number = unlist(col_orders),
  cluster  = rep(seq_along(col_orders), lengths(col_orders))
)

rownames(cluster_df)<- NULL

```

```{r}
cluster_df<- cluster_df %>% 
  mutate(id_number = ifelse(id_number<10,
                            paste0("MOBEAST00",id_number), 
                            paste0("MOBEAST0",id_number)))
#write_csv(cluster_df, here("data", "DNA", "24h_clusters.csv"))
```

```{r}
OTU_htm_24h %>% left_join(cluster_df) %>% 
  select(-kingdom, -class, -order, -genus, -long_name, -date_time, -date, -time) %>% 
  select(id_number, tank_number, treatment, date_time_string,cluster) %>% 
  distinct() %>% 
  arrange(cluster) %>% 
  mutate(cluster = factor(cluster, levels = c("1", "2", "3", "4"))) %>% 
  group_by(cluster, treatment) %>% 
  gt() %>% 
  cols_label(id_number = "ID Number",
             tank_number = "Tank Number", 
             date_time_string = " Date-Time") #%>% 
  #gtsave("24H_cluster_table.png")
```



### Cluster membership - code by Sean

Not useful for cluster membership but to see the dendrograms, it's useful!
```{r}
# choose a distance metric
# note: you could use vegdist to do something like Bray curtis distances instead
# by default Heatmap uses 'euclidean' distances
our_dist_method = "euclidean"

# choose a method for clustering
# note: by default Heatmap uses 'complete' clustering
#our_clust_method = "ward.D"
our_clust_method = "complete"

# cluster rows and columns
# note: dist generates a distance matrix and hclust does the clustering
row_dend = as.dendrogram(hclust(dist(zscore.data.otu.matrix, 
                                     method = our_dist_method),
                                method = our_clust_method))

col_dend = as.dendrogram(hclust(dist(t(zscore.data.otu.matrix), 
                                     method = our_dist_method),
                                method = our_clust_method))

# pick an appropriate number of groups for splitting rows and columns
row_k = 8
col_k = 4

# split into clusters
#note: sample clusts and analyte clusts will tell you which cluster each sample/analyte is in!
sample_clusts = cutree(row_dend,
                       row_k,
                       order_clusters_as_data = F)

analyte_clusts = cutree(col_dend,
                        col_k,
                        order_clusters_as_data = Fplot(analyte_clusts))

plot(col_dend)

# store sample clusters as a dataframe which can be merged with your other sample metadata
#analyte_clust_df = data.frame(id_number = names(analyte_clusts), cluster = analyte_clusts)

```



## 24h separated maps

### Preparing the OTU matrix for heatmap

```{r}
OTU_htm_24_12 <- OTU_htm %>% 
  filter(date == ymd("2024-06-02") & time_string == "12:00:00") %>% 
  filter(family != "Bradymonadales_fa")

OTU_htm_24_18 <- OTU_htm %>% 
  filter(time_string == "18:00:00") %>% 
  filter(family != "Bradymonadales_fa")

OTU_htm_24_00 <- OTU_htm %>% 
  filter(time_string == "00:00:00") %>% 
  filter(family != "Bradymonadales_fa")

OTU_htm_24_06 <- OTU_htm %>% 
  filter(time_string == "06:00:00") %>% 
  filter(family != "Bradymonadales_fa")

```

### 24h - 12:00 map

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_htm_24_12 %>%
  group_by(id_number, family) %>%
  summarize(abund = sum(abund, na.rm = TRUE))

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = family, values_from = abund)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% ungroup() %>% tibble() %>%  select(-id_number)

# Calculate Z-score on df/matrix
z.zscore.data.otu.subset <- as.data.frame(apply(culled.abund.OTU, 2, FUN=zscore.calculation))

# bring back the sample ID number in the df 
z.zscore.data.otu.subset$id_number = family.culled.abund$id_number


# Set the ID numbers as the rownames (to make a matrix)
rownames(z.zscore.data.otu.subset) = z.zscore.data.otu.subset$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
zscore.data.otu.matrix <- z.zscore.data.otu.subset %>% select(-id_number) %>% 
  as.matrix() %>% t()
```

### Generate Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number), 
                 "treatment"]

#date_time <- meta[match(colnames(zscore.data.otu.matrix),
                  #         meta$id_number),
                 # "date_time_string"]
```

```{r}
## 12PM Samples
ha = HeatmapAnnotation(
  treatment = treatment,
  #date_time = date_time,
  col = list(treatment = c("Inflow"="#00838F","Control"="grey", "Coral"="#03A9F4", 
                           "CCA"="#EF5350", "Algae"="#4CAF50","Endmember" = "black", "Rain"= "yellow"),
             date_time = c("2024-06-02 12:00:00" = "red"#,
                           #"2024-06-02 18:00:00" = "yellow", 
                           #"2024-06-03 00:00:00" = "green",
                           #"2024-06-03 06:00:00" = "blue"
                           ))
)


```

#### Printing heatmap

```{r, fig.height = 10, fig.width = 16}
#png(width=1800, height=1000, file= here("outputs", "12PM_24hdata.png"))

ht = Heatmap(zscore.data.otu.matrix, name = "z-scored", rect_gp = gpar(col = "white", lwd = 0.1),
        column_title = "id_number", column_title_side = "bottom",
        row_title = "family",
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10),
        row_km = 4, #number of forced segregation in rows 
        column_km = 4,#number of forced segregation in columns 
        top_annotation = ha
        )

draw(ht)

#dev.off()

```

### 24h - 06:00 map

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_htm_24_06 %>%
  group_by(id_number, family) %>%
  summarize(abund = sum(abund, na.rm = TRUE))

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = family, values_from = abund)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% ungroup() %>% tibble() %>%  select(-id_number)

# Calculate Z-score on df/matrix
z.zscore.data.otu.subset <- as.data.frame(apply(culled.abund.OTU, 2, FUN=zscore.calculation))

# bring back the sample ID number in the df 
z.zscore.data.otu.subset$id_number = family.culled.abund$id_number


# Set the ID numbers as the rownames (to make a matrix)
rownames(z.zscore.data.otu.subset) = z.zscore.data.otu.subset$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
zscore.data.otu.matrix <- z.zscore.data.otu.subset %>% 
  select(-id_number, - Moraxellaceae, -Idiomarinaceae ) %>% # removing NaN columns
  as.matrix() %>% t()
```

### Generate Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number), 
                 "treatment"]

#date_time <- meta[match(colnames(zscore.data.otu.matrix),
                  #         meta$id_number),
                  #"date_time_string"]
```

```{r}
## 12PM Samples
ha = HeatmapAnnotation(
  treatment = treatment,
  #date_time = date_time,
  col = list(treatment = c("Inflow"="#00838F","Control"="grey", "Coral"="#03A9F4", 
                           "CCA"="#EF5350", "Algae"="#4CAF50","Endmember" = "black", "Rain"= "yellow"),
             date_time = c(#"2024-06-02 12:00:00" = "red",
                           #"2024-06-02 18:00:00" = "yellow", 
                           #"2024-06-03 00:00:00" = "green",
                           "2024-06-03 06:00:00" = "blue"
                           ))
)


```

#### Printing heatmap

```{r, fig.height = 10, fig.width = 16}
#png(width=1800, height=1000, file= here("outputs", "6AM_24hdata.png"))

ht = Heatmap(zscore.data.otu.matrix, name = "z-scored", rect_gp = gpar(col = "white", lwd = 0.1),
        column_title = "id_number", column_title_side = "bottom",
        row_title = "family",
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10),
        row_km = 4, #number of forced segregation in rows 
        column_km = 2,#number of forced segregation in columns 
        top_annotation = ha
        )

draw(ht)

#dev.off()

```
