---
author: 'Keanu Rochette '
created: '18 April 2025'
updated: '`r format(Sys.Date())`'
title: "MOBEAST - Nutrients"
format: html
execute:
  echo: false
  message: false
fig-width: 8
fig-height: 4
fig-align: center
knitr:
  opts_chunk:
    out.width: "20%"
editor: visual
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE)
```

## To Do : Craig Comments 
- Try getting the two random samples that are clustering as outliers out of the analysis and see how it looks.  
- Look at how those samples show up on the ordination. Prep some slides to explore those samples, particularly the relative abundance (not z-scores) of the module of taxa enriched in those two samples, to see if they are transient rare taxa or abundant in those samples.
- Make a version of each violin plot where the points are colored by "Water Table" to see if there are any parameters or treatments or times where one table had some issues. Use "jitter" to spread the points out so you can see them.
- Consider doing outlier removals (should never be more than a few points, maybe 10% max for each parameter) before subtracting inlet mean values. This way if you get some random high value in an inlet sample it doesn't have an outsized impact on your analyses.
- If you are going to work with nitrate we need to look at distributions of the raw values of each of the nutrient parameters. You can make those graphs and/or send over your dataframe.



## Load Libraries {.hidden}

```{r}
#| output: false

library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
#library(GGally)
library(emmeans)
library(performance)
library(multcompView)

```

## Load Data 

```{r}
#| output: false
nutrients <- read_delim(here("data", "Nutrients", "MOBEAST_nutrients.csv"))
meta <- read_csv(here("data", "MOBEAST_metadata_V2.csv"))
```

## Data visualization theme {.hidden}

```{r}
#| output: false
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=32, face = "bold"), 
        plot.subtitle = element_text(size=30),
        axis.title = element_text(size = 28, face = "bold"),
        axis.text = element_text(size = 28),
        strip.text.x = element_text(size = 28, face = "bold"),
        legend.title=element_text(size=28, face = "bold"),
        legend.text=element_text(size=28))
```

```{r}
#| output: false
treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```

## Data Clean Up

```{r}
#| output: false

# nutrients are in umol/L 
nutrients<- nutrients %>% clean_names() %>% 
  rename(N_N = no3_no2_umol_l,
         NO2 = no2_umol_l,
         PO4 = po4_umol_l,
         SiO2 = si_o2_umol_l,
         NH3 = nh3_umol_l) %>% 
  mutate(across(N_N:NH3, ~ str_remove_all(.x, "[<>]"))) %>% 
  mutate(across(N_N:NH3, ~ as.numeric(.x))) %>% 
  mutate(NO3 = N_N - NO2)
```

```{r}
#| output: false

nutrients <- meta %>% left_join(nutrients) %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow", "Endmember")))

#write_csv(nutrients, here("data", "Nutrients", "nutrient_raw.csv"))
```

## Data exploration

```{r}
#| output: false
nutrients %>% 
  filter(treatment != "Endmember") %>%
  mutate(time_string = as.character(time_string),
         date_string = as.character(date_string),
         date_time_string = as.character(date_time_string)) %>% 
  select(-run_number) %>% 
  pivot_longer(cols = N_N:NO3,
               names_to = "nutrient",
               values_to = "conc") %>% 
  #mutate(inflow_color = ifelse(treatment == "Inflow", "Inflow", "Treatment")) %>% 
  #filter(conc <5) %>% 
  #filter(time_string == "06:00:00")%>% 
  ggplot(aes(x= treatment, y = conc, fill =treatment))+
  geom_violin()+
  geom_jitter()+
  stat_summary(color = "yellow", alpha =0.5)+
  labs(title = "Nutrient Concentration by Treatment",
       subtitle = "Trying to identify values that look off",
       x = "Treatment", 
       y = "Concentration (umol/L)")+
  #geom_text_repel(aes(label = tank_number), alpha = 0.4)+
  facet_wrap(~nutrient)+
  nunu_theme+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA","Algae", "Inflow"), values=c("grey", "#03A9F4", "#EF5350", "#4CAF50","#00838F"))+
  theme(legend.position = "none") 


```

```{r}

## Table of high concentration nutrient, maybe should be remeasured
nutrients %>% 
  filter(treatment != "Endmember") %>%
  mutate(time_string = as.character(time_string),
         date_string = as.character(date_string),
         date_time_string = as.character(date_time_string)) %>% 
  select(-run_number) %>% 
  pivot_longer(cols = N_N:NO3,
               names_to = "nutrient",
               values_to = "conc") %>% 
  filter(conc > 5) %>% 
  select(-time_string,-date_string, -date_time_string, -date, -time, - inflow_table,
         -tank_number) %>% 
  gt() %>% 
  cols_label(id_number = "ID Number",
             long_name = "Long Name", 
             treatment = "Treatment",
             date_time = "Date Time",
             nutrient = "Nutrient",
             conc = "Concentration umol/L") #%>% 
  #gtsave("nutrient_outliers.png")
         
```

```{r}
## removing outliers from the data set
nutrients%>%
  select(-run_number) %>% 
  pivot_longer(cols = N_N:NO3,
               names_to = "nutrient",
               values_to = "conc") %>% 
  filter(conc>0) %>% # one NO3 value that is negative due to subtraction
  filter(treatment != "Endmember") %>% 
  # removing outliers
  ## grouping by date, time and treatment
  group_by(nutrient, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(log10(conc), probs = 0.25, na.rm = T),
         quant3 = quantile(log10(conc), probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( log10(conc) > quant3 + (IQR*1.5) | log10(conc) < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% drop_na() %>% 
  
  ggplot(aes(x= treatment, y = conc))+
  geom_violin(aes(fill = treatment))+
  geom_jitter(aes(color = outlier), alpha = 0.5) + 
  facet_wrap(~nutrient)+
  scale_color_manual(values = c("red", "blue"))
  
  ## remove the outliers 
  #filter(outlier == "outlier") 


```


```{r}
## removing outliers from the data set
raw_flagged_outliers <- nutrients%>%
  select(-run_number) %>% 
  pivot_longer(cols = N_N:NO3,
               names_to = "nutrient",
               values_to = "conc") %>% 
  #filter(conc>0) %>% # one NO3 value that is negative due to subtraction
  filter(treatment != "Endmember") %>% 
  # removing outliers
  ## grouping by date, time and treatment
  group_by(nutrient, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(log10(conc), probs = 0.25, na.rm = T),
         quant3 = quantile(log10(conc), probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( log10(conc) > quant3 + (IQR*1.5) | log10(conc) < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() #%>% 
  
  ## remove the outliers 
  #filter(outlier != "outlier") 


```


```{}
# Normalization of data with the outliers 

## Normalization of the data to the inflow 
nutrients_norm <- nutrients %>% 
  filter(treatment != "Endmember") %>%
  mutate(time_string = as.character(time_string),
         date_string = as.character(date_string),
         date_time_string = as.character(date_time_string)) %>% 
  select(-run_number) %>% 
  pivot_longer(cols = N_N:NO3,
               names_to = "nutrient",
               values_to = "conc") %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table, nutrient) %>%
  mutate(inflow= conc[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # get the other table's value for that date
  group_by(date_time) %>% 
  mutate(inflow = ifelse(is.na(inflow),
      inflow[!is.na(inflow)][1],  
      inflow)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(conc_norm = conc - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))

#write_csv(nutrients_norm, here("data", "Nutrients", "nutrient_norm.csv"))
```


```{r}
# Normalization version 2
## averaging the inflows for normalization 
## remove the outliers a right before the normalization 

## Normalization of the data to the inflow 
#nutrients_norm <- 
nutrients_norm <- raw_flagged_outliers %>% 
  mutate(time_string = as.character(time_string),
         date_string = as.character(date_string),
         date_time_string = as.character(date_time_string)) %>% 
  select(-IQR, -quant1, -quant3) %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table, nutrient) %>%
  mutate(inflow = conc[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time, nutrient) %>% 
  mutate(inflow = ifelse(outlier == "outlier", inflow == NA, inflow)) %>% 
  mutate(inflow = mean(inflow, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(conc_norm = conc - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```


```{}
## Data frame with outliers for Craig
## removing outliers from the data set
nutrient_Craig <- nutrients_norm %>%
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment, nutrient) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(conc_norm, probs = 0.25, na.rm = T),
         quant3 = quantile(conc_norm, probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( conc_norm > quant3 + (IQR*1.5) | conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup()  

#write_csv(nutrient_Craig, here("data", "Nutrients", "nutrient_norm_w_outliers.csv"))
```

## 12PM Nutrient analysis

```{}

## removing outliers from the data set
nutrient_12PM <- nutrients_norm %>%
  filter(time_string == "12:00:00") %>% 
  filter(treatment != "Inflow") %>% 
  
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment, nutrient) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(conc_norm, probs = 0.25, na.rm = T),
         quant3 = quantile(conc_norm, probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( conc_norm > quant3 + (IQR*1.5) | conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") 

```

```{r}
# Version 2
## removing outliers from the data set
nutrient_12PM <- nutrients_norm %>%
  filter(time_string == "12:00:00") %>% 
  filter(treatment != "Inflow") %>% 
  
  ## remove the outliers 
  filter(outlier == "not outlier") 

```

```{r, fig.width=16, fig.height = 8}
#| output: false 

## Initial graph of the data distribution
nutrients_norm %>% 
  filter(time_string == "12:00:00") %>% 
  filter(treatment != "Inflow") %>% 
  filter(nutrient == "SiO2") %>% 
  ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
  geom_violin()+
  geom_point()+
  geom_hline(yintercept=0, linetype='dashed')+
  stat_summary(color = "yellow", alpha =0.5)+
  #facet_wrap(~nutrient, scales = "free")+
  labs(title = "12PM Silicate concentrations",
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Concentration  (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  #ylim(-20,120)+
  nunu_theme +
  theme(legend.position = "none") 
```

### Iterative statistical modeling {.hidden}

```{r}
# Suppose your dataframe is df with columns: date, time, resp, var1, var2, var3
# Independent variables:
nutrient_loop <- nutrient_12PM %>% select(nutrient) %>% distinct() %>% pull()

models_12PM <- vector(mode = "list", length = length(nutrient_loop))
names(models_12PM) <- nutrient_loop

for (var in nutrient_loop) {
  df <- nutrient_12PM %>% filter(nutrient == var)
  models_12PM[[var]] <- lmer(conc_norm ~ treatment + (1|tank_number), data = df)
}

```

### Iterative graphing {.hidden}

```{r}

nutrient_loop <- nutrient_12PM %>% select(nutrient) %>% distinct() %>% pull()

graphs_12PM <- vector(mode = "list", length = length(nutrient_loop))
names(graphs_12PM) <- nutrient_loop

for (var in nutrient_loop) {
  
  df <- nutrient_12PM %>% filter(nutrient == var)
  
  graphs_12PM[[var]] <- df %>%  
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("12PM", var, "concentrations"),
       #subtitle = "ANOVA: F = 0.01, Dunnett: p<0.05 (Algae)",
       x= "Treatment",
       y = "Concentration (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none")
}

```

#### Nitrate-Nitrite

```{r}
summary(models_12PM[[1]])
anova(models_12PM[[1]], type = 2)
emmeans(models_12PM[[1]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[[1]], ~ treatment)
```

```{r}
resid_panel(models_12PM[[1]])

```

```{r}
graphs_12PM[[1]]+ 
  geom_jitter(aes(color = inflow_table), size = 2)+
  scale_color_manual(values = c("red", "blue"))+
  theme(legend.position = "right")
```

```{r}
#emm <- emmeans(models_12PM[[1]], ~ treatment)
#cld_letters <- cld(models_12PM[[1]], Letters = letters)
```

#### Nitrite - Significant !

```{r}
summary(models_12PM[[2]])
anova(models_12PM[[2]], type = 2)
emmeans(models_12PM[[2]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[[2]], ~ treatment)
```

```{r}
resid_panel(models_12PM[[2]])
graphs_12PM[[2]] + 
  geom_jitter(aes(color = inflow_table), size = 2)+
  scale_color_manual(values = c("red", "blue"))+
  theme(legend.position = "right")
```

#### Nitrate - need to log the data

```{r}
summary(models_12PM[["NO3"]])
anova(models_12PM[["NO3"]], type = 2)
emmeans(models_12PM[["NO3"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[["NO3"]], ~ treatment)
```

```{r}
resid_panel(models_12PM[["NO3"]])
graphs_12PM[["NO3"]]+ 
  geom_jitter(aes(color = inflow_table), size = 2)+
  scale_color_manual(values = c("red", "blue"))+
  theme(legend.position = "right")
```

#### Ammonium

```{r}
summary(models_12PM[["NH3"]])
anova(models_12PM[["NH3"]], type = 2)
emmeans(models_12PM[["NH3"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[["NH3]], ~ treatment)
```

```{r}
resid_panel(models_12PM[["NH3"]])
graphs_12PM[["NH3"]]+ 
  geom_jitter(aes(color = inflow_table), size = 2)+
  scale_color_manual(values = c("red", "blue"))+
  theme(legend.position = "right")
```

#### Phosphate

```{r}
summary(models_12PM[["PO4"]])
anova(models_12PM[["PO4"]], type = 2)
emmeans(models_12PM[["PO4"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[["PO4]], ~ treatment)
```

```{r}
resid_panel(models_12PM[["PO4"]])
graphs_12PM[["PO4"]]+ 
  geom_jitter(aes(color = inflow_table), size = 2)+
  scale_color_manual(values = c("red", "blue"))+
  theme(legend.position = "right")
```

#### Silicate

```{r}
summary(models_12PM[["SiO2"]])
anova(models_12PM[["SiO2"]], type = 2)
emmeans(models_12PM[["SiO2"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[["SiO2]], ~ treatment)
```

```{r}
resid_panel(models_12PM[["SiO2"]])
graphs_12PM[["SiO2"]]+ 
  geom_jitter(aes(color = inflow_table), size = 2)+
  scale_color_manual(values = c("red", "blue"))+
  theme(legend.position = "right")
```

## 24h Nutrient Analysis

```{r}
## removing outliers from the data set
nutrient_24h <- nutrients_norm %>%
  filter(date_string == "2024-06-02"|date_string == "2024-06-03") %>% 
  filter(treatment != "Inflow") %>% 
  # removing outliers
  ## grouping by date, time and treatment
  group_by(date, time, treatment, nutrient) %>% 
  ## caluculating the Interquantile range and of the raw data
  mutate(quant1 = quantile(conc_norm, probs = 0.25, na.rm = T),
         quant3 = quantile(conc_norm, probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( conc_norm > quant3 + (IQR*1.5) | conc_norm < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup() %>% 
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00")))



```

```{r, fig.width=16, fig.height = 8}
## Initial graph of the data distribution
nutrient_24h %>% 
  filter(nutrient == "SiO2") %>% 
  ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
  geom_violin()+
  geom_point()+
  geom_hline(yintercept=0, linetype='dashed')+
  stat_summary(color = "yellow", alpha =0.5)+
  facet_wrap(~time_fac, ncol = 4 )+
  labs(title = "24h Silicate concentrations",
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Concentration  (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  #ylim(-20,120)+
  nunu_theme +
  theme(axis.text = element_text(size = 18),
        legend.position = "none") 
```

##### Data viz

```{r}
## Initial graph of the data distribution
nutrient_24h%>% 
  filter(nutrient == "NO3") %>% 
  ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
  geom_violin()+
  geom_point()+
  geom_hline(yintercept=0, linetype='dashed')+
  stat_summary(color = "yellow", alpha =0.5)+
  facet_wrap(~nutrient+time_fac, scales = "free", ncol = 4)+
  labs(#title = "12PM NO2 concentrations",
       #subtitle = "ANOVA: F <0.05 (treatment & time), Dunnett: p<0.05 (Algae) ",
       x= "Treatment",
       y = "Concentration (µmol/L)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))#+
  #ylim(-20,120)+
  #nunu_theme +
  #theme(axis.text = element_text(size = 18),
        #legend.position = "none") 

#ggsave(here("outputs","24H_NO2.png"), width = 16, height = 8)
```

### Iterative stats

```{r}
# Suppose your dataframe is df with columns: date, time, resp, var1, var2, var3
# Independent variables:
nutrient_loop <- nutrient_24h %>% select(nutrient) %>% distinct() %>% pull()

models_24h <- vector(mode = "list", length = length(nutrient_loop))
names(models_24h) <- nutrient_loop


for (var in nutrient_loop) {
  df <- nutrient_24h %>% filter(nutrient == var)
  models_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + (1|tank_number), data = df)
}

```

### Iterative graphing

```{r}

nutrient_loop <- nutrient_24h %>% select(nutrient) %>% distinct() %>% pull()

graphs_24h <- vector(mode = "list", length = length(nutrient_loop))
names(graphs_24h) <- nutrient_loop


for (var in nutrient_loop) {
  
  graphs_24h[[var]] <- nutrient_24h %>% 
    filter(nutrient == var) %>% 
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    facet_wrap(~time_fac, ncol = 4 )+
    labs(title = paste("24h", var, "concentrations"),
       #subtitle = "Norm DOC = Tank DOC - Inflow DOC",
       x= "Treatment",
       y = "Concentration  (µmol/L)")+
    scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
    nunu_theme +
    theme(axis.text = element_text(size = 18),
        legend.position = "none")
}

```

#### Nitrate-Nitrite

```{r}
summary(models_24h[["N_N"]])
anova(models_24h[["N_N"]], type = 2)
emmeans(models_24h[["N_N"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_24h[["N _N]], ~ treatment)
```

```{r}
resid_panel(models_24h[["N_N"]])
graphs_24h[["N_N"]]
```

#### Nitrite - Significant !

```{r}
summary(models_24h[["NO2"]])
anova(models_24h[["NO2"]], type = 2)
emmeans(models_24h[["NO2"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_24h[["NO2]], ~ treatment)
```

```{r}
resid_panel(models_24h[["NO2"]])
graphs_24h[["NO2"]]
```

#### Nitrate - Significant different in the CCA

```{r}
summary(models_24h[["NO3"]])
anova(models_24h[["NO3"]], type = 2)
emmeans(models_24h[["NO3"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_24h[["NO3]], ~ treatment)
```

```{r}
resid_panel(models_24h[["NO3"]])
graphs_24h[["NO3"]]
```

#### Ammonium

```{r}
summary(models_24h[["NH3"]])
anova(models_24h[["NH3"]], type = 2)
emmeans(models_24h[["NH3"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_24h[["NH3]], ~ treatment)
```

```{r}
resid_panel(models_24h[["NH3"]])
graphs_24h[["NH3"]]
```

::: ::::
#### Phosphate

```{r}
summary(models_24h[["PO4"]])
anova(models_24h[["PO4"]], type = 2)
emmeans(models_24h[["PO4"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_24h[["PO4]], ~ treatment)
```

```{r}
resid_panel(models_24h[["PO4"]])
graphs_24h[["PO4"]]
```

#### Silicate - May need to log transform

```{r}
summary(models_24h[["SiO2"]])
anova(models_24h[["SiO2"]], type = 2)
emmeans(models_24h[["SiO2"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_24h[["SiO2]], ~ treatment)
```

```{r}
resid_panel(models_24h[["SiO2"]])
graphs_24h[["SiO2"]]
```
:::