---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: ROTUI DOC verification
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
```


## Load data
```{r}
meta <- read_delim(here("data", "ROTUI_metadata.csv"))
DOC <- read_delim(here("data", "raw_data", "DOC", "ROTUI_NPOC_TN_241108.csv"))
```


```{r}
meta <- meta %>% clean_names() %>% filter(site != "River 1") %>% 
  mutate(date = ymd(date))  
```

```{r}
DOC <- DOC %>% clean_names() 
```


```{r}
DOC_join <- left_join(meta, DOC, by = "sample_id")
```


## Check data for NPOC (mg/L)
The values appear mostly consistent.  


However, **site B2** (2024-08-09) has higher value compared to other B2 measurements. B2 is located in the middle of the Tautira Bassin, close to the village. It is arguably the closest site to the **Water Treatment Plant (WTP)** outfall. Maybe we just caught a plume of outfall water, _OR_ it's an outlier point.  
```{r, fig.height= 6, fig.width=10}
DOC_join %>% drop_na(npoc_mg_l) %>% 
  ggplot(aes(x= site, y = npoc_mg_l)) +
  geom_point() + 
  geom_text_repel(aes(label = site)) +
  geom_point(data=DOC_join %>% filter(npoc_mg_l > 1.3), 
           pch=21, 
           size=4, 
           colour="red") +
  facet_wrap(.~date)+
  theme_bw()+
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle = 60, vjust = 0.5),
        panel.background = element_rect(fill = "azure1")) 
```

## Check data for TN (mg/L)
All the data point seem to be consistent with one another.  
The Water Treatment Plant (WTP) data point is the highest but I think that's normal.

```{r, fig.height= 6, fig.width=10}
DOC_join %>% drop_na(tn_mg_l) %>% 
  ggplot(aes(x= site, y = tn_mg_l)) +
  geom_point() + 
  geom_text_repel(aes(label = site)) +
  geom_point(data=DOC_join %>% filter(tn_mg_l > 2), 
           pch=21, 
           size=4, 
           colour="red") +
  facet_wrap(.~date)+
  theme_bw()+
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle = 60, vjust = 0.5),
        panel.background = element_rect(fill = "azure1")) 
```


## Some Stats
The NPOC standard deviations remain fairly small.  
The largest std dev occur for B2 (0.33), R8 (0.20) and S2 (0.14).  
B2 might be impacted by the treatment plant outfall.  
S2 is a stream mouth and it may vary more because I may have sampled more river water one time and more lagoon water another time.  
  
```{r}
DOC_join %>%
  group_by(site) %>% 
  summarise(avg_npoc_mgl = round(mean(npoc_mg_l, na.rm = T), digits = 3),
            std_dev_npoc_mgl = round(sd(npoc_mg_l, na.rm = T),digits = 3),
            avg_tn_mgl = round(mean(tn_mg_l, na.rm = T), digits = 3),
            std_dev_tn_mgl = round(sd(tn_mg_l, na.rm = T),digits = 3)) %>% 
  gt() %>% 
  cols_label(avg_npoc_mgl = "Average NPOC (mg/L)",
             std_dev_npoc_mgl = "Std.Dev.NPOC (mg/L)",
             avg_tn_mgl = "Average TN (mg/L)",
             std_dev_tn_mgl= "Std.Dev.TN (mg/L)")
  
```



