---
title: "MOBEAST DNA"
format: html
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)

library(ComplexHeatmap)
library(data.table)
library(dendextend)

# Rayna library
library(dendsort)
library(ggsci)
library(randomForest)
library(vegan)
library(ape)
library(devtools)
library(ggpubr)
library(viridis)
library(colorspace)
library(readr)
library(circlize)

```

Heatmap info and tutorial: https://jokergoo.github.io/ComplexHeatmap-reference/book/a-single-heatmap.html

## Functions

### Heatmap Sean's function

```{r}
# Heat map funciton
# function to sum by family
prep_ht_mat = function(otus,
                       tax_table,
                       tax_level = "family",
                       cutoff = 40,
                       min_sam = 2,
                       scale_vals = F){
  if(tax_level == "OTU"){
    otu_sum = otus
    
  }else{
    otu_key = merge(otus,
                    tax_table[,
                              .SD,
                              .SDcols = c(tax_level, "OTU")], ## Make sure the seq are "ASV"
                    by = "OTU",
                    all.x = T)
    
    otu_key[, OTU := NULL]
    
    # sum by tax level
    otu_sum = otu_key[, lapply(.SD, sum), by = tax_level] #### maybe code crashes here
        ## Aggregates data by family
  }
  
  # order by most abundant taxa and subset to top N
  otu_sum[ , tax_sums := rowSums(otu_sum[ , .SD, .SDcols = !tax_level])]
  otu_sum = otu_sum[tax_sums > 0]
  otu_sum = otu_sum[order(tax_sums, decreasing = T)]
  otu_sum[ , tax_sums := NULL]
  otu_sum = otu_sum[1:cutoff,]
  
  # transform data.table to matrix
  otu_mat = as.matrix(otu_sum[, .SD,
                              .SDcols = !tax_level],
                      rownames.value = otu_sum[[tax_level]])
  
  # drop taxa that don't meet min sample count
  keep_tax = apply(otu_mat, 1, function(x)
    length(x[x > 0]) > min_sam)
  otu_mat = otu_mat[keep_tax, ]
  
  # optionally z-score a.k.a. scale values
  if (isTRUE(scale_vals)) {
    otu_mat = t(scale(t(otu_mat)))
  }
  
  return(otu_mat)
}
```

### Rayna's z-score function

```{r, echo=FALSE, collapse=TRUE}

zscore.calculation = function(x) {
  #Inputs: x = dummy variable that function will be run on. Must be a vector of numbers.
  
  zscoredat=x #duplicate vector x, save as a new df zscoredat.
  
  for (i in 1:length(x)) { #for each value in vector x
    zscoredat[i]=(x[i]-mean(x))/sd(x) #subset the input data to the corresponding position, and make it equal to the zscored value of the original cell value at that position.
    #zscoring is done by subtracting the mean value for a vector from the value of the vector at position i, and then dividing by the SD for the vector.
  }
  zscoredat1 <<- zscoredat #save output as new df, zscoredat1
}

```

## Data visualization theme

```{r}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        panel.background = element_rect(fill = "azure1"))
```

### Color palette

```{r}
treatment_colors = c(
  #Treatments
  "Control" = "grey",
  "Coral" = "#03A9F4",
  "CCA" = "#EF5350",
  "Algae" = "#4CAF50"
)

```

## Load Data

```{r}
OTU <- read_csv(here("data", "DNA","MOBEAST_OTU_culled.csv"))
meta <- read_csv(here("data","MOBEAST_metadata_heatmaps.csv"))
tax <- read_csv(here("data","DNA", "MOBEAST_tax_culled.csv"))
OTU_long <- read_csv(here("data","MOBEAST_DNA_long.csv"))

# Note: DO NOT DELETE METADATA_HEATMAP.CSV !
# Somehow if we use a clean-up version of the metadata, 
# heatmap annotations don't work!

```

## Data clean-up

Reformating the date, time and date-time in the metadata file

```{r, meta data clean up}
meta<- meta %>% mutate(time = str_pad(time, width = 4, side = "left", pad = "0"),
               time = str_replace(time, "(\\d{2})(\\d{2})", "\\1:\\2")) %>% 
  # "(\\d{2})(\\d{2})" separates the string in 2 groups of 2 characters
  # "\\1:\\2", group1 + ":" + group2
  mutate(time = paste0(time, ":00"),
               date_time = paste(date, time),
               date = mdy(date), 
               date_time = mdy_hms(date_time),
               time = factor(time))
  
  
```

### Preparing the metadata to annotate the heatmap

```{r}
# renaming some of treatment so that environmental data have a label: endmember or rain
# Same for inflow table, environmental samples now have a label
## Convert tibble to df for later
meta <- meta %>% mutate(treatment = as.character(treatment)) %>% 
  mutate(treatment = ifelse(id_number == "MOBEAST115", "Rain", 
                            ifelse(is.na(treatment), "Endmember", treatment))) %>% 
  mutate(inflow_table = ifelse(is.na(inflow_table), "Environmental data", inflow_table)) %>% 
  mutate(date_string = as.character(date),
                time_string = as.character(time),
                date_time_string = paste(date_string, time_string)) %>% 
  as.data.frame()
```

Clean up the long format data to make it easier to merge with the meta data file

```{r}
OTU_long<- OTU_long %>% clean_names() %>% 
  mutate(id_number = sample_id) %>% 
  relocate(id_number, .before= otu_number) %>% 
  select(-sample_id, -date, -treatment, -time)
```

Fusion!

```{r}
OTU_htm <- left_join(OTU_long, meta)
#write_csv(OTU_htm, "OTU_heatmap_ready.csv", path = here("data", "OTU_heatmap_ready.csv"))
```

## Z-Scored Heatmap Reference

### Preparing the OTU matrix for heatmap

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_htm %>%
  group_by(id_number, family) %>%
  summarize(abund = sum(abund, na.rm = TRUE))

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = family, values_from = abund)

```

```{r, nunu's}
# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% ungroup() %>% tibble() %>%  select(-id_number)

```

```{r, nunu's}
# Calculate Z-score on df/matrix
z.zscore.data.otu.subset <- as.data.frame(apply(culled.abund.OTU, 2, FUN=zscore.calculation))

# bring back the sample ID number in the df 
z.zscore.data.otu.subset$id_number = family.culled.abund$id_number
```

```{r, nunu's}
# Set the ID numbers as the rownames (to make a matrix)
rownames(z.zscore.data.otu.subset) = z.zscore.data.otu.subset$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
zscore.data.otu.matrix <- z.zscore.data.otu.subset %>% select(-id_number) %>% 
  as.matrix() %>% t()
```

### Generate Heatmap

#### Annotations for the heatmap

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number), 
                 "treatment"]

inflow_table <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number),
                     "inflow_table"]

date_time <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number),
                  "date_time_string"]


```

```{r}

ha = HeatmapAnnotation(
  treatment = treatment,
  date_time = date_time,
  col = list(treatment = c("Inflow"="#00838F","Control"="grey", "Coral"="#03A9F4", 
                           "CCA"="#EF5350", "Algae"="#4CAF50","Endmember" = "black", "Rain"= "yellow"),
             date_time = c("2024-06-02 12:00:00" = "red",
                           "2024-06-02 18:00:00" = "orange", 
                           "2024-06-03 00:00:00" = "yellow",
                           "2024-06-03 06:00:00" = "green",
                           "2024-06-04 12:00:00" = "royalblue",
                           "2024-06-06 12:00:00" = "blue",
                           "2024-05-27 15:00:00" = "purple",
                           "2024-05-27 16:45:00" = "purple",
                           "2024-05-25 15:10:00" = "purple",
                           "2024-05-25 15:00:00" = "purple",
                           "2024-05-25 08:45:00" = "purple",
                           "2024-06-06 14:25:00" = "hotpink"
                           ))
)


```

#### Print the Heatmap

```{r, fig.height = 10, fig.width = 16}
#png(width=1800, height=1000, file= here("outputs", "zscore_alldata.png"))

ht = Heatmap(zscore.data.otu.matrix, name = "z-scored", rect_gp = gpar(col = "white", lwd = 0.1),
        column_title = "id_number", column_title_side = "bottom",
        row_title = "family",
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10),
        row_km = 10, #number of forced segregation in rows 
        column_km = 6,#number of forced segregation in columns 
        top_annotation = ha
        )

draw(ht)

#dev.off()

```

## 12PM Z-Scored Heatmap

### Preparing the OTU matrix for heatmap

```{r}
OTU_htm_12PM <- OTU_htm %>% filter(time_string == "12:00:00")
```

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_htm_12PM %>%
  group_by(id_number, family) %>%
  summarize(abund = sum(abund, na.rm = TRUE))

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = family, values_from = abund)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% ungroup() %>% tibble() %>%  select(-id_number)

# Calculate Z-score on df/matrix
z.zscore.data.otu.subset <- as.data.frame(apply(culled.abund.OTU, 2, FUN=zscore.calculation))

# bring back the sample ID number in the df 
z.zscore.data.otu.subset$id_number = family.culled.abund$id_number


# Set the ID numbers as the rownames (to make a matrix)
rownames(z.zscore.data.otu.subset) = z.zscore.data.otu.subset$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
zscore.data.otu.matrix <- z.zscore.data.otu.subset %>% select(-id_number) %>% 
  as.matrix() %>% t()
```

### Generate Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number), 
                 "treatment"]

date_time <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number),
                  "date_time_string"]
```

```{r}
## 12PM Samples
ha = HeatmapAnnotation(
  treatment = treatment,
  date_time = date_time,
  col = list(treatment = c("Inflow"="#00838F","Control"="grey", "Coral"="#03A9F4", 
                           "CCA"="#EF5350", "Algae"="#4CAF50","Endmember" = "black", "Rain"= "yellow"),
             date_time = c("2024-06-02 12:00:00" = "red",
                           "2024-06-04 12:00:00" = "yellow",
                           "2024-06-06 12:00:00" = "blue"
                           ))
)


```

#### Printing heatmap

```{r, fig.height = 10, fig.width = 16}
#png(width=1800, height=1000, file= here("outputs", "zscore_12PMdata.png"))

ht = Heatmap(zscore.data.otu.matrix, name = "z-scored", rect_gp = gpar(col = "white", lwd = 0.1),
        column_title = "id_number", column_title_side = "bottom",
        row_title = "family",
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10),
        row_km = 10, #number of forced segregation in rows 
        column_km = 6,#number of forced segregation in columns 
        top_annotation = ha
        )

draw(ht)

#dev.off()

```

### 12PM color by day

```{r}
OTU_htm_12PM <- OTU_htm %>% filter(time_string == "12:00:00") 
```

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_htm_12PM %>%
  group_by(id_number, family) %>%
  summarize(abund = sum(abund, na.rm = TRUE))

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = family, values_from = abund)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% ungroup() %>% tibble() %>%  select(-id_number)

# Calculate Z-score on df/matrix
z.zscore.data.otu.subset <- as.data.frame(apply(culled.abund.OTU, 2, FUN=zscore.calculation))

# bring back the sample ID number in the df 
z.zscore.data.otu.subset$id_number = family.culled.abund$id_number


# Set the ID numbers as the rownames (to make a matrix)
rownames(z.zscore.data.otu.subset) = z.zscore.data.otu.subset$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
zscore.data.otu.matrix <- z.zscore.data.otu.subset %>% select(-id_number) %>% 
  as.matrix() %>% t()
```

### Generate Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number), 
                 "treatment"]

date_time <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number),
                  "date_time_string"]
```

```{r}
## 12PM Samples
ha = HeatmapAnnotation(
  #treatment = treatment,
  date_time = date_time,
  col = list(#treatment = c("Inflow"="#00838F","Control"="grey", "Coral"="#03A9F4", 
                           #"CCA"="#EF5350", "Algae"="#4CAF50","Endmember" = "black", "Rain"= "yellow"),
             date_time = c("2024-06-02 12:00:00" = "red",
                           "2024-06-04 12:00:00" = "yellow",
                           "2024-06-06 12:00:00" = "blue"
                           ))
)


```

#### Printing heatmap

```{r, fig.height = 10, fig.width = 16}
#png(width=1800, height=1000, file= here("outputs", "zscore_12PMdata.png"))

ht = Heatmap(zscore.data.otu.matrix, name = "z-scored", rect_gp = gpar(col = "white", lwd = 0.1),
        column_title = "id_number", column_title_side = "bottom",
        row_title = "family",
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10),
        row_km = 10, #number of forced segregation in rows 
        column_km = 6,#number of forced segregation in columns 
        top_annotation = ha
        )

draw(ht)

#dev.off()

```

## June 4 and 6 map

```{r}
OTU_htm_June4_6 <- OTU_htm %>% filter(time_string == "12:00:00") %>% 
  filter(date_string == "2024-06-04" | date_string == "2024-06-06" ) 
```

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_htm_June4_6 %>%
  group_by(id_number, family) %>%
  summarize(abund = sum(abund, na.rm = TRUE))


# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = family, values_from = abund)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% ungroup() %>% tibble() %>%  select(-id_number)

# Sum the total abundance by families and only keep the name of the families that have a 
# total abundance greater than 0.
sum_abun_fam <- culled.abund.OTU %>% colSums() %>% as.data.frame() 
colnames(sum_abun_fam) <- "abund"
sum_abun_fam<- sum_abun_fam %>% mutate(family = rownames(sum_abun_fam)) %>% filter(abund >0) 
keep_family <- sum_abun_fam$family

## Then filter in the main data from such that only-only 0 families remain in the heatmap.
culled.abund.OTU <- culled.abund.OTU[, colnames(culled.abund.OTU) %in% keep_family]

# Calculate Z-score on df/matrix
z.zscore.data.otu.subset <- as.data.frame(apply(culled.abund.OTU, 2, FUN=zscore.calculation))

# bring back the sample ID number in the df 
z.zscore.data.otu.subset$id_number = family.culled.abund$id_number


# Set the ID numbers as the rownames (to make a matrix)
rownames(z.zscore.data.otu.subset) = z.zscore.data.otu.subset$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
zscore.data.otu.matrix <- z.zscore.data.otu.subset %>% select(-id_number) %>% 
  as.matrix() %>% t()
```

### Generate Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number), 
                 "treatment"]

date_time <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number),
                  "date_time_string"]
```

```{r}
## 12PM Samples
ha = HeatmapAnnotation(
  treatment = treatment,
  date_time = date_time,
  col = list(treatment = c("Inflow"="#00838F","Control"="grey", "Coral"="#03A9F4", 
                           "CCA"="#EF5350", "Algae"="#4CAF50","Endmember" = "black", "Rain"= "yellow"),
             date_time = c("2024-06-02 12:00:00" = "red",
                           "2024-06-04 12:00:00" = "yellow",
                           "2024-06-06 12:00:00" = "blue"
                           ))
)


```

#### Printing heatmap

```{r, fig.height = 10, fig.width = 16}
#png(width=1800, height=1000, file= here("outputs", "zscore_12PMdata.png"))

ht = Heatmap(zscore.data.otu.matrix, name = "z-scored", rect_gp = gpar(col = "white", lwd = 0.1),
        column_title = "id_number", column_title_side = "bottom",
        row_title = "family",
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10),
        row_km = 10, #number of forced segregation in rows 
        column_km = 4,#number of forced segregation in columns 
        top_annotation = ha
        )

draw(ht)

#dev.off()

```

## 24h Z-Scored Heatmap

### Preparing the OTU matrix for heatmap

```{r}
OTU_htm_24h <- OTU_htm %>% 
  filter(date == ymd("2024-06-02")| date == ymd("2024-06-03")) 
OTU_htm_24h <- OTU_htm_24h %>% filter(family != "Bradymonadales_fa")
```

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_htm_24h %>%
  group_by(id_number, family) %>%
  summarize(abund = sum(abund, na.rm = TRUE))

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = family, values_from = abund)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% ungroup() %>% tibble() %>%  select(-id_number)

# Calculate Z-score on df/matrix
z.zscore.data.otu.subset <- as.data.frame(apply(culled.abund.OTU, 2, FUN=zscore.calculation))

# bring back the sample ID number in the df 
z.zscore.data.otu.subset$id_number = family.culled.abund$id_number


# Set the ID numbers as the rownames (to make a matrix)
rownames(z.zscore.data.otu.subset) = z.zscore.data.otu.subset$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
zscore.data.otu.matrix <- z.zscore.data.otu.subset %>% select(-id_number) %>% 
  as.matrix() %>% t()
```

### Generate Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number), 
                 "treatment"]

date_time <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number),
                  "date_time_string"]
```

```{r}
## 12PM Samples
ha = HeatmapAnnotation(
  treatment = treatment,
  date_time = date_time,
  col = list(treatment = c("Inflow"="#00838F","Control"="grey", "Coral"="#03A9F4", 
                           "CCA"="#EF5350", "Algae"="#4CAF50",
                           "Endmember" = "black", "Rain"= "yellow"),
             date_time = c("2024-06-02 12:00:00" = "red",
                           "2024-06-02 18:00:00" = "yellow", 
                           "2024-06-03 00:00:00" = "green",
                           "2024-06-03 06:00:00" = "blue"
                           ))
)


```

#### Printing heatmap

```{r, fig.height = 10, fig.width = 16}
#png(width=1800, height=1000, file= here("outputs", "zscore_24hdata.png"))

ht = Heatmap(zscore.data.otu.matrix, name = "z-scored", 
             #rect_gp = gpar(col = "white", lwd = 0.1),
        
        column_title = "24h Sampling - All samples", 
        column_title_gp = gpar(fontsize = 16, fontface = "bold"),
        #column_title_side = "bottom",
        row_title = "Family",
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10),
        row_km = 8, #number of forced segregation in rows 
        column_km = 4,#number of forced segregation in columns 
        top_annotation = ha
        )
set.seed(1999)
ht_drawn <- draw(ht)

#dev.off()

```

### Cluster membership

```{r}
col_orders<- column_order(ht_drawn)

cluster_df <- data.frame(
  id_number = unlist(col_orders),
  cluster  = rep(seq_along(col_orders), lengths(col_orders))
)

rownames(cluster_df)<- NULL

```

```{r}
cluster_df<- cluster_df %>% 
  mutate(id_number = ifelse(id_number<10,
                            paste0("MOBEAST00",id_number), 
                            paste0("MOBEAST0",id_number)))
#write_csv(cluster_df, here("data", "DNA", "24h_clusters.csv"))
```

### Cluster membership - code by Sean

Not useful for cluster membership but to see the dendrograms, it's useful!

```         
# choose a distance metric
# note: you could use vegdist to do something like Bray curtis distances instead
# by default Heatmap uses 'euclidean' distances
our_dist_method = "euclidean"

#the distance measure to be used. This must be one of "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski".

# choose a method for clustering
# note: by default Heatmap uses 'complete' clustering
#our_clust_method = "ward.D"
our_clust_method = "complete"


#plot(analyte_clusts)

# cluster rows and columns
# note: dist generates a distance matrix and hclust does the clustering
row_dend = as.dendrogram(hclust(dist(zscore.data.otu.matrix, 
                                     method = our_dist_method),
                                method = our_clust_method))

col_dend = as.dendrogram(hclust(dist(t(zscore.data.otu.matrix), 
                                     method = our_dist_method),
                                method = our_clust_method))

# pick an appropriate number of groups for splitting rows and columns
row_k = 8
col_k = 4

# split into clusters
#note: sample clusts and analyte clusts will tell you which cluster each sample/analyte is in!
sample_clusts = cutree(row_dend,
                       row_k,
                       order_clusters_as_data = F)

analyte_clusts = cutree(col_dend,
                        col_k,
                        order_clusters_as_data = Fplot(analyte_clusts))

# store sample clusters as a dataframe which can be merged with your other sample metadata
analyte_clust_df = data.frame(id_number = names(analyte_clusts), cluster = analyte_clusts)



# you can pick custom colors if you want to
#row_dend_k = color_branches(row_dend,
                            #clusters = sample_clusts)

#col_dend_k = color_branches(col_dend,
                            #clusters = analyte_clusts)

#ht_k = Heatmap(scale_mat,
             #cluster_rows = row_dend_k,
             #cluster_columns = col_dend_k,
             #row_split = row_k,
             #column_split = col_k
             #)
             
#print(ht_k)
```

```{r, fig.width= 16, fig.height= 4}
par(mfrow = c(2,1))
plot(col_dend)
```

## 24h separated maps

### Preparing the OTU matrix for heatmap

```{r}
OTU_htm_24_12 <- OTU_htm %>% 
  filter(date == ymd("2024-06-02") & time_string == "12:00:00") %>% 
  filter(family != "Bradymonadales_fa")

OTU_htm_24_18 <- OTU_htm %>% 
  filter(time_string == "18:00:00") %>% 
  filter(family != "Bradymonadales_fa")

OTU_htm_24_00 <- OTU_htm %>% 
  filter(time_string == "00:00:00") %>% 
  filter(family != "Bradymonadales_fa")

OTU_htm_24_06 <- OTU_htm %>% 
  filter(time_string == "06:00:00") %>% 
  filter(family != "Bradymonadales_fa")

```

### 24h - 12:00 map

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_htm_24_12 %>%
  group_by(id_number, family) %>%
  summarize(abund = sum(abund, na.rm = TRUE))

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = family, values_from = abund)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% ungroup() %>% tibble() %>%  select(-id_number)

# Calculate Z-score on df/matrix
z.zscore.data.otu.subset <- as.data.frame(apply(culled.abund.OTU, 2, FUN=zscore.calculation))

# bring back the sample ID number in the df 
z.zscore.data.otu.subset$id_number = family.culled.abund$id_number


# Set the ID numbers as the rownames (to make a matrix)
rownames(z.zscore.data.otu.subset) = z.zscore.data.otu.subset$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
zscore.data.otu.matrix <- z.zscore.data.otu.subset %>% select(-id_number) %>% 
  as.matrix() %>% t()
```

### Generate Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number), 
                 "treatment"]

#date_time <- meta[match(colnames(zscore.data.otu.matrix),
                  #         meta$id_number),
                 # "date_time_string"]
```

```{r}
## 12PM Samples
ha = HeatmapAnnotation(
  treatment = treatment,
  #date_time = date_time,
  col = list(treatment = c("Inflow"="#00838F","Control"="grey", "Coral"="#03A9F4", 
                           "CCA"="#EF5350", "Algae"="#4CAF50","Endmember" = "black", "Rain"= "yellow"),
             date_time = c("2024-06-02 12:00:00" = "red"#,
                           #"2024-06-02 18:00:00" = "yellow", 
                           #"2024-06-03 00:00:00" = "green",
                           #"2024-06-03 06:00:00" = "blue"
                           ))
)


```

#### Printing heatmap

```{r, fig.height = 10, fig.width = 16}
#png(width=1800, height=1000, file= here("outputs", "zscore_24hdata.png"))

ht = Heatmap(zscore.data.otu.matrix, name = "z-scored", rect_gp = gpar(col = "white", lwd = 0.1),
        column_title = "id_number", column_title_side = "bottom",
        row_title = "family",
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10),
        row_km = 10, #number of forced segregation in rows 
        column_km = 4,#number of forced segregation in columns 
        top_annotation = ha
        )

draw(ht)

#dev.off()

```

### 24h - 06:00 map

```{r, echo=TRUE, collapse=TRUE}
#Make a heatmap of z-scored relative abundance data after it was summed by family
#https://jokergoo.github.io/ComplexHeatmap-reference/book/heatmap-annotations.html

# some the abundance of bacteria by family
family.culled.long.table <- OTU_htm_24_06 %>%
  group_by(id_number, family) %>%
  summarize(abund = sum(abund, na.rm = TRUE))

# Make wide format table to make a matrix later
family.culled.abund <- pivot_wider(family.culled.long.table, names_from = family, values_from = abund)

# Remove ID number from the data frame
## Z-score function only works on df that has a matrix format
culled.abund.OTU <- family.culled.abund %>% ungroup() %>% tibble() %>%  select(-id_number)

# Calculate Z-score on df/matrix
z.zscore.data.otu.subset <- as.data.frame(apply(culled.abund.OTU, 2, FUN=zscore.calculation))

# bring back the sample ID number in the df 
z.zscore.data.otu.subset$id_number = family.culled.abund$id_number


# Set the ID numbers as the rownames (to make a matrix)
rownames(z.zscore.data.otu.subset) = z.zscore.data.otu.subset$id_number

# Remove the ID number column, convert the df to a matrix and transpose to make the final matrix.
zscore.data.otu.matrix <- z.zscore.data.otu.subset %>% 
  select(-id_number, - Moraxellaceae, -Idiomarinaceae ) %>% # removing NaN columns
  as.matrix() %>% t()
```

### Generate Heatmap

#### Annotation

```{r}
## Create a vector that will have information of a variable and its associate sample 
treatment <- meta[match(colnames(zscore.data.otu.matrix),
                           meta$id_number), 
                 "treatment"]

#date_time <- meta[match(colnames(zscore.data.otu.matrix),
                  #         meta$id_number),
                  #"date_time_string"]
```

```{r}
## 12PM Samples
ha = HeatmapAnnotation(
  treatment = treatment,
  #date_time = date_time,
  col = list(treatment = c("Inflow"="#00838F","Control"="grey", "Coral"="#03A9F4", 
                           "CCA"="#EF5350", "Algae"="#4CAF50","Endmember" = "black", "Rain"= "yellow"),
             date_time = c(#"2024-06-02 12:00:00" = "red",
                           #"2024-06-02 18:00:00" = "yellow", 
                           #"2024-06-03 00:00:00" = "green",
                           "2024-06-03 06:00:00" = "blue"
                           ))
)


```

#### Printing heatmap

```{r, fig.height = 10, fig.width = 16}
#png(width=1800, height=1000, file= here("outputs", "zscore_24hdata.png"))

ht = Heatmap(zscore.data.otu.matrix, name = "z-scored", rect_gp = gpar(col = "white", lwd = 0.1),
        column_title = "id_number", column_title_side = "bottom",
        row_title = "family",
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10),
        row_km = 10, #number of forced segregation in rows 
        column_km = 6,#number of forced segregation in columns 
        top_annotation = ha
        )

draw(ht)

#dev.off()

```
