---
title: "MOBEAST fDOM V3"
format: 
  html:
    toc: true
    toc-location: left
editor: source
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

## Load Libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggrepel)
library(ghibli)
library(gt)
library(patchwork)
library(RColorBrewer)
library(ggridges)

library(car) # ANOVA package
library(ggeffects)
library(effects) #dependency for ggeffects
library(ggResidpanel) #residual panel
library(lme4)
library(lmerTest)
library(sjPlot)
#library(GGally)
library(emmeans)
library(performance)


```




## Data visualization theme

```{r}
nunu_theme <- theme_bw()+ 
  theme(plot.title = element_text(size=24, face = "bold"), 
        plot.subtitle = element_text(size=20),
        axis.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 16, face = "bold"),
        legend.title=element_text(size=18, face = "bold"),
        legend.text=element_text(size=16),
        strip.background = element_rect(fill = "white"))
```

```{}
nunu_theme <- theme_bw() + 
  theme(plot.title = element_text(size=32, face = "bold"), 
        plot.subtitle = element_text(size=30),
        axis.title = element_text(size = 28, face = "bold"),
        axis.text = element_text(size = 28),
        strip.text.x = element_text(size = 28, face = "bold"),
        legend.title=element_text(size=28, face = "bold"),
        legend.text=element_text(size=28))
```

```{r}

treatment_colors = c("Inflow"="#00838F",
                     "Control"="grey", 
                     "Coral"="#03A9F4", 
                     "CCA"="#EF5350", 
                     "Algae"="#4CAF50",
                     "Endmember" = "black", 
                     "Rain"= "yellow")
```
## Load Data

```{r}
fDOM <- read_csv(here("data", "fDOM", "fDOM_sorted.csv"))

fDOM <- fDOM %>% mutate(treatment = ifelse(treatment == "Rubble", "CCA", treatment)) %>%
  select(-sample_type)

table_id <- read_csv(here("data", "Laurel", "TableID.csv"))
```

## Data Clean-Up
### Sub-catergorization of fDOM types
```{r}
fDOM_desc <- tibble(fdom = c("coble_a","coble_b","coble_c","coble_m","coble_t",
  "m_to_c","bix","hix","fpeak","fi",
  "lignin"),
  fdom_alt_name = c("UV-Humic","Tyrosine","Visible-Humic","Marine Humic","Tryptophan",
  "m_to_c","bix","hix","Phenylalanine","Fluvic Acid",
  "lignin"),
  fdom_cat= c("Humic","Proteinaceous","Humic","Humic","Proteinaceous",
  "Humic","Humic","Humic","Proteinaceous","Humic",
  "Other"))
```

### Assign Table ID to samples
```{r}
table_id <- clean_names(table_id)

table_id <- table_id %>% rename(tank_number=tankid) %>% select(-treatment) %>% 
  mutate(tank_number = paste0("T", tank_number)) %>% 
  mutate(inflow_table = paste("Table", inflow_table))
fDOM <- left_join(fDOM, table_id, by= "tank_number")
```

```{r}
fDOM <- fDOM %>% mutate(inflow_table = ifelse(is.na(inflow_table), tank_number, inflow_table)) %>% 
   mutate(inflow_table = ifelse(inflow_table == "F1", "Table 1", 
                                ifelse(inflow_table == "F2", "Table 2", inflow_table))) %>% 
  mutate(inflow_table = factor(inflow_table)) %>% 
  relocate(inflow_table, .after = "treatment") %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow", "Endmember")))
```

### Create long format df to make data analysis easier later
```{r}
fDOM_long <- fDOM %>% pivot_longer(cols = coble_a:lignin,
                      names_to = "fdom", 
                      values_to = "fdom_conc") %>% 
  select(-long_name)
```

Attaching the subcategories of fDOM to the long df
```{r}
fdom_long <- left_join(fDOM_long, fDOM_desc, by = "fdom")
```


## Outlier detection 

```{, log10 data for IQR}
# ATTENTION !!!!!!!!!!!!! Data was log transform to do the IQR method. 

## removing outliers from the data set
raw_flagged_outliers <- fdom_long%>%
  filter(treatment != "Endmember") %>% 
  # removing outliers
  ## grouping by date, time and treatment
  group_by(fdom, treatment) %>% 
  ## caluculating the Interquantile range and of the raw data
  
# !!!!!!!!!!!!!  ATTENTION REQUIRE BELOW !!!!!!!!!!!!! 

  mutate(quant1 = quantile(log10(fdom_conc), probs = 0.25, na.rm = T),
         quant3 = quantile(log10(fdom_conc), probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse( log10(fdom_conc) > quant3 + (IQR*1.5) | log10(fdom_conc) < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup()  
```


```{r}
## removing outliers from the data set
raw_flagged_outliers <- fdom_long%>%
  filter(treatment != "Endmember") %>% 
  # removing outliers
  ## grouping by date, time and treatment
  group_by(fdom, treatment) %>% 
  ## calculating the Interquantile range and of the raw data

  mutate(quant1 = quantile(fdom_conc, probs = 0.25, na.rm = T),
         quant3 = quantile(fdom_conc, probs = 0.75, na.rm = T),
         IQR = quant3-quant1) %>% 
  ## removing data that are under Q1-1.5* or over 1.5+Q3*IQR, and labeling them outlier/not outliers
  ## Important to filter later to understand why they are outliers. 
  mutate(outlier = ifelse(fdom_conc > quant3 + (IQR*1.5) | fdom_conc < quant1 - (IQR*1.5), "outlier", "not outlier")) %>% 
  ungroup()  
```


```{r}
# Normalization version 2
## averaging the inflows for normalization 
## remove the outliers a right before the normalization 

## Normalization of the data to the inflow 
#nutrients_norm <- 
fDOM_norm <- raw_flagged_outliers %>% 
  mutate(time_string = as.character(time),
         date_string = as.character(date),
         date_time_string = paste(date_string, time_string),
         date_time = ymd_hms(date_time_string)) %>% 
  select(-IQR, -quant1, -quant3) %>% 
  
  filter(treatment != "Rain") %>% 
  
  # Isolate inflow for each date_time and table
  group_by(date_time, inflow_table, fdom) %>%
  mutate(inflow = fdom_conc[treatment == "Inflow"]) %>%  
  ungroup() %>%
  
  # averaging the inflow values 
  group_by(date_time, fdom) %>% 
  mutate(inflow = ifelse(outlier == "outlier", inflow == NA, inflow)) %>% 
  mutate(inflow = mean(inflow, na.rm = TRUE)) %>%
  ungroup() %>% 
  
  # Normalize for treatments
  mutate(conc_norm = fdom_conc - inflow)  %>%  
  relocate(inflow, .after = "conc_norm") %>% 
  relocate(date_time, .before = "date") %>%
  
  #organize the treatment levels
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae", "Inflow")))
```

```{r}

## Check outlier detection 
 fDOM_norm %>% filter(outlier == "outlier") %>% 
  select(id_number, tank_number, treatment, date_time, fdom, fdom_conc, conc_norm) %>% 
  group_by(fdom) %>% 
  summarise(number = n())


```

```{r, warning=FALSE}
fDOM_norm %>% 
  #filter(outlier == "outlier") %>% 
  ggplot(aes(treatment, fdom_conc, fill = treatment))+
  #geom_violin()+
  geom_point(aes(color = outlier), size = 1.7)+
  facet_wrap(.~fdom, scales = "free")+
  scale_fill_manual(values = treatment_colors)+

  scale_color_manual(values = c("white", "red"))
  
```


## 12PM data
```{r}
# Version 2
## removing outliers from the data set
fDOM_12PM <- fDOM_norm %>%
  filter(time_string == "12:00:00") %>% 
  filter(treatment != "Inflow") %>% 
  
  ## remove the outliers 
  filter(outlier == "not outlier") 

```

### Iterative statistical modeling

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

fDOM_loop <- fDOM_12PM %>% select(fdom) %>% distinct() %>% pull()

models_12PM <- vector(mode = "list", length = length(fDOM_loop))
names(models_12PM) <- fDOM_loop

for (var in fDOM_loop) {
  df <- fDOM_12PM %>% filter(fdom == var)
  models_12PM[[var]] <- lmer(conc_norm ~ treatment + (1|tank_number), data = df)
}

```

### Iterative graphing

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## --> In the for-loop, filter the df by nutrient and make a graph that will display nutrient data
## Save the outputs of the model in the list such that each nutrient has its stats

fDOM_loop <- fDOM_12PM %>% select(fdom) %>% distinct() %>% pull()

graphs_12PM <- vector(mode = "list", length = length(fDOM_loop))
names(graphs_12PM) <- fDOM_loop

for (var in fDOM_loop) {
  
  df <- fDOM_12PM %>% filter(fdom == var)
  
  graphs_12PM[[var]] <- df %>%  
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("12PM", var, "concentrations"),
       x= "Treatment",
       y = "Concentration (R.U.)")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none")
}

```

### Humics
#### CobleA

```{r}
summary(models_12PM[["coble_a"]])
anova(models_12PM[["coble_a"]], type = 2)
emmeans(models_12PM[["coble_a"]], specs = trt.vs.ctrl ~ treatment)
```

```{r}
resid_panel(models_12PM[["coble_a"]])
graphs_12PM[["coble_a"]]
```

##### !!!! T-test  Coral sig diff
 
```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_12PM %>%
  filter(fdom == "coble_a") %>% 
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```



#### CobleM

```{r}
summary(models_12PM[["coble_m"]])
anova(models_12PM[["coble_m"]], type = 2)
emmeans(models_12PM[["coble_m"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[[1]], ~ treatment)
```

```{r}
resid_panel(models_12PM[["coble_m"]])
graphs_12PM[["coble_m"]]
```

##### !!!! T-test  Coral sig diff
 
```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_12PM %>%
  filter(fdom == "coble_m") %>% 
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```

#### CobleC *
Sig ANOVA with log10 data outlier detection but not contrast
```{r}
summary(models_12PM[["coble_c"]])
anova(models_12PM[["coble_c"]], type = 2)
emmeans(models_12PM[["coble_c"]], specs = trt.vs.ctrl ~ treatment)

```

```{r, fig.width= 12, fig.height = 6}
resid_panel(models_12PM[["coble_c"]])
graphs_12PM[["coble_c"]]
```

##### !!!! T-test sig 
```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_12PM %>%
  filter(fdom == "coble_c") %>% 
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```



#### Sig! M to C

```{r}
summary(models_12PM[["m_to_c"]])
anova(models_12PM[["m_to_c"]], type = 2)
emmeans(models_12PM[["m_to_c"]], specs = trt.vs.ctrl ~ treatment)

```

```{r,fig.width= 12, fig.height = 6 }
resid_panel(models_12PM[["m_to_c"]])
graphs_12PM[["m_to_c"]]+
  labs(y = "Marine:Colored Humics Ratio")

ggsave("MtoC_12PM.png", width = 16, height = 8, path = here("outputs"))
```

##### T-test Sig 
```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_12PM %>%
  filter(fdom == "m_to_c") %>% 
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```

#### FI

```{r}
summary(models_12PM[["fi"]])
anova(models_12PM[["fi"]], type = 2)
emmeans(models_12PM[["fi"]], specs = trt.vs.ctrl ~ treatment)

```

```{r,fig.width= 12, fig.height = 6}
resid_panel(models_12PM[["fi"]])
graphs_12PM[["fi"]]+
  labs(title = "12PM Fluvic Acid")
```


##### !!!! Sig diff coral 

```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_12PM %>%
  filter(fdom == "fi") %>% 
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```

#### BIX

```{r}
summary(models_12PM[["bix"]])
anova(models_12PM[["bix"]], type = 2)
emmeans(models_12PM[["bix"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["bix"]])
graphs_12PM[["bix"]]
```

##### !!! All sig diff compared to 0 

CONTROL IS NOT SUPPOSED TO BE DIFFERENT !!! ?????? WHAT ????
```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_12PM %>%
  filter(fdom == "bix") %>% 
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```


#### HIX

```{r}
summary(models_12PM[["hix"]])
anova(models_12PM[["hix"]], type = 2)
emmeans(models_12PM[["hix"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["hix"]])
graphs_12PM[["hix"]]
```

##### !!! Control & Coral sig diff

```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_12PM %>%
  filter(fdom == "hix") %>% 
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```


### Proteinaceous
#### CobleB

```{r}
summary(models_12PM[["coble_b"]])
anova(models_12PM[["coble_b"]], type = 2)
emmeans(models_12PM[["coble_b"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["coble_b"]])
graphs_12PM[["coble_b"]]
```

##### Sig in Coral and Algae
```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_12PM %>%
  filter(fdom == "coble_b") %>% 
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```

#### CobleT

```{r}
summary(models_12PM[["coble_t"]])
anova(models_12PM[["coble_t"]], type = 2)
emmeans(models_12PM[["coble_t"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["coble_t"]])
graphs_12PM[["coble_t"]]
```

##### Sig in Coral, CCA and Algae
```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_12PM %>%
  filter(fdom == "coble_t") %>% 
  group_by(treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```

#### Fpeak

```{r}
summary(models_12PM[["fpeak"]])
anova(models_12PM[["fpeak"]], type = 2)
emmeans(models_12PM[["fpeak"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_12PM[["fpeak"]])
graphs_12PM[["fpeak"]]
```

##### T-test NO sig

```{r}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_12PM %>%
  filter(fdom == "fpeak") %>% 
  group_by( treatment) %>%
  summarize(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value
  )

# p<0.05, reject the null hypothesis, the difference in mean is not due to chance
```

#### Graphs 12PM 

```{r, fig.width=6, fig.height = 10}
fDOM_12PM %>%  
  filter(fdom_cat == "Humic") %>% 
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("12PM concentrations"),
       x= "Treatment",
       y = "Concentration (R.U.")+
    facet_grid(fdom~., scales = "free")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme + theme_bw()+
  theme(legend.position = "none")

```


### Heatmap summary 

```{, fig.width= 14, fig.height= 8}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_12PM %>%
  
  group_by(fdom_alt_name, treatment) %>%
  reframe(
    t_test = list(t.test(conc_norm, mu = 0)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value,
    fdom_cat=fdom_cat) %>% distinct() %>% 
  select(-t_test) %>% 
  
  mutate(significance = ifelse(p_value <0.05, "Enrichment", "No Difference"),
         significance = ifelse(mean <0 & p_value <0.05, "Depletion", significance)#,
         #significance = as.character(significance)
         ) %>% 
    ungroup() %>% 
  
  arrange(fdom_cat, fdom_alt_name) %>%   # ensures ordering in the data frame
  mutate(fdom_alt_name = factor(fdom_alt_name, levels = c("UV-Humic", "Marine Humic","Visible-Humic", "m_to_c", "Fluvic Acid", "bix", "hix", "Tyrosine", "Tryptophan", "Phenylalanine", "lignin")),
         fdom_alt_name = fct_rev(fdom_alt_name)# lock that order in y-axis
  ) %>% 
   mutate(fdom_cat = factor(fdom_cat, levels = c("Other", "Proteinaceous", "Humic"))) %>% 
  filter(fdom_alt_name != "m_to_c",
         fdom_alt_name != "lignin") %>% 
         #fdom_alt_name != "bix",
         #fdom_alt_name != "hix") %>% 
  
  ggplot(aes(x = treatment, y= fdom_alt_name, fill = significance))+
  #geom_tile(aes(color = fdom_cat), size = 1)+
  geom_tile(color = "white", size = 1)+
  labs(title = "fDOM Concentration Change in Mesocosms",
       x= "Treatment",
       y = "fDOM Type")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  #scale_fill_distiller(palette = "RdBu", direction = 1)+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw()+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"))

ggsave("fdom_tiles_12PM.png",width = 14, height = 8, path = here("outputs"))
```

## 24h data 
```{r}
# Version 2
## removing outliers from the data set
fDOM_24h <- fDOM_norm %>%
  filter(date_string == "2024-06-02" | date_string == "2024-06-03" ) %>% 
  filter(treatment != "Inflow") %>% 
  
  ## remove the outliers 
  filter(outlier == "not outlier") %>% 
  mutate(time_fac = factor(time_string, levels = c("12:00:00", "18:00:00",
                                                   "00:00:00","06:00:00"))) %>% 
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae")))

```

### Iterative statistical modeling

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## In the for-loop, filter the df by nutrient and apply a statistical model with lmer
## Save the outputs of the model in the list such that each nutrient has its stats

fDOM_loop <- fDOM_24h %>% select(fdom) %>% distinct() %>% pull()

models_24h <- vector(mode = "list", length = length(fDOM_loop))
names(models_24h) <- fDOM_loop

for (var in fDOM_loop) {
  df <- fDOM_24h %>% filter(fdom == var)
  models_24h[[var]] <- lmer(conc_norm ~ treatment*time_fac + (1|tank_number), data = df)
}

```

### Iterative graphing

```{r}
# Loop Function:
## Take the 12PM samples and isolate the types of nutrient present on the df in a vector
## Create a list where the loop results will be save, named after the nutrients of the df
## --> In the for-loop, filter the df by nutrient and make a graph that will display nutrient data
## Save the outputs of the model in the list such that each nutrient has its stats

fDOM_loop <- fDOM_24h %>% select(fdom) %>% distinct() %>% pull()

graphs_24h <- vector(mode = "list", length = length(fDOM_loop))
names(graphs_24h) <- fDOM_loop

for (var in fDOM_loop) {
  
  df <- fDOM_24h %>% filter(fdom == var)
  
  graphs_24h[[var]] <- df %>%  
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("24h", var, "concentrations"),
       x= "Treatment",
       y = "Concentration (R.U.)")+
    facet_wrap(~time_fac, ncol = 4 )+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme +
  theme(legend.position = "none")
}

```


### Pairwise comparison function

```{r}
nunu_pairwise <- function(data, type = "coble_a", time_point = 6){

  ## Get a list of the treatments for the for loop   
  treatment <- data %>% mutate(treatment = as.character(treatment)) %>%
  select(treatment) %>%  distinct() %>% pull()
  
  ## Create un empty data frame
pairwise_results <- tibble(
  time_point = vector(),
  treatment= vector(),
  mean = vector(),
  p_value = vector())

# this is the pairwise comparison for the 6 pairs (6AM-6PM)
  if(time_point == 6){
    for(var in treatment){
      
      ## only keep the 6AM and 6PM
      df <- data %>% 
        filter(fdom == type) %>% 
        filter(treatment == var) %>% 
  filter(time_string == "18:00:00" | time_string == "06:00:00") %>% 
        ## Rename the times so they can become column names
  mutate(time_US = ifelse(time_string == "18:00:00", "T18", "T06")) %>% 
        ## Only keep the column of interest for the stats
  select(tank_number, time_US, conc_norm) %>% 
  pivot_wider(names_from = time_US, values_from = conc_norm) 
      
      ## Pairwise function 
      paired <- t.test(df$T18, df$T06, paired = TRUE, na.action = na.pass)
      
      ## Save the result in at data frame
      df2 <- tibble(
        time_point = "06AM-06PM",
        fdom = type,
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      # Code will loop oveer all treatments and save the results in a data frame
      pairwise_results <-rbind(pairwise_results,df2)
      
      # Save the results in RStudio
      pairwise_results <<- pairwise_results
    }  
  

# This is for the 12 samples (12AM-12PM)
  }else if(time_point == 12){
    for(var in treatment){
      
      ## Same as before but for the 12AM-12PM
      df <- data %>% 
        filter(fdom == type) %>% 
        filter(treatment == var) %>% 
  filter(time_string == "12:00:00" | time_string == "00:00:00") %>% 
  mutate(time_US = ifelse(time_string == "12:00:00", "T12", "T00")) %>% 
  select(tank_number, time_US, conc_norm) %>% 
  pivot_wider(names_from = time_US, values_from = conc_norm) 
      
      paired <- t.test(df$T12, df$T00, paired = TRUE, na.action = na.pass)
      
      df2 <- tibble(
        time_point = "12AM-12PM",
        fdom = type,
        treatment = var,
        mean = paired$estimate, 
        p_value = paired$p.value)
      
      pairwise_results <- rbind(pairwise_results, df2)
      pairwise_results <<- pairwise_results
    }
}

# Display the results of the code 
print(pairwise_results)
}

```

### Humics
#### CobleA

```{r}
summary(models_24h[["coble_a"]])
anova(models_24h[["coble_a"]], type = 2)
emmeans(models_24h[["coble_a"]], specs = trt.vs.ctrl ~ treatment)
```

```{r}
resid_panel(models_24h[["coble_a"]])
graphs_24h[["coble_a"]]
```




##### Pairwise

```{r}
nunu_pairwise(data = fDOM_24h, time_point = 6, type = "coble_a")

pairwise_summary <- nunu_pairwise(data = fDOM_24h, time_point = 12, type = "coble_a")
```




#### CobleM

```{r}
summary(models_24h[["coble_m"]])
anova(models_24h[["coble_m"]], type = 2)
emmeans(models_24h[["coble_m"]], specs = trt.vs.ctrl ~ treatment)
#emmeans(models_12PM[[1]], ~ treatment)
```

```{r}
resid_panel(models_24h[["coble_m"]])
graphs_24h[["coble_m"]]
```

##### Pairwise

```{r}
nunu_pairwise(data = fDOM_24h, time_point = 6, type = "coble_m")

nunu_pairwise(data = fDOM_24h, time_point = 12, type = "coble_m")

pairwise_summary<- rbind(pairwise_summary, nunu_pairwise(data = fDOM_24h, time_point = 12, type = "coble_m"))
```

#### Sig! CobleC 
Sig ANOVA and Contrasts with and without log10 outlier detection method
```{r}
summary(models_24h[["coble_c"]])
anova(models_24h[["coble_c"]], type = 2)
emmeans(models_24h[["coble_c"]], specs = trt.vs.ctrl ~ treatment)

```

```{r,fig.width= 12, fig.height = 6}
resid_panel(models_24h[["coble_c"]])
graphs_24h[["coble_c"]]
```

##### Pairwise

```{r}
nunu_pairwise(data = fDOM_24h, time_point = 6, type = "coble_c")

nunu_pairwise(data = fDOM_24h, time_point = 12, type = "coble_c")

pairwise_summary<- rbind(pairwise_summary, 
                         nunu_pairwise(data = fDOM_24h, time_point = 12, type = "coble_c"))
```

#### Sig! M to C 
Sig ANOVA and Contrasts with and without log10 outlier detection method

```{r}
summary(models_24h[["m_to_c"]])
anova(models_24h[["m_to_c"]], type = 2)
emmeans(models_24h[["m_to_c"]], specs = trt.vs.ctrl ~ treatment)

```

```{r,fig.width= 12, fig.height = 6}
resid_panel(models_24h[["m_to_c"]])
graphs_24h[["m_to_c"]]+
  labs(y = "Marine:Colored Humics Ratio")

ggsave("MtoC_24h.png", width = 16, height = 8, path = here("outputs"))
```


##### Pairwise

```{r}
nunu_pairwise(data = fDOM_24h, time_point = 6, type = "m_to_c")

nunu_pairwise(data = fDOM_24h, time_point = 12, type = "m_to_c")

pairwise_summary<- rbind(pairwise_summary, 
                         nunu_pairwise(data = fDOM_24h, time_point = 12, type = "m_to_c"))
```

#### Sig! FI  
Sig ANOVA and Contrasts with and without log10 outlier detection method

```{r}
summary(models_24h[["fi"]])
anova(models_24h[["fi"]], type = 2)
emmeans(models_24h[["fi"]], specs = trt.vs.ctrl ~ treatment)

```

```{r,fig.width= 12, fig.height = 6}
resid_panel(models_24h[["fi"]])
graphs_24h[["fi"]]+
  labs(title = "24h Fluvic Acid")
```

##### Pairwise

```{r}
nunu_pairwise(data = fDOM_24h, time_point = 6, type = "fi")

nunu_pairwise(data = fDOM_24h, time_point = 12, type = "fi")

pairwise_summary<- rbind(pairwise_summary, 
                         nunu_pairwise(data = fDOM_24h, time_point = 12, type = "fi"))
```

#### BIX

```{r}
summary(models_24h[["bix"]])
anova(models_24h[["bix"]], type = 2)
emmeans(models_24h[["bix"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["bix"]])
graphs_24h[["bix"]]
```

##### Pairwise

```{r}
nunu_pairwise(data = fDOM_24h, time_point = 6, type = "bix")

nunu_pairwise(data = fDOM_24h, time_point = 12, type = "bix")

pairwise_summary<- rbind(pairwise_summary, 
                         nunu_pairwise(data = fDOM_24h, time_point = 12, type = "bix"))
```


#### HIX *Sig ANOVA
Sig ANOVA with log10 outlier detection method but not the contrasts
Not sig without log10

```{r}
summary(models_24h[["hix"]])
anova(models_24h[["hix"]], type = 2)
emmeans(models_24h[["hix"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["hix"]])
graphs_24h[["hix"]]
```


##### Pairwise

```{r}
HIX_pairwise <- fDOM_24h %>% 
        filter(fdom == "hix") %>% 
  filter(time_string == "18:00:00" | time_string == "06:00:00") %>% 
        ## Rename the times so they can become column names
  mutate(time_US = ifelse(time_string == "18:00:00", "T18", "T06")) %>% 
        ## Only keep the column of interest for the stats
  select(tank_number, time_US, conc_norm) %>% 
  pivot_wider(names_from = time_US, values_from = conc_norm) %>% drop_na()
```

```{r}
t.test(HIX_pairwise$T06, HIX_pairwise$T18, paired = TRUE, na.action = na.pass)
```

```{r}
#nunu_pairwise(data = fDOM_24h, time_point = 6, type = "hix")

nunu_pairwise(data = fDOM_24h, time_point = 12, type = "hix")

pairwise_summary<- rbind(pairwise_summary, 
                         nunu_pairwise(data = fDOM_24h, time_point = 12, type = "hix"))
```

### Graphs 24h fDOM 

```{r,  fig.width=6, fig.height = 10}
fDOM_24h %>%  
  filter(fdom_cat == "Humic") %>% 
  filter(time_string == "06:00:00"| time_string == "18:00:00") %>% 
    ggplot(aes(x= treatment, y= conc_norm, fill = treatment))+
    geom_violin()+
    geom_point()+
    geom_hline(yintercept=0, linetype='dashed')+
    stat_summary(color = "yellow", alpha =0.5)+
    #coord_trans(y="log10")+
  #facet_wrap(~date_string, scales = "free")+
    labs(title = paste("24h concentrations"),
       x= "Treatment",
       y = "X")+
    facet_grid(fdom~time_fac, scales = "free")+
  scale_fill_manual(breaks = c("Control", "Coral", "CCA", "Algae"), 
                       values=c("grey", "#03A9F4", "#EF5350", "#4CAF50"))+
  nunu_theme + theme_bw()+
  theme(legend.position = "none")

```



### Proteinaceous
#### CobleB

```{r}
summary(models_24h[["coble_b"]])
anova(models_24h[["coble_b"]], type = 2)
emmeans(models_24h[["coble_b"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["coble_b"]])
graphs_24h[["coble_b"]]
```

##### Pairwise

```{r}
nunu_pairwise(data = fDOM_24h, time_point = 6, type = "coble_b")

nunu_pairwise(data = fDOM_24h, time_point = 12, type = "coble_b")

pairwise_summary<- rbind(pairwise_summary, 
                         nunu_pairwise(data = fDOM_24h, time_point = 12, type = "coble_b"))
```

#### CobleT

```{r}
summary(models_24h[["coble_t"]])
anova(models_24h[["coble_t"]], type = 2)
emmeans(models_24h[["coble_t"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["coble_t"]])
graphs_24h[["coble_t"]]
```

##### Pairwise

```{r}
nunu_pairwise(data = fDOM_24h, time_point = 6, type = "coble_t")

nunu_pairwise(data = fDOM_24h, time_point = 12, type = "coble_t")

pairwise_summary<- rbind(pairwise_summary, 
                         nunu_pairwise(data = fDOM_24h, time_point = 12, type = "coble_t"))
```

#### Fpeak

```{r}
summary(models_24h[["fpeak"]])
anova(models_24h[["fpeak"]], type = 2)
emmeans(models_24h[["fpeak"]], specs = trt.vs.ctrl ~ treatment)

```

```{r}
resid_panel(models_24h[["fpeak"]])
graphs_24h[["fpeak"]]
```

##### Pairwise

```{r}
nunu_pairwise(data = fDOM_24h, time_point = 6, type = "fpeak")

nunu_pairwise(data = fDOM_24h, time_point = 12, type = "fpeak")

pairwise_summary<- rbind(pairwise_summary, 
                         nunu_pairwise(data = fDOM_24h, time_point = 12, type = "fpeak"))
```


## Heat map summary of 12PM pairwise


```{}
fdom_desc <- fDOM_24h %>% select(fdom, fdom_alt_name, fdom_cat) %>% distinct()
pairwise_summary <- left_join(pairwise_summary,fdom_desc)
```

```{, fig.width= 14, fig.height= 8}
# We are testing if the average of the treatments at 12PM is significantly different to 0
## H0 : mean of treatments is equal to 0
## HA: mean is different from 0 
fDOM_24h %>%
  filter(fdom_alt_name != "hix") %>% # remove hix bcs at 6AM, only 1 observation in CCA can't do t-test on that
  
  group_by(fdom_alt_name, treatment, time_fac) %>%
  reframe(
    t_test = list(t.test(conc_norm, mu = 0, na.action= na.pass)),
    mean = t_test[[1]]$estimate,
    p_value = t_test[[1]]$p.value,
    fdom_cat=fdom_cat) %>% distinct() %>% 
  select(-t_test) %>% 
  
  mutate(significance = ifelse(p_value <0.05, "Enrichment", "No Difference"),
         significance = ifelse(mean <0 & p_value <0.05, "Depletion", significance)#,
         #significance = as.character(significance)
         ) %>% 
    ungroup() %>% 
  
  arrange(fdom_cat, fdom_alt_name) %>%   # ensures ordering in the data frame
  mutate(fdom_alt_name = factor(fdom_alt_name, levels = c("UV-Humic", "Marine Humic","Visible-Humic", "m_to_c", "Fluvic Acid", "bix", "hix", "Tyrosine", "Tryptophan", "Phenylalanine", "lignin")),
         fdom_alt_name = fct_rev(fdom_alt_name)# lock that order in y-axis
  ) %>% 
   mutate(fdom_cat = factor(fdom_cat, levels = c("Other", "Proteinaceous", "Humic"))) %>% 
  filter(fdom_alt_name != "m_to_c",
         fdom_alt_name != "lignin") %>% 
         #fdom_alt_name != "bix",
         #fdom_alt_name != "hix") %>% 
  
  ggplot(aes(x = treatment, y= fdom_alt_name, fill = significance))+
  #geom_tile(aes(color = fdom_cat), size = 1)+
  geom_tile(color = "white", size = 1)+
  facet_wrap(.~time_fac, nrow = 1)+
  labs(title = " 24h fDOM Variations",
       x= "Treatment",
       y = "fDOM Type")+
  guides(fill=guide_legend(title="p-value interpretation"))+
  #scale_fill_distiller(palette = "RdBu", direction = 1)+
  scale_fill_manual(breaks = c("Enrichment", "No Difference", "Depletion"),
                    values = c("#43A047", "#90CAF9", "#F44336"))+
  theme_bw()+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")

ggsave("fdom_tiles_24h_t_test.png",width = 16, height = 8, path = here("outputs"))
```

```{}
pairwise_summary %>%
  mutate(fdom_cat = factor(fdom_cat, levels = c("Other", "Proteinaceous", "Humic"))) %>%
  mutate(treatment = factor(treatment, levels = c("Control", "Coral", "CCA", "Algae"))) %>% 
  mutate(significance = ifelse(p_value <0.05, "Diel Variations", "No Variations")#,
         #significance = as.character(significance)
         ) %>% 
    ungroup() %>% 
  
  arrange(fdom_cat, fdom_alt_name) %>%   # ensures ordering in the data frame
  mutate(fdom_alt_name = factor(fdom_alt_name, levels = c("UV-Humic", "Marine Humic","Visible-Humic", "m_to_c", "Fluvic Acid", "bix", "hix", "Tyrosine", "Tryptophan", "Phenylalanine", "lignin")),
         fdom_alt_name = fct_rev(fdom_alt_name)# lock that order in y-axis
  ) %>% 
    
  filter(fdom_alt_name != "m_to_c",
         fdom_alt_name != "lignin") %>%
  
  ggplot(aes(x = treatment, y= fdom_alt_name, fill = significance))+
  #geom_tile(aes(color = fdom_cat), size = 1)+
  geom_tile(color = "white", size = 1)+
  labs(title= "fDOM Diel Variations", 
       x= "Treatment",
       y = "fDOM Type")+
  scale_fill_manual(breaks = c("No Variations", "Diel Variations"),
                    values = c( "#90CAF9", "#43A047"))+
  theme_bw()+ 
  theme(plot.title = element_text(size=30, face = "bold"), 
        plot.subtitle = element_text(size=24),
        axis.title = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 16),
        strip.text.x = element_text(size = 20, face = "bold"),
        legend.title=element_text(size=16, face = "bold"),
        legend.text=element_text(size=15),
        strip.background = element_rect(fill = "white"),
        legend.position = "top")

ggsave("fdom_tiles_24h_pairwise.png",width = 10, height = 8, path = here("outputs"))
```

